### 2015-10-21 paper.tex
first commit

\documentclass[11pt]{article}

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{amsmath, amsthm, stmaryrd, scalerel}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{fontspec}
\usepackage{geometry}
\usepackage{proof}
\usepackage{supertabular}
\usepackage{unicode-math}
\usepackage{xltxtra}
\usepackage{xunicode}

\usepackage{setspace,url,hyperref}

\definecolor{dark-gray}{gray}{0.20}

% fonts
% \setromanfont[Mapping=tex-text, Ligatures=Rare, Numbers=OldStyle]{Linux Libertine O}
\setmathfont{Asana Math}

\newtheorem{thm}{Theorem}[section]
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[thm]{Definition}
\newtheorem{example}[thm]{Example}
\theoremstyle{remark}
\newtheorem{remark}[thm]{Remark}
\numberwithin{equation}{section}

\def\IModeColorName{MidnightBlue}
\def\OModeColorName{Maroon}
\newcommand\IMode[1]{{\color{\IModeColorName}{#1}}}
\newcommand\OMode[1]{{\color{\OModeColorName}{#1}}}
\newcommand\Quote[1]{{\color{Gray}{«}}#1{\color{Gray}{»}}}

\newcommand\HypJ[2]{#1\ \ (#2)}
\newcommand\GenJ[2]{\vert_{\IMode{#1}}\; #2}
\newcommand\JJ{J}
\newcommand\RSet{\mathfrak{R}}
\newcommand\JSet{\mathfrak{J}}

\newcommand\IsTac[1]{\IMode{#1}\ \textit{tactic}}

\newcommand\MkSet[1]{\left\{#1\right\}}
\newcommand\MkFam[3]{\MkSet{#1_{#2}}_{#2\in #3}}
\newcommand\Member[2]{\IMode{#1}\in\IMode{#2}}

\newcommand\MkValence[3]{#1\parallel #2\,.\, #3}
\newcommand\MkArity[2]{\left(#1\right) #2}
\newcommand\IsArity[1]{\IMode{#1}\ \textit{arity}}
\newcommand\IsValence[1]{\IMode{#1}\ \textit{valence}}
\newcommand\IsSort[1]{\IMode{#1}\ \textit{sort}}
\newcommand\Leq[2]{\IMode{#1}\leq\IMode{#2}}

\newcommand\Symbols{\mathbb{I}}
\newcommand\Finsets{\mathbb{F}}
\newcommand\Sorts{\mathcal{S}}
\newcommand\Valences{\mathcal{V}}
\newcommand\Arities{\mathcal{A}}
\newcommand\Operators{\mathcal{O}}

\newcommand\ParamsOf{\pi_\mathfrak{p}}
\newcommand\ArityOf{\pi_\mathfrak{a}}
\newcommand\Sets{\mathbf{Set}}

\newcommand\IsOperator[3]{\IMode{#1}\vdash \IMode{#2} : \IMode{#3}}

\newcommand\IsMetaCtx[1]{\IMode{#1}\ \textit{mctx}}
\newcommand\IsVarCtx[1]{\IMode{#1}\ \textit{vctx}}
\newcommand\IsSymCtx[1]{\IMode{#1}\ \textit{sctx}}

\newcommand\Lookup[3]{\IMode{#1}\ni\IMode{#2}:\OMode{#3}}

\newcommand\IsAbt[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \IMode{#3}\vdash%
  \IMode{#4}:\OMode{#5}%
}

\newcommand\MV[1]{\textsc{#1}}
\newcommand\App[3]{#1\left\{#2\right\}\left(#2;\; #3\right)}
\newcommand\Dom[1]{\left\vert #1\right\vert}
\newcommand\Fresh[2]{\IMode{#1}\notin\IMode{\Dom{#2}}}

\newcommand\SortExp{\mathsf{exp}}

\begin{document}

\title{Abstract Binding Trees}
\date{}
\author{Jonathan Sterling}
\maketitle

\section{Preliminaries}

Fix a set $\Sorts$ of \emph{sorts}. We will say $\IsSort{s}$ when
$\Member{s}{S}$. A valence $\MkValence{\vec{p}}{\vec{q}}{s}$ specifies an
expression of sort $s$ which binds symbols in $\vec{p}$ and variables in
$\vec{q}$.

\[
  \infer{
    \IsValence{\MkValence{p_0,\dots,p_m}{q_0,\dots,q_n}{s}}
  }{
    \IsSort{s} &
    \HypJ{\IsSort{p_i}}{\Leq{i}{m}} &
    \HypJ{\IsSort{q_i}}{\Leq{i}{n}} &
  }
\]

An arity $\MkArity{\vec{v}}{s}$ specifies an operator of sort $s$ with
arguments of valences $\vec{v}$. We will call the set of valences $\Valences$,
and the set of arities $\Arities$.

\[
  \infer{
    \IsArity{\MkArity{v_0,...,v_n}{s}}
  }{
    \IsSort{s} &
    \HypJ{\IsValence{v_i}}{\Leq{i}{n}}
  }
\]

Let $\Symbols$ be the infinite set of symbols; let $\Finsets$ be the free
cocartesian category over $\Symbols$. Then, fix a covariant presheaf of operators
$\Operators:\Finsets\times\Arities\to\Sets$ such that the arrows in $\Finsets$
lift to renamings of operators' parameters; via the Grothendieck construction,
we can also consider the set $\int\Operators$ of operators
$(U,\varrho,\vartheta)$ for $\Member{\vartheta}{\Operators(U,\varrho)}$.

\[
  \infer{
    \IsOperator{U}{\vartheta}{\varrho}
  }{
    \Member{\vartheta}{\Operators(U,\varrho)}
  }
\]

The judgment $\IsOperator{U}{\vartheta}{\varrho}$ supports the structural
principles of weakening and exchange, because of the functoriality of
$\Operators$.

\paragraph{Examples}

Operators are defined by specifying the fibres of $\int\Operators$ in which
they reside. For instance, consider the lambda calculus with a single sort,
$\SortExp$, about whose operators we may assert the following:

\begin{gather*}
  \IsOperator{U}{\lambda}{
    \MkArity{\MkValence{\cdot}{\SortExp}{\SortExp}}{\SortExp})
  }\\
  \IsOperator{U}{\mathsf{ap}}{
    \MkArity{
      \MkValence{\cdot}{\cdot}{\SortExp},
      \MkValence{\cdot}{\cdot}{\SortExp}
    }{\SortExp}
  }
\end{gather*}

However, consider the extension of the calculus with assignables. Then, we
shall have the following:

\begin{gather*}
  \IsOperator{U}{\mathsf{decl}}{
    \MkArity{
      \MkValence{\cdot}{\cdot}{\SortExp},
      \MkValence{\SortExp}{\cdot}{\SortExp}
    }{
      \SortExp
    }
  }\\
  \IsOperator{U,u}{\mathsf{get}\{u\}}{
    \MkArity{
      \cdot
    }{
      \SortExp
    }
  }\\
  \IsOperator{U,u}{\mathsf{set}\{u\}}{
    \MkArity{
      \MkValence{\cdot}{\cdot}{\SortExp}
    }{
      \SortExp
    }
  }
\end{gather*}

Weakening can be seen as inducing a ``degeneracy map'' on operators, whereas a
renaming $u\mapsto v$ will take $\mathsf{get}\{u\}$ to $\mathsf{get}\{v\}$.

\section{Contexts}

In general, we will have three kinds of context: metacontexts, variable
contexts, and parameter contexts. A metacontext $\Omega$ consists of bindings
of valences to metavariables; a variable context $\Gamma$ is a collection of
bindings of sorts to variables, and a parameter context $\Upsilon$ is a
collection of bindings of sorts to symbols.

\begin{gather*}
  \infer{
    \IsMetaCtx{\cdot}
  }{
  }\qquad
  \infer{
    \IsMetaCtx{\Omega,\MV{m}:v}
  }{
    \IsMetaCtx{\Omega} &
    \IsValence{v} &
    \Fresh{\MV{m}}{\Omega}
  }
\end{gather*}

\begin{gather*}
  \infer{
    \IsVarCtx{\cdot}
  }{
  }\qquad
  \infer{
    \IsVarCtx{\Gamma,x:s}
  }{
    \IsVarCtx{\Gamma} &
    \IsSort{s} &
    \Fresh{x}{\Gamma}
  }
\end{gather*}

\begin{gather*}
  \infer{
    \IsSymCtx{\cdot}
  }{
  }\qquad
  \infer{
    \IsSymCtx{\Upsilon,u:s}
  }{
    \IsVarCtx{\Upsilon} &
    \IsSort{s} &
    \Fresh{u}{\Upsilon}
  }
\end{gather*}

\section{Abstract Binding Trees}

Let the judgment $\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{s}$ presuppose
$\IsMetaCtx{\Omega}$, $\IsSymCtx\Upsilon$, $\IsVarCtx\Gamma$ and $\IsSort{s}$,
meaning that $M$ is an abstract binding tree of sort $s$, with metavariables in
$\Omega$, parameters in $\Upsilon$, and variables in $\Gamma$.

\begin{gather*}
  \infer[\textit{var}]{
    \IsAbt{\Omega}{\Upsilon}{\Gamma}{x}{s}
  }{
    \Lookup{\Gamma}{x}{s}
  }\\[6pt]
  %
  \infer[\textit{mvar}]{
    \IsAbt{\Omega}{\Upsilon}{\Gamma}{
      \App{\MV{m}}{u_0,\dots,u_m}{M_0,\dots,M_n}
    }{s}
  }{
    \begin{array}{l}
      \Lookup{\Omega}{\MV{m}}{\MkValence{p_0,\dots,p_m}{q_0,\dots,q_n}{s}}\\
      \HypJ{\Lookup{\Upsilon}{u_i}{p_i}}{\Leq{i}{m}}\\
      \HypJ{
        \IsAbt{\Omega}{\Upsilon}{\Gamma}{M_i}{q_i}
      }{
        \Leq{i}{n}
      }
    \end{array}
  }\\[6pt]
  %
  \infer[\textit{app}]{
    \IsAbt{\Omega}{\Upsilon}{\Gamma}{
      \App{\vartheta}{u_0,\dots,u_m}{M_0,\dots,M_n}
    }{s}
  }{
    \begin{array}{l}
      \IsOperator{\Dom{\Upsilon}}{\vartheta}{
        \MkValence{p_0,\dots,p_m}{q_0,\dots,q_n}{s}
      }\\
      \HypJ{\Lookup{\Upsilon}{u_i}{p_i}}{\Leq{i}{m}}\\
      \HypJ{
        \IsAbt{\Omega}{\Upsilon}{\Gamma}{M_i}{q_i}
      }{
        \Leq{i}{n}
      }
    \end{array}
  }
\end{gather*}

\end{document}

### 2015-10-21 paper.tex
wordsmithing

@@
 
-\newcommand\IsOperator[3]{\IMode{#1}\vdash \IMode{#2} : \IMode{#3}}
+\newcommand\IsOperator[3]{\IMode{#1}\vdash \IMode{#2} : \OMode{#3}}
 
@@
 
-Let $\Symbols$ be the infinite set of symbols; let $\Finsets$ be the free
+Let $\Symbols$ be an infinite set of symbols; let $\Finsets$ be the free
 cocartesian category over $\Symbols$. Then, fix a covariant presheaf of operators
@@
 
-The judgment $\IsOperator{U}{\vartheta}{\varrho}$ supports the structural
-principles of weakening and exchange, because of the functoriality of
+The judgment $\IsOperator{U}{\vartheta}{\varrho}$ enjoys the structural
+properties of weakening and exchange via the functoriality of
 $\Operators$.
@@
 they reside. For instance, consider the lambda calculus with a single sort,
-$\SortExp$, about whose operators we may assert the following:
-
+$\SortExp$; we give its signature by asserting the following about its
+operators:
 \begin{gather*}
@@
 
-However, consider the extension of the calculus with assignables. Then, we
-shall have the following:
+So far, we have made no use of symbols and parameters; however, consider the
+extension of the calculus with assignables (references):
 
@@
 
+Declaring a new assignable consists in providing an initial value, and an
+expression with a free symbol (which shall represent the assignable in scope).
 Weakening can be seen as inducing a ``degeneracy map'' on operators, whereas a
@@
 
-In general, we will have three kinds of context: metacontexts, variable
-contexts, and parameter contexts. A metacontext $\Omega$ consists of bindings
-of valences to metavariables; a variable context $\Gamma$ is a collection of
-bindings of sorts to variables, and a parameter context $\Upsilon$ is a
-collection of bindings of sorts to symbols.
+In general, we will have three kinds of context: metavariable contexts,
+variable contexts, and symbol (parameter) contexts. A metavariable context
+$\Omega$ consists of bindings of valences to metavariables; a variable context
+$\Gamma$ is a collection of bindings of sorts to variables, and a parameter
+context $\Upsilon$ is a collection of bindings of sorts to symbols.
 
@@
 meaning that $M$ is an abstract binding tree of sort $s$, with metavariables in
-$\Omega$, parameters in $\Upsilon$, and variables in $\Gamma$.
+$\Omega$, parameters in $\Upsilon$, and variables in $\Gamma$. Then, the syntax
+of abstract binding trees is inductively defined in three rules:
 

### 2015-10-21 paper.tex
notation
fix bug in App notation
Fix incorrect App rule
Fix typo (unbalanced parens)

@@
 
-\newcommand\MkValence[3]{#1\parallel #2\,.\, #3}
+\newcommand\MkValence[3]{#1;#2\,.\, #3}
+\newcommand\MkBTm[3]{#1;#2\,.\, #3}
 \newcommand\MkArity[2]{\left(#1\right) #2}
@@
 
-\newcommand\IsOperator[3]{\IMode{#1}\vdash \IMode{#2} : \OMode{#3}}
+\newcommand\IsOperator[3]{\IMode{#1}\Vdash \IMode{#2} : \OMode{#3}}
 
@@
 
+\newcommand\IsBTm[5]{%
+  \IMode{#1}\triangleright%
+  \IMode{#2}\parallel%
+  \IMode{#3}\vdash%
+  \IMode{#4}:\IMode{#5}%
+}
+
 \newcommand\MV[1]{\textsc{#1}}
-\newcommand\App[3]{#1\left\{#2\right\}\left(#2;\; #3\right)}
+\newcommand\MApp[3]{#1\left\{#2\right\}\left(#3\right)}
+\newcommand\App[2]{#1\left(#2\right)}
 \newcommand\Dom[1]{\left\vert #1\right\vert}
@@
   \IsOperator{U}{\lambda}{
-    \MkArity{\MkValence{\cdot}{\SortExp}{\SortExp}}{\SortExp})
+    \MkArity{\MkValence{\cdot}{\SortExp}{\SortExp}}{\SortExp}
   }\\
@@
   }\\
-  \IsOperator{U,u}{\mathsf{get}\{u\}}{
+  \IsOperator{U,u}{\mathsf{get}[u]}{
     \MkArity{
@@
   }\\
-  \IsOperator{U,u}{\mathsf{set}\{u\}}{
+  \IsOperator{U,u}{\mathsf{set}[u]}{
     \MkArity{
@@
 Weakening can be seen as inducing a ``degeneracy map'' on operators, whereas a
-renaming $u\mapsto v$ will take $\mathsf{get}\{u\}$ to $\mathsf{get}\{v\}$.
+renaming $u\mapsto v$ will take $\mathsf{get}[u]$ to $\mathsf{get}[v]$.
 
@@
 meaning that $M$ is an abstract binding tree of sort $s$, with metavariables in
-$\Omega$, parameters in $\Upsilon$, and variables in $\Gamma$. Then, the syntax
-of abstract binding trees is inductively defined in three rules:
+$\Omega$, parameters in $\Upsilon$, and variables in $\Gamma$. Let the judgment
+$\IsBTm{\Omega}{\Upsilon}{\Gamma}{E}{v}$ presuppose $\IsValence{v}$.  Then, the
+syntax of abstract binding trees is inductively defined in four rules:
 
@@
     \IsAbt{\Omega}{\Upsilon}{\Gamma}{
-      \App{\MV{m}}{u_0,\dots,u_m}{M_0,\dots,M_n}
+      \MApp{\MV{m}}{u_0,\dots,u_m}{M_0,\dots,M_n}
     }{s}
@@
     \IsAbt{\Omega}{\Upsilon}{\Gamma}{
-      \App{\vartheta}{u_0,\dots,u_m}{M_0,\dots,M_n}
+      \App{\vartheta}{E_0,\dots,E_n}
     }{s}
@@
       \IsOperator{\Dom{\Upsilon}}{\vartheta}{
-        \MkValence{p_0,\dots,p_m}{q_0,\dots,q_n}{s}
+        v_1,\dots,v_n
       }\\
-      \HypJ{\Lookup{\Upsilon}{u_i}{p_i}}{\Leq{i}{m}}\\
       \HypJ{
-        \IsAbt{\Omega}{\Upsilon}{\Gamma}{M_i}{q_i}
+        \IsBTm{\Omega}{\Upsilon}{\Gamma}{E_i}{q_i}
       }{
@@
     \end{array}
+  }\\[6pt]
+  \infer[\textit{abs}]{
+    \IsBTm{\Omega}{\Upsilon}{\Gamma}{
+      (\MkBTm{\vec{u}}{\vec{x}}{E})
+    }{
+      (\MkValence{\vec{p}}{\vec{q}}{s})
+    }
+  }{
+    \IsAbt{\Omega}{\Upsilon, \vec{u}:\vec{p}}{\Gamma,\vec{x}:{\vec{q}}}{
+      E
+    }{
+      s
+    }
   }

### 2015-10-21 paper.tex
Fix math font
Fix forgotten rule
Change notation for valences and bound terms
Add todos
correct the size of some delimiters; not sure if this is an improvement
Try and resolve #6

@@
 % \setromanfont[Mapping=tex-text, Ligatures=Rare, Numbers=OldStyle]{Linux Libertine O}
-\setmathfont{Asana Math}
+\setmathfont{Asana-Math.otf}
 
@@
 
-\newcommand\MkValence[3]{#1;#2\,.\, #3}
-\newcommand\MkBTm[3]{#1;#2\,.\, #3}
+\newcommand\MkValence[3]{\left\{#1\right\}\left[#2\right]\,.\, #3}
+\newcommand\MkBTm[3]{\left\{#1\right\}\left[#2\right]\,.\, #3}
 \newcommand\MkArity[2]{\left(#1\right) #2}
@@
 Let $\Symbols$ be an infinite set of symbols; let $\Finsets$ be the free
-cocartesian category over $\Symbols$. Then, fix a covariant presheaf of operators
+cocartesian category over $\Symbols\times\Sorts$. Then, fix a covariant presheaf of operators
 $\Operators:\Finsets\times\Arities\to\Sets$ such that the arrows in $\Finsets$
@@
 we can also consider the set $\int\Operators$ of operators
-$(U,\varrho,\vartheta)$ for $\Member{\vartheta}{\Operators(U,\varrho)}$.
+$(U,\varrho,\vartheta)$ for $\Member{\vartheta}{\Operators(\Upsilon,\varrho)}$.
 
@@
   \infer{
-    \IsOperator{U}{\vartheta}{\varrho}
+    \IsOperator{\Upsilon}{\vartheta}{\varrho}
   }{
-    \Member{\vartheta}{\Operators(U,\varrho)}
+    \Member{\vartheta}{\Operators(\Upsilon,\varrho)}
   }
@@
 
-The judgment $\IsOperator{U}{\vartheta}{\varrho}$ enjoys the structural
+The judgment $\IsOperator{\Upsilon}{\vartheta}{\varrho}$ enjoys the structural
 properties of weakening and exchange via the functoriality of
@@
 \begin{gather*}
-  \IsOperator{U}{\lambda}{
+  \IsOperator{\Upsilon}{\lambda}{
     \MkArity{\MkValence{\cdot}{\SortExp}{\SortExp}}{\SortExp}
   }\\
-  \IsOperator{U}{\mathsf{ap}}{
+  \IsOperator{\Upsilon}{\mathsf{ap}}{
     \MkArity{
@@
 \begin{gather*}
-  \IsOperator{U}{\mathsf{decl}}{
+  \IsOperator{\Upsilon}{\mathsf{decl}}{
     \MkArity{
@@
   }\\
-  \IsOperator{U,u}{\mathsf{get}[u]}{
+  \IsOperator{\Upsilon,u:\SortExp}{\mathsf{get}[u]}{
     \MkArity{
@@
   }\\
-  \IsOperator{U,u}{\mathsf{set}[u]}{
+  \IsOperator{\Upsilon,u:\SortExp}{\mathsf{set}[u]}{
     \MkArity{
@@
     \begin{array}{l}
-      \IsOperator{\Dom{\Upsilon}}{\vartheta}{
+      \IsOperator{\Upsilon}{\vartheta}{
         v_1,\dots,v_n

### 2015-10-21 paper.tex
Fix one I missed

@@
 Let $\Symbols$ be an infinite set of symbols; let $\Finsets$ be the free
-cocartesian category over $\Symbols\times\Sorts$. Then, fix a covariant presheaf of operators
-$\Operators:\Finsets\times\Arities\to\Sets$ such that the arrows in $\Finsets$
-lift to renamings of operators' parameters; via the Grothendieck construction,
-we can also consider the set $\int\Operators$ of operators
-$(U,\varrho,\vartheta)$ for $\Member{\vartheta}{\Operators(\Upsilon,\varrho)}$.
+cocartesian category over $\Symbols\times\Sorts$. Then, fix a covariant
+presheaf of operators $\Operators:\Finsets\times\Arities\to\Sets$ such that the
+arrows in $\Finsets$ lift to renamings of operators' parameters; via the
+Grothendieck construction, we can also consider the set $\int\Operators$ of
+operators $(\Upsilon,\varrho,\vartheta)$ for
+$\Member{\vartheta}{\Operators(\Upsilon,\varrho)}$.
 

### 2015-10-21 paper.tex
Use \oint for grothendieck construction

--- 2015-10-21 paper.tex
@@
 
+\newcommand\Grothendieck[1]{\oint #1}
+
 \newcommand\IsAbt[5]{%
@@
 Let $\Symbols$ be an infinite set of symbols; let $\Finsets$ be the free
-cocartesian category over $\Symbols\times\Sorts$. Then, fix a covariant presheaf of operators
-$\Operators:\Finsets\times\Arities\to\Sets$ such that the arrows in $\Finsets$
-lift to renamings of operators' parameters; via the Grothendieck construction,
-we can also consider the set $\int\Operators$ of operators
-$(U,\varrho,\vartheta)$ for $\Member{\vartheta}{\Operators(\Upsilon,\varrho)}$.
+cocartesian category over $\Symbols\times\Sorts$. Then, fix a covariant
+presheaf of operators $\Operators:\Finsets\times\Arities\to\Sets$ such that the
+arrows in $\Finsets$ lift to renamings of operators' parameters; via the
+Grothendieck construction, we can also consider the set $\Grothendieck\Operators$ of
+operators $(\Upsilon,\varrho,\vartheta)$ for
+$\Member{\vartheta}{\Operators(\Upsilon,\varrho)}$.
 
@@
 
-Operators are defined by specifying the fibres of $\int\Operators$ in which
+Operators are defined by specifying the fibres of $\Grothendieck\Operators$ in which
 they reside. For instance, consider the lambda calculus with a single sort,

### 2015-10-22 paper.tex
implement variable substitution (close #2)
Fix the category of symbol contexts

@@
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-\usepackage{amsmath, amsthm, stmaryrd, scalerel}
+\usepackage{amsmath, amsthm, stmaryrd, scalerel, amssymb}
 \usepackage{mathtools}
@@
 
-\newcommand\MkValence[3]{\left\{#1\right\}\left[#2\right]\,.\, #3}
-\newcommand\MkBTm[3]{\left\{#1\right\}\left[#2\right]\,.\, #3}
+\newcommand\MkValence[3]{\left\{#1\right\}\left[#2\right].\, #3}
+\newcommand\MkBTm[3]{\left\{#1\right\}\left[#2\right].\, #3}
 \newcommand\MkArity[2]{\left(#1\right) #2}
@@
 \newcommand\Sets{\mathbf{Set}}
+\newcommand\SCtx{\mathbf{SCtx}}
 
@@
 \newcommand\Grothendieck[1]{\oint #1}
+\newcommand\Comma[2]{#1 \downarrow #2}
 
@@
 \newcommand\Dom[1]{\left\vert #1\right\vert}
-\newcommand\Fresh[2]{\IMode{#1}\notin\IMode{\Dom{#2}}}
+\newcommand\FV[1]{\mathbf{FV}\left(#1\right)}
+\newcommand\FS[1]{\mathbf{FS}\left(#1\right)}
+\newcommand\NotIn[2]{\IMode{#1}\notin\IMode{#2}}
+\newcommand\SortExp{\mathsf{exp}}
 
-\newcommand\SortExp{\mathsf{exp}}
+\newcommand\Hom[3]{#1\left[#2,#3\right]}
+
+\newcommand\Subst[3]{
+  \left[#1\, /\, #2\right]#3%
+}
+
+\newcommand\IsRenaming[3]{
+  \IMode{#2}\upharpoonright\IMode{#1}%
+  \leadsto\OMode{#3}%
+}
+
+\newcommand\IsSubst[4]{%
+  \Subst{\IMode{#1}}{\IMode{#2}}{\IMode{#3}}%
+  \leadsto\OMode{#4}%
+}
+
+\newcommand\IsEq[2]{\IMode{#1} = \IMode{#2}}
+\newcommand\IsApart[2]{\IMode{#1} \mathrel{\#} \IMode{#2}}
+
+\newcommand\Define[2]{\IMode{#1}\triangleq\OMode{#2}}
 
@@
 
-Let $\Symbols$ be an infinite set of symbols; let $\Finsets$ be the free
-cocartesian category over $\Symbols\times\Sorts$. Then, fix a covariant
-presheaf of operators $\Operators:\Finsets\times\Arities\to\Sets$ such that the
-arrows in $\Finsets$ lift to renamings of operators' parameters; via the
-Grothendieck construction, we can also consider the set $\Grothendieck\Operators$ of
-operators $(\Upsilon,\varrho,\vartheta)$ for
+Let $\Symbols$ be an infinite set of symbols. Let $\Finsets$ be the category of
+finite subsets of $\Symbols$ and their injective maps; then the comma construction
+$\Define{\SCtx}{\Comma{\Finsets}{\Sorts}}$, with $\Sorts$ regarded as a discrete category, is
+the category of contexts of symbols, whose objects are finite sets of symbols
+$U$ and sort-assignments $\mathfrak{s}:U\to\Sorts$, and whose morphisms are
+sort-preserving renamings; we will write $\Upsilon$ for a symbol context
+$(U,\mathfrak{s})$.
+
+Then, fix a covariant presheaf of operators
+$\Operators:\SCtx\times\Arities\to\Sets$ such that the arrows in $\SCtx$ lift
+to renamings of operators' parameters; via the Grothendieck construction, we
+can also consider the set $\Grothendieck\Operators$ of operators
+$(\Upsilon,\varrho,\vartheta)$ for
 $\Member{\vartheta}{\Operators(\Upsilon,\varrho)}$.
@@
     \IsValence{v} &
-    \Fresh{\MV{m}}{\Omega}
+    \NotIn{\MV{m}}{\Dom\Omega}
   }
@@
     \IsSort{s} &
-    \Fresh{x}{\Gamma}
+    \NotIn{x}{\Dom\Gamma}
   }
@@
     \IsSort{s} &
-    \Fresh{u}{\Upsilon}
+    \NotIn{u}{\Dom\Upsilon}
   }
@@
     \IsBTm{\Omega}{\Upsilon}{\Gamma}{
-      (\MkBTm{\vec{u}}{\vec{x}}{E})
+      \left(\MkBTm{\vec{u}}{\vec{x}}{M}\right)
     }{
-      (\MkValence{\vec{p}}{\vec{q}}{s})
+      \left(\MkValence{\vec{p}}{\vec{q}}{s}\right)
     }
@@
     \IsAbt{\Omega}{\Upsilon, \vec{u}:\vec{p}}{\Gamma,\vec{x}:{\vec{q}}}{
-      E
+      M
     }{
@@
 
+Abstract binding trees are identified up to $\alpha$-equivalence.
+
+%%% \subsection{Renaming of Symbols}
+%%%
+%%% The renaming of symbols in abstract binding trees is defined inductively by the
+%%% judgment $\IsRenaming{\psi}{M}{M'}$, where
+%%% $\Member{\psi}{\Hom{\SCtx}{\FS{M}}{-}}$
+%%%
+%%% \begin{gather*}
+%%% \end{gather*}
+
+\subsection{Substitution of Variables}
+
+Variable substitution in abstract binding trees is defined inductively by a
+pair of judgments, $\IsSubst{N}{x}{M}{M'}$ and $\IsSubst{N}{x}{E}{E'}$:
+
+\begin{gather*}
+  \infer{
+    \IsSubst{N}{x}{y}{N}
+  }{
+    \IsEq{x}{y}
+  }\qquad
+  \infer{
+    \IsSubst{N}{x}{y}{y}
+  }{
+    \IsApart{x}{y}
+  }\\[6pt]
+  \infer{
+    \IsSubst{N}{x}{
+      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
+    }{
+      \MApp{\MV{m}}{\vec{u}}{M'_0,\dots,M'_n}
+    }
+  }{
+    \HypJ{
+      \IsSubst{N}{x}{M_i}{M'_i}
+    }{
+      \Leq{i}{n}
+    }
+  }\\[6pt]
+  \infer{
+    \IsSubst{N}{x}{
+      \App{\vartheta}{E_0,\dots,E_n}
+    }{
+      \App{\vartheta}{E'_0,\dots,E'_n}
+    }
+  }{
+    \HypJ{
+      \IsSubst{N}{x}{E_i}{E'_i}
+    }{
+      \Leq{i}{n}
+    }
+  }
+\end{gather*}
+\begin{gather*}
+  \infer{
+    \IsSubst{N}{x}{
+      \MkBTm{\vec{u}}{\vec{y}}{M}
+    }{
+      \MkBTm{\vec{u}}{\vec{y}}{M'}
+    }
+  }{
+    \NotIn{x}{\vec{y}} &
+    \IsApart{\vec{y}}{\FV{N}} &
+    \IsSubst{N}{x}{M}{M'}
+  }\qquad
+  \infer{
+    \IsSubst{N}{x}{
+      \MkBTm{\vec{u}}{\vec{y}}{M}
+    }{
+      \MkBTm{\vec{u}}{\vec{y}}{M}
+    }
+  }{
+    \Member{x}{\vec{y}} &
+    \IsApart{\vec{y}}{\FV{N}}
+  }
+\end{gather*}
+
+Because terms are identified up to $\alpha$-equivalence, the variable
+substitution judgment is functional in its inputs, and so we are justified in
+writing $\Subst{N}{x}{M}$ for $M'$ when $\IsSubst{N}{x}{M}{M'}$. We write
+$\Subst{\vec{N}}{\vec{x}}{M}$ for the simultaneous substitution of $\vec{N}$
+for $\vec{x}$ in $M$.
+
+\subsection{Substitution of Metavariables}
+
+Metavariable substitution is defined inductively by the judgment
+$\IsSubst{E}{\MV{m}}{M}{M'}$:
+
+\begin{gather*}
+  \infer{
+    \IsSubst{E}{\MV{m}}{x}{x}
+  }{
+  }\\[6pt]
+  \infer{
+    \IsSubst{E}{\MV{m}}{
+      \MApp{\MV{n}}{\vec{u}}{M_0,\dots,M_n}
+    }{
+      \MApp{\MV{n}}{\vec{u}}{M'_0,\dots,M'_n}
+    }
+  }{
+    \IsApart{\MV{m}}{\MV{n}} &
+    \HypJ{
+      \IsSubst{E}{\MV{n}}{M_i}{M'_i}
+    }{
+      \Leq{i}{n}
+    }
+  }\\[6pt]
+  %%% \infer{
+  %%%   \IsSubst{
+  %%%     \MkBTm{\vec{u}}{x_0,\dots,x_n}{N}
+  %%%   }{\MV{m}}{
+  %%%     \MApp{\MV{n}}{\vec{v}}{M_0,\dots,M_n}
+  %%%   }{
+  %%%     %TODO!!!
+  %%%     %\Rename{\vec{v}}{\vec{u}}{
+  %%%     \Subst{\vec{M}}{\vec{x}}{N}
+  %%%     % }
+  %%%   }
+  %%% }{
+  %%%   \IsEq{\MV{m}}{\MV{n}}
+  %%% }\\[6pt]
+  %%% \infer{
+  %%%   \IsSubst{E}{\MV{m}}{
+  %%%     \App{\vartheta}{F_0,\dots,F_n}
+  %%%   }{
+  %%%   }
+  %%% }{
+  %%% }
+\end{gather*}
+
 \end{document}

### 2015-10-22 paper.tex
Integrate @freebroccolo's clarifications, thanks! close #9

@@
 
-\newcommand\Grothendieck[1]{\oint #1}
+\newcommand\Grothendieck[2]{\oint^{#1} #2#1}
+\newcommand\Pair[2]{\langle #1, #2 \rangle}
+
 \newcommand\Comma[2]{#1 \downarrow #2}
@@
 
-Then, fix a covariant presheaf of operators
-$\Operators:\SCtx\times\Arities\to\Sets$ such that the arrows in $\SCtx$ lift
-to renamings of operators' parameters; via the Grothendieck construction, we
-can also consider the set $\Grothendieck\Operators$ of operators
-$(\Upsilon,\varrho,\vartheta)$ for
-$\Member{\vartheta}{\Operators(\Upsilon,\varrho)}$.
+Via the Grothendieck construction\footnote{Category of elements.}
+$\Grothendieck{}{}$ on the operator presheaf $\Operators$ we also have a
+category with objects
+$\Member{\Pair{\Pair{\Upsilon}{\varrho}}{\vartheta}}{\Grothendieck{}{\Operators}}$
+for $\Member{\vartheta}{\Operators\Pair{\Upsilon}{\varrho}}$ and morphisms
+$\Hom{\Grothendieck{}{\Operators}}{\Pair{\Pair{\Upsilon}{\varrho}}{\vartheta}}{\Pair{\Pair{\Upsilon'}{\varrho'}}{\vartheta'}}$
+for $f : \Pair{\Upsilon}{\varrho}\to\Pair{\Upsilon'}{\varrho'}$ such that
+$\Member{\vartheta}{f^{*}\Operators} = \vartheta'$.
 
@@
   }{
-    \Member{\vartheta}{\Operators(\Upsilon,\varrho)}
+    \Member{\vartheta}{\Operators\langle\Upsilon,\varrho\rangle}
   }
@@
 
-Operators are defined by specifying the fibres of $\Grothendieck\Operators$ in which
+Operators are defined by specifying the fibres of $\Grothendieck{}{\Operators}$ in which
 they reside. For instance, consider the lambda calculus with a single sort,

### 2015-10-22 paper.tex
Add @freebroccolo as author
Add missing pair

@@
 \date{}
-\author{Jonathan Sterling}
+\author{Jon Sterling and Darin Morrison}
 \maketitle
@@
 for $f : \Pair{\Upsilon}{\varrho}\to\Pair{\Upsilon'}{\varrho'}$ such that
-$\Member{\vartheta}{f^{*}\Operators} = \vartheta'$.
+$\Member{\vartheta}{f^{*}\Operators}\Pair{\Upsilon'}{\varrho'} = \vartheta'$.
 
@@
 
-%%% \subsection{Renaming of Symbols}
-%%%
-%%% The renaming of symbols in abstract binding trees is defined inductively by the
-%%% judgment $\IsRenaming{\psi}{M}{M'}$, where
-%%% $\Member{\psi}{\Hom{\SCtx}{\FS{M}}{-}}$
-%%%
-%%% \begin{gather*}
-%%% \end{gather*}
-
 \subsection{Substitution of Variables}

### 2015-10-22 paper.tex
restore sentence I accidentally deleted
Fix morphism part for Grothendieck construction

@@
 \newcommand\Pair[2]{\langle #1, #2 \rangle}
+\newcommand\Restriction[1]{#1^{*}}
 
@@
 
-Via the Grothendieck construction\footnote{Category of elements.}
-$\Grothendieck{}{}$ on the operator presheaf $\Operators$ we also have a
-category with objects
+Then, fix a covariant presheaf of operators
+$\Operators:\SCtx\times\Arities\to\Sets$ such that the arrows in $\SCtx$ lift to
+renamings of operators' parameters. Via the Grothendieck
+construction\footnote{Category of elements.} $\Grothendieck{}{}$ on the operator
+presheaf $\Operators$ we also have a category with objects
 $\Member{\Pair{\Pair{\Upsilon}{\varrho}}{\vartheta}}{\Grothendieck{}{\Operators}}$
@@
 for $f : \Pair{\Upsilon}{\varrho}\to\Pair{\Upsilon'}{\varrho'}$ such that
-$\Member{\vartheta}{f^{*}\Operators}\Pair{\Upsilon'}{\varrho'} = \vartheta'$.
+$\Member{\Restriction{f}\vartheta =
+  \vartheta'}{\Operators\Pair{\Upsilon'}{\varrho'}}$.
 

### 2015-10-22 paper.tex
formatting

@@
 \newcommand\Member[2]{\IMode{#1}\in\IMode{#2}}
+\newcommand\EqMember[3]{\IMode{#1}=\IMode{#2}\in\IMode{#3}}
 
@@
 for $f : \Pair{\Upsilon}{\varrho}\to\Pair{\Upsilon'}{\varrho'}$ such that
-$\Member{\Restriction{f}\vartheta =
-  \vartheta'}{\Operators\Pair{\Upsilon'}{\varrho'}}$.
+$\EqMember{\Restriction{f}\vartheta}{\vartheta'}{\Operators\Pair{\Upsilon'}{\varrho'}}$.
 

### 2015-10-22 paper.tex
Elaborate a bit on the Grothendieck construction

@@
 \newcommand\ArityOf{\pi_\mathfrak{a}}
+\newcommand\Cats{\mathbf{Cat}}
 \newcommand\Sets{\mathbf{Set}}
@@
 renamings of operators' parameters. Via the Grothendieck
-construction\footnote{Category of elements.} $\Grothendieck{}{}$ on the operator
-presheaf $\Operators$ we also have a category with objects
+construction\footnote{In this case, $C\Grothendieck{}{\Psi}$ represents the
+  category of elements of a copresheaf $\Psi : C\to\Sets$ but we keep the $C$
+  implicit and simply refer to it as the Grothendieck construction.
+  Alternatively, this construction can be understood as a coend
+  $C\Grothendieck{}{\Psi} \cong \int^{c \in C}c / C \otimes |\Psi(c)|$.}
+$\Grothendieck{}{(-)} : \Sets^{C}\to\Cats$ on operators we have a category of
+objects
 $\Member{\Pair{\Pair{\Upsilon}{\varrho}}{\vartheta}}{\Grothendieck{}{\Operators}}$

### 2015-10-22 paper.tex
Use A_{≡} for discrete cat on set A

@@
 \newcommand\Cats{\mathbf{Cat}}
+\newcommand\Discrete[1]{#1_{\equiv}}
 \newcommand\Sets{\mathbf{Set}}
@@
 Let $\Symbols$ be an infinite set of symbols. Let $\Finsets$ be the category of
-finite subsets of $\Symbols$ and their injective maps; then the comma construction
-$\Define{\SCtx}{\Comma{\Finsets}{\Sorts}}$, with $\Sorts$ regarded as a discrete category, is
-the category of contexts of symbols, whose objects are finite sets of symbols
-$U$ and sort-assignments $\mathfrak{s}:U\to\Sorts$, and whose morphisms are
+finite subsets of $\Symbols$ and their injective maps; then the comma
+construction $\Define{\SCtx}{\Comma{\Finsets}{\Discrete{\Sorts}}}$, with
+$\Discrete{\Sorts}$ the discrete category on the set $\mathcal{S}$, is the
+category of contexts of symbols, whose objects are finite sets of symbols $U$
+and sort-assignments $\mathfrak{s}:U\to\Sorts$, and whose morphisms are
 sort-preserving renamings; we will write $\Upsilon$ for a symbol context
@@
   Alternatively, this construction can be understood as a coend
-  $C\Grothendieck{}{\Psi} \cong \int^{c \in C}c / C \otimes |\Psi(c)|$.}
+  $C\Grothendieck{}{\Psi} \cong \int^{c \in C}c / C \otimes \Discrete{\Psi(c)}$.}
 $\Grothendieck{}{(-)} : \Sets^{C}\to\Cats$ on operators we have a category of

### 2015-10-22 paper.tex

@@
 \usepackage{supertabular}
+\usepackage{tikz-cd}
 \usepackage{unicode-math}
@@
 \newcommand\Restriction[1]{#1^{*}}
+\newcommand\Singleton{\{\bullet\}}
 
@@
   Alternatively, this construction can be understood as a coend
-  $C\Grothendieck{}{\Psi} \cong \int^{c \in C}c / C \otimes \Discrete{\Psi(c)}$.}
-$\Grothendieck{}{(-)} : \Sets^{C}\to\Cats$ on operators we have a category of
-objects
+  $C\Grothendieck{}{\Psi} \cong \int^{c \in C}c / C \otimes
+  \Discrete{\Psi(c)}$.} $\Grothendieck{}{(-)} : \Sets^{C}\to\Cats$ on operators
+we have a category of objects
 $\Member{\Pair{\Pair{\Upsilon}{\varrho}}{\vartheta}}{\Grothendieck{}{\Operators}}$
@@
 $\EqMember{\Restriction{f}\vartheta}{\vartheta'}{\Operators\Pair{\Upsilon'}{\varrho'}}$.
+Equivalently, $\Grothendieck{}{\Operators}$ is the pullback of $\Operators$
+along the universal $\Sets$-bundle where $\pi_{\Operators}$ is a discrete
+Grothendieck fibration and $ap_{\Singleton}$ is the forgetful functor from pointed sets:
+
+\[
+\begin{tikzcd}[ampersand replacement = \&, sep = large]
+  {\Grothendieck{}{\Operators}}
+    \arrow[r, "{\pi_{\Singleton}}"]
+    \arrow[d, "{\pi_{\Operators}}" left]
+  \& {\Singleton / \Sets}
+    \arrow[d, "{ap_{\Singleton}}"] \\
+  {\SCtx\times\Arities}
+     \arrow[r, "{\Operators}" below]
+  \& {\Sets} \\
+\end{tikzcd}
+\]
 

### 2015-10-22 paper.tex
clarification

@@
 
-Then, fix a covariant presheaf of operators
+Then, fix a covariant presheaf (copresheaf) of operators
 $\Operators:\SCtx\times\Arities\to\Sets$ such that the arrows in $\SCtx$ lift to
@@
 
-Operators are defined by specifying the fibres of $\Grothendieck{}{\Operators}$ in which
+Operators are defined by specifying the fibers of $\pi_{\Operators}$ in which
 they reside. For instance, consider the lambda calculus with a single sort,

### 2015-10-22 paper.tex
Use existing macro
fibration ⇒ opfibration

@@
 \newcommand\Restriction[1]{#1^{*}}
-\newcommand\Singleton{\{\bullet\}}
+\newcommand\Singleton{\MkSet{\bullet}}
 
@@
 along the universal $\Sets$-bundle where $\pi_{\Operators}$ is a discrete
-Grothendieck fibration and $ap_{\Singleton}$ is the forgetful functor from pointed sets:
+Grothendieck opfibration and $ap_{\Singleton}$ is the forgetful functor from pointed sets:
 

### 2015-10-22 paper.tex
Rename π{●} to π!
Just use ● instead of {●}

--- 2015-10-22 paper.tex
@@
 \newcommand\Restriction[1]{#1^{*}}
-\newcommand\Singleton{\{\bullet\}}
+\newcommand\Singleton{\bullet}
 
 \newcommand\Comma[2]{#1 \downarrow #2}
+\newcommand\Coslice[2]{#1\, /\, #2}
+\newcommand\Slice[2]{#1\, /\, #2}
 
@@
   Alternatively, this construction can be understood as a coend
-  $C\Grothendieck{}{\Psi} \cong \int^{c \in C}c / C \otimes
+  $C\Grothendieck{}{\Psi} \cong \int^{c \in C}\Coslice{c}{C} \otimes
   \Discrete{\Psi(c)}$.} $\Grothendieck{}{(-)} : \Sets^{C}\to\Cats$ on operators
@@
 along the universal $\Sets$-bundle where $\pi_{\Operators}$ is a discrete
-Grothendieck fibration and $ap_{\Singleton}$ is the forgetful functor from pointed sets:
+Grothendieck opfibration and $ap_{\bullet}$ is the forgetful functor from
+pointed sets:
 
@@
   {\Grothendieck{}{\Operators}}
-    \arrow[r, "{\pi_{\Singleton}}"]
+    \arrow[r, "{\pi_{!}}"]
     \arrow[d, "{\pi_{\Operators}}" left]
-  \& {\Singleton / \Sets}
+  \& {\Coslice{\Singleton}{\Sets}}
     \arrow[d, "{ap_{\Singleton}}"] \\

### 2015-10-23 paper.tex
notation, macros

@@
 \newcommand\EqMember[3]{\IMode{#1}=\IMode{#2}\in\IMode{#3}}
+\newcommand\Of[2]{\IMode{#1} : \IMode{#2}}
 
@@
 category of contexts of symbols, whose objects are finite sets of symbols $U$
-and sort-assignments $\mathfrak{s}:U\to\Sorts$, and whose morphisms are
+and sort-assignments $\Of{\mathfrak{s}}{U\to\Sorts}$, and whose morphisms are
 sort-preserving renamings; we will write $\Upsilon$ for a symbol context
@@
 Then, fix a covariant presheaf (copresheaf) of operators
-$\Operators:\SCtx\times\Arities\to\Sets$ such that the arrows in $\SCtx$ lift to
-renamings of operators' parameters. Via the Grothendieck
+$\Of{\Operators}{\SCtx\times\Discrete{\Arities}\to\Sets}$ such that the arrows
+in $\SCtx$ lift to renamings of operators' parameters. Via the Grothendieck
 construction\footnote{In this case, $C\Grothendieck{}{\Psi}$ represents the
-  category of elements of a copresheaf $\Psi : C\to\Sets$ but we keep the $C$
-  implicit and simply refer to it as the Grothendieck construction.
+  category of elements of a copresheaf $\Of{\Psi}{C\to\Sets}$ but we keep the
+  $C$ implicit and simply refer to it as the Grothendieck construction.
   Alternatively, this construction can be understood as a coend
-  $C\Grothendieck{}{\Psi} \cong \int^{c \in C}\Coslice{c}{C} \otimes
-  \Discrete{\Psi(c)}$.} $\Grothendieck{}{(-)} : \Sets^{C}\to\Cats$ on operators
-we have a category of objects
+$\IMode{C\Grothendieck{}{\Psi}}\cong \IMode{\int^{c \in C}\Coslice{c}{C}
+\otimes \Discrete{\Psi(c)}}$.} $\Of{\Grothendieck{}{(-)}}{\Sets^{C}\to\Cats}$
+on operators we have a category of objects
 $\Member{\Pair{\Pair{\Upsilon}{\varrho}}{\vartheta}}{\Grothendieck{}{\Operators}}$
@@
 $\Hom{\Grothendieck{}{\Operators}}{\Pair{\Pair{\Upsilon}{\varrho}}{\vartheta}}{\Pair{\Pair{\Upsilon'}{\varrho'}}{\vartheta'}}$
-for $f : \Pair{\Upsilon}{\varrho}\to\Pair{\Upsilon'}{\varrho'}$ such that
+for $\Of{f}{\Pair{\Upsilon}{\varrho}\to\Pair{\Upsilon'}{\varrho'}}$ such that
 $\EqMember{\Restriction{f}\vartheta}{\vartheta'}{\Operators\Pair{\Upsilon'}{\varrho'}}$.

### 2015-10-23 paper.tex
Notation

@@
 \newcommand\Singleton{\bullet}
+\newcommand\Forget[1]{|\, #1\, |}
 
@@
 along the universal $\Sets$-bundle where $\pi_{\Operators}$ is a discrete
-Grothendieck opfibration and $ap_{\bullet}$ is the forgetful functor from
-pointed sets:
+Grothendieck opfibration and $\Forget{-}_{\Singleton}$ is the forgetful functor
+from pointed sets:
 
@@
   {\Grothendieck{}{\Operators}}
-    \arrow[r, "{\pi_{!}}"]
+    \arrow[r, "{\pi_{\in}}"]
     \arrow[d, "{\pi_{\Operators}}" left]
   \& {\Coslice{\Singleton}{\Sets}}
-    \arrow[d, "{ap_{\Singleton}}"] \\
+    \arrow[d, "{\Forget{-}_{\Singleton}}"] \\
   {\SCtx\times\Arities}

### 2015-10-24 paper.tex
Use discrete for arities in pullback diagram
start talking a little bit about renaming (#10)
- [notation] Let 'a' range over arities rather than 'ρ'

@@
 on operators we have a category of objects
-$\Member{\Pair{\Pair{\Upsilon}{\varrho}}{\vartheta}}{\Grothendieck{}{\Operators}}$
-for $\Member{\vartheta}{\Operators\Pair{\Upsilon}{\varrho}}$ and morphisms
-$\Hom{\Grothendieck{}{\Operators}}{\Pair{\Pair{\Upsilon}{\varrho}}{\vartheta}}{\Pair{\Pair{\Upsilon'}{\varrho'}}{\vartheta'}}$
-for $\Of{f}{\Pair{\Upsilon}{\varrho}\to\Pair{\Upsilon'}{\varrho'}}$ such that
-$\EqMember{\Restriction{f}\vartheta}{\vartheta'}{\Operators\Pair{\Upsilon'}{\varrho'}}$.
+$\Member{\Pair{\Pair{\Upsilon}{a}}{\vartheta}}{\Grothendieck{}{\Operators}}$
+for $\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$ and morphisms
+$\Hom{\Grothendieck{}{\Operators}}{\Pair{\Pair{\Upsilon}{a}}{\vartheta}}{\Pair{\Pair{\Upsilon'}{a'}}{\vartheta'}}$
+for $\Of{f}{\Pair{\Upsilon}{a}\to\Pair{\Upsilon'}{a'}}$ such that
+$\EqMember{\Restriction{f}\vartheta}{\vartheta'}{\Operators\Pair{\Upsilon'}{a'}}$.
 Equivalently, $\Grothendieck{}{\Operators}$ is the pullback of $\Operators$
@@
     \arrow[d, "{\Forget{-}_{\Singleton}}"] \\
-  {\SCtx\times\Arities}
+  {\SCtx\times\Discrete{\Arities}}
      \arrow[r, "{\Operators}" below]
@@
   \infer{
-    \IsOperator{\Upsilon}{\vartheta}{\varrho}
+    \IsOperator{\Upsilon}{\vartheta}{a}
   }{
-    \Member{\vartheta}{\Operators\langle\Upsilon,\varrho\rangle}
+    \Member{\vartheta}{\Operators\langle\Upsilon,a\rangle}
   }
@@
 
-The judgment $\IsOperator{\Upsilon}{\vartheta}{\varrho}$ enjoys the structural
+The judgment $\IsOperator{\Upsilon}{\vartheta}{a}$ enjoys the structural
 properties of weakening and exchange via the functoriality of
@@
 expression with a free symbol (which shall represent the assignable in scope).
-Weakening can be seen as inducing a ``degeneracy map'' on operators, whereas a
-renaming $u\mapsto v$ will take $\mathsf{get}[u]$ to $\mathsf{get}[v]$.
+Note that the functoriality of $\Operators$ guarantees for any renaming
+$\Of{\varrho}{\Upsilon\to\Upsilon'}$ a restriction map $\Restriction{\varrho}$
+such that $\IsOperator{\Upsilon'}{\Restriction{\varrho}(\vartheta)}{a}$ when
+$\IsOperator{\Upsilon}{\vartheta}{a}$. In particular, the renaming
+$\Upsilon,u\mapsto\Upsilon,v$ shall take $\mathsf{get}[u]$ to
+$\mathsf{get}[v]$.
 

### 2015-10-24 paper.tex
Try and implement symbol/parameter renaming (resolve #10)

@@
 
+\newcommand\ArrDom[1]{\mathsf{dom}(#1)}
+\newcommand\Match[2]{\IMode{#1}\equiv\OMode{#2}}
+
 \newcommand\Symbols{\mathbb{I}}
@@
 
+\newcommand\Renaming[2]{
+  #2\upharpoonright #1
+}
+\newcommand\IsRenaming[3]{
+  \Renaming{\IMode{#1}}{\IMode{#2}}%
+  \leadsto\OMode{#3}%
+}
+
 \newcommand\Subst[3]{
   \left[#1\, /\, #2\right]#3%
-}
-
-\newcommand\IsRenaming[3]{
-  \IMode{#2}\upharpoonright\IMode{#1}%
-  \leadsto\OMode{#3}%
 }
@@
 
+\subsection{Renaming of Symbols}
+
+The only place that symbols appear in our calculus is as parameters to
+operators (unlike variables, symbols are not terms). Therefore, the functorial
+action of the operator presheaf can be lifted into terms by recursion on their
+structure, via a pair of judgments $\IsRenaming{\varrho}{M}{N}$ and
+$\IsRenaming{\varrho}{E}{F}$, presupposing
+$\Of{\varrho}{\Upsilon\to\Upsilon'}$, and $\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{s}$
+and $\IsAbt{\Omega}{\Upsilon}{\Gamma}{E}{v}$ respectively:
+
+\begin{gather*}
+  \infer{
+    \IsRenaming{\varrho}{x}{x}
+  }{
+  }\\[6pt]
+  \infer{
+    \IsRenaming{\varrho}{
+      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
+    }{
+      \MApp{\MV{m}}{\vec{v}}{N_0,\dots,N_n}
+    }
+  }{
+    \Match{\varrho(\vec{u})}{\vec{v}} &
+    \HypJ{
+      \IsRenaming{\varrho}{M_i}{N_i}
+    }{
+      \Leq{i}{n}
+    }
+  }\\[6pt]
+  \infer{
+    \IsRenaming{\varrho}{
+      \App{\vartheta}{E_0,\dots,E_n}
+    }{
+      \App{\vartheta'}{F_0,\dots,F_n}
+    }
+  }{
+    \Match{\Restriction{\varrho}(\vartheta)}{\vartheta'} &
+    \HypJ{
+      \IsRenaming{\varrho}{E_i}{F_i}
+    }{
+      \Leq{i}{n}
+    }
+  }\\[6pt]
+  \infer{
+    \IsRenaming{\varrho}{
+      \left(\MkBTm{\vec{u}}{\vec{x}}{M}\right)
+    }{
+      \left(\MkBTm{\vec{u}}{\vec{x}}{N}\right)
+    }
+  }{
+    \IsRenaming{\varrho\setminus\vec{u}}{M}{N}
+  }
+\end{gather*}
+
+Above, the notation $\varrho\setminus\vec{u}$ means the omission of the
+variables $\vec{u}$ from the renaming $\varrho$. Because terms are identified
+up to $\alpha$-equivalence, the renaming judgment is functional in its input,
+and so we are justified in writing $\Renaming{\varrho}{M}$ for $N$ when
+$\IsRenaming{\varrho}{M}{N}$.
+
 \subsection{Substitution of Variables}
@@
 
-Because terms are identified up to $\alpha$-equivalence, the variable
-substitution judgment is functional in its inputs, and so we are justified in
-writing $\Subst{N}{x}{M}$ for $M'$ when $\IsSubst{N}{x}{M}{M'}$. We write
-$\Subst{\vec{N}}{\vec{x}}{M}$ for the simultaneous substitution of $\vec{N}$
-for $\vec{x}$ in $M$.
+Going forward, we will write $\Subst{N}{x}{M}$ for $M'$ when
+$\IsSubst{N}{x}{M}{M'}$, and $\Subst{\vec{N}}{\vec{x}}{M}$ for the
+simultaneous substitution of $\vec{N}$ for $\vec{x}$ in $M$.
 

### 2015-10-24 paper.tex
Add a TODO comment
Clarification (really close #10)

@@
 
-\newcommand\ArrDom[1]{\mathsf{dom}(#1)}
+\newcommand\ArrId{\mathbf{id}}
 \newcommand\Match[2]{\IMode{#1}\equiv\OMode{#2}}
@@
 Note that the functoriality of $\Operators$ guarantees for any renaming
-$\Of{\varrho}{\Upsilon\to\Upsilon'}$ a restriction map $\Restriction{\varrho}$
-such that $\IsOperator{\Upsilon'}{\Restriction{\varrho}(\vartheta)}{a}$ when
-$\IsOperator{\Upsilon}{\vartheta}{a}$. In particular, the renaming
+$\Of{\varrho}{\Upsilon\to\Upsilon'}$ and $\Of{f}{a\to a'}$, a restriction map
+$\Restriction{\Pair{\varrho}{f}}$ such that
+$\IsOperator{\Upsilon'}{\Restriction{\Pair{\varrho}{f}}(\vartheta)}{a}$ when
+$\IsOperator{\Upsilon}{\vartheta}{a'}$. Because the category
+$\Discrete{\Arities}$ is discrete and has only identity arrows, we will often
+write $\Restriction{\varrho}$ to mean $\Restriction{\Pair{\varrho}{\ArrId}}$
+through an abuse of notation. In particular, the renaming
 $\Upsilon,u\mapsto\Upsilon,v$ shall take $\mathsf{get}[u]$ to
@@
 
+%%% TODO: show that renaming preserves wellformedness
+
 \subsection{Substitution of Variables}

### 2015-10-24 paper.tex
brief metavariable subst implementation (resolve #3)

@@
 
-\newcommand\Renaming[2]{
+\newcommand\Rename[2]{
   #2\upharpoonright #1
@@
 \newcommand\IsRenaming[3]{
-  \Renaming{\IMode{#1}}{\IMode{#2}}%
+  \Rename{\IMode{#1}}{\IMode{#2}}%
   \leadsto\OMode{#3}%
@@
 up to $\alpha$-equivalence, the renaming judgment is functional in its input,
-and so we are justified in writing $\Renaming{\varrho}{M}$ for $N$ when
+and so we are justified in writing $\Rename{\varrho}{M}$ for $N$ when
 $\IsRenaming{\varrho}{M}{N}$.
@@
 Variable substitution in abstract binding trees is defined inductively by a
-pair of judgments, $\IsSubst{N}{x}{M}{M'}$ and $\IsSubst{N}{x}{E}{E'}$:
+pair of judgments, $\IsSubst{N}{x}{M}{M'}$ and $\IsSubst{N}{x}{E}{F}$:
 
@@
     }{
-      \App{\vartheta}{E'_0,\dots,E'_n}
+      \App{\vartheta}{F_0,\dots,F_n}
     }
@@
     \HypJ{
-      \IsSubst{N}{x}{E_i}{E'_i}
+      \IsSubst{N}{x}{E_i}{F_i}
     }{
@@
 
-Metavariable substitution is defined inductively by the judgment
-$\IsSubst{E}{\MV{m}}{M}{M'}$:
+Metavariables are substituted by bound terms; since a metavariable may only
+appear in an application expression $\MApp{\MV{m}}{\cdots}{\cdots}$, we will
+instantiate the bound term at the supplied parameters and arguments.
+Substitution for metavariables is defined inductively by the judgments
+$\IsSubst{E}{\MV{m}}{M}{N}$ and $\IsSubst{E}{\MV{m}}{F}{F'}$:
 
@@
     }{
-      \MApp{\MV{n}}{\vec{u}}{M'_0,\dots,M'_n}
+      \MApp{\MV{n}}{\vec{u}}{N_0,\dots,N_n}
     }
@@
     \HypJ{
-      \IsSubst{E}{\MV{n}}{M_i}{M'_i}
+      \IsSubst{E}{\MV{n}}{M_i}{N_i}
     }{
@@
   }\\[6pt]
-  %%% \infer{
-  %%%   \IsSubst{
-  %%%     \MkBTm{\vec{u}}{x_0,\dots,x_n}{N}
-  %%%   }{\MV{m}}{
-  %%%     \MApp{\MV{n}}{\vec{v}}{M_0,\dots,M_n}
-  %%%   }{
-  %%%     %TODO!!!
-  %%%     %\Rename{\vec{v}}{\vec{u}}{
-  %%%     \Subst{\vec{M}}{\vec{x}}{N}
-  %%%     % }
-  %%%   }
-  %%% }{
-  %%%   \IsEq{\MV{m}}{\MV{n}}
-  %%% }\\[6pt]
-  %%% \infer{
-  %%%   \IsSubst{E}{\MV{m}}{
-  %%%     \App{\vartheta}{F_0,\dots,F_n}
-  %%%   }{
-  %%%   }
-  %%% }{
-  %%% }
+  \infer{
+    \IsSubst{
+      \MkBTm{\vec{u}}{\vec{x}}{N}
+    }{\MV{m}}{
+      \MApp{\MV{n}}{\vec{v}}{\vec{M}}
+    }{
+      N''
+    }
+  }{
+    \IsEq{\MV{m}}{\MV{n}} &
+    \IsSubst{\vec{M}}{\vec{x}}{N}{N'} &
+    \IsRenaming{\MkSet{\vec{u}\mapsto\vec{v}}}{N'}{N''}
+  }\\[6pt]
+  \infer{
+    \IsSubst{E}{\MV{m}}{
+      \App{\vartheta}{F_0,\dots,F_n}
+    }{
+      \App{\vartheta}{F'_0,\dots,F'_n}
+    }
+  }{
+    \HypJ{
+      \IsSubst{E}{\MV{m}}{F_i}{F'_i}
+    }{
+      \Leq{i}{n}
+    }
+  }\\[6pt]
+  \infer{
+    \IsSubst{E}{\MV{m}}{
+      \MkBTm{\vec{u}}{\vec{x}}{M}
+    }{
+      \MkBTm{\vec{u}}{\vec{x}}{N}
+    }
+  }{
+    \IsApart{\vec{u}}{\FV{E}} &
+    \IsSubst{E}{\MV{m}}{M}{N}
+  }
 \end{gather*}
 
+As before, we will write $\Subst{E}{\MV{m}}{M}$ for $N$ when
+$\IsSubst{E}{\MV{m}}{M}{N}$.
+
 \end{document}

### 2015-10-24 paper.tex
fix improper handling of free symbols
wording

@@
 
-Abstract binding trees are identified up to $\alpha$-equivalence.
+Abstract binding trees are identified up to $\alpha$-equivalence. Let $\FS{M}$
+be the collection of symbols free in $M$, and let $\FV{M}$ be the collection of
+variables free $\pi_\Operators$ to operators in $M$. %% TODO: actually define these
 
@@
     \NotIn{x}{\vec{y}} &
+    \IsApart{\vec{u}}{\FS{N}} &
     \IsApart{\vec{y}}{\FV{N}} &
@@
     \Member{x}{\vec{y}} &
+    \IsApart{\vec{u}}{\FS{N}} &
     \IsApart{\vec{y}}{\FV{N}}
@@
   }{
-    \IsApart{\vec{u}}{\FV{E}} &
+    \IsApart{\vec{u}}{\FS{E}} &
+    \IsApart{\vec{x}}{\FV{E}} &
     \IsSubst{E}{\MV{m}}{M}{N}
@@
 
-As before, we will write $\Subst{E}{\MV{m}}{M}$ for $N$ when
+As usual, we will write $\Subst{E}{\MV{m}}{M}$ for $N$ when
 $\IsSubst{E}{\MV{m}}{M}{N}$.

### 2015-10-24 paper.tex
wordsmithing

@@
 \otimes \Discrete{\Psi(c)}}$.} $\Of{\Grothendieck{}{(-)}}{\Sets^{C}\to\Cats}$
-on operators we have a category of objects
+on operators, we have a category of objects
 $\Member{\Pair{\Pair{\Upsilon}{a}}{\vartheta}}{\Grothendieck{}{\Operators}}$
@@
 Declaring a new assignable consists in providing an initial value, and an
-expression with a free symbol (which shall represent the assignable in scope).
+expression binding a symbol (which shall represent the assignable in scope).
 Note that the functoriality of $\Operators$ guarantees for any renaming
@@
 $\IsOperator{\Upsilon}{\vartheta}{a'}$. Because the category
-$\Discrete{\Arities}$ is discrete and has only identity arrows, we will often
-write $\Restriction{\varrho}$ to mean $\Restriction{\Pair{\varrho}{\ArrId}}$
-through an abuse of notation. In particular, the renaming
+$\Discrete{\Arities}$ is discrete and has only identity arrows, through an
+abuse of notation we will often write $\Restriction{\varrho}$ to mean
+$\Restriction{\Pair{\varrho}{\ArrId}}$. In particular, the renaming
 $\Upsilon,u\mapsto\Upsilon,v$ shall take $\mathsf{get}[u]$ to

### 2015-10-24 paper.tex
remove unused macros
notation
Use lift/map instead of *

@@
 \usepackage{geometry}
+\usepackage{ifthen}
 \usepackage{proof}
@@
 \def\OModeColorName{Maroon}
+
 \newcommand\IMode[1]{{\color{\IModeColorName}{#1}}}
 \newcommand\OMode[1]{{\color{\OModeColorName}{#1}}}
-\newcommand\Quote[1]{{\color{Gray}{«}}#1{\color{Gray}{»}}}
 
@@
 \newcommand\GenJ[2]{\vert_{\IMode{#1}}\; #2}
-\newcommand\JJ{J}
-\newcommand\RSet{\mathfrak{R}}
-\newcommand\JSet{\mathfrak{J}}
 
@@
 \newcommand\MkSet[1]{\left\{#1\right\}}
-\newcommand\MkFam[3]{\MkSet{#1_{#2}}_{#2\in #3}}
 \newcommand\Member[2]{\IMode{#1}\in\IMode{#2}}
@@
 
-\newcommand\MkValence[3]{\left\{#1\right\}\left[#2\right].\, #3}
-\newcommand\MkBTm[3]{\left\{#1\right\}\left[#2\right].\, #3}
-\newcommand\MkArity[2]{\left(#1\right) #2}
+\newcommand\MkValence[3]{\{#1\}[#2].\, #3}
+\newcommand\MkBTm[3]{\{#1\}[#2].\, #3}
+\newcommand\MkArity[2]{(#1) #2}
 \newcommand\IsArity[1]{\IMode{#1}\ \textit{arity}}
@@
 
-\newcommand\ArrId{\mathbf{id}}
+\newcommand\ArrId[1]{1_{#1}}
 \newcommand\Match[2]{\IMode{#1}\equiv\OMode{#2}}
@@
 
-\newcommand\ParamsOf{\pi_\mathfrak{p}}
-\newcommand\ArityOf{\pi_\mathfrak{a}}
 \newcommand\Cats{\mathbf{Cat}}
@@
 \newcommand\Pair[2]{\langle #1, #2 \rangle}
-\newcommand\Restriction[1]{#1^{*}}
+\newcommand\Lift[2]{
+  \Operators\langle#1\ifthenelse{\equal{#2}{}}{}{, #2}\rangle
+}
+\newcommand\PairTimes[2]{\langle #1\, {\scriptstyle {{\times}}}\, #2 \rangle}
 \newcommand\Singleton{\bullet}
@@
 \newcommand\Coslice[2]{#1\, /\, #2}
-\newcommand\Slice[2]{#1\, /\, #2}
 
@@
 \newcommand\MV[1]{\textsc{#1}}
-\newcommand\MApp[3]{#1\left\{#2\right\}\left(#3\right)}
-\newcommand\App[2]{#1\left(#2\right)}
+\newcommand\MApp[3]{#1\{#2\}(#3)}
+\newcommand\App[2]{#1(#2)}
 \newcommand\Dom[1]{\left\vert #1\right\vert}
@@
 for $\Of{f}{\Pair{\Upsilon}{a}\to\Pair{\Upsilon'}{a'}}$ such that
-$\EqMember{\Restriction{f}\vartheta}{\vartheta'}{\Operators\Pair{\Upsilon'}{a'}}$.
+$\EqMember{\Lift{f}{}(\vartheta)}{\vartheta'}{\Operators\Pair{\Upsilon'}{a'}}$.
 Equivalently, $\Grothendieck{}{\Operators}$ is the pullback of $\Operators$
@@
 Note that the functoriality of $\Operators$ guarantees for any renaming
-$\Of{\varrho}{\Upsilon\to\Upsilon'}$ and $\Of{f}{a\to a'}$, a restriction map
-$\Restriction{\Pair{\varrho}{f}}$ such that
-$\IsOperator{\Upsilon'}{\Restriction{\Pair{\varrho}{f}}(\vartheta)}{a}$ when
+$\Of{\varrho}{\Upsilon\to\Upsilon'}$ and $\Of{f}{a\to a'}$, a lifted map
+$\Lift{\varrho}{f}$ such that
+$\IsOperator{\Upsilon'}{\Lift{\varrho}{f}(\vartheta)}{a}$ when
 $\IsOperator{\Upsilon}{\vartheta}{a'}$. Because the category
-$\Discrete{\Arities}$ is discrete and has only identity arrows, through an
-abuse of notation we will often write $\Restriction{\varrho}$ to mean
-$\Restriction{\Pair{\varrho}{\ArrId}}$. In particular, the renaming
-$\Upsilon,u\mapsto\Upsilon,v$ shall take $\mathsf{get}[u]$ to
-$\mathsf{get}[v]$.
+$\Discrete{\Arities}$ is discrete and has only identity arrows, through an abuse
+of notation we will often write $\Lift{\varrho}{}$ to mean
+$\Lift{\varrho}{\ArrId{}}$. In particular, the renaming
+$\Upsilon,u\mapsto\Upsilon,v$ shall take $\mathsf{get}[u]$ to $\mathsf{get}[v]$.
 
@@
   }{
-    \Match{\Restriction{\varrho}(\vartheta)}{\vartheta'} &
+    \Match{\Lift{\varrho}{}(\vartheta)}{\vartheta'} &
     \HypJ{

### 2015-10-24 paper.tex
Use subscript for pointed sets
Fix types; annotate lifted map

@@
     \arrow[d, "{\pi_{\Operators}}" left]
-  \& {\Coslice{\Singleton}{\Sets}}
+  \& {\Sets_{\Singleton}}
     \arrow[d, "{\Forget{-}_{\Singleton}}"] \\
@@
 $\Of{\varrho}{\Upsilon\to\Upsilon'}$ and $\Of{f}{a\to a'}$, a lifted map
-$\Lift{\varrho}{f}$ such that
-$\IsOperator{\Upsilon'}{\Lift{\varrho}{f}(\vartheta)}{a}$ when
-$\IsOperator{\Upsilon}{\vartheta}{a'}$. Because the category
+$\Of{\Lift{\varrho}{f}}{\Operators\Pair{\Upsilon}{a}\to\Operators\Pair{\Upsilon'}{a'}}$
+such that $\IsOperator{\Upsilon'}{\Lift{\varrho}{f}(\vartheta)}{a'}$ when
+$\IsOperator{\Upsilon}{\vartheta}{a}$. Because the category
 $\Discrete{\Arities}$ is discrete and has only identity arrows, through an abuse

### 2015-10-25 paper.tex
Split f into two explicit maps
Begin implementing case study (#4)

@@
 
-\newcommand\MkValence[3]{\{#1\}[#2].\, #3}
+\newcommand\MkValence[3]{%
+  \ifthenelse{\equal{#1}{}}{}{%
+    \{#1\}%
+  }%
+  \ifthenelse{\equal{#2}{}}{}{%
+    [#2]%
+  }.\, #3
+}
 \newcommand\MkBTm[3]{\{#1\}[#2].\, #3}
@@
 \newcommand\Lift[2]{
-  \Operators\langle#1\ifthenelse{\equal{#2}{}}{}{, #2}\rangle
+  \Operators\langle#1%
+  \ifthenelse{\equal{#2}{}}{}{, #2}\rangle
 }
@@
   $\IMode{C\Grothendieck{}{\Psi}}\cong \IMode{\int^{c \in C}\Coslice{c}{C}
-\otimes \Discrete{\Psi(c)}}$.} $\Of{\Grothendieck{}{(-)}}{\Sets^{C}\to\Cats}$
-on operators, we have a category of objects
-$\Member{\Pair{\Pair{\Upsilon}{a}}{\vartheta}}{\Grothendieck{}{\Operators}}$
-for $\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$ and morphisms
+    \otimes \Discrete{\Psi(c)}}$.}
+$\Of{\Grothendieck{}{(-)}}{\Sets^{C}\to\Cats}$ on operators, we have a category
+of objects
+$\Member{\Pair{\Pair{\Upsilon}{a}}{\vartheta}}{\Grothendieck{}{\Operators}}$ for
+$\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$ and morphisms
 $\Hom{\Grothendieck{}{\Operators}}{\Pair{\Pair{\Upsilon}{a}}{\vartheta}}{\Pair{\Pair{\Upsilon'}{a'}}{\vartheta'}}$
-for $\Of{f}{\Pair{\Upsilon}{a}\to\Pair{\Upsilon'}{a'}}$ such that
-$\EqMember{\Lift{f}{}(\vartheta)}{\vartheta'}{\Operators\Pair{\Upsilon'}{a'}}$.
+for $\Of{\Pair{\varrho}{f}}{\Pair{\Upsilon}{a}\to\Pair{\Upsilon'}{a'}}$ such
+that
+$\EqMember{\Lift{\varrho}{f}(\vartheta)}{\vartheta'}{\Operators\Pair{\Upsilon'}{a'}}$.
 Equivalently, $\Grothendieck{}{\Operators}$ is the pullback of $\Operators$
@@
   \IsOperator{\Upsilon}{\lambda}{
-    \MkArity{\MkValence{\cdot}{\SortExp}{\SortExp}}{\SortExp}
+    \MkArity{\MkValence{}{\SortExp}{\SortExp}}{\SortExp}
   }\\
@@
     \MkArity{
-      \MkValence{\cdot}{\cdot}{\SortExp},
-      \MkValence{\cdot}{\cdot}{\SortExp}
+      \MkValence{}{}{\SortExp},
+      \MkValence{}{}{\SortExp}
     }{\SortExp}
@@
     \MkArity{
-      \MkValence{\cdot}{\cdot}{\SortExp},
-      \MkValence{\SortExp}{\cdot}{\SortExp}
+      \MkValence{}{}{\SortExp},
+      \MkValence{\SortExp}{}{\SortExp}
     }{
@@
   \IsOperator{\Upsilon,u:\SortExp}{\mathsf{get}[u]}{
-    \MkArity{
-      \cdot
-    }{
-      \SortExp
-    }
+    \MkArity{}{\SortExp}
   }\\
@@
     \MkArity{
-      \MkValence{\cdot}{\cdot}{\SortExp}
+      \MkValence{}{}{\SortExp}
     }{
@@
 $\IsBTm{\Omega}{\Upsilon}{\Gamma}{E}{v}$ presuppose $\IsValence{v}$.  Then, the
-syntax of abstract binding trees is inductively defined in four rules:
+syntax of abstract binding trees (abts) is inductively defined in four rules:
 
@@
 
-Abstract binding trees are identified up to $\alpha$-equivalence. Let $\FS{M}$
+Abts are identified up to $\alpha$-equivalence. Let $\FS{M}$
 be the collection of symbols free in $M$, and let $\FV{M}$ be the collection of
@@
 
-Variable substitution in abstract binding trees is defined inductively by a
+Variable substitution in abts is defined inductively by a
 pair of judgments, $\IsSubst{N}{x}{M}{M'}$ and $\IsSubst{N}{x}{E}{F}$:
@@
 
+\section{Case Study: Wellformed Sequents}
+
+The representation of telescopes and sequents in a logical framework is
+notoriously difficult; whilst it is possible to use higher-order abstract
+syntax or abts to encode the binding-structure of telescopes
+and sequents, the encoding is sufficiently laborious and obscure that it is not
+used in practice.
+
+Crary has demonstrated a first-order encoding of contexts in the logical
+framework in bijection with actual LF-contexts, which has been successfully
+used in large-scale mechanization efforts, including that of Standard ML and
+the Edinburgh Logical Framework itself.
+
+%%% TODO: Cite Crary; cite mechaniziation of Standard ML; cite Martens & Crary
+%%% TODO: talk about context encodings in Abella
+
+We will approach the problem of encoding telescopes and sequents from the
+\emph{refinements} perspective, where a conservative approximation of the
+grammar is first given using the abt logical framework, and then the
+correctness of a code is expressed separately in a judgment that refines the
+existing specification.
+
+Because we have not committed to using the built-in binding machinery to
+express the well-scopedness of telescopes and sequents, we are free to use
+\emph{symbols} in order to model the variables in the context. This is actually
+quite sensible if we are trying to actually faithfully represent the syntax of
+telescopes and sequents, rather than replace them with their counterparts on the
+meta-level. It is certainly not intended that we should be able to take a
+context $\Match{\Gamma}{x:A}$ and apply the substitution
+$\IsSubst{M}{x}{\Gamma}{M:A}$, which is nonsensical; on the other hand, if $x$
+had been a \emph{symbol}, all we could do is rename it in the context, since
+substitution is not defined for symbols.
+
+This insight leads the way to a simple abt signature for the theory of telescopes
+and sequents.
+%
+\newcommand\SortCtx{\mathsf{ctx}}
+\newcommand\SortJdg{\mathsf{jdg}}
+\newcommand\SortType{\mathsf{type}}
+\newcommand\SortExpr{\mathsf{exp}}
+\newcommand\OpNil{\mathsf{nil}}
+\newcommand\OpSnoc[1]{\mathsf{snoc}[#1]}
+\newcommand\OpVar[1]{\mathsf{var}[#1]}
+\newcommand\OpSequent{\mathsf{sequent}}
+%
+\[
+  \begin{array}{l}
+    \IsSort{\SortCtx}\\
+    \IsSort{\SortExpr}\\
+    \IsSort{\SortType}\\
+    \IsSort{\SortJdg}\\\\
+    \IsOperator{\Upsilon, u:\SortExpr}{\OpVar{u}}{
+      \MkArity{}{\SortExpr}
+    }\\\\
+    \IsOperator{\Upsilon}{\OpNil}{
+      \MkArity{}{\SortCtx}
+    }\\
+    \IsOperator{\Upsilon, u:\SortExpr}{\OpSnoc{u}}{
+      \MkArity{
+        \MkValence{}{}{\SortCtx},
+        \MkValence{}{}{\SortType}
+      }{\SortCtx}
+    }\\\\
+    \IsOperator{\Upsilon}{\OpSequent}{
+      \MkArity{
+        \MkValence{}{}{\SortCtx},
+        \MkValence{}{}{\SortType}
+      }{\SortJdg}
+    }
+  \end{array}
+\]
+
+\newcommand\OpPi{\mathsf{pi}}
+\newcommand\OpSg{\mathsf{sg}}
+\newcommand\OpBool{\mathsf{bool}}
+\newcommand\OpSo{\mathsf{isTrue}}
+
+Suppose we have defined encoded a fragment of type theory as well:
+\[
+  \begin{array}{l}
+    \IsOperator{\Upsilon}{\top}{
+      \MkArity{}{\SortType}
+    }\\
+    \IsOperator{\Upsilon}{\bot}{
+      \MkArity{}{\SortType}
+    }\\
+    \IsOperator{\Upsilon}{\OpBool}{
+      \MkArity{}{\SortType}
+    }\\
+    \IsOperator{\Upsilon}{\OpSo}{
+      \MkArity{
+        \MkValence{}{}{\SortExp}
+      }{\SortType}
+    }\\
+    \IsOperator{\Upsilon}{\OpPi}{
+      \MkArity{
+        \MkValence{}{}{\SortType},
+        \MkValence{}{\SortExp}{\SortType}
+      }{\SortType}
+    }\\
+    \IsOperator{\Upsilon}{\OpSg}{
+      \MkArity{
+        \MkValence{}{}{\SortType},
+        \MkValence{}{\SortExp}{\SortType}
+      }{\SortType}
+    }\\
+  \end{array}
+\]
+
+Terms written using the abstract syntax will be difficult to read, so let us
+define some notation:
+\begin{align*}
+  \IMode{\diamond} &\triangleq \OMode{\OpNil}\\
+  \IMode{\Gamma, u:P} &\triangleq \OMode{
+    \App{\OpSnoc{u}}{\Gamma, P}
+  }\\
+  \IMode{\Gamma\gg P} &\triangleq \OMode{
+    \App{\OpSequent}{
+      \Gamma, P
+    }
+  }\\
+  \IMode{`u} &\triangleq \OMode{\OpVar{u}}
+\end{align*}
+
+Now, we have the following well-formed sequent:
+\[
+  \IsAbt{\cdot}{
+    u:\SortExp,
+    v:\SortExp
+  }{\cdot}{
+    \diamond,u:\OpBool,v:\App{\OpSo}{`u}%
+    \gg%
+    \App{\OpSo}{`u}
+  }{\SortJdg}
+\]
+
+%% TODO: add \nabla operator to close the context
+%% TODO: implement refinement judgment for wellscopedness
+
 \end{document}

### 2015-10-25 paper.tex
wording/notation
Add parametric object-judgment
Add first paragraph of refinements section

@@
   \IMode{#4}:\OMode{#5}%
+}
+\newcommand\IsWf[5]{%
+  \IMode{#1}\triangleright%
+  \IMode{#2}\parallel%
+  \IMode{#3}\vdash%
+  \IMode{#4}\in_{\mathbf{wf}}\OMode{#5}%
 }
@@
 %
-\newcommand\SortCtx{\mathsf{ctx}}
+\newcommand\SortTele{\mathsf{tele}}
 \newcommand\SortJdg{\mathsf{jdg}}
@@
   \begin{array}{l}
-    \IsSort{\SortCtx}\\
+    \IsSort{\SortTele}\\
     \IsSort{\SortExpr}\\
@@
     \IsOperator{\Upsilon}{\OpNil}{
-      \MkArity{}{\SortCtx}
+      \MkArity{}{\SortTele}
     }\\
@@
       \MkArity{
-        \MkValence{}{}{\SortCtx},
+        \MkValence{}{}{\SortTele},
         \MkValence{}{}{\SortType}
-      }{\SortCtx}
+      }{\SortTele}
     }\\\\
@@
       \MkArity{
-        \MkValence{}{}{\SortCtx},
+        \MkValence{}{}{\SortTele},
         \MkValence{}{}{\SortType}
@@
 
-%% TODO: add \nabla operator to close the context
+\newcommand\OpNabla[1]{\nabla[#1]}
+
+The above sequent has free symbols, but we can close over them by adding a form
+of parametric higher-order judgment to our object language, indexed by a
+collection of sorts $\vec{s}$:
+\[
+  \IsOperator{\Upsilon}{\OpNabla{\vec{s}}}{
+    \MkArity{
+      \MkValence{\vec{s}}{}{\SortJdg}
+    }{\SortJdg}
+  }
+\]
+
+Then, we may write a closed sequent judgment as follows:
+\[
+  \IsAbt{\cdot}{\cdot}{\cdot}{
+    \App{\OpNabla{\SortExp,\SortExp}}{
+      \MkBTm{u,v}{}{
+        \diamond,u:\OpBool,v:\App{\OpSo}{`u}%
+        \gg%
+        \App{\OpSo}{`u}
+      }
+    }
+  }{\SortJdg}
+\]
+
+\subsection{Refinements for wellformedness}
+
+Having specified an approximation of the grammar of telescopes and sequents in
+the abt logical framework, we can proceed to define proper wellformedness via
+\emph{inductive refinement}. The basic idea is to introduce a new form of
+(meta)-judgment $\IsWf{\Omega}{\Upsilon}{\Gamma}{M}{s}$ which expresses the
+extrinsic wellformedness properties we wish to verify, presupposing
+$\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{s}$.
+
 %% TODO: implement refinement judgment for wellscopedness

### 2015-10-25 paper.tex
Add wf-refinement rules for judgments and telescopes

@@
 meta-level. It is certainly not intended that we should be able to take a
-context $\Match{\Gamma}{x:A}$ and apply the substitution
-$\IsSubst{M}{x}{\Gamma}{M:A}$, which is nonsensical; on the other hand, if $x$
+context $\Match{H}{x:A}$ and apply the substitution
+$\IsSubst{M}{x}{H}{M:A}$, which is nonsensical; on the other hand, if $x$
 had been a \emph{symbol}, all we could do is rename it in the context, since
@@
   \IMode{\diamond} &\triangleq \OMode{\OpNil}\\
-  \IMode{\Gamma, u:P} &\triangleq \OMode{
-    \App{\OpSnoc{u}}{\Gamma, P}
+  \IMode{H, u:P} &\triangleq \OMode{
+    \App{\OpSnoc{u}}{H, P}
   }\\
-  \IMode{\Gamma\gg P} &\triangleq \OMode{
+  \IMode{H\gg A} &\triangleq \OMode{
     \App{\OpSequent}{
-      \Gamma, P
+      H, A
     }
@@
 
+For the sake of this example, we will just stipulate that anything of sort $\SortExp$ or $\SortType$ is
+grammatical:
+\begin{align*}
+  \infer{
+    \IsWf{\Omega}{\Upsilon}{\Gamma}{A}{\SortType}
+  }{
+  }\qquad
+  \infer{
+    \IsWf{\Omega}{\Upsilon}{\Gamma}{M}{\SortExp}
+  }{
+  }
+\end{align*}
+
+The refinements for parametric judgment and sequents simply delegate to their
+subterms:
+\begin{gather*}
+  \infer{
+    \IsWf{\Omega}{\Upsilon}{\Gamma}{
+      \App{\OpNabla{\vec{s}}}{
+        \MkBTm{\vec{u}}{}{J}
+      }
+    }{\SortJdg}
+  }{
+    \IsWf{\Omega}{\Upsilon,\vec{u}:\vec{s}}{\Gamma}{J}{\SortJdg}
+  }\\[6pt]
+  \infer{
+    \IsWf{\Omega}{\Upsilon}{\Gamma}{
+      H\gg A
+    }{\SortJdg}
+  }{
+    \IsWf{\Omega}{\Upsilon}{\Gamma}{H}{\SortTele} &
+    \IsWf{\Omega}{\Upsilon}{\Gamma}{A}{\SortType}
+  }
+\end{gather*}
+
+The refinement for telescopes proceeds by induction:
+\begin{gather*}
+  \infer{
+    \IsWf{\Omega}{\Upsilon}{\Gamma}{\diamond}{\SortTele}
+  }{
+  }
+  \qquad
+  \infer{
+    \IsWf{\Omega}{\Upsilon}{\Gamma}{H,u:A}{\SortTele}
+  }{
+    \begin{array}{l}
+      \IsWf{\Omega}{\Upsilon\setminus\MkSet{u}}{\Gamma}{H}{\SortTele}\\
+      \IsWf{\Omega}{\Upsilon\setminus\MkSet{u}}{\Gamma}{A}{\SortType}
+    \end{array}
+  }
+\end{gather*}
+
 \end{document}

### 2015-10-25 paper.tex
Bold the category C so it doesn't look like an object

@@
 in $\SCtx$ lift to renamings of operators' parameters. Via the Grothendieck
-construction\footnote{In this case, $C\Grothendieck{}{\Psi}$ represents the
-  category of elements of a copresheaf $\Of{\Psi}{C\to\Sets}$ but we keep the
-  $C$ implicit and simply refer to it as the Grothendieck construction.
-  Alternatively, this construction can be understood as a coend
-  $\IMode{C\Grothendieck{}{\Psi}}\cong \IMode{\int^{c \in C}\Coslice{c}{C}
-    \otimes \Discrete{\Psi(c)}}$.}
-$\Of{\Grothendieck{}{(-)}}{\Sets^{C}\to\Cats}$ on operators, we have a category
-of objects
-$\Member{\Pair{\Pair{\Upsilon}{a}}{\vartheta}}{\Grothendieck{}{\Operators}}$ for
-$\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$ and morphisms
+construction\footnote{In this case, $\mathbf{C}\Grothendieck{}{\Psi}$
+  represents the category of elements of a copresheaf
+  $\Of{\Psi}{\mathbf{C}\to\Sets}$ but we keep the $\mathbf{C}$ implicit and
+  simply refer to it as the Grothendieck construction.  Alternatively, this
+  construction can be understood as a coend
+$\IMode{\mathbf{C}\Grothendieck{}{\Psi}}\cong \IMode{\int^{c \in
+\mathbf{C}}\Coslice{c}{\mathbf{C}} \otimes \Discrete{\Psi(c)}}$.}
+$\Of{\Grothendieck{}{(-)}}{\Sets^{\mathbf{C}}\to\Cats}$ on operators, we have a
+category of objects
+$\Member{\Pair{\Pair{\Upsilon}{a}}{\vartheta}}{\Grothendieck{}{\Operators}}$
+for $\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$ and morphisms
 $\Hom{\Grothendieck{}{\Operators}}{\Pair{\Pair{\Upsilon}{a}}{\vartheta}}{\Pair{\Pair{\Upsilon'}{a'}}{\vartheta'}}$
@@
 express the well-scopedness of telescopes and sequents, we are free to use
-\emph{symbols} in order to model the variables in the context. This is actually
-quite sensible if we are trying to actually faithfully represent the syntax of
-telescopes and sequents, rather than replace them with their counterparts on the
-meta-level. It is certainly not intended that we should be able to take a
-context $\Match{H}{x:A}$ and apply the substitution
-$\IsSubst{M}{x}{H}{M:A}$, which is nonsensical; on the other hand, if $x$
-had been a \emph{symbol}, all we could do is rename it in the context, since
-substitution is not defined for symbols.
+\emph{symbols} in order to model the variables in the context. This is in fact
+quite sensible if we are actually trying to faithfully represent the syntax of
+telescopes and sequents, rather than replace them with their counterparts on
+the meta-level. It is certainly not intended that we should be able to take a
+context $\Match{H}{x:A}$ and apply the substitution $\IsSubst{M}{x}{H}{M:A}$,
+which is nonsensical; on the other hand, if $x$ had been a \emph{symbol}, all
+we could do is rename it in the context, since substitution is not defined for
+symbols.
 

### 2015-10-25 paper.tex
remove silly statement

@@
 telescopes and sequents, rather than replace them with their counterparts on
-the meta-level. It is certainly not intended that we should be able to take a
-context $\Match{H}{x:A}$ and apply the substitution $\IsSubst{M}{x}{H}{M:A}$,
-which is nonsensical; on the other hand, if $x$ had been a \emph{symbol}, all
-we could do is rename it in the context, since substitution is not defined for
-symbols.
+the meta-level.
 

### 2015-10-25 paper.tex
add refinements for variables & metavariables

@@
   \IMode{#3}\vdash%
-  \IMode{#4}\in_{\mathbf{wf}}\OMode{#5}%
+  \IMode{#4}\in_{\mathbf{wf}}\IMode{#5}%
 }
@@
 extrinsic wellformedness properties we wish to verify, presupposing
-$\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{s}$.
+$\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{s}$. Additionally, we introduce an
+analogous judgment on bound terms, $\IsWf{\Omega}{\Upsilon}{\Gamma}{E}{v}$
+presupposing $\IsAbt{\Omega}{\Upsilon}{\Gamma}{E}{v}$, defined uniformly as
+follows:
+\[
+  \infer{
+    \IsWf{\Omega}{\Upsilon}{\Gamma}{
+      \MkBTm{\vec{u}}{\vec{x}}{M}
+    }{
+      \MkValence{\vec{p}}{\vec{q}}{s}
+    }
+  }{
+    \IsWf{\Omega}{\Upsilon,\vec{u}:\vec{p}}{\Gamma,\vec{x}:\vec{q}}{M}{s}
+  }
+\]
 
-%% TODO: implement refinement judgment for wellscopedness
+Likewise, wellformedness for variables and metavariables is defined uniformly:
+\begin{gather*}
+  \infer{
+    \IsWf{\Omega}{\Upsilon}{\Gamma}{x}{s}
+  }{
+  }\qquad
+  \infer{
+    \IsWf{\Omega}{\Upsilon}{\Gamma}{
+      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
+    }{s}
+  }{
+    \begin{array}{l}
+      \Lookup{\Omega}{\MV{m}}{\MkValence{\vec{p}}{q_0,\dots,q_n}{s}}\\
+      \HypJ{
+        \IsWf{\Omega}{\Upsilon}{\Gamma}{M_i}{q_i}
+      }{
+        \Leq{i}{n}
+      }
+    \end{array}
+  }
+\end{gather*}
 
-For the sake of this example, we will just stipulate that anything of sort $\SortExp$ or $\SortType$ is
-grammatical:
+\begin{remark}
+  Note that the refinement for variables $x$ is not trivial, since it is only
+  defined in case the presupposition $\IsAbt{\Omega}{\Upsilon}{\Gamma}{x}{s}$
+  is satisfied.
+\end{remark}
+
+The remainder of the definition of refinement proceeds by induction on sorts
+and operators. For the sake of this example, we will just stipulate that
+anything of sort $\SortExp$ or $\SortType$ is grammatical:
+
 \begin{align*}

### 2015-10-25 paper.tex
further fixes to refinements
typos

@@
 
-Abts are identified up to $\alpha$-equivalence. Let $\FS{M}$
-be the collection of symbols free in $M$, and let $\FV{M}$ be the collection of
-variables free $\pi_\Operators$ to operators in $M$. %% TODO: actually define these
+Abts are identified up to $\alpha$-equivalence. Let $\FS{M}$ be the collection
+of symbols free in $M$, and let $\FV{M}$ be the collection of variables free in
+$M$. %% TODO: actually define these
 
@@
 
-Suppose we have defined encoded a fragment of type theory as well:
+Suppose we have encoded a fragment of type theory as well:
 \[
@@
 and operators. For the sake of this example, we will just stipulate that
-anything of sort $\SortExp$ or $\SortType$ is grammatical:
+anything of sort $\SortExp$ or $\SortType$ is grammatical if its subterms are
+grammatical:
 
-\begin{align*}
-  \infer{
-    \IsWf{\Omega}{\Upsilon}{\Gamma}{A}{\SortType}
+\[
+  \infer[
+    \text{for $\Member{s}{\MkSet{\SortExp,\SortType}}$}
+  ]{
+    \IsWf{\Omega}{\Upsilon}{\Gamma}{
+      \App{\vartheta}{E_0,\dots,E_n}
+    }{s}
   }{
-  }\qquad
-  \infer{
-    \IsWf{\Omega}{\Upsilon}{\Gamma}{M}{\SortExp}
+    \begin{array}{l}
+      \IsOperator{\Upsilon}{\vartheta}{
+        \MkArity{
+          v_0,\dots,v_n
+        }{s}
+      }\\
+      \HypJ{
+        \IsWf{\Omega}{\Upsilon}{\Gamma}{E_i}{v_i}
       }{
+        \Leq{i}{n}
       }
-\end{align*}
+    \end{array}
+  }
+\]
 
 The refinements for parametric judgment and sequents simply delegate to their
-subterms:
+subterms as well:
 \begin{gather*}

### 2015-10-26 paper.tex
Fix up font reference
Use microtype package
Adjust spacing for arities
Sort packages
Disable xelatex-specific package
Enable linux libertine font
Add some references (#5)
Add a couple of citations

@@
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-\usepackage{amsmath, amsthm, stmaryrd, scalerel, amssymb}
-\usepackage{mathtools}
-\usepackage{graphicx}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{amsthm}
 \usepackage{fontspec}
 \usepackage{geometry}
+\usepackage{graphicx}
 \usepackage{ifthen}
+\usepackage{mathtools}
 \usepackage{proof}
+\usepackage{scalerel}
+\usepackage{stmaryrd}
 \usepackage{supertabular}
@@
 \usepackage{unicode-math}
-\usepackage{xltxtra}
 \usepackage{xunicode}
@@
 % fonts
-% \setromanfont[Mapping=tex-text, Ligatures=Rare, Numbers=OldStyle]{Linux Libertine O}
-\setmathfont{Asana-Math.otf}
+\usepackage[
+  activate={true,nocompatibility},
+  final,
+  tracking=true
+]{microtype}
+\setmainfont{Linux Libertine O}
+\setmathfont{Asana Math}
 
@@
 \newcommand\MkBTm[3]{\{#1\}[#2].\, #3}
-\newcommand\MkArity[2]{(#1) #2}
+\newcommand\MkArity[2]{(#1)\, #2}
 \newcommand\IsArity[1]{\IMode{#1}\ \textit{arity}}
@@
 Crary has demonstrated a first-order encoding of contexts in the logical
-framework in bijection with actual LF-contexts, which has been successfully
-used in large-scale mechanization efforts, including that of Standard ML and
-the Edinburgh Logical Framework itself.
+framework in bijection with actual LF-contexts \cite{crary:2009}, which has
+been successfully used in large-scale mechanization efforts, including that of
+Standard ML \cite{lee-crary-harper:2007} and the Edinburgh Logical Framework
+itself \cite{martens-crary:2012}.
 
-%%% TODO: Cite Crary; cite mechaniziation of Standard ML; cite Martens & Crary
 %%% TODO: talk about context encodings in Abella
@@
 
+\nocite{*}
+\bibliographystyle{abbrv}
+\bibliography{refs}
+
 \end{document}

### 2015-10-26 paper.tex
remove some unnecessary parentheses
Renaming

@@
 \newcommand\Pair[2]{\langle #1, #2 \rangle}
-\newcommand\Lift[2]{
+\newcommand\FLift[2]{
   \Operators\langle#1%
@@
 \newcommand\PairTimes[2]{\langle #1\, {\scriptstyle {{\times}}}\, #2 \rangle}
-\newcommand\Singleton{\bullet}
+\newcommand\SSingleton{\bullet}
 \newcommand\Forget[1]{|\, #1\, |}
@@
 that
-$\EqMember{\Lift{\varrho}{f}(\vartheta)}{\vartheta'}{\Operators\Pair{\Upsilon'}{a'}}$.
+$\EqMember{\FLift{\varrho}{f}(\vartheta)}{\vartheta'}{\Operators\Pair{\Upsilon'}{a'}}$.
 Equivalently, $\Grothendieck{}{\Operators}$ is the pullback of $\Operators$
 along the universal $\Sets$-bundle where $\pi_{\Operators}$ is a discrete
-Grothendieck opfibration and $\Forget{-}_{\Singleton}$ is the forgetful functor
+Grothendieck opfibration and $\Forget{-}_{\SSingleton}$ is the forgetful functor
 from pointed sets:
@@
     \arrow[d, "{\pi_{\Operators}}" left]
-  \& {\Sets_{\Singleton}}
-    \arrow[d, "{\Forget{-}_{\Singleton}}"] \\
+  \& {\Sets_{\SSingleton}}
+    \arrow[d, "{\Forget{-}_{\SSingleton}}"] \\
   {\SCtx\times\Discrete{\Arities}}
@@
 $\Of{\varrho}{\Upsilon\to\Upsilon'}$ and $\Of{f}{a\to a'}$, a lifted map
-$\Of{\Lift{\varrho}{f}}{\Operators\Pair{\Upsilon}{a}\to\Operators\Pair{\Upsilon'}{a'}}$
-such that $\IsOperator{\Upsilon'}{\Lift{\varrho}{f}(\vartheta)}{a'}$ when
+$\Of{\FLift{\varrho}{f}}{\Operators\Pair{\Upsilon}{a}\to\Operators\Pair{\Upsilon'}{a'}}$
+such that $\IsOperator{\Upsilon'}{\FLift{\varrho}{f}(\vartheta)}{a'}$ when
 $\IsOperator{\Upsilon}{\vartheta}{a}$. Because the category
 $\Discrete{\Arities}$ is discrete and has only identity arrows, through an abuse
-of notation we will often write $\Lift{\varrho}{}$ to mean
-$\Lift{\varrho}{\ArrId{}}$. In particular, the renaming
+of notation we will often write $\FLift{\varrho}{}$ to mean
+$\FLift{\varrho}{\ArrId{}}$. In particular, the renaming
 $\Upsilon,u\mapsto\Upsilon,v$ shall take $\mathsf{get}[u]$ to $\mathsf{get}[v]$.
@@
     \IsBTm{\Omega}{\Upsilon}{\Gamma}{
-      \left(\MkBTm{\vec{u}}{\vec{x}}{M}\right)
+      \MkBTm{\vec{u}}{\vec{x}}{M}
     }{
-      \left(\MkValence{\vec{p}}{\vec{q}}{s}\right)
+      \MkValence{\vec{p}}{\vec{q}}{s}
     }
@@
   }{
-    \Match{\Lift{\varrho}{}(\vartheta)}{\vartheta'} &
+    \Match{\FLift{\varrho}{}(\vartheta)}{\vartheta'} &
     \HypJ{
@@
     \IsRenaming{\varrho}{
-      \left(\MkBTm{\vec{u}}{\vec{x}}{M}\right)
+      \MkBTm{\vec{u}}{\vec{x}}{M}
     }{
-      \left(\MkBTm{\vec{u}}{\vec{x}}{N}\right)
+      \MkBTm{\vec{u}}{\vec{x}}{N}
     }

### 2015-10-27 paper.tex
Sort packages
Add support for including agda code
[agda] add variable context; symbols are not terms, but vars are
Add definition of substitution monoidal structures & monoid
@freebroccolo Would you check my work? In generalizing this to our
setting, I sort of just guessed what it "had" to be in order to work,
but you will likely have a better perspective on this.

@@
 
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
 \usepackage{amsmath}
@@
 \usepackage{amsthm}
+\usepackage{agda} % this should come after ams*
+\usepackage{catchfilebetweentags}
 \usepackage{fontspec}
@@
 \usepackage{graphicx}
+\usepackage{hyperref}
 \usepackage{ifthen}
@@
 \usepackage{scalerel}
+\usepackage{setspace}
 \usepackage{stmaryrd}
@@
 \usepackage{unicode-math}
+\usepackage{url}
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
 \usepackage{xunicode}
-
-\usepackage{setspace,url,hyperref}
 
@@
 \setmathfont{Asana Math}
+
+% make lualatex play nice with agda
+\makeatletter
+\ifluatex
+   \newwrite\CATCHPATCHOUT
+\fi
+\def\CatchFBT@Fin@l#1[#2]{%
+   \begingroup
+      \endlinechar\m@ne \makeatletter #2%
+      \ifluatex
+         \immediate\openout\CATCHPATCHOUT=\jobname.catchpatch
+         \immediate\write\CATCHPATCHOUT {\CatchFBT@tok{\the\CatchFBT@tok}}%
+         \immediate\closeout\CATCHPATCHOUT
+         \@@input \jobname.catchpatch
+      \else
+      \scantokens\expandafter{%
+         \expandafter\CatchFBT@tok\expandafter{\the\CatchFBT@tok}}%
+      \fi
+      \CatchFBT@IsAToken{#1}
+         {\global#1\expandafter{\the\CatchFBT@tok}}
+         {\xdef#1{\the\CatchFBT@tok}}%
+      \ifx\CatchFBT@tok#1\else\global\CatchFBT@tok{}\fi
+   \endgroup
+}% \CatchFBT@Final
+\makeatother
 
@@
 
-\newcommand\Symbols{\mathbb{I}}
-\newcommand\Finsets{\mathbb{F}}
+\newcommand\Symbols{\mathbf{I}}
+\newcommand\SymSets{\mathbb{F}_\Symbols}
+\newcommand\FinSets{\mathbb{F}}
 \newcommand\Sorts{\mathcal{S}}
@@
 \newcommand\SCtx{\mathbf{SCtx}}
+\newcommand\Ctx{\mathbf{Ctx}}
+\newcommand\Seq{\mathbf{Seq}}
+
+\newcommand\OpCat[1]{#1^{op}}
 
@@
 \newcommand\Pair[2]{\langle #1, #2 \rangle}
+\newcommand\Tuple[1]{\langle #1 \rangle}
 \newcommand\FLift[2]{
@@
 
+\DeclareMathOperator\Yoneda{\mathbf{y}}
+
 \begin{document}
@@
 
-Let $\Symbols$ be an infinite set of symbols. Let $\Finsets$ be the category of
+Let $\Symbols$ be an infinite set of symbols. Let $\SymSets$ be the category of
 finite subsets of $\Symbols$ and their injective maps; then the comma
-construction $\Define{\SCtx}{\Comma{\Finsets}{\Discrete{\Sorts}}}$, with
+construction $\Define{\SCtx}{\Comma{\SymSets}{\Discrete{\Sorts}}}$, with
 $\Discrete{\Sorts}$ the discrete category on the set $\mathcal{S}$, is the
@@
 $\Upsilon,u\mapsto\Upsilon,v$ shall take $\mathsf{get}[u]$ to $\mathsf{get}[v]$.
+
+\begin{definition}
+  A \emph{signature} $\Sigma$ is a set of sorts $\Sorts$ together with a copresheaf $\Operators$ of operators.
+\end{definition}
 
@@
 
+\section{Model Theory}
+
+Let $\FinSets$ be the free cocartesian category on a single object (that is,
+the category of finite cardinals and functions between them); then the comma
+construction $\Define{\Ctx}{\Comma{\FinSets}{\Sorts}}$ is the category of
+variable contexts, and $\Define{\Seq}{\SCtx\times\Ctx\times\Sorts}$ is the category of
+sequents $\Upsilon\parallel\Gamma\vdash s$.  Then, we fix the presheaf category $\Sets^\Seq$ as
+our semantic universe.
+
+\subsection{Signatures and substitution monoidal structures}
+
+Any two presheaves $A,B : \Sets^\Seq$ have a tensor $A\bullet B$,
+defined as a coend in the following way:
+\[
+  \Define{(A\bullet B)(\Upsilon\parallel\Gamma\vdash s)}{
+    \int^{\Delta\in\Ctx}
+      A(\Upsilon\parallel\Delta\vdash s)
+        \otimes \prod_{x\in\Dom\Delta} B(\Upsilon\parallel\Gamma\vdash\Delta(x))
+  }
+\]
+
+The presheaf of variables $\Define{V(\Upsilon\parallel\Gamma\vdash s)}{\MkSet{x \in \Dom\Gamma
+\mid \Gamma(x) = s}}$ is the unit to this tensor, and together they comprise a
+substitution monoidal structure. Then, a presheaf $A$ is a monoid if it is
+equipped with the following natural transformations:
+
+\[
+  \begin{tikzcd}[ampersand replacement = \&]
+    {\IMode{V}} \arrow[r, "\IMode\nu"]
+    \& {\IMode{A}}
+    \& {\IMode{A\bullet A}}
+    \arrow[l, "\IMode\varsigma", swap]
+  \end{tikzcd}
+\]
+
+Intuitively, $\nu$ embeds variables into $A$ and $\varsigma$ equips $A$ with an
+operation for simultaneous substitutions of variables.
+
 \section{Case Study: Wellformed Sequents}
@@
 
+\section{Code}
+
+\CatchFileBetweenTags\tmp{latex/Code.tex}{agda}
+
+\subsection{Nat}
+
+\ExecuteMetaData[latex/Code.tex]{nat}
+
+\subsection{Fin}
+
+\ExecuteMetaData[latex/Code.tex]{fin}
+
+\subsection{Name}
+
+\ExecuteMetaData[latex/Code.tex]{name}
+
+\subsection{Ctx}
+
+\ExecuteMetaData[latex/Code.tex]{ctx}
+
+\subsection{Signatures}
+
+\ExecuteMetaData[latex/Code.tex]{sign}
+
+\subsection{Trees}
+
+\ExecuteMetaData[latex/Code.tex]{trees}
+
 \nocite{*}

### 2015-10-27 paper.tex
Formatting
Add explanation of signature endofunctor & initial algebra

@@
 
-\DeclareMathOperator\Yoneda{\mathbf{y}}
+\newcommand\SigFun[1]{\mathcal{F}_{#1}}
 
@@
 
-\subsection{Signatures and substitution monoidal structures}
+\subsection{Substitution monoidal structures}
 
@@
 \mid \Gamma(x) = s}}$ is the unit to this tensor, and together they comprise a
-substitution monoidal structure. Then, a presheaf $A$ is a monoid if it is
+\emph{substitution monoidal structure}. Then, a presheaf $A$ is a monoid if it is
 equipped with the following natural transformations:
@@
 operation for simultaneous substitutions of variables.
+
+\subsection{The signature endofunctor and its initial algebras}
+
+For each signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we have an
+endofunctor $\Of{\SigFun{\Sigma}}{\Sets^\Seq\to\Sets^\Seq}$, which is defined
+as follows:
+
+\begin{align*}
+  \IMode{\SigFun{\Sigma}(X)(\Upsilon\parallel\Gamma\vdash s)}
+    &\triangleq
+    \OMode{
+      \coprod_{\vartheta\in\Operators\Pair{\Upsilon}{\MkArity{\vec{v}}{s}}}
+      \prod_{\MkValence{\vec{p}}{\vec{q}}{s'}\in\vec{v}}
+      X(\Upsilon,\vec{p}\parallel\Gamma,\vec{q}\vdash s')
+    }
+\end{align*}
+
+%% TODO: in the above, \vec{q} is a list of sorts, but we need to pick out symbol
+%% names for these before we can extend \Upsilon! Or do we?
+
+Then, a $\Sigma$-model is a monoid $A$ which is equipped with an initial
+algebra $\Of{\alpha}{\SigFun{\Sigma}(A)\to A}$, which shall interpret
+applications of each operator.
 

### 2015-10-28 paper.tex
Try just papering over notation w/ a "remark"
@freebroccolo I wonder if this is fine? It got really gnarly-looking
when I tried to explicitly do the quantification.

@@
 
+\begin{remark}
+  A \emph{relative monad} on a functor $\Of{J}{\mathbf{J}\to\mathbf{C}}$ is a
+  monoid in the functor category $\mathbf{C}^\mathbf{J}$
+  \cite{altenkirch-chapman-uustalu:2015}. A presheaf $A$ with a substitution
+  monoidal structure $\Pair{\nu}{\varsigma}$ can be viewed as a relative monad
+  on the presheaf $\Of{V}{\Sets^\Seq}$ of variables.
+\end{remark}
+
 \subsection{The signature endofunctor and its initial algebras}
@@
 
-%% TODO: in the above, \vec{q} is a list of sorts, but we need to pick out symbol
-%% names for these before we can extend \Upsilon! Or do we?
+\begin{remark}
+  By $\Upsilon,\vec{p}$ we mean any extension of $\Upsilon$ by fresh symbols $\vec{u}$ of sorts $\vec{p}$.
+\end{remark}
 

### 2015-10-29 paper.tex
Rearrange semantic universe; start defining interpretation

--- 2015-10-27 paper.tex
@@
 \newcommand\Seq{\mathbf{Seq}}
-
-\newcommand\OpCat[1]{#1^{op}}
+\newcommand\HypCat{\mathbf{H}}
+
+\newcommand\OpCat[1]{#1^{\mathsf{op}}}
 
@@
 \newcommand\Define[2]{\IMode{#1}\triangleq\OMode{#2}}
-
-\DeclareMathOperator\Yoneda{\mathbf{y}}
+\newcommand\ADefine[2]{\IMode{#1}&\triangleq\OMode{#2}}
+
+\newcommand\SigFun[1]{\mathcal{F}_{#1}}
+
+\newcommand\SemBrackets[1]{\left\llbracket #1\right\rrbracket}
+\newcommand\Yoneda[1]{\boldsymbol{y}(#1)}
 
@@
 construction $\Define{\Ctx}{\Comma{\FinSets}{\Sorts}}$ is the category of
-variable contexts, and $\Define{\Seq}{\SCtx\times\Ctx\times\Sorts}$ is the category of
-sequents $\Upsilon\parallel\Gamma\vdash s$.  Then, we fix the presheaf category $\Sets^\Seq$ as
-our semantic universe.
-
-\subsection{Signatures and substitution monoidal structures}
-
-Any two presheaves $A,B : \Sets^\Seq$ have a tensor $A\bullet B$,
+variable contexts. Let $\Define{\HypCat}{\OpCat{(\SCtx\times\Ctx)}}$; then we
+fix the functor category $\widehat{\HypCat}^\Sorts$ as our semantic universe.
+
+\subsection{Substitution monoidal structures}
+
+For a presheaf $A:\widehat{\HypCat}$ and a sort-indexed family of presheaves
+$P:\widehat{\HypCat}^\Sorts$, we have an operation $A\bullet P$,
 defined as a coend in the following way:
 \[
-  \Define{(A\bullet B)(\Upsilon\parallel\Gamma\vdash s)}{
-    \int^{\Delta\in\Ctx}
-      A(\Upsilon\parallel\Delta\vdash s)
-        \otimes \prod_{x\in\Dom\Delta} B(\Upsilon\parallel\Gamma\vdash\Delta(x))
-  }
-\]
-
-The presheaf of variables $\Define{V(\Upsilon\parallel\Gamma\vdash s)}{\MkSet{x \in \Dom\Gamma
-\mid \Gamma(x) = s}}$ is the unit to this tensor, and together they comprise a
-substitution monoidal structure. Then, a presheaf $A$ is a monoid if it is
-equipped with the following natural transformations:
-
+  \Define{(A\bullet P)(\Upsilon\parallel\Gamma)}{
+    \int^{\Delta\in\OpCat\Ctx}
+      A(\Upsilon\parallel\Delta)
+      \times \prod_{x\in\Dom\Delta} P_{\Delta(x)}(\Upsilon\parallel\Gamma)
+  }
+\]
+
+Using this, we can define a tensor $P\odot Q$ for
+$\Of{P,Q}{\widehat{\HypCat}^\Sorts}$ as follows:
+\[
+  \HypJ{
+    \Define{(P\odot Q)_s}{
+      P_s\bullet Q
+    }
+    \quad
+  }{
+    \Member{s}{\Sorts}
+  }
+\]
+
+Then, variables object
+$\Define{V_s(\Upsilon\parallel\Gamma)}{\MkSet{x\in\Dom\Gamma\mid \Gamma(x)=s}}$
+is the unit to this tensor. We will say that an object
+$\Of{P}{\widehat{\HypCat}^\Sorts}$ is a $\Sigma$-monoid in case it is equipped
+with the following natural transformations:
 \[
@@
     {\IMode{V}} \arrow[r, "\IMode\nu"]
-    \& {\IMode{A}}
-    \& {\IMode{A\bullet A}}
+    \& {\IMode{P}}
+    \& {\IMode{P\odot P}}
     \arrow[l, "\IMode\varsigma", swap]
@@
 \]
-
-Intuitively, $\nu$ embeds variables into $A$ and $\varsigma$ equips $A$ with an
+Intuitively, $\nu$ embeds variables into $P$ and $\varsigma$ equips $P$ with an
 operation for simultaneous substitutions of variables.
+
+\begin{remark}
+  A \emph{relative monad} on a functor $\Of{J}{\mathbf{J}\to\mathbf{C}}$ is a
+  monoid in the functor category $\mathbf{C}^\mathbf{J}$
+  \cite{altenkirch-chapman-uustalu:2015}. A presheaf $P$ with a substitution
+  monoidal structure $\Pair{\nu}{\varsigma}$ can be viewed as a relative monad
+  on the functor $\Of{V}{\widehat{\HypCat}^\Sorts}$ of variables.
+\end{remark}
+
+For any context $\Gamma$, we have an induced map
+$\Of{\nu_\Gamma}{\left(\prod_{x\in\Dom\Gamma}V_{\Gamma(x)}\right)\to
+\left(\prod_{x\in\Dom\Gamma}P_{\Gamma(x)}\right)}$.
+
+\subsection{The signature endofunctor and its initial algebras}
+
+For each signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we have an
+endofunctor
+$\Of{\SigFun{\Sigma}}{\widehat{\HypCat}^\Sorts\to\widehat{\HypCat}^\Sorts}$,
+which is defined as follows:
+
+\begin{align*}
+  \IMode{\SigFun{\Sigma}(X)_s(\Upsilon\parallel\Gamma)}
+    &\triangleq
+    \OMode{
+      \coprod_{\vartheta\in\Operators\Pair{\Upsilon}{\MkArity{\vec{v}}{s}}}
+      \prod_{\MkValence{\vec{p}}{\vec{q}}{s'}\in\vec{v}}
+      X_{s'}(\Upsilon,\vec{p}\parallel\Gamma,\vec{q})
+    }
+\end{align*}
+
+\begin{remark}
+  By $\Upsilon,\vec{p}$ we mean any extension of $\Upsilon$ by fresh symbols $\vec{u}$ of sorts $\vec{p}$.
+\end{remark}
+
+Then, a $\Sigma$-model is a $\Sigma$-monoid $P$ which is equipped with an initial
+algebra $\Of{\alpha}{\SigFun{\Sigma}(P)\to P}$, which shall interpret
+applications of each operator.
+
+\subsection{Interpretations of terms}
+
+\[
+  \Define{
+    \SemBrackets{
+      \Omega\triangleright\Upsilon\parallel\Gamma
+    }_P
+  }{
+    \left(\prod_{(\MV{m}:\MkValence{\vec{p}}{\vec{q}}{s})\in\Omega}
+    P_{s}^{\Yoneda{\vec{p}\parallel\vec{q}})}\right)
+    \times
+    \prod_{x\in\Dom\Gamma}V_{\Gamma(x)}
+  } %% ???
+\]
+
+Let $\Of{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash
+M:s}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P\to P_s}$ be
+defined as follows:
+\begin{align*}
+  \ADefine{
+    \SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x_j : s}_P
+  }{
+    \begin{tikzcd}[ampersand replacement = \&]
+      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P} \arrow[r, "\pi_2"]
+      \& { \prod_{x\in\Dom\Gamma}V_{\Gamma(x)} } \arrow[r, "\nu_{\Gamma}"]
+      \& {\prod_{x\in\Dom\Gamma} P_{\Gamma(x)}} \arrow[r, "\pi_j"]
+      \& P_s
+    \end{tikzcd}
+  }
+\end{align*}
 

### 2015-10-30 paper.tex
Describe relative monads as skew monoids

@@
   A \emph{relative monad} on a functor $\Of{J}{\mathbf{J}\to\mathbf{C}}$ is a
-  monoid in the functor category $\mathbf{C}^\mathbf{J}$
-  \cite{altenkirch-chapman-uustalu:2015}. A presheaf $P$ with a substitution
-  monoidal structure $\Pair{\nu}{\varsigma}$ can be viewed as a relative monad
-  on the functor $\Of{V}{\widehat{\HypCat}^\Sorts}$ of variables.
+  skew-monoid in the skew-monoidal structure induced by the functor category
+  $\mathbf{C}^\mathbf{J}$ \cite{altenkirch-chapman-uustalu:2015}. In contrast to
+  monoids, relative monads can also be understood as a generalization of the
+  \emph{Kleisli structure} or \emph{extension system} presentation of monads. A
+  presheaf $P$ with a substitution monoidal structure $\Pair{\nu}{\varsigma}$
+  can be viewed as a relative monad on the functor
+  $\Of{V}{\widehat{\HypCat}^\Sorts}$ of variables.
 \end{remark}

### 2015-10-30 paper.tex
Expand on relative monad part

--- 2015-10-29 paper.tex
@@
 $\Of{P}{\widehat{\HypCat}^\Sorts}$ is a $\Sigma$-monoid in case it is equipped
-with the following natural transformations:
+with the following natural transformations where $\nu$ embeds variables into $P$
+and $\varsigma$ equips $P$ with an operation for simultaneous substitutions of
+variables.
+
 \[
@@
 \]
-Intuitively, $\nu$ embeds variables into $P$ and $\varsigma$ equips $P$ with an
-operation for simultaneous substitutions of variables.
 
 \begin{remark}
-  A \emph{relative monad} on a functor $\Of{J}{\mathbf{J}\to\mathbf{C}}$ is a
-  monoid in the functor category $\mathbf{C}^\mathbf{J}$
-  \cite{altenkirch-chapman-uustalu:2015}. A presheaf $P$ with a substitution
-  monoidal structure $\Pair{\nu}{\varsigma}$ can be viewed as a relative monad
-  on the functor $\Of{V}{\widehat{\HypCat}^\Sorts}$ of variables.
+  A presheaf $P$ with a substitution monoidal structure $\Pair{\nu}{\varsigma}$
+  also forms a \emph{relative monad} on the functor
+  $\Of{V}{\widehat{\HypCat}^\Sorts}$ of variables.\footnote{Relative monads are
+    skew-monoids in the skew-monoidal structure induced by a functor category
+    $\mathbf{C}^\mathbf{J}$ \cite{altenkirch-chapman-uustalu:2015}. However,
+    relative monads can also be understood without reference to monoids by
+    viewing them as a generalization of \emph{extension systems} or
+    \emph{Kleisli structures}. This latter presentation of monads, popular in
+    computer science and logic, does not involve iteration of the monad
+    functor.} Consequently, the monoid multiplication $\odot$ can be replaced
+  with an extension operation:
+  \[
+    \begin{tikzcd}[ampersand replacement = \&]
+      {\IMode{P(\,=\,)^{V(\,-\,)}}} \arrow[r, "(\,-\,)^{P}"]
+      \& {\IMode{P(\,=\,)^{P(\,-\,)}}}
+    \end{tikzcd}
+  \]
 \end{remark}

### 2015-10-30 paper.tex
Move footnote; formatting; notation

@@
   A presheaf $P$ with a substitution monoidal structure $\Pair{\nu}{\varsigma}$
-  also forms a \emph{relative monad} on the functor
-  $\Of{V}{\widehat{\HypCat}^\Sorts}$ of variables.\footnote{Relative monads are
-    skew-monoids in the skew-monoidal structure induced by a functor category
+  also forms a \emph{relative monad}\footnote{Relative monads are skew-monoids
+    in the skew-monoidal structure induced by a functor category
     $\mathbf{C}^\mathbf{J}$ \cite{altenkirch-chapman-uustalu:2015}. However,
@@
     computer science and logic, does not involve iteration of the monad
-    functor.} Consequently, the monoid multiplication $\odot$ can be replaced
-  with an extension operation:
-  \[
-    \begin{tikzcd}[ampersand replacement = \&]
-      {\IMode{P(\,=\,)^{V(\,-\,)}}} \arrow[r, "(\,-\,)^{P}"]
+    functor.} on the functor $\Of{V}{\widehat{\HypCat}^\Sorts}$ of variables.
+  Consequently, the monoid multiplication $\odot$ can be replaced with an
+  extension operation:
+  \[
+    \begin{tikzcd}[ampersand replacement = \&, sep = huge]
+      {\IMode{P(\,=\,)^{V(\,-\,)}}} \arrow[r, "(\,-\,)^{\circledast}"]
       \& {\IMode{P(\,=\,)^{P(\,-\,)}}}

### 2015-11-01 paper.tex
Add ends and Kan extensions to agda code
Define substitution tensor in Agda
Fix part about replacing multiplication
Fix up rendering
notation changes; add symbols to interp; tentative interp of metavars

@@
   activate={true,nocompatibility},
-  final,
+  draft,
   tracking=true
@@
 \newcommand\Grothendieck[2]{\oint^{#1} #2#1}
-\newcommand\Pair[2]{\langle #1, #2 \rangle}
-\newcommand\Tuple[1]{\langle #1 \rangle}
+\newcommand\Pair[2]{\left\langle #1, #2 \right\rangle}
+\newcommand\Tuple[1]{\left\langle #1 \right\rangle}
 \newcommand\FLift[2]{
@@
 
-Then, variables object
+Then, the variables object
 $\Define{V_s(\Upsilon\parallel\Gamma)}{\MkSet{x\in\Dom\Gamma\mid \Gamma(x)=s}}$
-is the unit to this tensor. We will say that an object
+is the unit to this tensor; additionally, we have a presheaf of symbols
+$\Define{Y_s(\Upsilon\parallel\Gamma)}{\MkSet{u\in\Dom\Upsilon\mid
+\Upsilon(x)=s}}$. We will say that an object
 $\Of{P}{\widehat{\HypCat}^\Sorts}$ is a $\Sigma$-monoid in case it is equipped
-with the following natural transformations where $\nu$ embeds variables into $P$
-and $\varsigma$ equips $P$ with an operation for simultaneous substitutions of
-variables.
+with the following natural transformations where $\nu$ embeds variables into
+$P$ and $\varsigma$ equips $P$ with an operation for simultaneous substitutions
+of variables.
 
@@
 \]
+
+For an object $\Of{P}{\widehat{\HypCat}^\Sorts}$, we will use the notation
+$P^{[\Gamma]}$ to mean $\prod_{x\in\Dom\Gamma}P_{\Gamma(x)}$; likewise,
+$Y^{\{\Upsilon\}}$ shall mean $\prod_{u\in\Dom\Upsilon}Y_{\Upsilon(u)}$.
+Additionally, we have an induced map $\Of{\nu_\Gamma}{V^{[\Gamma]}\to
+P^{[\Gamma]}}$.
 
@@
     functor.} on the functor $\Of{V}{\widehat{\HypCat}^\Sorts}$ of variables.
-  Consequently, the monoid multiplication $\odot$ can be replaced with an
+  Consequently, the monoid multiplication $\varsigma$ can be replaced with an
   extension operation:
@@
 \end{remark}
-
-For any context $\Gamma$, we have an induced map
-$\Of{\nu_\Gamma}{\left(\prod_{x\in\Dom\Gamma}V_{\Gamma(x)}\right)\to
-\left(\prod_{x\in\Dom\Gamma}P_{\Gamma(x)}\right)}$.
 
@@
 
-\subsection{Interpretations of terms}
+\subsection{Interpretation of terms}
 
@@
     \times
-    \prod_{x\in\Dom\Gamma}V_{\Gamma(x)}
-  } %% ???
-\]
-
-Let $\Of{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash
-M:s}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P\to P_s}$ be
-defined as follows:
-\begin{align*}
-  \ADefine{
-    \SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x_j : s}_P
-  }{
+    Y^{\{\Upsilon\}}
+    \times
+    V^{[\Gamma]}
+  }
+\]
+
+We define the interpretation of a term $M$ in the model $P$ as a morphism
+$\Of{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash
+M:s}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P\to P_s}$.
+The interpretation of a variable
+$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x_j:s}_P$ is
+the following composite:
+
+\[
   \begin{tikzcd}[ampersand replacement = \&]
-      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P} \arrow[r, "\pi_2"]
-      \& { \prod_{x\in\Dom\Gamma}V_{\Gamma(x)} } \arrow[r, "\nu_{\Gamma}"]
-      \& {\prod_{x\in\Dom\Gamma} P_{\Gamma(x)}} \arrow[r, "\pi_j"]
-      \& P_s
+    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}} \arrow[r, "\IMode{\pi_3}"]
+    \& {\IMode{V^{[\Gamma]}}} \arrow[r, "\IMode{\nu_{\Gamma}}"]
+    \& {\IMode{P^{[\Gamma]}}} \arrow[r, "\IMode{\pi_j}"]
+    \& \IMode{P_s}
   \end{tikzcd}
-  }
-\end{align*}
+  \tag{var}
+\]
+
+The interpretation of metavariable applications
+$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash
+\MApp{\MV{m}_i}{u_0,\dots,u_m}{M_0,\dots,M_n} : s}_P$, where
+$\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{p}}{\vec{q}}{s}}$, is the following composite
+\[
+  \begin{tikzcd}[ampersand replacement = \&, sep=large]
+    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}} \arrow[r, "\IMode{\Tuple{\pi_i\,\pi_1, \phi}}"]
+    \& {\IMode{P_{s'}^{\Yoneda{\vec{p}\parallel\vec{q}}} \times P^{[\vec{q}]}}} \arrow[r, "\IMode{\varsigma_{\vec{q}}}"]
+    \& \IMode{P_s}
+  \end{tikzcd}
+  \tag{metavar}
+\]
+where
+\[
+  \Define{\phi}{
+    \Tuple{
+      \SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M_j:q_j}_P
+    }_{j\leq n}
+  }
+\]
 
@@
 
-\subsection{Nat}
-
-\ExecuteMetaData[latex/Code.tex]{nat}
-
-\subsection{Fin}
-
-\ExecuteMetaData[latex/Code.tex]{fin}
-
-\subsection{Name}
-
-\ExecuteMetaData[latex/Code.tex]{name}
-
-\subsection{Ctx}
-
-\ExecuteMetaData[latex/Code.tex]{ctx}
+\subsection{Kan Extensions}
+
+\subsubsection{Lan}
+
+\ExecuteMetaData[latex/Code.tex]{lang}\\
+
+\ExecuteMetaData[latex/Code.tex]{lan}
+
+\subsubsection{Ran}
+
+\ExecuteMetaData[latex/Code.tex]{rang}\\
+
+\ExecuteMetaData[latex/Code.tex]{ran}
+
+\subsection{Variables and Symbols}
+
+\ExecuteMetaData[latex/Code.tex]{var}\\
+
+\ExecuteMetaData[latex/Code.tex]{sym}
+
+\subsection{Contexts}
+
+\ExecuteMetaData[latex/Code.tex]{ctx}\\
+
+\ExecuteMetaData[latex/Code.tex]{sctx}\\
+
+\ExecuteMetaData[latex/Code.tex]{ctxdom}\\
+
+\ExecuteMetaData[latex/Code.tex]{ctxidx}\\
+
+\ExecuteMetaData[latex/Code.tex]{elem}
 
@@
 \ExecuteMetaData[latex/Code.tex]{trees}
+
+\subsection{Substitution Monoidal Structure}
+
+\ExecuteMetaData[latex/Code.tex]{H}\\
+
+\ExecuteMetaData[latex/Code.tex]{HHat}\\
+
+\ExecuteMetaData[latex/Code.tex]{V}\\
+
+\ExecuteMetaData[latex/Code.tex]{tensor0}\\
+
+\ExecuteMetaData[latex/Code.tex]{tensor1}\\
+
+\ExecuteMetaData[latex/Code.tex]{extension}
 

### 2015-11-01 paper.tex
Fix symbol-related bugs

@@
 
+Let $\Define{V_s(\Upsilon\parallel\Gamma)}{\MkSet{x\in\Dom\Gamma\mid
+\Gamma(x)=s}}$ be called the presheaf of variables; additionally, we have a
+presheaf of symbols
+$\Define{Y_s(\Upsilon\parallel\Gamma)}{\MkSet{u\in\Dom\Upsilon\mid
+\Upsilon(x)=s}}$. For an object $\Of{P}{\widehat{\HypCat}^\Sorts}$, we will
+use the notation $P^{[\Gamma]}$ to mean $\prod_{x\in\Dom\Gamma}P_{\Gamma(x)}$;
+likewise, $Y^{\{\Upsilon\}}$ shall mean
+$\prod_{u\in\Dom\Upsilon}Y_{\Upsilon(u)}$.
+
 For a presheaf $A:\widehat{\HypCat}$ and a sort-indexed family of presheaves
@@
   \Define{(A\bullet P)(\Upsilon\parallel\Gamma)}{
-    \int^{\Delta\in\OpCat\Ctx}
-      A(\Upsilon\parallel\Delta)
-      \times \prod_{x\in\Dom\Delta} P_{\Delta(x)}(\Upsilon\parallel\Gamma)
+    \int^{(\Upsilon'\parallel\Delta)\in\HypCat}
+      A(\Upsilon'\parallel\Delta)
+      \times Y^{\{\Upsilon'\}}(\Upsilon\parallel\Gamma)
+      \times P^{[\Delta]}(\Upsilon\parallel\Gamma)
   }
@@
 
-Then, the variables object
-$\Define{V_s(\Upsilon\parallel\Gamma)}{\MkSet{x\in\Dom\Gamma\mid \Gamma(x)=s}}$
-is the unit to this tensor; additionally, we have a presheaf of symbols
-$\Define{Y_s(\Upsilon\parallel\Gamma)}{\MkSet{u\in\Dom\Upsilon\mid
-\Upsilon(x)=s}}$. We will say that an object
+Then, $V$ is the unit to this tensor. We will say that an object
 $\Of{P}{\widehat{\HypCat}^\Sorts}$ is a $\Sigma$-monoid in case it is equipped
@@
 
-For an object $\Of{P}{\widehat{\HypCat}^\Sorts}$, we will use the notation
-$P^{[\Gamma]}$ to mean $\prod_{x\in\Dom\Gamma}P_{\Gamma(x)}$; likewise,
-$Y^{\{\Upsilon\}}$ shall mean $\prod_{u\in\Dom\Upsilon}Y_{\Upsilon(u)}$.
-Additionally, we have an induced map $\Of{\nu_\Gamma}{V^{[\Gamma]}\to
-P^{[\Gamma]}}$.
+%%TODO
+We also have an induced maps $\Of{\nu_\Gamma}{V^{[\Gamma]}\to P^{[\Gamma]}}$
+and
+$\Of{\varsigma_{\Upsilon\parallel\Gamma}}{A^{\Yoneda{\Upsilon\parallel\Gamma}} \times
+Y^{\{\Upsilon\}}\times A^{[\Gamma]}}$.
 
@@
 
+The metavariable, symbol and variable contexts are interpreted in a model $P$ as an
+environment presheaf in the following way:
 \[
@@
 
-We define the interpretation of a term $M$ in the model $P$ as a morphism
+Then, the interpretation of a term in a model $P$ is a map from its environment
+to $P$, that is,
 $\Of{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash
@@
 The interpretation of a variable
-$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x_j:s}_P$ is
-the following composite:
-
+$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x_j:s}_P$
+simply projects it of the environment and embeds it into the
+model, as follows:
 \[
@@
 \MApp{\MV{m}_i}{u_0,\dots,u_m}{M_0,\dots,M_n} : s}_P$, where
-$\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{p}}{\vec{q}}{s}}$, is the following composite
-\[
-  \begin{tikzcd}[ampersand replacement = \&, sep=large]
-    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}} \arrow[r, "\IMode{\Tuple{\pi_i\,\pi_1, \phi}}"]
-    \& {\IMode{P_{s'}^{\Yoneda{\vec{p}\parallel\vec{q}}} \times P^{[\vec{q}]}}} \arrow[r, "\IMode{\varsigma_{\vec{q}}}"]
+$\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{p}}{\vec{q}}{s}}$, proceeds by projecting the metavariable's interpretation from the environment and intantiating it via substitution:
+\[
+  \begin{tikzcd}[ampersand replacement = \&, sep=huge]
+    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
+          \arrow[r, "\IMode{\Tuple{\pi_i\,\pi_1, \phi\,\pi_2, \psi}}"]
+    \& {\IMode{P_{s'}^{\Yoneda{\vec{p}\parallel\vec{q}}}\times Y^{\{\vec{p}\}} \times P^{[\vec{q}]}}}
+          \arrow[r, "\IMode{\varsigma_{\vec{p}\parallel\vec{q}}}"]
     \& \IMode{P_s}
@@
 where
-\[
-  \Define{\phi}{
+\begin{align*}
+  \ADefine{\phi}{
+    \Dom{\vec{p}}\mapsto \vec{u}
+  }\\
+  \ADefine{\psi}{
     \Tuple{
@@
   }
-\]
+\end{align*}
 

### 2015-11-01 paper.tex
update pdf
fix typo

@@
   activate={true,nocompatibility},
-  draft,
+  final,
   tracking=true
@@
 and
-$\Of{\varsigma_{\Upsilon\parallel\Gamma}}{A^{\Yoneda{\Upsilon\parallel\Gamma}} \times
-Y^{\{\Upsilon\}}\times A^{[\Gamma]}}$.
+$\Of{\varsigma_{\Upsilon\parallel\Gamma}}{P_s^{\Yoneda{\Upsilon\parallel\Gamma}} \times
+Y^{\{\Upsilon\}}\times P^{[\Gamma]}\to P_s}$.
 

### 2015-11-01 paper.tex
slight change of notation
Y looks too much like Upsilon

@@
 presheaf of symbols
-$\Define{Y_s(\Upsilon\parallel\Gamma)}{\MkSet{u\in\Dom\Upsilon\mid
+$\Define{S_s(\Upsilon\parallel\Gamma)}{\MkSet{u\in\Dom\Upsilon\mid
 \Upsilon(x)=s}}$. For an object $\Of{P}{\widehat{\HypCat}^\Sorts}$, we will
 use the notation $P^{[\Gamma]}$ to mean $\prod_{x\in\Dom\Gamma}P_{\Gamma(x)}$;
-likewise, $Y^{\{\Upsilon\}}$ shall mean
-$\prod_{u\in\Dom\Upsilon}Y_{\Upsilon(u)}$.
+likewise, $S^{\{\Upsilon\}}$ shall mean
+$\prod_{u\in\Dom\Upsilon}S_{\Upsilon(u)}$.
 
@@
       A(\Upsilon'\parallel\Delta)
-      \times Y^{\{\Upsilon'\}}(\Upsilon\parallel\Gamma)
+      \times S^{\{\Upsilon'\}}(\Upsilon\parallel\Gamma)
       \times P^{[\Delta]}(\Upsilon\parallel\Gamma)
@@
 $\Of{\varsigma_{\Upsilon\parallel\Gamma}}{P_s^{\Yoneda{\Upsilon\parallel\Gamma}} \times
-Y^{\{\Upsilon\}}\times P^{[\Gamma]}\to P_s}$.
+S^{\{\Upsilon\}}\times P^{[\Gamma]}\to P_s}$.
 
@@
     \times
-    Y^{\{\Upsilon\}}
+    S^{\{\Upsilon\}}
     \times
@@
           \arrow[r, "\IMode{\Tuple{\pi_i\,\pi_1, \phi\,\pi_2, \psi}}"]
-    \& {\IMode{P_{s'}^{\Yoneda{\vec{p}\parallel\vec{q}}}\times Y^{\{\vec{p}\}} \times P^{[\vec{q}]}}}
+    \& {\IMode{P_{s'}^{\Yoneda{\vec{p}\parallel\vec{q}}}\times S^{\{\vec{p}\}} \times P^{[\vec{q}]}}}
           \arrow[r, "\IMode{\varsigma_{\vec{p}\parallel\vec{q}}}"]

### 2015-11-04 paper.tex
formatting
Add agda interpretation to paper
Had to change notation to get it to render properly, weird
Fix signature endofunctor
In a presheaf category, the exponential is not defined pointwise, what
we had before probably wasn't OK.
Use "final" globally, not just for microtype

--- 2015-11-01 paper.tex
@@
-\documentclass[11pt]{article}
+\documentclass[final,11pt]{article}
 
@@
   activate={true,nocompatibility},
-  draft,
   tracking=true
@@
 \newcommand\Grothendieck[2]{\oint^{#1} #2#1}
-\newcommand\Pair[2]{\left\langle #1, #2 \right\rangle}
-\newcommand\Tuple[1]{\left\langle #1 \right\rangle}
+\newcommand\Pair[2]{\langle #1, #2 \rangle}
+\newcommand\Tuple[1]{\langle #1 \rangle}
 \newcommand\FLift[2]{
@@
 
-Let $\Define{V_s(\Upsilon\parallel\Gamma)}{\MkSet{x\in\Dom\Gamma\mid
-\Gamma(x)=s}}$ be called the presheaf of variables; additionally, we have a
-presheaf of symbols
-$\Define{Y_s(\Upsilon\parallel\Gamma)}{\MkSet{u\in\Dom\Upsilon\mid
-\Upsilon(x)=s}}$. For an object $\Of{P}{\widehat{\HypCat}^\Sorts}$, we will
-use the notation $P^{[\Gamma]}$ to mean $\prod_{x\in\Dom\Gamma}P_{\Gamma(x)}$;
-likewise, $Y^{\{\Upsilon\}}$ shall mean
-$\prod_{u\in\Dom\Upsilon}Y_{\Upsilon(u)}$.
-
 For a presheaf $A:\widehat{\HypCat}$ and a sort-indexed family of presheaves
@@
   \Define{(A\bullet P)(\Upsilon\parallel\Gamma)}{
-    \int^{(\Upsilon'\parallel\Delta)\in\HypCat}
-      A(\Upsilon'\parallel\Delta)
-      \times Y^{\{\Upsilon'\}}(\Upsilon\parallel\Gamma)
-      \times P^{[\Delta]}(\Upsilon\parallel\Gamma)
+    \int^{\Delta\in\OpCat\Ctx}
+      A(\Upsilon\parallel\Delta)
+      \times \prod_{x\in\Dom\Delta} P_{\Delta(x)}(\Upsilon\parallel\Gamma)
   }
@@
 
-Then, $V$ is the unit to this tensor. We will say that an object
+Then, variables object
+$\Define{V_s(\Upsilon\parallel\Gamma)}{\MkSet{x\in\Dom\Gamma\mid \Gamma(x)=s}}$
+is the unit to this tensor. We will say that an object
 $\Of{P}{\widehat{\HypCat}^\Sorts}$ is a $\Sigma$-monoid in case it is equipped
-with the following natural transformations where $\nu$ embeds variables into
-$P$ and $\varsigma$ equips $P$ with an operation for simultaneous substitutions
-of variables.
+with the following natural transformations where $\nu$ embeds variables into $P$
+and $\varsigma$ equips $P$ with an operation for simultaneous substitutions of
+variables.
 
@@
 \]
-
-%%TODO
-We also have an induced maps $\Of{\nu_\Gamma}{V^{[\Gamma]}\to P^{[\Gamma]}}$
-and
-$\Of{\varsigma_{\Upsilon\parallel\Gamma}}{A^{\Yoneda{\Upsilon\parallel\Gamma}} \times
-Y^{\{\Upsilon\}}\times A^{[\Gamma]}}$.
 
@@
 
+For any context $\Gamma$, we have an induced map
+$\Of{\nu_\Gamma}{\left(\prod_{x\in\Dom\Gamma}V_{\Gamma(x)}\right)\to
+\left(\prod_{x\in\Dom\Gamma}P_{\Gamma(x)}\right)}$.
+
 \subsection{The signature endofunctor and its initial algebras}
@@
 
-\subsection{Interpretation of terms}
-
-The metavariable, symbol and variable contexts are interpreted in a model $P$ as an
-environment presheaf in the following way:
+\subsection{Interpretations of terms}
+
 \[
@@
     \times
-    Y^{\{\Upsilon\}}
-    \times
-    V^{[\Gamma]}
-  }
-\]
-
-Then, the interpretation of a term in a model $P$ is a map from its environment
-to $P$, that is,
-$\Of{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash
-M:s}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P\to P_s}$.
-The interpretation of a variable
-$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x_j:s}_P$
-simply projects it of the environment and embeds it into the
-model, as follows:
-\[
+    \prod_{x\in\Dom\Gamma}V_{\Gamma(x)}
+  } %% ???
+\]
+
+Let $\Of{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash
+M:s}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P\to P_s}$ be
+defined as follows:
+\begin{align*}
+  \ADefine{
+    \SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x_j : s}_P
+  }{
     \begin{tikzcd}[ampersand replacement = \&]
-    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}} \arrow[r, "\IMode{\pi_3}"]
-    \& {\IMode{V^{[\Gamma]}}} \arrow[r, "\IMode{\nu_{\Gamma}}"]
-    \& {\IMode{P^{[\Gamma]}}} \arrow[r, "\IMode{\pi_j}"]
-    \& \IMode{P_s}
+      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P} \arrow[r, "\pi_2"]
+      \& { \prod_{x\in\Dom\Gamma}V_{\Gamma(x)} } \arrow[r, "\nu_{\Gamma}"]
+      \& {\prod_{x\in\Dom\Gamma} P_{\Gamma(x)}} \arrow[r, "\pi_j"]
+      \& P_s
     \end{tikzcd}
-  \tag{var}
-\]
-
-The interpretation of metavariable applications
-$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash
-\MApp{\MV{m}_i}{u_0,\dots,u_m}{M_0,\dots,M_n} : s}_P$, where
-$\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{p}}{\vec{q}}{s}}$, proceeds by projecting the metavariable's interpretation from the environment and intantiating it via substitution:
-\[
-  \begin{tikzcd}[ampersand replacement = \&, sep=huge]
-    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
-          \arrow[r, "\IMode{\Tuple{\pi_i\,\pi_1, \phi\,\pi_2, \psi}}"]
-    \& {\IMode{P_{s'}^{\Yoneda{\vec{p}\parallel\vec{q}}}\times Y^{\{\vec{p}\}} \times P^{[\vec{q}]}}}
-          \arrow[r, "\IMode{\varsigma_{\vec{p}\parallel\vec{q}}}"]
-    \& \IMode{P_s}
-  \end{tikzcd}
-  \tag{metavar}
-\]
-where
-\begin{align*}
-  \ADefine{\phi}{
-    \Dom{\vec{p}}\mapsto \vec{u}
-  }\\
-  \ADefine{\psi}{
-    \Tuple{
-      \SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M_j:q_j}_P
-    }_{j\leq n}
   }

### 2015-11-05 paper.tex
Comment line endings
Set mode to draft by default
Update paths to agda latex output

--- 2015-11-01 paper.tex
@@
-\documentclass[11pt]{article}
+\documentclass[draft,11pt]{article}
 
@@
   activate={true,nocompatibility},
-  final,
   tracking=true
@@
 \def\CatchFBT@Fin@l#1[#2]{%
-   \begingroup
+   \begingroup%
       \endlinechar\m@ne \makeatletter #2%
-      \ifluatex
-         \immediate\openout\CATCHPATCHOUT=\jobname.catchpatch
+      \ifluatex%
+         \immediate\openout\CATCHPATCHOUT=\jobname.catchpatch%
          \immediate\write\CATCHPATCHOUT {\CatchFBT@tok{\the\CatchFBT@tok}}%
-         \immediate\closeout\CATCHPATCHOUT
-         \@@input \jobname.catchpatch
-      \else
+         \immediate\closeout\CATCHPATCHOUT%
+         \@@input \jobname.catchpatch%
+      \else%
       \scantokens\expandafter{%
          \expandafter\CatchFBT@tok\expandafter{\the\CatchFBT@tok}}%
-      \fi
-      \CatchFBT@IsAToken{#1}
-         {\global#1\expandafter{\the\CatchFBT@tok}}
+      \fi%
+      \CatchFBT@IsAToken{#1}%
+         {\global#1\expandafter{\the\CatchFBT@tok}}%
          {\xdef#1{\the\CatchFBT@tok}}%
-      \ifx\CatchFBT@tok#1\else\global\CatchFBT@tok{}\fi
-   \endgroup
+      \ifx\CatchFBT@tok#1\else\global\CatchFBT@tok{}\fi%
+   \endgroup%
 }% \CatchFBT@Final
@@
     [#2]%
-  }.\, #3
+  }.\, #3%
 }
@@
   \Operators\langle#1%
-  \ifthenelse{\equal{#2}{}}{}{, #2}\rangle
+  \ifthenelse{\equal{#2}{}}{}{, #2}\rangle%
 }
@@
 
-\newcommand\Rename[2]{
-  #2\upharpoonright #1
-}
-\newcommand\IsRenaming[3]{
+\newcommand\Rename[2]{%
+  #2\upharpoonright #1%
+}
+\newcommand\IsRenaming[3]{%
   \Rename{\IMode{#1}}{\IMode{#2}}%
@@
 
-\newcommand\Subst[3]{
+\newcommand\Subst[3]{%
   \left[#1\, /\, #2\right]#3%
@@
 
-\subsection{Substitution monoidal structures}
-
 Let $\Define{V_s(\Upsilon\parallel\Gamma)}{\MkSet{x\in\Dom\Gamma\mid
@@
 $\Define{S_s(\Upsilon\parallel\Gamma)}{\MkSet{u\in\Dom\Upsilon\mid
-\Upsilon(x)=s}}$. For an object $\Of{P}{\widehat{\HypCat}^\Sorts}$, we will
-use the notation $P^{[\Gamma]}$ to mean $\prod_{x\in\Dom\Gamma}P_{\Gamma(x)}$;
-likewise, $S^{\{\Upsilon\}}$ shall mean
-$\prod_{u\in\Dom\Upsilon}S_{\Upsilon(u)}$.
+\Upsilon(x)=s}}$. Lastly, we have the presheaf of operators with arity $a$,
+$\Define{\Operators_a(\Upsilon\parallel\Gamma)}{\Operators\Pair{\Upsilon}{a}}$
+
+\subsection{Substitution monoidal structures}
+
+For an object $\Of{P}{\widehat{\HypCat}^\Sorts}$, we will use the notation
+$P^{[\Gamma]}$ to mean $\prod_{x\in\Dom\Gamma}P_{\Gamma(x)}$; likewise,
+$S^{\{\Upsilon\}}$ shall mean $\prod_{u\in\Dom\Upsilon}S_{\Upsilon(u)}$.
 
@@
 which is defined as follows:
-
 \begin{align*}
-  \IMode{\SigFun{\Sigma}(X)_s(\Upsilon\parallel\Gamma)}
+  \IMode{\SigFun{\Sigma}(X)_s}
     &\triangleq
     \OMode{
-      \coprod_{\vartheta\in\Operators\Pair{\Upsilon}{\MkArity{\vec{v}}{s}}}
+      \coprod_{\vartheta\in \Operators_{\MkArity{\vec{v}}{s}}}
       \prod_{\MkValence{\vec{p}}{\vec{q}}{s'}\in\vec{v}}
-      X_{s'}(\Upsilon,\vec{p}\parallel\Gamma,\vec{q})
+      X_{s'}^{\Yoneda{\vec{p}\parallel\vec{q}}}
     }
 \end{align*}
-
-\begin{remark}
-  By $\Upsilon,\vec{p}$ we mean any extension of $\Upsilon$ by fresh symbols $\vec{u}$ of sorts $\vec{p}$.
-\end{remark}
 
@@
 
-\CatchFileBetweenTags\tmp{latex/Code.tex}{agda}
+\CatchFileBetweenTags\tmp{Code.tex}{agda}
 
@@
 
-\ExecuteMetaData[latex/Code.tex]{lang}\\
-
-\ExecuteMetaData[latex/Code.tex]{lan}
+\ExecuteMetaData[Code.tex]{lang}\\
+
+\ExecuteMetaData[Code.tex]{lan}
 
@@
 
-\ExecuteMetaData[latex/Code.tex]{rang}\\
-
-\ExecuteMetaData[latex/Code.tex]{ran}
+\ExecuteMetaData[Code.tex]{rang}\\
+
+\ExecuteMetaData[Code.tex]{ran}
 
@@
 
-\ExecuteMetaData[latex/Code.tex]{var}\\
-
-\ExecuteMetaData[latex/Code.tex]{sym}
+\ExecuteMetaData[Code.tex]{var}\\
+
+\ExecuteMetaData[Code.tex]{sym}
 
@@
 
-\ExecuteMetaData[latex/Code.tex]{ctx}\\
-
-\ExecuteMetaData[latex/Code.tex]{sctx}\\
-
-\ExecuteMetaData[latex/Code.tex]{ctxdom}\\
-
-\ExecuteMetaData[latex/Code.tex]{ctxidx}\\
-
-\ExecuteMetaData[latex/Code.tex]{elem}
+\ExecuteMetaData[Code.tex]{ctx}\\
+
+\ExecuteMetaData[Code.tex]{sctx}\\
+
+\ExecuteMetaData[Code.tex]{ctxdom}\\
+
+\ExecuteMetaData[Code.tex]{ctxidx}\\
+
+\ExecuteMetaData[Code.tex]{elem}
 
@@
 
-\ExecuteMetaData[latex/Code.tex]{sign}
+\ExecuteMetaData[Code.tex]{sign}
 
@@
 
-\ExecuteMetaData[latex/Code.tex]{trees}
+\ExecuteMetaData[Code.tex]{trees}
 
@@
 
-\ExecuteMetaData[latex/Code.tex]{H}\\
-
-\ExecuteMetaData[latex/Code.tex]{HHat}\\
-
-\ExecuteMetaData[latex/Code.tex]{V}\\
-
-\ExecuteMetaData[latex/Code.tex]{tensor0}\\
-
-\ExecuteMetaData[latex/Code.tex]{tensor1}\\
-
-\ExecuteMetaData[latex/Code.tex]{extension}
+\ExecuteMetaData[Code.tex]{H}\\
+
+\ExecuteMetaData[Code.tex]{HHat}\\
+
+\ExecuteMetaData[Code.tex]{V}\\
+
+\ExecuteMetaData[Code.tex]{tensor0}\\
+
+\ExecuteMetaData[Code.tex]{tensor1}\\
+
+\ExecuteMetaData[Code.tex]{extension}
 

### 2015-11-05 paper.tex
Remove code from document for now until it settles down
formatting of metavariables [try using fraktur]

@@
 \makeatletter
-\ifluatex
-   \newwrite\CATCHPATCHOUT
+\ifluatex%
+   \newwrite\CATCHPATCHOUT%
 \fi
@@
    \begingroup%
-      \endlinechar\m@ne \makeatletter #2%
+      \endlinechar\m@ne\makeatletter #2%
       \ifluatex%
          \immediate\openout\CATCHPATCHOUT=\jobname.catchpatch%
-         \immediate\write\CATCHPATCHOUT {\CatchFBT@tok{\the\CatchFBT@tok}}%
+         \immediate\write\CATCHPATCHOUT{\CatchFBT@tok{\the\CatchFBT@tok}}%
          \immediate\closeout\CATCHPATCHOUT%
-         \@@input \jobname.catchpatch%
+         \@@input\jobname.catchpatch%
       \else%
@@
 \newcommand\EqMember[3]{\IMode{#1}=\IMode{#2}\in\IMode{#3}}
-\newcommand\Of[2]{\IMode{#1} : \IMode{#2}}
+\newcommand\Of[2]{\IMode{#1}: \IMode{#2}}
 
@@
 
-\newcommand\IsOperator[3]{\IMode{#1}\Vdash \IMode{#2} : \OMode{#3}}
+\newcommand\IsOperator[3]{\IMode{#1}\Vdash\IMode{#2}: \OMode{#3}}
 
@@
 \newcommand\Grothendieck[2]{\oint^{#1} #2#1}
-\newcommand\Pair[2]{\left\langle #1, #2 \right\rangle}
-\newcommand\Tuple[1]{\left\langle #1 \right\rangle}
+\newcommand\Pair[2]{\left\langle#1, #2 \right\rangle}
+\newcommand\Tuple[1]{\left\langle#1 \right\rangle}
 \newcommand\FLift[2]{
@@
 }
-\newcommand\PairTimes[2]{\langle #1\, {\scriptstyle {{\times}}}\, #2 \rangle}
+\newcommand\PairTimes[2]{\langle#1\, {\scriptstyle{{\times}}}\, #2 \rangle}
 \newcommand\SSingleton{\bullet}
@@
 
-\newcommand\Comma[2]{#1 \downarrow #2}
+\newcommand\Comma[2]{#1 \downarrow#2}
 \newcommand\Coslice[2]{#1\, /\, #2}
@@
 \newcommand\App[2]{#1(#2)}
-\newcommand\Dom[1]{\left\vert #1\right\vert}
+\newcommand\Dom[1]{\left\vert#1\right\vert}
 \newcommand\FV[1]{\mathbf{FV}\left(#1\right)}
@@
 \newcommand\Rename[2]{%
-  #2\upharpoonright #1%
-}
+  #2\upharpoonright#1%
+}
+
 \newcommand\IsRenaming[3]{%
@@
 
-\newcommand\SemBrackets[1]{\left\llbracket #1\right\rrbracket}
-\newcommand\Yoneda[1]{\boldsymbol{y}(#1)}
+\newcommand\SemBrackets[1]{\left\llbracket#1\right\rrbracket}
+\newcommand\Yoneda[1]{\boldsymbol{y} (#1)}
 
@@
   \infer{
-    \IsArity{\MkArity{v_0,...,v_n}{s}}
+    \IsArity{\MkArity{v_0,\ldots,v_n}{s}}
   }{
@@
   \HypJ{
-    \Define{(P\odot Q)_s}{
+    \Define{{(P\odot Q)}_{s}}{
       P_s\bullet Q
@@
     in the skew-monoidal structure induced by a functor category
-    $\mathbf{C}^\mathbf{J}$ \cite{altenkirch-chapman-uustalu:2015}. However,
+    $\mathbf{C}^\mathbf{J}$~\cite{altenkirch-chapman-uustalu:2015}. However,
     relative monads can also be understood without reference to monoids by
@@
     \begin{tikzcd}[ampersand replacement = \&, sep = huge]
-      {\IMode{P(\,=\,)^{V(\,-\,)}}} \arrow[r, "(\,-\,)^{\circledast}"]
-      \& {\IMode{P(\,=\,)^{P(\,-\,)}}}
+      {\IMode{{P(\,=\,)}^{V(\,-\,)}}} \arrow[r, "{(\,-\,)}^{\circledast}"]
+      \& {\IMode{{P(\,=\,)}^{P(\,-\,)}}}
     \end{tikzcd}
@@
 \begin{align*}
-  \IMode{\SigFun{\Sigma}(X)_s}
+  \IMode{{\SigFun{\Sigma}(X)}_{s}}
     &\triangleq
@@
   }{
-    \left(\prod_{(\MV{m}:\MkValence{\vec{p}}{\vec{q}}{s})\in\Omega}
-    P_{s}^{\Yoneda{\vec{p}\parallel\vec{q}})}\right)
+    \left(\prod_{(m:\MkValence{\vec{p}}{\vec{q}}{s})\in\Omega}P_{s}^{\Yoneda{\vec{p}\parallel\vec{q}}}\right)
     \times
@@
 Crary has demonstrated a first-order encoding of contexts in the logical
-framework in bijection with actual LF-contexts \cite{crary:2009}, which has
+framework in bijection with actual LF-contexts~\cite{crary:2009}, which has
 been successfully used in large-scale mechanization efforts, including that of
-Standard ML \cite{lee-crary-harper:2007} and the Edinburgh Logical Framework
-itself \cite{martens-crary:2012}.
+Standard ML~\cite{lee-crary-harper:2007} and the Edinburgh Logical Framework
+itself~\cite{martens-crary:2012}.
 
@@
 
-\section{Code}
-
-\CatchFileBetweenTags\tmp{Code.tex}{agda}
-
-\subsection{Kan Extensions}
-
-\subsubsection{Lan}
-
-\ExecuteMetaData[Code.tex]{lang}\\
-
-\ExecuteMetaData[Code.tex]{lan}
-
-\subsubsection{Ran}
-
-\ExecuteMetaData[Code.tex]{rang}\\
-
-\ExecuteMetaData[Code.tex]{ran}
-
-\subsection{Variables and Symbols}
-
-\ExecuteMetaData[Code.tex]{var}\\
-
-\ExecuteMetaData[Code.tex]{sym}
-
-\subsection{Contexts}
-
-\ExecuteMetaData[Code.tex]{ctx}\\
-
-\ExecuteMetaData[Code.tex]{sctx}\\
-
-\ExecuteMetaData[Code.tex]{ctxdom}\\
-
-\ExecuteMetaData[Code.tex]{ctxidx}\\
-
-\ExecuteMetaData[Code.tex]{elem}
-
-\subsection{Signatures}
-
-\ExecuteMetaData[Code.tex]{sign}
-
-\subsection{Trees}
-
-\ExecuteMetaData[Code.tex]{trees}
-
-\subsection{Substitution Monoidal Structure}
-
-\ExecuteMetaData[Code.tex]{H}\\
-
-\ExecuteMetaData[Code.tex]{HHat}\\
-
-\ExecuteMetaData[Code.tex]{V}\\
-
-\ExecuteMetaData[Code.tex]{tensor0}\\
-
-\ExecuteMetaData[Code.tex]{tensor1}\\
-
-\ExecuteMetaData[Code.tex]{extension}
-
 \nocite{*}

### 2015-11-06 paper.tex
Fix script option passing behavior
Apparently options to the class are only globally inherited by loaded
packages if the options are written directly on the class line within
the document. If the options are specified with `\PassOptionsToClass`
they only affect the class itself so we need to also provide a
`\PassOptionsToPackage` for each individual package.
Use draft for more pkgs; don't run bibtex in draft
Fix some warnings
Fix some overfulls

@@
-\documentclass[draft,11pt]{article}
+\documentclass[11pt]{article}
 
@@
 \usepackage{hyperref}
+\usepackage{ifdraft}
 \usepackage{ifthen}
-\usepackage{mathtools}
+%\usepackage{mathtools}
 \usepackage{proof}
@@
 Then, the interpretation of a term in a model $P$ is a map from its environment
-to $P$, that is,
-$\Of{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash
-M:s}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P\to P_s}$.
+to $P$:
+\[%
+  \Of{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash%
+      M:s}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P\to P_s}%
+\]%
 The interpretation of a variable
-$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x_j:s}_P$
-simply projects it of the environment and embeds it into the
-model, as follows:
+$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x_j:s}_P$ simply
+projects it of the environment and embeds it into the model, as follows:
 \[
@@
 
-The interpretation of metavariable applications
-$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash
-\MApp{\MV{m}_i}{u_0,\dots,u_m}{M_0,\dots,M_n} : s}_P$, where
-$\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{p}}{\vec{q}}{s}}$, proceeds by projecting the metavariable's interpretation from the environment and intantiating it via substitution:
+Next we have the interpretation of metavariable applications
+\[%
+  \SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash%
+    \MApp{\MV{m}_i}{u_0,\dots,u_m}{M_0,\dots,M_n} : s}_P%
+  \qquad(\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{p}}{\vec{q}}{s}})
+\]
+which proceeds by projecting the metavariable's interpretation from the
+environment and intantiating it via substitution:
 \[
@@
 
+\ifdraft{}{
   \nocite{*}
@@
   \bibliography{refs}
+}
 

### 2015-11-07 paper.tex
Use lualatex-math
[paper] fix up metavariable interpretation

@@
 \usepackage{ifthen}
+\usepackage{lualatex-math}
 %\usepackage{mathtools}
@@
 
-For a presheaf $A:\widehat{\HypCat}$ and a sort-indexed family of presheaves
-$P:\widehat{\HypCat}^\Sorts$, we have an operation $A\bullet P$,
-defined as a coend in the following way:
+For a presheaf $\Of{A}{\widehat{\HypCat}}$ and a sort-indexed family of
+presheaves $\Of{P}{\widehat{\HypCat}^\Sorts}$, we have an operation $A\bullet
+P$, defined as a coend in the following way:
 \[
@@
 %%TODO
-We also have an induced maps $\Of{\nu_\Gamma}{V^{[\Gamma]}\to P^{[\Gamma]}}$
+We also have the induced maps $\Of{\nu_\Gamma}{V^{[\Gamma]}\to P^{[\Gamma]}}$
 and
@@
   \SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash%
-    \MApp{\MV{m}_i}{u_0,\dots,u_m}{M_0,\dots,M_n} : s}_P%
+    \MApp{\MV{m}_i}{\vec{u}}{\vec{M}} : s}_P%
   \qquad(\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{p}}{\vec{q}}{s}})
@@
   \ADefine{\phi}{
-    \Dom{\vec{p}}\mapsto \vec{u}
+  \begin{tikzcd}[ampersand replacement = \&, sep=huge]
+    {S^{\{\Upsilon\}}} \arrow[r, "\Tuple{\pi_u}_{u\in\vec{u}}"]
+    \& {S^{\{\vec{p}\}}}
+  \end{tikzcd}
   }\\
   \ADefine{\psi}{
-    \Tuple{
-      \SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M_j:q_j}_P
-    }_{j\leq n}
+   \begin{tikzcd}[ampersand replacement = \&, sep=28ex]
+     {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}
+        \arrow[r, "\Tuple{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M:q}_P}_{(M,q)\in \vec{M}\otimes\vec{q}}"]
+     \& {P^{[\vec{q}]}}
+   \end{tikzcd}
   }

### 2015-11-07 paper.tex
[paper] notation changes to comport with Agda development
let σ, τ range over sorts; use fraktur for metavariables

@@
 
-\newcommand\MV[1]{\textsc{#1}}
+\newcommand\MV[1]{\mathfrak{#1}}
 \newcommand\MApp[3]{#1\{#2\}(#3)}
@@
 
-Fix a set $\Sorts$ of \emph{sorts}. We will say $\IsSort{s}$ when
-$\Member{s}{S}$. A valence $\MkValence{\vec{p}}{\vec{q}}{s}$ specifies an
-expression of sort $s$ which binds symbols in $\vec{p}$ and variables in
-$\vec{q}$.
-
-\[
-  \infer{
-    \IsValence{\MkValence{p_0,\dots,p_m}{q_0,\dots,q_n}{s}}
-  }{
-    \IsSort{s} &
-    \HypJ{\IsSort{p_i}}{\Leq{i}{m}} &
-    \HypJ{\IsSort{q_i}}{\Leq{i}{n}} &
-  }
-\]
-
-An arity $\MkArity{\vec{v}}{s}$ specifies an operator of sort $s$ with
+Fix a set $\Sorts$ of \emph{sorts}. We will say $\IsSort{\tau}$ when
+$\Member{\tau}{\Sorts}$. A valence $\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}$ specifies an
+expression of sort $\tau$ which binds symbols in $\vec{\sigma}$ and variables in
+$\vec{\tau}$.
+
+\[
+  \infer{
+    \IsValence{\MkValence{\sigma_0,\dots,\sigma_m}{\tau_0,\dots,\tau_n}{\tau}}
+  }{
+    \IsSort{\tau} &
+    \HypJ{\IsSort{\sigma_i}}{\Leq{i}{m}} &
+    \HypJ{\IsSort{\tau_i}}{\Leq{i}{n}} &
+  }
+\]
+
+An arity $\MkArity{\vec{v}}{\tau}$ specifies an operator of sort $\tau$ with
 arguments of valences $\vec{v}$. We will call the set of valences $\Valences$,
@@
   \infer{
-    \IsArity{\MkArity{v_0,\ldots,v_n}{s}}
-  }{
-    \IsSort{s} &
+    \IsArity{\MkArity{v_0,\ldots,v_n}{\tau}}
+  }{
+    \IsSort{\tau} &
     \HypJ{\IsValence{v_i}}{\Leq{i}{n}}
@@
   \infer{
-    \IsVarCtx{\Gamma,x:s}
+    \IsVarCtx{\Gamma,x:\tau}
   }{
     \IsVarCtx{\Gamma} &
-    \IsSort{s} &
+    \IsSort{\tau} &
     \NotIn{x}{\Dom\Gamma}
@@
   \infer{
-    \IsSymCtx{\Upsilon,u:s}
+    \IsSymCtx{\Upsilon,u:\tau}
   }{
     \IsVarCtx{\Upsilon} &
-    \IsSort{s} &
+    \IsSort{\tau} &
     \NotIn{u}{\Dom\Upsilon}
@@
 
-Let the judgment $\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{s}$ presuppose
-$\IsMetaCtx{\Omega}$, $\IsSymCtx\Upsilon$, $\IsVarCtx\Gamma$ and $\IsSort{s}$,
+Let the judgment $\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$ presuppose
+$\IsMetaCtx{\Omega}$, $\IsSymCtx\Upsilon$, $\IsVarCtx\Gamma$ and $\IsSort{\tau}$,
 meaning that $M$ is an abstract binding tree of sort $s$, with metavariables in
@@
   \infer[\textit{var}]{
-    \IsAbt{\Omega}{\Upsilon}{\Gamma}{x}{s}
-  }{
-    \Lookup{\Gamma}{x}{s}
+    \IsAbt{\Omega}{\Upsilon}{\Gamma}{x}{\tau}
+  }{
+    \Lookup{\Gamma}{x}{\tau}
   }\\[6pt]
@@
       \MApp{\MV{m}}{u_0,\dots,u_m}{M_0,\dots,M_n}
-    }{s}
+    }{\tau}
   }{
     \begin{array}{l}
-      \Lookup{\Omega}{\MV{m}}{\MkValence{p_0,\dots,p_m}{q_0,\dots,q_n}{s}}\\
-      \HypJ{\Lookup{\Upsilon}{u_i}{p_i}}{\Leq{i}{m}}\\
+      \Lookup{\Omega}{\MV{m}}{\MkValence{\sigma_0,\dots,\sigma_m}{\tau_0,\dots,\tau_n}{\tau}}\\
+      \HypJ{\Lookup{\Upsilon}{u_i}{\sigma_i}}{\Leq{i}{m}}\\
       \HypJ{
-        \IsAbt{\Omega}{\Upsilon}{\Gamma}{M_i}{q_i}
+        \IsAbt{\Omega}{\Upsilon}{\Gamma}{M_i}{\tau_i}
       }{
@@
       \App{\vartheta}{E_0,\dots,E_n}
-    }{s}
+    }{\tau}
   }{
@@
       \IsOperator{\Upsilon}{\vartheta}{
-        v_1,\dots,v_n
+        \MkArity{v_1,\dots,v_n}{\tau}
       }\\
       \HypJ{
-        \IsBTm{\Omega}{\Upsilon}{\Gamma}{E_i}{q_i}
+        \IsBTm{\Omega}{\Upsilon}{\Gamma}{E_i}{v_i}
       }{
@@
     }{
-      \MkValence{\vec{p}}{\vec{q}}{s}
-    }
-  }{
-    \IsAbt{\Omega}{\Upsilon, \vec{u}:\vec{p}}{\Gamma,\vec{x}:{\vec{q}}}{
+      \MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}
+    }
+  }{
+    \IsAbt{\Omega}{\Upsilon, \vec{u}:\vec{\sigma}}{\Gamma,\vec{x}:{\vec{\tau}}}{
       M
     }{
-      s
+      \tau
     }
@@
 $\IsRenaming{\varrho}{E}{F}$, presupposing
-$\Of{\varrho}{\Upsilon\to\Upsilon'}$, and $\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{s}$
+$\Of{\varrho}{\Upsilon\to\Upsilon'}$, and $\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$
 and $\IsAbt{\Omega}{\Upsilon}{\Gamma}{E}{v}$ respectively:
@@
 
-Let $\Define{V_s(\Upsilon\parallel\Gamma)}{\MkSet{x\in\Dom\Gamma\mid
-\Gamma(x)=s}}$ be called the presheaf of variables; additionally, we have a
+Let $\Define{V_\tau(\Upsilon\parallel\Gamma)}{\MkSet{x\in\Dom\Gamma\mid
+\Gamma(x)=\tau}}$ be called the presheaf of variables; additionally, we have a
 presheaf of symbols
-$\Define{S_s(\Upsilon\parallel\Gamma)}{\MkSet{u\in\Dom\Upsilon\mid
-\Upsilon(x)=s}}$. Lastly, we have the presheaf of operators with arity $a$,
+$\Define{S_\tau(\Upsilon\parallel\Gamma)}{\MkSet{u\in\Dom\Upsilon\mid
+\Upsilon(x)=\tau}}$. Lastly, we have the presheaf of operators with arity $a$,
 $\Define{\Operators_a(\Upsilon\parallel\Gamma)}{\Operators\Pair{\Upsilon}{a}}$
@@
   \HypJ{
-    \Define{{(P\odot Q)}_{s}}{
-      P_s\bullet Q
+    \Define{{(P\odot Q)}_{\tau}}{
+      P_\tau\bullet Q
     }
@@
   }{
-    \Member{s}{\Sorts}
+    \Member{\tau}{\Sorts}
   }
@@
 and
-$\Of{\varsigma_{\Upsilon\parallel\Gamma}}{P_s^{\Yoneda{\Upsilon\parallel\Gamma}} \times
-S^{\{\Upsilon\}}\times P^{[\Gamma]}\to P_s}$.
+$\Of{\varsigma_{\Upsilon\parallel\Gamma}}{P_\tau^{\Yoneda{\Upsilon\parallel\Gamma}}
+\times S^{\{\Upsilon\}}\times P^{[\Gamma]}\to P_\tau}$.
 
@@
 \begin{align*}
-  \IMode{{\SigFun{\Sigma}(X)}_{s}}
+  \IMode{{\SigFun{\Sigma}(X)}_{\tau}}
     &\triangleq
     \OMode{
-      \coprod_{\vartheta\in \Operators_{\MkArity{\vec{v}}{s}}}
-      \prod_{\MkValence{\vec{p}}{\vec{q}}{s'}\in\vec{v}}
-      X_{s'}^{\Yoneda{\vec{p}\parallel\vec{q}}}
+      \coprod_{\vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}}
+      \prod_{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau'}\in\vec{v}}
+      X_{\tau'}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}
     }
@@
   }{
-    \left(\prod_{(m:\MkValence{\vec{p}}{\vec{q}}{s})\in\Omega}P_{s}^{\Yoneda{\vec{p}\parallel\vec{q}}}\right)
+    \left(\prod_{(m:\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau})\in\Omega}P_{\tau}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}\right)
     \times
@@
   \Of{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash%
-      M:s}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P\to P_s}%
+      M:\tau}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P\to P_\tau}%
 \]%
 The interpretation of a variable
-$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x_j:s}_P$ simply
+$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x:\tau}_P$ simply
 projects it of the environment and embeds it into the model, as follows:
@@
     \& {\IMode{V^{[\Gamma]}}} \arrow[r, "\IMode{\nu_{\Gamma}}"]
-    \& {\IMode{P^{[\Gamma]}}} \arrow[r, "\IMode{\pi_j}"]
-    \& \IMode{P_s}
+    \& {\IMode{P^{[\Gamma]}}} \arrow[r, "\IMode{\pi_x}"]
+    \& \IMode{P_\tau}
   \end{tikzcd}
@@
   \SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash%
-    \MApp{\MV{m}_i}{\vec{u}}{\vec{M}} : s}_P%
-  \qquad(\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{p}}{\vec{q}}{s}})
+    \MApp{\MV{m}}{\vec{u}}{\vec{M}} : \tau}_P%
+  \qquad(\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}})
 \]
@@
     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
-          \arrow[r, "\IMode{\Tuple{\pi_i\,\pi_1, \phi\,\pi_2, \psi}}"]
-    \& {\IMode{P_{s'}^{\Yoneda{\vec{p}\parallel\vec{q}}}\times S^{\{\vec{p}\}} \times P^{[\vec{q}]}}}
-          \arrow[r, "\IMode{\varsigma_{\vec{p}\parallel\vec{q}}}"]
+          \arrow[r, "\IMode{\Tuple{\pi_\MV{m}\,\pi_1, \phi\,\pi_2, \psi}}"]
+    \& {\IMode{P_{\tau}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}\times S^{\{\vec{\sigma}\}} \times P^{[\vec{\tau}]}}}
+          \arrow[r, "\IMode{\varsigma_{\vec{\sigma}\parallel\vec{\tau}}}"]
     \& \IMode{P_s}
@@
     {S^{\{\Upsilon\}}} \arrow[r, "\Tuple{\pi_u}_{u\in\vec{u}}"]
-    \& {S^{\{\vec{p}\}}}
+    \& {S^{\{\vec{\sigma}\}}}
   \end{tikzcd}
@@
      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}
-        \arrow[r, "\Tuple{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M:q}_P}_{(M,q)\in \vec{M}\otimes\vec{q}}"]
-     \& {P^{[\vec{q}]}}
+        \arrow[r, "\Tuple{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M:q}_P}_{(M,q)\in \vec{M}\otimes\vec{\tau}}"]
+     \& {P^{[\vec{\tau}]}}
    \end{tikzcd}
@@
 of parametric higher-order judgment to our object language, indexed by a
-collection of sorts $\vec{s}$:
-\[
-  \IsOperator{\Upsilon}{\OpNabla{\vec{s}}}{
+collection of sorts $\vec{\sigma}$:
+\[
+  \IsOperator{\Upsilon}{\OpNabla{\vec{\sigma}}}{
     \MkArity{
-      \MkValence{\vec{s}}{}{\SortJdg}
+      \MkValence{\vec{\sigma}}{}{\SortJdg}
     }{\SortJdg}
@@
 \emph{inductive refinement}. The basic idea is to introduce a new form of
-(meta)-judgment $\IsWf{\Omega}{\Upsilon}{\Gamma}{M}{s}$ which expresses the
+(meta)-judgment $\IsWf{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$ which expresses the
 extrinsic wellformedness properties we wish to verify, presupposing
-$\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{s}$. Additionally, we introduce an
+$\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$. Additionally, we introduce an
 analogous judgment on bound terms, $\IsWf{\Omega}{\Upsilon}{\Gamma}{E}{v}$
@@
     }{
-      \MkValence{\vec{p}}{\vec{q}}{s}
-    }
-  }{
-    \IsWf{\Omega}{\Upsilon,\vec{u}:\vec{p}}{\Gamma,\vec{x}:\vec{q}}{M}{s}
+      \MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}
+    }
+  }{
+    \IsWf{\Omega}{\Upsilon,\vec{u}:\vec{\sigma}}{\Gamma,\vec{x}:\vec{\tau}}{M}{\tau}
   }
@@
   \infer{
-    \IsWf{\Omega}{\Upsilon}{\Gamma}{x}{s}
+    \IsWf{\Omega}{\Upsilon}{\Gamma}{x}{\tau}
   }{
@@
       \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
-    }{s}
+    }{\tau}
   }{
     \begin{array}{l}
-      \Lookup{\Omega}{\MV{m}}{\MkValence{\vec{p}}{q_0,\dots,q_n}{s}}\\
+      \Lookup{\Omega}{\MV{m}}{\MkValence{\vec{\sigma}}{\tau_0,\dots,\tau_n}{\tau}}\\
       \HypJ{
-        \IsWf{\Omega}{\Upsilon}{\Gamma}{M_i}{q_i}
+        \IsWf{\Omega}{\Upsilon}{\Gamma}{M_i}{\tau_i}
       }{
@@
   Note that the refinement for variables $x$ is not trivial, since it is only
-  defined in case the presupposition $\IsAbt{\Omega}{\Upsilon}{\Gamma}{x}{s}$
+  defined in case the presupposition $\IsAbt{\Omega}{\Upsilon}{\Gamma}{x}{\tau}$
   is satisfied.
@@
   \infer[
-    \text{for $\Member{s}{\MkSet{\SortExp,\SortType}}$}
+    \text{for $\Member{\tau}{\MkSet{\SortExp,\SortType}}$}
   ]{
@@
           v_0,\dots,v_n
-        }{s}
+        }{\tau}
       }\\
@@
     \IsWf{\Omega}{\Upsilon}{\Gamma}{
-      \App{\OpNabla{\vec{s}}}{
+      \App{\OpNabla{\vec{\sigma}}}{
         \MkBTm{\vec{u}}{}{J}
@@
   }{
-    \IsWf{\Omega}{\Upsilon,\vec{u}:\vec{s}}{\Gamma}{J}{\SortJdg}
+    \IsWf{\Omega}{\Upsilon,\vec{u}:\vec{\sigma}}{\Gamma}{J}{\SortJdg}
   }\\[6pt]

### 2015-11-07 paper.tex
[paper] fix typo

@@
 and
-$\Of{\varsigma_{\Upsilon\parallel\Gamma}}{P_\tau^{\Yoneda{\Upsilon\parallel\Gamma}}
+$\Of{\varsigma^\tau_{\Upsilon\parallel\Gamma}}{P_\tau^{\Yoneda{\Upsilon\parallel\Gamma}}
 \times S^{\{\Upsilon\}}\times P^{[\Gamma]}\to P_\tau}$.
@@
      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}
-        \arrow[r, "\Tuple{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M:q}_P}_{(M,q)\in \vec{M}\otimes\vec{\tau}}"]
+        \arrow[r, "\Tuple{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M:\tau}_P}_{(M,\tau)\in \vec{M}\otimes\vec{\tau}}"]
      \& {P^{[\vec{\tau}]}}

### 2015-11-07 paper.tex
Add natural transformation to rename operators' parameters in interpretation

@@
      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}
-        \arrow[r, "\Tuple{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M:\tau}_P}_{(M,\tau)\in \vec{M}\otimes\vec{\tau}}"]
+     \arrow[r, "\Tuple{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M:\tau}_P}_{(M,\tau)\in (\vec{M},\vec{\tau})}"]
      \& {P^{[\vec{\tau}]}}
@@
 \end{align*}
+
+The interpretation of operator applications is most complicated. Recall that,
+unlike in standard treatments of algebraic theories, our operators are indexed
+by symbol collections; therefore, operators must pass through suitable
+renamings in order to be used in the interpretation. Let us begin by
+constructing for each operator $\IsOperator{\Upsilon}{\vartheta}{a}$ the
+natural transformation
+$\Of{\SemBrackets{\vartheta}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
+\to\Operators_{a}}$, which shall rename the parameters of the operator using
+the environment.
+
+\[
+  \Define{
+    \SemBrackets{\vartheta}_P
+  }{
+   \begin{tikzcd}[ampersand replacement = \&, sep=large]
+     {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P} \arrow[r, "\pi_2"]
+     \& {S^{\{\Upsilon\}}} \arrow[r, "\FLift{\pi_{(-)}}{}"]
+     \& {\Operators_a}
+   \end{tikzcd}
+ }
+\]
 

### 2015-11-07 paper.tex
[paper] Rough implementation of the interpretation
@freebroccolo This seems like it should work; could you give me a second
pair of eyes, though? Incidentally, I think that the factoring I have
proposed here will make the Agda proof more elegant if we should adopt
it.

This probably should be reorganized a bit so it's not just a block of
definitions(?), but this is what "hole-driven-development" in LaTeX
looks like :)

@@
       \coprod_{\vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}}
-      \prod_{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau'}\in\vec{v}}
-      X_{\tau'}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}
+      \prod_{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau_i}\in\vec{v}}
+      X_{\tau_i}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}
     }
@@
           \arrow[r, "\IMode{\varsigma_{\vec{\sigma}\parallel\vec{\tau}}}"]
-    \& \IMode{P_s}
+    \& \IMode{P_\tau}
   \end{tikzcd}
@@
 The interpretation of operator applications is most complicated. Recall that,
-unlike in standard treatments of algebraic theories, our operators are indexed
+unlike in standard treatments of universal algebra, our operators are indexed
 by symbol collections; therefore, operators must pass through suitable
@@
 \]
+
+We will proceed using the initial $\SigFun{\Sigma}$-algebra $\alpha$, as
+follows.
+\[
+   \begin{tikzcd}[ampersand replacement = \&, sep=large]
+     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}} \arrow[r, "\OMode{\beta}"]
+     \& {\IMode{\SigFun{\Sigma}(P)_\tau}} \arrow[r, "\IMode{\alpha}"]
+     \& {\IMode{P_\tau}}
+   \end{tikzcd}
+\]
+
+\begin{align*}
+  \ADefine{
+    \beta
+  }{
+   \begin{tikzcd}[ampersand replacement = \&, sep=huge]
+      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P} \arrow[r, "\Tuple{\SemBrackets{\vartheta}_P, \gamma}"]
+      \& {
+        \coprod_{\vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}}
+        \prod_{\MkValence{\vec{\sigma}_i}{\vec{\tau}_i}{\tau_i}\in\vec{v}}
+        P_{\tau_i}^{\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}}
+      }
+   \end{tikzcd}
+  }\\
+  \ADefine{
+    \gamma
+  }{
+   \begin{tikzcd}[ampersand replacement = \&, sep=6em]
+      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P} \arrow[r, "\Tuple{\lambda x. \eta_i(-,x))}_i"]
+      \& {
+        \prod_{\MkValence{\vec{\sigma}_i}{\vec{\tau}_i}{\tau_i}\in\vec{v}}
+        P_{\tau_i}^{\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}}
+      }
+   \end{tikzcd}
+  }\\
+  \ADefine{
+    \eta_i
+  }{
+   \begin{tikzcd}[ampersand replacement = \&, sep=7em]
+      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
+        \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_u}
+      } \arrow[r, "\Tuple{\pi_1\circ\pi_1,\phi_i,\nu_{\Gamma,\vec{\tau}_i}\circ\psi_i}"]
+      \& {\SemBrackets{\Omega\triangleright\Upsilon,\vec{\sigma}_i\parallel\Gamma,\vec{\tau}_i}_P}
+      \arrow[r, "\SemBrackets{\Omega\triangleright\Upsilon,\vec{\sigma}_i\parallel\Gamma,\vec{\tau}_i\vdash M_i:\tau_i}_P"]
+      \& {P_{\tau_i}}
+   \end{tikzcd}
+  }\\
+  \ADefine{
+    \phi_i
+  }{
+   \begin{tikzcd}[ampersand replacement = \&, sep=12em]
+      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
+        \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_u}
+      } \arrow[r, "\Tuple{\pi_u\circ\pi_1}_{u\in\Dom\Upsilon}\oplus\Tuple{\pi_u\circ\pi_1\circ\pi_2}_{u\in\Dom{\vec{\sigma}_i}}"]
+      \& {S^{\{\Upsilon,\vec{\sigma}_i\}}}
+   \end{tikzcd}
+  }\\
+  \ADefine{
+    \psi_i
+  }{
+   \begin{tikzcd}[ampersand replacement = \&, sep=12em]
+      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
+        \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_u}
+      } \arrow[r, "\Tuple{\pi_u\circ\pi_1}_{u\in\Dom\Upsilon}\oplus\Tuple{\pi_u\circ\pi_1\circ\pi_2}_{u\in\Dom{\vec{\sigma}_i}}"]
+      \& {V^{[\Gamma,\vec{\tau}_i]}}
+   \end{tikzcd}
+  }
+\end{align*}
 

### 2015-11-07 paper.tex
[paper] fix bug in interpretation
[paper] typos
Switch to codomain fibration

@@
 \newcommand\Discrete[1]{#1_{\equiv}}
+\newcommand\Arr[1]{\mathbf{{#1}^{\rightarrow}}}
 \newcommand\Sets{\mathbf{Set}}
@@
 in $\SCtx$ lift to renamings of operators' parameters. Via the Grothendieck
-construction\footnote{In this case, $\mathbf{C}\Grothendieck{}{\Psi}$
-  represents the category of elements of a copresheaf
-  $\Of{\Psi}{\mathbf{C}\to\Sets}$ but we keep the $\mathbf{C}$ implicit and
-  simply refer to it as the Grothendieck construction.  Alternatively, this
-  construction can be understood as a coend
+construction\footnote{In this case, $\mathbf{C}\Grothendieck{}{\Psi}$ represents
+  the category of elements of a copresheaf $\Of{\Psi}{\mathbf{C}\to\Sets}$ but
+  we keep the $\mathbf{C}$ implicit and simply refer to it as the Grothendieck
+  construction. Alternatively, this construction can be understood as a coend
   $\IMode{\mathbf{C}\Grothendieck{}{\Psi}}\cong \IMode{\int^{c \in
@@
 category of objects
-$\Member{\Pair{\Pair{\Upsilon}{a}}{\vartheta}}{\Grothendieck{}{\Operators}}$
-for $\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$ and morphisms
+$\Member{\Pair{\Pair{\Upsilon}{a}}{\vartheta}}{\Grothendieck{}{\Operators}}$ for
+$\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$ and morphisms
 $\Hom{\Grothendieck{}{\Operators}}{\Pair{\Pair{\Upsilon}{a}}{\vartheta}}{\Pair{\Pair{\Upsilon'}{a'}}{\vartheta'}}$
@@
 Equivalently, $\Grothendieck{}{\Operators}$ is the pullback of $\Operators$
-along the universal $\Sets$-bundle where $\pi_{\Operators}$ is a discrete
-Grothendieck opfibration and $\Forget{-}_{\SSingleton}$ is the forgetful functor
-from pointed sets:
+along the the codomain Grothendieck bifibration $\pi_{cod}$ where
+$\pi_{\Operators}$ is also a discrete Grothendieck opfibration:
 
@@
   {\Grothendieck{}{\Operators}}
-    \arrow[r, "{\pi_{\in}}"]
+    \arrow[rr, "{\Pair{\Pair{\Upsilon}{a}}{\vartheta}\ \mapsto\ \ArrId{\Operators\Pair{\Upsilon}{a}}}"]
     \arrow[d, "{\pi_{\Operators}}" left]
-  \& {\Sets_{\SSingleton}}
-    \arrow[d, "{\Forget{-}_{\SSingleton}}"] \\
+  \&
+  \& {\Arr{\Sets}}
+    \arrow[d, "{\pi_{cod}}"] \\
   {\SCtx\times\Discrete{\Arities}}
-     \arrow[r, "{\Operators}" below]
+     \arrow[rr, "{\Operators}" below]
+  \&
   \& {\Sets} \\
@@
      {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}} \arrow[r, "\OMode{\beta}"]
-     \& {\IMode{\SigFun{\Sigma}(P)_\tau}} \arrow[r, "\IMode{\alpha}"]
+     \& {\IMode{\SigFun{\Sigma}(P)_\tau}} \arrow[r, "\IMode{\alpha_\tau}"]
      \& {\IMode{P_\tau}}
@@
         \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_u}
-      } \arrow[r, "\Tuple{\pi_1\circ\pi_1,\phi_i,\nu_{\Gamma,\vec{\tau}_i}\circ\psi_i}"]
+      } \arrow[r, "\Tuple{\pi_1\circ\pi_1,\phi_i,\psi_i}"]
       \& {\SemBrackets{\Omega\triangleright\Upsilon,\vec{\sigma}_i\parallel\Gamma,\vec{\tau}_i}_P}
@@
       {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
-        \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_u}
+        \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}
       } \arrow[r, "\Tuple{\pi_u\circ\pi_1}_{u\in\Dom\Upsilon}\oplus\Tuple{\pi_u\circ\pi_1\circ\pi_2}_{u\in\Dom{\vec{\sigma}_i}}"]
@@
       {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
-        \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_u}
-      } \arrow[r, "\Tuple{\pi_u\circ\pi_1}_{u\in\Dom\Upsilon}\oplus\Tuple{\pi_u\circ\pi_1\circ\pi_2}_{u\in\Dom{\vec{\sigma}_i}}"]
+        \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}
+      } \arrow[r, "\Tuple{\pi_x\circ\pi_1}_{x\in\Dom\Gamma}\oplus\Tuple{\pi_x\circ\pi_1\circ\pi_2}_{x\in\Dom{\vec{\tau}_i}}"]
       \& {V^{[\Gamma,\vec{\tau}_i]}}

### 2015-11-07 paper.tex
[paper] fix typo

@@
 Equivalently, $\Grothendieck{}{\Operators}$ is the pullback of $\Operators$
-along the the codomain Grothendieck bifibration $\pi_{cod}$ where
+along the codomain Grothendieck bifibration $\pi_{cod}$ where
 $\pi_{\Operators}$ is also a discrete Grothendieck opfibration:

### 2015-11-07 paper.tex
[paper] Use element notation for map
[paper] notation adjustment in pullback square
- make the diagram well-moded (\oint O is the *result* of pulling back O
  along ...)

- use the tuple notation for elements consistently
Revert part of the previous notation change
Revamp interpretation presentation
I've arranged this to feel a bit more natural/readable & less
artificial.

@@
 \begin{tikzcd}[ampersand replacement = \&, sep = large]
-  {\Grothendieck{}{\Operators}}
-    \arrow[rr, "{\Pair{\Pair{\Upsilon}{a}}{\vartheta}\ \mapsto\ \ArrId{\Operators\Pair{\Upsilon}{a}}}"]
-    \arrow[d, "{\pi_{\Operators}}" left]
-  \&
-  \& {\Arr{\Sets}}
-    \arrow[d, "{\pi_{cod}}"] \\
-  {\SCtx\times\Discrete{\Arities}}
-     \arrow[rr, "{\Operators}" below]
-  \&
-  \& {\Sets} \\
+  {\OMode{\Grothendieck{}{\Operators}}}
+  \arrow[rr,
+    "{\IMode{\vartheta\in\Operators\Pair{\Upsilon}{a}\ \mapsto\
+      \ArrId{\Operators\Pair{\Upsilon}{a}}}}"] \arrow[d, "\IMode{\pi_{\Operators}}" left]
+  \& \& {\IMode{\Arr{\Sets}}}
+    \arrow[d, "\IMode{\pi_{cod}}"] \\
+  {\IMode{\SCtx\times\Discrete{\Arities}}} \arrow[rr, "\IMode{\Operators}" below] \&
+  \& {\IMode{\Sets}} \\
 \end{tikzcd}
@@
 Next we have the interpretation of metavariable applications
-\[%
-  \SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash%
-    \MApp{\MV{m}}{\vec{u}}{\vec{M}} : \tau}_P%
-  \qquad(\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}})
-\]
-which proceeds by projecting the metavariable's interpretation from the
-environment and intantiating it via substitution:
-\[
-  \begin{tikzcd}[ampersand replacement = \&, sep=huge]
+$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash\MApp{\MV{m}}{\vec{u}}{\vec{M}}
+: \tau}_P$ where
+$\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$, which
+proceeds by projecting the metavariable's interpretation from the environment
+and intantiating it via substitution:
+\[
+  \begin{tikzcd}[ampersand replacement = \&, sep=7em]
     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
@@
     \& {\IMode{P_{\tau}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}\times S^{\{\vec{\sigma}\}} \times P^{[\vec{\tau}]}}}
-          \arrow[r, "\IMode{\varsigma_{\vec{\sigma}\parallel\vec{\tau}}}"]
+          \arrow[r, "\IMode{\varsigma^\tau_{\vec{\sigma}\parallel\vec{\tau}}}"]
     \& \IMode{P_\tau}
@@
 where
-\begin{align*}
-  \ADefine{\phi}{
-  \begin{tikzcd}[ampersand replacement = \&, sep=huge]
-    {S^{\{\Upsilon\}}} \arrow[r, "\Tuple{\pi_u}_{u\in\vec{u}}"]
-    \& {S^{\{\vec{\sigma}\}}}
+\begin{gather*}
+  \begin{tikzcd}[ampersand replacement = \&, sep=8em]
+    {\IMode{S^{\{\Upsilon\}}}} \arrow[r, "\Define{\phi\,}{\,\Tuple{\pi_u}_{u\in\vec{u}}}"]
+    \& {\IMode{S^{\{\vec{\sigma}\}}}}
+  \end{tikzcd}\\
+  \begin{tikzcd}[ampersand replacement = \&, sep=14em]
+    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
+      \arrow[r, "\Define{\psi\,}{\,\Tuple{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M:\tau}_P}_{(M,\tau)\in (\vec{M},\vec{\tau})}}"]
+    \& {\IMode{P^{[\vec{\tau}]}}}
   \end{tikzcd}
-  }\\
-  \ADefine{\psi}{
-   \begin{tikzcd}[ampersand replacement = \&, sep=28ex]
-     {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}
-     \arrow[r, "\Tuple{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M:\tau}_P}_{(M,\tau)\in (\vec{M},\vec{\tau})}"]
-     \& {P^{[\vec{\tau}]}}
-   \end{tikzcd}
-  }
-\end{align*}
+\end{gather*}
 
@@
 constructing for each operator $\IsOperator{\Upsilon}{\vartheta}{a}$ the
-natural transformation
+morphism
 $\Of{\SemBrackets{\vartheta}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
 \to\Operators_{a}}$, which shall rename the parameters of the operator using
-the environment.
-
-\[
-  \Define{
-    \SemBrackets{\vartheta}_P
-  }{
+the environment. The natural transformation $\SemBrackets{\vartheta}_P$ is the
+following composite:
+\[
   \begin{tikzcd}[ampersand replacement = \&, sep=large]
-     {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P} \arrow[r, "\pi_2"]
-     \& {S^{\{\Upsilon\}}} \arrow[r, "\FLift{\pi_{(-)}}{}"]
-     \& {\Operators_a}
+    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}} \arrow[r, "\IMode{\pi_2}"]
+    \& {\IMode{S^{\{\Upsilon\}}}} \arrow[r, "\IMode{\FLift{\pi_{(-)}}{}}"]
+    \& {\IMode{\Operators_a}}
   \end{tikzcd}
- }
 \]
@@
 We will proceed using the initial $\SigFun{\Sigma}$-algebra $\alpha$, as
-follows.
+follows, by postcomposing it with a morphism $\beta$ from the environment into
+the signature endofunctor, which interprets the syntax of operator
+applications:
 \[
    \begin{tikzcd}[ampersand replacement = \&, sep=large]
-     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}} \arrow[r, "\OMode{\beta}"]
+     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}} \arrow[r, "\IMode{\beta}"]
      \& {\IMode{\SigFun{\Sigma}(P)_\tau}} \arrow[r, "\IMode{\alpha_\tau}"]
@@
 
-\begin{align*}
-  \ADefine{
-    \beta
-  }{
-   \begin{tikzcd}[ampersand replacement = \&, sep=huge]
-      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P} \arrow[r, "\Tuple{\SemBrackets{\vartheta}_P, \gamma}"]
-      \& {
-        \coprod_{\vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}}
-        \prod_{\MkValence{\vec{\sigma}_i}{\vec{\tau}_i}{\tau_i}\in\vec{v}}
+The construction of $\beta$ proceeds by renaming the parameters of the operator
+$\vartheta$ and constructing the (bound) arguments $\gamma$ of the operator.
+
+\[
+  \begin{tikzcd}[ampersand replacement = \&, sep=10em]
+    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
+       \arrow[r, "\Define{\beta\,}{\,\Tuple{\SemBrackets{\vartheta}_P, \gamma}}"]
+    \& {\IMode{%
+       \coprod_{\vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}}%
+       \prod_{\MkValence{\vec{\sigma}_i}{\vec{\tau}_i}{\tau_i}\in\vec{v}}%
+       P_{\tau_i}^{\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}}%
+    }}
+  \end{tikzcd}
+\]
+
+Each argument
+%
+\begin{tikzcd}[ampersand replacement = \&]
+  {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
+    \arrow[r, "\IMode{\gamma_i}"]
+  \& {\IMode{%
     P_{\tau_i}^{\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}}
-      }
+  }}
 \end{tikzcd}
-  }\\
-  \ADefine{
-    \gamma
-  }{
-   \begin{tikzcd}[ampersand replacement = \&, sep=6em]
-      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P} \arrow[r, "\Tuple{\lambda x. \eta_i(-,x))}_i"]
-      \& {
-        \prod_{\MkValence{\vec{\sigma}_i}{\vec{\tau}_i}{\tau_i}\in\vec{v}}
-        P_{\tau_i}^{\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}}
-      }
+%
+is the exponential transpose of the composite,
+\[
+  \begin{tikzcd}[ampersand replacement = \&, sep=8em]
+    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
+       \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_u}
+    }}
+      \arrow[r, "\IMode{\Tuple{\pi_1\circ\pi_1,\phi_i,\psi_i}}"]
+    \& {\IMode{\SemBrackets{\Omega\triangleright\Upsilon,\vec{\sigma}_i\parallel\Gamma,\vec{\tau}_i}_P}}
+      \arrow[r, "\IMode{\SemBrackets{\Omega\triangleright\Upsilon,\vec{\sigma}_i\parallel\Gamma,\vec{\tau}_i\vdash M_i:\tau_i}_P}"]
+    \& {\IMode{P_{\tau_i}}}
   \end{tikzcd}
-  }\\
-  \ADefine{
-    \eta_i
-  }{
-   \begin{tikzcd}[ampersand replacement = \&, sep=7em]
-      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
-        \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_u}
-      } \arrow[r, "\Tuple{\pi_1\circ\pi_1,\phi_i,\psi_i}"]
-      \& {\SemBrackets{\Omega\triangleright\Upsilon,\vec{\sigma}_i\parallel\Gamma,\vec{\tau}_i}_P}
-      \arrow[r, "\SemBrackets{\Omega\triangleright\Upsilon,\vec{\sigma}_i\parallel\Gamma,\vec{\tau}_i\vdash M_i:\tau_i}_P"]
-      \& {P_{\tau_i}}
+\]
+where $\phi_i,\psi_i$ are defined as follows:
+\begin{gather*}
+  \begin{tikzcd}[ampersand replacement = \&, sep=14em]
+    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
+       \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}
+    }}
+      \arrow[r, "\Define{\phi_i\,}{\,\Tuple{\pi_u\circ\pi_1}_{u\in\Dom{\Upsilon}},\, \Tuple{\pi_u\circ\pi_1\circ\pi_2}_{u\in\Dom{\vec{\sigma}_i}}}"]
+    \& {\IMode{S^{\{\Upsilon,\vec{\sigma}_i\}}}}
+  \end{tikzcd}\\
+  \begin{tikzcd}[ampersand replacement = \&, sep=14em]
+    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
+       \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}
+    }}
+      \arrow[r, "\Define{\psi_i\,}{\,\Tuple{\pi_x\circ\pi_1}_{x\in\Dom{\Gamma}},\, \Tuple{\pi_x\circ\pi_1\circ\pi_2}_{x\in\Dom{\vec{\tau}_i}}}"]
+    \& {\IMode{V^{[\Gamma,\vec{\tau}_i]}}}
   \end{tikzcd}
-  }\\
-  \ADefine{
-    \phi_i
-  }{
-   \begin{tikzcd}[ampersand replacement = \&, sep=12em]
-      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
-        \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}
-      } \arrow[r, "\Tuple{\pi_u\circ\pi_1}_{u\in\Dom\Upsilon}\oplus\Tuple{\pi_u\circ\pi_1\circ\pi_2}_{u\in\Dom{\vec{\sigma}_i}}"]
-      \& {S^{\{\Upsilon,\vec{\sigma}_i\}}}
-   \end{tikzcd}
-  }\\
-  \ADefine{
-    \psi_i
-  }{
-   \begin{tikzcd}[ampersand replacement = \&, sep=12em]
-      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
-        \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}
-      } \arrow[r, "\Tuple{\pi_x\circ\pi_1}_{x\in\Dom\Gamma}\oplus\Tuple{\pi_x\circ\pi_1\circ\pi_2}_{x\in\Dom{\vec{\tau}_i}}"]
-      \& {V^{[\Gamma,\vec{\tau}_i]}}
-   \end{tikzcd}
-  }
-\end{align*}
+\end{gather*}
+
+This concludes the interpretation of well-sorted terms into any $\Sigma$-model.
 

### 2015-11-08 paper.tex
fix typo
taking a stab at an intro

@@
 
-\title{Abstract Binding Trees}
+\title{Syntax and Semantics of Indexed Abstract Binding Trees}
 \date{}
@@
 \maketitle
+
+Abstract binding trees (abts) are a generalization of abstract syntax trees
+where operators may express variable binding structure as part of their
+arities.  Originally inspired by Martin-L\"of's theory of arities
+\cite{martin-lof:1984}, unisorted abts have been deployed successfully as the
+uniform syntactic framework for several implementations of Constructive Type
+Theory, including Nuprl \cite{constable:1986}, MetaPRL \cite{hickey:2003} and
+JonPRL \cite{jonprl:2015}.
+
+In \emph{Practical Foundations for Programming Languages} \cite{harper:2016},
+Robert Harper develops the multi-sorted version of abstract binding trees and
+proposes an extension to include families of operators indexed by
+\emph{symbols}, which are, unlike variables, subject to only
+distinctness-preserving renaming and not substitution; furthermore, symbols do
+not appear in the syntax of abts, and are only introduced as parameters to
+operators. This generalization to \emph{indexed abts} is essential for a
+correct treatment of programming languages with open sums (e.g. ML's
+\texttt{exn} type), as well as assignable references.
+
+In parallel, M.\ Fiore and his collaborators have developed the categorical
+semantics for several variations of second-order algebraic theories
+\cite{fiore:1999, fiore:2005, fiore-hur:2010, fiore-mamoud:2010}; of these, the
+simply sorted second order algebraic theories treated by Fiore et al.\ are
+equivalent to abstract binding trees augmented with a notion of second-order
+variable (metavariable).
+
+The contribution of this paper is the development of the syntax and semantics
+of \emph{indexed multi-sorted abts}, an extension of second order universal
+algebra to support symbol-indexed families of operators. Additionally, we have
+developed the categorical semantics for indexed abts formally in Constructive
+Type Theory using the Agda proof assistant.
 
@@
 $\IsBTm{\Omega}{\Upsilon}{\Gamma}{E}{v}$ presuppose $\IsValence{v}$.  Then, the
-syntax of abstract binding trees (abts) is inductively defined in four rules:
+syntax of abstract binding trees is inductively defined in four rules:
 
@@
     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
-       \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_u}
+       \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}
     }}

### 2015-11-08 paper.tex
wordsmithing

@@
 \cite{fiore:1999, fiore:2005, fiore-hur:2010, fiore-mamoud:2010}; of these, the
-simply sorted second order algebraic theories treated by Fiore et al.\ are
-equivalent to abstract binding trees augmented with a notion of second-order
-variable (metavariable).
+simply sorted variants are equivalent to Harper's abstract binding trees
+augmented with a notion of second-order variable (metavariable).
 

### 2015-11-08 paper.tex
cite Ulf's thesis for agda

@@
 developed the categorical semantics for indexed abts formally in Constructive
-Type Theory using the Agda proof assistant.
+Type Theory using the Agda proof assistant \cite{norell:2007}.
 

### 2015-11-08 paper.tex
restore Aczel's priority for unisorted abts

@@
 where operators may express variable binding structure as part of their
-arities.  Originally inspired by Martin-L\"of's theory of arities
-\cite{martin-lof:1984}, unisorted abts have been deployed successfully as the
-uniform syntactic framework for several implementations of Constructive Type
-Theory, including Nuprl \cite{constable:1986}, MetaPRL \cite{hickey:2003} and
-JonPRL \cite{jonprl:2015}.
+arities.  Originally formulated by Peter Aczel \cite{aczel:1978}, unisorted
+abts have been deployed successfully as the uniform syntactic framework for
+several implementations of Constructive Type Theory, including Nuprl
+\cite{constable:1986}, MetaPRL \cite{hickey:2003} and JonPRL
+\cite{jonprl:2015}.
 

### 2015-11-08 paper.tex

@@
 
-\title{Syntax and Semantics of Indexed Abstract Binding Trees}
+\title{Syntax and Semantics of Abstract Binding Trees}
 \date{}
@@
 not appear in the syntax of abts, and are only introduced as parameters to
-operators. This generalization to \emph{indexed abts} is essential for a
+operators. This extension to support symbols is essential for a
 correct treatment of programming languages with open sums (e.g. ML's
@@
 The contribution of this paper is the development of the syntax and semantics
-of \emph{indexed multi-sorted abts}, an extension of second order universal
+of multi-sorted abts, an extension of second order universal
 algebra to support symbol-indexed families of operators. Additionally, we have
-developed the categorical semantics for indexed abts formally in Constructive
+developed the categorical semantics for abts formally in Constructive
 Type Theory using the Agda proof assistant \cite{norell:2007}.

### 2015-11-09 paper.tex
fix typo (presheaf -> copresheaf)

@@
 operators (unlike variables, symbols are not terms). Therefore, the functorial
-action of the operator presheaf can be lifted into terms by recursion on their
+action of the operator copresheaf can be lifted into terms by recursion on their
 structure, via a pair of judgments $\IsRenaming{\varrho}{M}{N}$ and

### 2015-11-09 paper.tex
cite pfpl
Simplify Grothendieck construction explanation

@@
 \newcommand\ArrId[1]{1_{#1}}
+\newcommand\Shriek[1]{\,{{{\mathexclam}}}_{#1}}
 \newcommand\Match[2]{\IMode{#1}\equiv\OMode{#2}}
@@
 
+\newcommand\Iso[2]{#1\mathrel{{{\cong}}}#2}
 \newcommand\Cats{\mathbf{Cat}}
@@
 $\Of{\Operators}{\SCtx\times\Discrete{\Arities}\to\Sets}$ such that the arrows
-in $\SCtx$ lift to renamings of operators' parameters. Via the Grothendieck
-construction\footnote{In this case, $\mathbf{C}\Grothendieck{}{\Psi}$ represents
-  the category of elements of a copresheaf $\Of{\Psi}{\mathbf{C}\to\Sets}$ but
-  we keep the $\mathbf{C}$ implicit and simply refer to it as the Grothendieck
-  construction. Alternatively, this construction can be understood as a coend
-  $\IMode{\mathbf{C}\Grothendieck{}{\Psi}}\cong \IMode{\int^{c \in
-      \mathbf{C}}\Coslice{c}{\mathbf{C}} \otimes \Discrete{\Psi(c)}}$.}
-$\Of{\Grothendieck{}{(-)}}{\Sets^{\mathbf{C}}\to\Cats}$ on operators, we have a
-category of objects
-$\Member{\Pair{\Pair{\Upsilon}{a}}{\vartheta}}{\Grothendieck{}{\Operators}}$ for
-$\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$ and morphisms
-$\Hom{\Grothendieck{}{\Operators}}{\Pair{\Pair{\Upsilon}{a}}{\vartheta}}{\Pair{\Pair{\Upsilon'}{a'}}{\vartheta'}}$
-for $\Of{\Pair{\varrho}{f}}{\Pair{\Upsilon}{a}\to\Pair{\Upsilon'}{a'}}$ such
-that
-$\EqMember{\FLift{\varrho}{f}(\vartheta)}{\vartheta'}{\Operators\Pair{\Upsilon'}{a'}}$.
-Equivalently, $\Grothendieck{}{\Operators}$ is the pullback of $\Operators$
-along the codomain Grothendieck bifibration $\pi_{cod}$ where
-$\pi_{\Operators}$ is also a discrete Grothendieck opfibration:
-
-\[
-\begin{tikzcd}[ampersand replacement = \&, sep = large]
-  {\OMode{\Grothendieck{}{\Operators}}}
-  \arrow[rr,
-    "{\IMode{\vartheta\in\Operators\Pair{\Upsilon}{a}\ \mapsto\
-      \ArrId{\Operators\Pair{\Upsilon}{a}}}}"] \arrow[d, "\IMode{\pi_{\Operators}}" left]
-  \& \& {\IMode{\Arr{\Sets}}}
-    \arrow[d, "\IMode{\pi_{cod}}"] \\
-  {\IMode{\SCtx\times\Discrete{\Arities}}} \arrow[rr, "\IMode{\Operators}" below] \&
-  \& {\IMode{\Sets}} \\
+in $\SCtx$ lift to renamings of operators' parameters. Then by the Grothendieck
+construction $\Iso{\Grothendieck{}{\Operators}}{\Comma{𝟙}{\Operators}}$
+we have a category of elements of $\Operators$ where objects are triples
+$\Pair{\Pair{\Upsilon}{a}}{\vartheta}$, interpreted as
+$\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$, and morphisms are diagrams:
+
+\[
+  \begin{tikzcd}[ampersand replacement = \&]
+    \& 𝟙
+      \arrow[dl, "\vartheta", swap]
+      \arrow[dr, "\vartheta'"]
+    \&
+    \\
+    {\Operators\Pair{\Upsilon}{a}}
+      \arrow[rr, "\Operators\Pair{\rho}{f}", swap]
+    \&
+    \& {\Operators\Pair{\Upsilon'}{a'}}
+    \\
+    {\Pair{\Upsilon}{a}}
+      \arrow[rr, "\Pair{\rho}{f}", swap]
+    \&
+    \& {\Pair{\Upsilon'}{a'}}
   \end{tikzcd}
-\]
+\]\\
 
@@
 the abt logical framework, we can proceed to define proper wellformedness via
-\emph{inductive refinement}. The basic idea is to introduce a new form of
+\emph{inductive refinement} \cite{harper:2016}. The basic idea is to introduce a new form of
 (meta)-judgment $\IsWf{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$ which expresses the

### 2015-11-09 paper.tex
[paper] wording

@@
 $\Of{\Operators}{\SCtx\times\Discrete{\Arities}\to\Sets}$ such that the arrows
-in $\SCtx$ lift to renamings of operators' parameters. Then by the Grothendieck
-construction $\Iso{\Grothendieck{}{\Operators}}{\Comma{𝟙}{\Operators}}$
-we have a category of elements of $\Operators$ where objects are triples
+in $\SCtx$ lift to renamings of operators' parameters. By the Grothendieck
+construction $\Iso{\Grothendieck{}{\Operators}}{\Comma{𝟙}{\Operators}}$ we have
+a category of elements of $\Operators$ where objects are triples
 $\Pair{\Pair{\Upsilon}{a}}{\vartheta}$, interpreted as

### 2015-11-10 paper.tex
[paper] annotate the Grothendieck opfibration
Switch back to pdflatex

@@
 
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
 \usepackage{amsmath}
@@
 \usepackage{amsthm}
-\usepackage{agda} % this should come after ams*
+%\usepackage{agda} % this should come after ams*
 \usepackage{catchfilebetweentags}
-\usepackage{fontspec}
+%\usepackage{fontspec}
 \usepackage{geometry}
@@
 \usepackage{ifthen}
-\usepackage{lualatex-math}
-%\usepackage{mathtools}
+%\usepackage{lualatex-math}
+\usepackage{mathtools}
 \usepackage{proof}
@@
 \usepackage{tikz-cd}
-\usepackage{unicode-math}
+%\usepackage{unicode-math}
 \usepackage{url}
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-\usepackage{xunicode}
+%\usepackage{xunicode}
 
@@
 % fonts
+%\usepackage{pxfonts}
+%\usepackage{eulervm}
 \usepackage[
   activate={true,nocompatibility},
+  kerning=true,
+  spacing=true,
   tracking=true
 ]{microtype}
-\setmainfont{Linux Libertine O}
-\setmathfont{Asana Math}
+\microtypecontext{spacing=nonfrench}
+% \setmainfont{Linux Libertine O}
+% \setmathfont{Asana Math}
 
@@
 in $\SCtx$ lift to renamings of operators' parameters. By the Grothendieck
-construction $\Iso{\Grothendieck{}{\Operators}}{\Comma{𝟙}{\Operators}}$ we have
+construction $\Iso{\Grothendieck{}{\Operators}}{\Comma{*}{\Operators}}$ we have
 a category of elements of $\Operators$ where objects are triples
@@
   \begin{tikzcd}[ampersand replacement = \&]
-    \& 𝟙
+    \& *
       \arrow[dl, "\vartheta", swap]
@@
 
-Operators are defined by specifying the fibers of $\pi_{\Operators}$ in which
-they reside. For instance, consider the lambda calculus with a single sort,
-$\SortExp$; we give its signature by asserting the following about its
+Operators are defined by specifying the fibers of
+$\Of{\pi_{\Operators}}{\Grothendieck{}{\Operators}\to\SCtx\times\Discrete{\Arities}}$ in
+which they reside. For instance, consider the lambda calculus with a single
+sort, $\SortExp$; we give its signature by asserting the following about its
 operators:

### 2015-11-11 paper.tex
font tweaks
properly introduce \IsOperator judgment
Clarify operator/fiber part

@@
 % fonts
-%\usepackage{pxfonts}
-%\usepackage{eulervm}
+\usepackage{pxfonts}
+\usepackage{eulervm}
 \usepackage[
@@
 
-\newcommand\OpCat[1]{#1^{\mathsf{op}}}
-
-\newcommand\IsOperator[3]{\IMode{#1}\Vdash\IMode{#2}: \OMode{#3}}
+\newcommand\OpCat[1]{#1^{\mathtt{op}}}
+
+\newcommand\IsOperator[3]{\IMode{#1}\,{\Vdash}_{\scriptscriptstyle{\Operators}}\,\IMode{#2}: \OMode{#3}}
 
@@
 \newcommand\Grothendieck[2]{\oint^{#1} #2#1}
-\newcommand\Pair[2]{\left\langle#1, #2 \right\rangle}
+\newcommand\Pair[2]{\left\langle\,#1,#2\,\right\rangle}
 \newcommand\Tuple[1]{\left\langle#1 \right\rangle}
@@
   \IMode{#2}\parallel%
-  \IMode{#3}\vdash%
+  \,{\IMode{#3}\vdash}_{\scriptscriptstyle{\Operators}}\,%
   \IMode{#4}:\OMode{#5}%
@@
   \IMode{#2}\parallel%
-  \IMode{#3}\vdash%
+  \,{\IMode{#3}\vdash}_{\scriptscriptstyle{\Operators}}\,%
   \IMode{#4}:\IMode{#5}%
@@
 \newcommand\NotIn[2]{\IMode{#1}\notin\IMode{#2}}
-\newcommand\SortExp{\mathsf{exp}}
+\newcommand\SortExp{\mathtt{exp}}
 
@@
 $\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$, and morphisms are diagrams:
-
 \[
@@
   \end{tikzcd}
-\]\\
-
-\[
-  \infer{
-    \IsOperator{\Upsilon}{\vartheta}{a}
-  }{
-    \Member{\vartheta}{\Operators\langle\Upsilon,a\rangle}
-  }
-\]
-
-The judgment $\IsOperator{\Upsilon}{\vartheta}{a}$ enjoys the structural
-properties of weakening and exchange via the functoriality of
-$\Operators$.
+\]
+We will write $\IsOperator{\Upsilon}{\vartheta}{a}$ in case
+$\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$ and note that this judgment
+enjoys the structural properties of weakening and exchange via functoriality of
+$\Operators$. Operators $\vartheta_{0}, \ldots, \vartheta_{n}$ are defined by
+specifying the fibers of
+$\Of{\pi_{\Operators}}{\Grothendieck{}{\Operators}\to\SCtx\times\Discrete{\Arities}}$
+in which they reside:
+\begin{alignat*}{3}
+  &{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}&&\,\mathrel{\ {{:}}}\,&&\Sets\\
+  &{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}&&\,=\,\ && %
+    \{\,\vartheta \in \Pair{\Upsilon}{a}%
+      {{|}}\ {\pi_{\Operators}}{\Pair{\Pair{\Upsilon}{a}}{\vartheta}} = \Pair{\Upsilon}{a}%
+  \}
+\end{alignat*}
 
@@
 
-Operators are defined by specifying the fibers of
-$\Of{\pi_{\Operators}}{\Grothendieck{}{\Operators}\to\SCtx\times\Discrete{\Arities}}$ in
-which they reside. For instance, consider the lambda calculus with a single
-sort, $\SortExp$; we give its signature by asserting the following about its
-operators:
+For instance, consider the lambda calculus with a single sort, $\SortExp$; we
+give its signature by asserting the following about its operators:
 \begin{gather*}
@@
   }\\
-  \IsOperator{\Upsilon}{\mathsf{ap}}{
+  \IsOperator{\Upsilon}{\mathtt{ap}}{
     \MkArity{
@@
 \end{gather*}
-
+These rules correspond to objects of $\Grothendieck{}{\Operators}$ and we
+collect their fibers into sets of operators:
+\begin{alignat*}{5}
+{\pi_{\Operators}}^{-1}&\Pair{\Upsilon}{{\MkArity{\MkValence{}{\SortExp}{\SortExp}}{\SortExp}}}\quad&&=
+  &&\quad\{\,\lambda\,\}\\
+{\pi_{\Operators}}^{-1}&\Pair{\Upsilon}{\MkArity{\MkValence{}{}{\SortExp},\MkValence{}{}{\SortExp}}{\SortExp}}\quad&&=
+  &&\quad\{\,\mathtt{ap}\,\}\\
+\bigcup_{\Pair{\Upsilon}{a}}&{{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}}\quad&&=
+  &&\quad\{\,\lambda,\ \mathtt{ap}\,\}\\
+\end{alignat*}
 So far, we have made no use of symbols and parameters; however, consider the
@@
 \begin{gather*}
-  \IsOperator{\Upsilon}{\mathsf{decl}}{
+  \IsOperator{\Upsilon}{\mathtt{decl}}{
     \MkArity{
@@
   }\\
-  \IsOperator{\Upsilon,u:\SortExp}{\mathsf{get}[u]}{
+  \IsOperator{\Upsilon,u:\SortExp}{\mathtt{get}[u]}{
     \MkArity{}{\SortExp}
   }\\
-  \IsOperator{\Upsilon,u:\SortExp}{\mathsf{set}[u]}{
+  \IsOperator{\Upsilon,u:\SortExp}{\mathtt{set}[u]}{
     \MkArity{
@@
 $\FLift{\varrho}{\ArrId{}}$. In particular, the renaming
-$\Upsilon,u\mapsto\Upsilon,v$ shall take $\mathsf{get}[u]$ to $\mathsf{get}[v]$.
+$\Upsilon,u\mapsto\Upsilon,v$ shall take $\mathtt{get}[u]$ to $\mathtt{get}[v]$.
 
@@
 %
-\newcommand\SortTele{\mathsf{tele}}
-\newcommand\SortJdg{\mathsf{jdg}}
-\newcommand\SortType{\mathsf{type}}
-\newcommand\SortExpr{\mathsf{exp}}
-\newcommand\OpNil{\mathsf{nil}}
-\newcommand\OpSnoc[1]{\mathsf{snoc}[#1]}
-\newcommand\OpVar[1]{\mathsf{var}[#1]}
-\newcommand\OpSequent{\mathsf{sequent}}
+\newcommand\SortTele{\mathtt{tele}}
+\newcommand\SortJdg{\mathtt{jdg}}
+\newcommand\SortType{\mathtt{type}}
+\newcommand\SortExpr{\mathtt{exp}}
+\newcommand\OpNil{\mathtt{nil}}
+\newcommand\OpSnoc[1]{\mathtt{snoc}[#1]}
+\newcommand\OpVar[1]{\mathtt{var}[#1]}
+\newcommand\OpSequent{\mathtt{sequent}}
 %
@@
 
-\newcommand\OpPi{\mathsf{pi}}
-\newcommand\OpSg{\mathsf{sg}}
-\newcommand\OpBool{\mathsf{bool}}
-\newcommand\OpSo{\mathsf{isTrue}}
+\newcommand\OpPi{\mathtt{pi}}
+\newcommand\OpSg{\mathtt{sg}}
+\newcommand\OpBool{\mathtt{bool}}
+\newcommand\OpSo{\mathtt{isTrue}}
 

### 2015-11-11 paper.tex
Move definition of signature upwards; fix a few layout issues

@@
 
-\newcommand\MkSet[1]{\left\{#1\right\}}
+\newcommand\MkSet[1]{\left\{\,#1\,\right\}}
 \newcommand\Member[2]{\IMode{#1}\in\IMode{#2}}
@@
 
-\newcommand\IsOperator[3]{\IMode{#1}\,{\Vdash}_{\scriptscriptstyle{\Operators}}\,\IMode{#2}: \OMode{#3}}
+\newcommand\IsOperator[3]{
+  \IMode{#1}\,{\Vdash}_{\Sigma}\,\IMode{#2}: \OMode{#3}
+}
 
@@
 $\Of{\Operators}{\SCtx\times\Discrete{\Arities}\to\Sets}$ such that the arrows
-in $\SCtx$ lift to renamings of operators' parameters. By the Grothendieck
-construction $\Iso{\Grothendieck{}{\Operators}}{\Comma{*}{\Operators}}$ we have
-a category of elements of $\Operators$ where objects are triples
+in $\SCtx$ lift to renamings of operators' parameters. Together, $\Sorts$ and
+$\Operators$ form a \emph{signature} $\Define{\Sigma}{\Pair{\Sorts}{\Operators}}$.
+
+By the Grothendieck construction
+$\Iso{\Grothendieck{}{\Operators}}{\Comma{*}{\Operators}}$ we have a category
+of elements of $\Operators$ where objects are triples
 $\Pair{\Pair{\Upsilon}{a}}{\vartheta}$, interpreted as
@@
 {\pi_{\Operators}}^{-1}&\Pair{\Upsilon}{{\MkArity{\MkValence{}{\SortExp}{\SortExp}}{\SortExp}}}\quad&&=
-  &&\quad\{\,\lambda\,\}\\
+  &&\quad\MkSet{\lambda}\\
 {\pi_{\Operators}}^{-1}&\Pair{\Upsilon}{\MkArity{\MkValence{}{}{\SortExp},\MkValence{}{}{\SortExp}}{\SortExp}}\quad&&=
-  &&\quad\{\,\mathtt{ap}\,\}\\
+  &&\quad\MkSet{\mathtt{ap}}\\
 \bigcup_{\Pair{\Upsilon}{a}}&{{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}}\quad&&=
-  &&\quad\{\,\lambda,\ \mathtt{ap}\,\}\\
+  &&\quad\MkSet{\lambda,\ \mathtt{ap}}\\
 \end{alignat*}
@@
 $\Upsilon,u\mapsto\Upsilon,v$ shall take $\mathtt{get}[u]$ to $\mathtt{get}[v]$.
-
-\begin{definition}
-  A \emph{signature} $\Sigma$ is a set of sorts $\Sorts$ together with a copresheaf $\Operators$ of operators.
-\end{definition}
 
@@
       M:\tau}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P\to P_\tau}%
-\]%
+\]
+
 The interpretation of a variable
@@
 
-Next we have the interpretation of metavariable applications
+Next interpret metavariable applications
 $\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash\MApp{\MV{m}}{\vec{u}}{\vec{M}}
@@
 \[
-  \begin{tikzcd}[ampersand replacement = \&, sep=8em]
+  \begin{tikzcd}[ampersand replacement = \&, sep=7.6em]
     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P

### 2015-11-11 paper.tex
Use Sigma subscript and update PDF
Don't use elem in fib; looks less trivial this way
[paper] change fiber definition back
[paper] Use a symbol to make abs easier to read
[paper] further clarify fiber/operator stuff

@@
 \usepackage{eulervm}
+\usepackage{bbold}
 \usepackage[
@@
 }
-\newcommand\MkBTm[3]{\{#1\}[#2].\, #3}
+\newcommand\MkBTm[3]{\mathbb{\lambda}\{#1\}[#2].\, #3}
 \newcommand\MkArity[2]{(#1)\, #2}
@@
 \newcommand\IsOperator[3]{
-  \IMode{#1}\,{\Vdash}_{\Sigma}\,\IMode{#2}: \OMode{#3}
+  \IMode{#1}\,{\Vdash}\,\IMode{#2}: \OMode{#3}
 }
@@
   \IMode{#2}\parallel%
-  \,{\IMode{#3}\vdash}_{\scriptscriptstyle{\Operators}}\,%
+  \,{\IMode{#3}\vdash}\,%
   \IMode{#4}:\OMode{#5}%
@@
   \IMode{#2}\parallel%
-  \,{\IMode{#3}\vdash}_{\scriptscriptstyle{\Operators}}\,%
+  \,{\IMode{#3}\vdash}\,%
   \IMode{#4}:\IMode{#5}%
@@
 \begin{alignat*}{3}
-  &{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}&&\,\mathrel{\ {{:}}}\,&&\Sets\\
+  &{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}&&\,\mathrel{\ {{:}}}\,&&\SCtx\times\Discrete{\Arities}\to\Sets\\
   &{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}&&\,=\,\ && %
-    \{\,\vartheta \in \Pair{\Upsilon}{a}%
-      {{|}}\ {\pi_{\Operators}}{\Pair{\Pair{\Upsilon}{a}}{\vartheta}} = \Pair{\Upsilon}{a}%
-  \}
+    \{\ \ \Member{\vartheta}{\Pair{\Upsilon}{a}}\ \ |\ \
+    {\pi_{\Operators}}{\Pair{\Pair{\Upsilon}{a}}{\vartheta}} =
+    \Pair{\Upsilon}{a}\,\}
 \end{alignat*}
@@
 
-For instance, consider the lambda calculus with a single sort, $\SortExp$; we
-give its signature by asserting the following about its operators:
-\begin{gather*}
-  \IsOperator{\Upsilon}{\lambda}{
+For instance, consider the $\lambda$-calculus with a single sort, $\SortExp$; we
+give its signature $\Sigma_{\lambda}$ by asserting the following about its
+operators:
+\begin{gather*}
+  \IsOperator{\Upsilon}{\mathtt{lam}}{
     \MkArity{\MkValence{}{\SortExp}{\SortExp}}{\SortExp}
@@
 \end{gather*}
-These rules correspond to objects of $\Grothendieck{}{\Operators}$ and we
+These rules correspond to objects of $\Grothendieck{}{{\Operators}_{\lambda}}$ and we
 collect their fibers into sets of operators:
 \begin{alignat*}{5}
-{\pi_{\Operators}}^{-1}&\Pair{\Upsilon}{{\MkArity{\MkValence{}{\SortExp}{\SortExp}}{\SortExp}}}\quad&&=
-  &&\quad\MkSet{\lambda}\\
-{\pi_{\Operators}}^{-1}&\Pair{\Upsilon}{\MkArity{\MkValence{}{}{\SortExp},\MkValence{}{}{\SortExp}}{\SortExp}}\quad&&=
+{\pi_{{\Operators}_{\lambda}}}^{-1}&\Pair{\Upsilon}{{\MkArity{\MkValence{}{\SortExp}{\SortExp}}{\SortExp}}}\quad&&=
+  &&\quad\MkSet{\mathtt{lam}}\\
+{\pi_{{\Operators}_{\lambda}}}^{-1}&\Pair{\Upsilon}{\MkArity{\MkValence{}{}{\SortExp},\MkValence{}{}{\SortExp}}{\SortExp}}\quad&&=
   &&\quad\MkSet{\mathtt{ap}}\\
-\bigcup_{\Pair{\Upsilon}{a}}&{{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}}\quad&&=
-  &&\quad\MkSet{\lambda,\ \mathtt{ap}}\\
+{\pi_{{\Sigma}_{\lambda}}}^{-1}&=\bigcup_{\Pair{\Upsilon}{a}}{{\pi_{{\Operators}_{\lambda}}}^{-1}{\Pair{\Upsilon}{a}}}\quad&&=
+  &&\quad\MkSet{\mathtt{lam},\ \mathtt{ap}}\\
 \end{alignat*}

### 2015-11-11 paper.tex
[paper] Improve layout of diagrams

@@
 \[
-  \begin{tikzcd}[ampersand replacement = \&]
+  \begin{tikzcd}[cramped, ampersand replacement = \&]
+    {\Pair{\Upsilon}{a}}
+      \arrow[r, "\Pair{\rho}{f}", swap]
+    \& {\Pair{\Upsilon'}{a'}}
+  \end{tikzcd}
+  \qquad\begin{tikzcd}[cramped, sep = small, ampersand replacement = \&]
     \& *
@@
     \& {\Operators\Pair{\Upsilon'}{a'}}
-    \\
-    {\Pair{\Upsilon}{a}}
-      \arrow[rr, "\Pair{\rho}{f}", swap]
-    \&
-    \& {\Pair{\Upsilon'}{a'}}
   \end{tikzcd}
@@
 $P$ and $\varsigma$ equips $P$ with an operation for simultaneous substitutions
-of variables.
-
-\[
-  \begin{tikzcd}[ampersand replacement = \&]
+of variables. Furthermore, $\nu$ and $\varsigma$ induce maps ${\nu}_{\Gamma}$ and
+${{\varsigma}^{\tau}}_{\Upsilon\parallel\Gamma}$:
+\vspace{-0.5em}
+\[
+  \begin{tikzcd}[cramped, ampersand replacement = \&]
     {\IMode{V}} \arrow[r, "\IMode\nu"]
@@
   \end{tikzcd}
-\]
-
-%%TODO
-We also have the induced maps $\Of{\nu_\Gamma}{V^{[\Gamma]}\to P^{[\Gamma]}}$
-and
-$\Of{\varsigma^\tau_{\Upsilon\parallel\Gamma}}{P_\tau^{\Yoneda{\Upsilon\parallel\Gamma}}
-\times S^{\{\Upsilon\}}\times P^{[\Gamma]}\to P_\tau}$.
+  \qquad\begin{tikzcd}[cramped, ampersand replacement = \&]
+    {\IMode{{V}^{[\Gamma]}}}
+      \arrow[r, "\IMode{{\nu}_{\Gamma}}"]
+    \& {\IMode{{P}^{[\Gamma]}}}
+  \end{tikzcd}
+  \qquad\begin{tikzcd}[cramped, sep = large, ampersand replacement = \&]
+    {\IMode{{P_\tau^{\Yoneda{\Upsilon\parallel\Gamma}}\times S^{\{\Upsilon\}}\times P^{[\Gamma]}}}}
+      \arrow[r, "\IMode{\varsigma^\tau_{\Upsilon\parallel\Gamma}}"]
+    \& {\IMode{P_\tau}}
+  \end{tikzcd}
+\]
 
@@
   Consequently, the monoid multiplication $\varsigma$ can be replaced with an
-  extension operation:
-  \[
-    \begin{tikzcd}[ampersand replacement = \&, sep = huge]
-      {\IMode{{P(\,=\,)}^{V(\,-\,)}}} \arrow[r, "{(\,-\,)}^{\circledast}"]
-      \& {\IMode{{P(\,=\,)}^{P(\,-\,)}}}
-    \end{tikzcd}
-  \]
+  extension operation $\Of{{(\,-\,)}^{\circledast}}{{\IMode{{P(\,=\,)}^{V(\,-\,)}}}\to{\IMode{{P(\,=\,)}^{P(\,-\,)}}}}$.
 \end{remark}
@@
   \Of{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash%
-      M:\tau}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P\to P_\tau}%
+      M:\tau}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P\longrightarrow P_\tau}%
 \]
@@
 \[
-  \begin{tikzcd}[ampersand replacement = \&]
+  \begin{tikzcd}[cramped, ampersand replacement = \&]
     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}} \arrow[r, "\IMode{\pi_3}"]
@@
 \[
-  \begin{tikzcd}[ampersand replacement = \&, sep=7em]
+  \begin{tikzcd}[cramped, ampersand replacement = \&]
     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
           \arrow[r, "\IMode{\Tuple{\pi_\MV{m}\,\pi_1, \phi\,\pi_2, \psi}}"]
-    \& {\IMode{P_{\tau}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}\times S^{\{\vec{\sigma}\}} \times P^{[\vec{\tau}]}}}
+    \&[4.0em] {\IMode{P_{\tau}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}\times S^{\{\vec{\sigma}\}} \times P^{[\vec{\tau}]}}}
           \arrow[r, "\IMode{\varsigma^\tau_{\vec{\sigma}\parallel\vec{\tau}}}"]
-    \& \IMode{P_\tau}
+    \&[1.0em] \IMode{P_\tau}
   \end{tikzcd}
@@
 where
-\begin{gather*}
-  \begin{tikzcd}[ampersand replacement = \&, sep=8em]
+\vspace{-0.5em}
+\begin{gather*}
+  \begin{tikzcd}[cramped, ampersand replacement = \&]
     {\IMode{S^{\{\Upsilon\}}}} \arrow[r, "\Define{\phi\,}{\,\Tuple{\pi_u}_{u\in\vec{u}}}"]
-    \& {\IMode{S^{\{\vec{\sigma}\}}}}
-  \end{tikzcd}\\
-  \begin{tikzcd}[ampersand replacement = \&, sep=14em]
+    \&[3.0em] {\IMode{S^{\{\vec{\sigma}\}}}}
+  \end{tikzcd}
+  \qquad\begin{tikzcd}[cramped, ampersand replacement = \&]
     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
       \arrow[r, "\Define{\psi\,}{\,\Tuple{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M:\tau}_P}_{(M,\tau)\in (\vec{M},\vec{\tau})}}"]
-    \& {\IMode{P^{[\vec{\tau}]}}}
+    \&[11.0em] {\IMode{P^{[\vec{\tau}]}}}
   \end{tikzcd}
@@
 \[
-  \begin{tikzcd}[ampersand replacement = \&, sep=large]
+  \begin{tikzcd}[cramped, ampersand replacement = \&]
     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}} \arrow[r, "\IMode{\pi_2}"]
     \& {\IMode{S^{\{\Upsilon\}}}} \arrow[r, "\IMode{\FLift{\pi_{(-)}}{}}"]
-    \& {\IMode{\Operators_a}}
+    \&[1.5em] {\IMode{\Operators_a}}
   \end{tikzcd}
@@
 \[
-   \begin{tikzcd}[ampersand replacement = \&, sep=large]
+   \begin{tikzcd}[cramped, ampersand replacement = \&]
      {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}} \arrow[r, "\IMode{\beta}"]
@@
 $\vartheta$ and constructing the (bound) arguments $\gamma$ of the operator.
-
-\[
-  \begin{tikzcd}[ampersand replacement = \&, sep=10em]
+\[
+  \begin{tikzcd}[cramped, ampersand replacement = \&]
     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
        \arrow[r, "\Define{\beta\,}{\,\Tuple{\SemBrackets{\vartheta}_P, \gamma}}"]
-    \& {\IMode{%
+    \&[3.0em] {\IMode{%
        \coprod_{\vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}}%
@@
 \]
-
 Each argument
 %
-\begin{tikzcd}[ampersand replacement = \&]
+\begin{tikzcd}[cramped, ampersand replacement = \&]
   {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
@@
 \[
-  \begin{tikzcd}[ampersand replacement = \&, sep=7.6em]
+  \begin{tikzcd}[cramped, ampersand replacement = \&]
     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
@@
       \arrow[r, "\IMode{\Tuple{\pi_1\circ\pi_1,\phi_i,\psi_i}}"]
-    \& {\IMode{\SemBrackets{\Omega\triangleright\Upsilon,\vec{\sigma}_i\parallel\Gamma,\vec{\tau}_i}_P}}
+    \&[4.0em] {\IMode{\SemBrackets{\Omega\triangleright\Upsilon,\vec{\sigma}_i\parallel\Gamma,\vec{\tau}_i}_P}}
       \arrow[r, "\IMode{\SemBrackets{\Omega\triangleright\Upsilon,\vec{\sigma}_i\parallel\Gamma,\vec{\tau}_i\vdash M_i:\tau_i}_P}"]
-    \& {\IMode{P_{\tau_i}}}
+    \&[7.5em] {\IMode{P_{\tau_i}}}
   \end{tikzcd}
@@
 \begin{gather*}
-  \begin{tikzcd}[ampersand replacement = \&, sep=14em]
+  \begin{tikzcd}[cramped, ampersand replacement = \&]
     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
@@
       \arrow[r, "\Define{\phi_i\,}{\,\Tuple{\pi_u\circ\pi_1}_{u\in\Dom{\Upsilon}},\, \Tuple{\pi_u\circ\pi_1\circ\pi_2}_{u\in\Dom{\vec{\sigma}_i}}}"]
-    \& {\IMode{S^{\{\Upsilon,\vec{\sigma}_i\}}}}
-  \end{tikzcd}\\
-  \begin{tikzcd}[ampersand replacement = \&, sep=14em]
+    \&[12.0em] {\IMode{S^{\{\Upsilon,\vec{\sigma}_i\}}}}\\[-1.0em]
     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
@@
       \arrow[r, "\Define{\psi_i\,}{\,\Tuple{\pi_x\circ\pi_1}_{x\in\Dom{\Gamma}},\, \Tuple{\pi_x\circ\pi_1\circ\pi_2}_{x\in\Dom{\vec{\tau}_i}}}"]
-    \& {\IMode{V^{[\Gamma,\vec{\tau}_i]}}}
+    \&[12.0em] {\IMode{V^{[\Gamma,\vec{\tau}_i]}}}
   \end{tikzcd}

### 2015-11-11 paper.tex
[paper] describe morphism via commuting diagram

@@
 The interpretation of operator applications is most complicated. Recall that,
-unlike in standard treatments of universal algebra, our operators are indexed
-by symbol collections; therefore, operators must pass through suitable
-renamings in order to be used in the interpretation. Let us begin by
-constructing for each operator $\IsOperator{\Upsilon}{\vartheta}{a}$ the
-morphism
-$\Of{\SemBrackets{\vartheta}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
-\to\Operators_{a}}$, which shall rename the parameters of the operator using
-the environment. The natural transformation $\SemBrackets{\vartheta}_P$ is the
-following composite:
+unlike in standard treatments of universal algebra, our operators are indexed by
+symbol collections; therefore, operators must pass through suitable renamings in
+order to be used in the interpretation. Let us begin by constructing for each
+operator $\IsOperator{\Upsilon}{\vartheta}{a}$ the morphism
+$\SemBrackets{\vartheta}_P$ which shall rename the parameters of the operator
+using the environment:
 \[
   \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}} \arrow[r, "\IMode{\pi_2}"]
-    \& {\IMode{S^{\{\Upsilon\}}}} \arrow[r, "\IMode{\FLift{\pi_{(-)}}{}}"]
-    \&[1.5em] {\IMode{\Operators_a}}
+    |[alias = Src]|
+    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
+      \arrow[r, swap, "\IMode{\pi_2}"]
+    \&
+    {\IMode{S^{\{\Upsilon\}}}}
+      \arrow[r, swap, "\IMode{\FLift{\pi_{(-)}}{}}"]
+    \&[1.5em]
+    |[alias = Tgt]|
+    {\IMode{\Operators_a}}
+    \arrow[from = Src, to = Tgt, bend left, dashed, "\IMode{\SemBrackets{\vartheta}_{P}}"]
   \end{tikzcd}

### 2015-11-11 paper.tex
formatting
Fix possible typo?
@freebroccolo was this intentional? If so, I'll revert
Revert "Fix possible typo?"
This reverts commit f4a26616a843bec5d6be0030506e87daae804cf1.
[paper] clarify some details

@@
 By the Grothendieck construction
-$\Iso{\Grothendieck{}{\Operators}}{\Comma{*}{\Operators}}$ we have a category
-of elements of $\Operators$ where objects are triples
+$\Iso{\Grothendieck{}{\Operators}}{\Comma{*}{\Operators}}$ we have a category of
+elements of $\Operators$ where objects are triples
 $\Pair{\Pair{\Upsilon}{a}}{\vartheta}$, interpreted as
-$\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$, and morphisms are diagrams:
+$\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$, and morphisms are paired
+diagrams:
 \[
@@
 \begin{alignat*}{3}
-  &{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}&&\,\mathrel{\ {{:}}}\,&&\SCtx\times\Discrete{\Arities}\to\Sets\\
-  &{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}&&\,=\,\ && %
-    \{\ \ \Member{\vartheta}{\Pair{\Upsilon}{a}}\ \ |\ \
+  &\IMode{{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}}&&\,\mathrel{\ {{:}}}\,&&\IMode{\SCtx\times\Discrete{\Arities}\to\Sets}\\
+  &\IMode{{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}}&&\,=\,\ && %
+    \OMode{%
+      \MkSet{
+        \vartheta\in\Pair{\Upsilon}{a}
+        \mid
         {\pi_{\Operators}}{\Pair{\Pair{\Upsilon}{a}}{\vartheta}} =
-    \Pair{\Upsilon}{a}\,\}
+        \Pair{\Upsilon}{a}
+      }
+    }
 \end{alignat*}
@@
 \begin{alignat*}{5}
-{\pi_{{\Operators}_{\lambda}}}^{-1}&\Pair{\Upsilon}{{\MkArity{\MkValence{}{\SortExp}{\SortExp}}{\SortExp}}}\quad&&=
-  &&\quad\MkSet{\mathtt{lam}}\\
-{\pi_{{\Operators}_{\lambda}}}^{-1}&\Pair{\Upsilon}{\MkArity{\MkValence{}{}{\SortExp},\MkValence{}{}{\SortExp}}{\SortExp}}\quad&&=
-  &&\quad\MkSet{\mathtt{ap}}\\
-{\pi_{{\Sigma}_{\lambda}}}^{-1}&=\bigcup_{\Pair{\Upsilon}{a}}{{\pi_{{\Operators}_{\lambda}}}^{-1}{\Pair{\Upsilon}{a}}}\quad&&=
-  &&\quad\MkSet{\mathtt{lam},\ \mathtt{ap}}\\
+\IMode{{\pi_{{\Operators}_{\lambda}}}^{-1}}&\IMode{\Pair{\Upsilon}{{\MkArity{\MkValence{}{\SortExp}{\SortExp}}{\SortExp}}}}\quad&&=
+  &&\quad\OMode{\MkSet{\mathtt{lam}}}\\
+\IMode{{\pi_{{\Operators}_{\lambda}}}^{-1}}&\IMode{\Pair{\Upsilon}{\MkArity{\MkValence{}{}{\SortExp},\MkValence{}{}{\SortExp}}{\SortExp}}}\quad&&=
+  &&\quad\OMode{\MkSet{\mathtt{ap}}}\\
+\IMode{{\pi_{{\Sigma}_{\lambda}}}^{-1}}&\triangleq\IMode{\bigcup_{\Pair{\Upsilon}{a}}{{\pi_{{\Operators}_{\lambda}}}^{-1}{\Pair{\Upsilon}{a}}}}\quad&&=
+  &&\quad\OMode{\MkSet{\mathtt{lam},\ \mathtt{ap}}}\\
 \end{alignat*}

### 2015-11-11 paper.tex
[paper] improve interpretation diagrams

@@
 The construction of $\beta$ proceeds by renaming the parameters of the operator
-$\vartheta$ and constructing the (bound) arguments $\gamma$ of the operator.
+$\vartheta$ and constructing the (bound) exponentiated arguments $\gamma$ of the
+operator.
 \[
@@
     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
-       \arrow[r, "\Define{\beta\,}{\,\Tuple{\SemBrackets{\vartheta}_P, \gamma}}"]
-    \&[3.0em] {\IMode{%
+       \arrow[r, "\Define{\beta\,}{\,\Tuple{\SemBrackets{\vartheta}_P, \lambda\gamma}}"]
+    \&[4.0em] {\IMode{%
        \coprod_{\vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}}%
@@
 \]
-Each argument
-%
+Argument $\lambda\IMode{{\gamma}_{i}}$ are the exponential transposes (curried
+form) of the composites $\IMode{{\gamma}_{i}}$:
+\[
   \begin{tikzcd}[cramped, ampersand replacement = \&]
-  {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
-    \arrow[r, "\IMode{\gamma_i}"]
-  \& {\IMode{%
-    P_{\tau_i}^{\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}}
-  }}
-\end{tikzcd}
-%
-is the exponential transpose of the composite,
-\[
+    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_{P}}}
+      \arrow[d, dashed, swap, "\IMode{\lambda{\gamma}_{i}}"]
+    \&
+    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_{P}
+       \times\Yoneda{\vec{\sigma}_{i}\parallel\vec{\tau}_{i}}}
+    }
+      \arrow[r, "\IMode{\Tuple{{\pi}_{[1,1]},{\phi}_{i},{\psi}_{i}}}"]
+      \arrow[d, dashed, swap, "\IMode{{\gamma}_{i}}"]
+    \&[3.5em]
+    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon,\vec{\sigma}_i\parallel\Gamma,\vec{\tau}_{i}}_{P}}}
+      \ar{dl}[
+        , rotate = +19.25
+        , xshift = -4.5em
+        , yshift = -0.2em
+      ]{\IMode{\SemBrackets{\Omega\triangleright\Upsilon,\vec{\sigma}_{i}\parallel\Gamma,\vec{\tau}_{i}\vdash M_i:\tau_i}_P}}
+    \\[2.0em]
+    {\IMode{{P}_{{\tau}_{i}}^{\Yoneda{\vec{\sigma}_{i}\parallel\vec{\tau}_{i}}}}}
+    \& {\IMode{P_{{\tau}_{i}}}}
+    \&
+  \end{tikzcd}
+\]
+where $\phi_i,\psi_i$ are defined as follows:
+\begin{gather*}
   \begin{tikzcd}[cramped, ampersand replacement = \&]
@@
     }}
-      \arrow[r, "\IMode{\Tuple{\pi_1\circ\pi_1,\phi_i,\psi_i}}"]
-    \&[4.0em] {\IMode{\SemBrackets{\Omega\triangleright\Upsilon,\vec{\sigma}_i\parallel\Gamma,\vec{\tau}_i}_P}}
-      \arrow[r, "\IMode{\SemBrackets{\Omega\triangleright\Upsilon,\vec{\sigma}_i\parallel\Gamma,\vec{\tau}_i\vdash M_i:\tau_i}_P}"]
-    \&[7.5em] {\IMode{P_{\tau_i}}}
-  \end{tikzcd}
-\]
-where $\phi_i,\psi_i$ are defined as follows:
-\begin{gather*}
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
-       \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}
-    }}
-      \arrow[r, "\Define{\phi_i\,}{\,\Tuple{\pi_u\circ\pi_1}_{u\in\Dom{\Upsilon}},\, \Tuple{\pi_u\circ\pi_1\circ\pi_2}_{u\in\Dom{\vec{\sigma}_i}}}"]
+      \arrow[r, "\Define{\phi_i\,}{\,\Pair{\Tuple{\pi_u\circ\pi_1}_{u\in\Dom{\Upsilon}}}{\,\Tuple{\pi_u\circ\pi_{[2,1]}}_{u\in\Dom{\vec{\sigma}_i}}}}"]
     \&[12.0em] {\IMode{S^{\{\Upsilon,\vec{\sigma}_i\}}}}\\[-1.0em]
@@
     }}
-      \arrow[r, "\Define{\psi_i\,}{\,\Tuple{\pi_x\circ\pi_1}_{x\in\Dom{\Gamma}},\, \Tuple{\pi_x\circ\pi_1\circ\pi_2}_{x\in\Dom{\vec{\tau}_i}}}"]
+      \arrow[r, "\Define{\psi_i\,}{\,\Pair{\Tuple{\pi_x\circ\pi_1}_{x\in\Dom{\Gamma}}}{\,\Tuple{\pi_x\circ\pi_{[2,1]}}_{x\in\Dom{\vec{\tau}_i}}}}"]
     \&[12.0em] {\IMode{V^{[\Gamma,\vec{\tau}_i]}}}

### 2015-11-12 paper.tex
Sucky workaround for Euler lacking \varsigma, \varrho
[paper] consolidate the diagrams a bit more

@@
 
+%% Euler doesn't have \varsigma and \varrho
+\DeclareSymbolFont{greekletters}{OML}{cmr}{m}{it}
+\DeclareMathSymbol{\varrho}{\mathalpha}{greekletters}{"25}
+\DeclareMathSymbol{\varsigma}{\mathalpha}{greekletters}{"26}
+
 \newcommand\IMode[1]{{\color{\IModeColorName}{#1}}}
@@
 
-The interpretation of a variable
-$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x:\tau}_P$ simply
-projects it of the environment and embeds it into the model, as follows:
+Variables are interpreted by the map
+$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x:\tau}_P$ which
+projects them from the environment and embeds them into the model. Metavariables
+are resolved by the map
+$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash\MApp{\MV{m}}{\vec{u}}{\vec{M}}
+  : \tau}_P$ (where
+$\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$) which
+projects their interpretation from the environment and intantiates it via
+substitution:
 \[
   \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}} \arrow[r, "\IMode{\pi_3}"]
-    \& {\IMode{V^{[\Gamma]}}} \arrow[r, "\IMode{\nu_{\Gamma}}"]
-    \& {\IMode{P^{[\Gamma]}}} \arrow[r, "\IMode{\pi_x}"]
-    \& \IMode{P_\tau}
+    {\IMode{P_\tau}}
+    \&[8.0em] {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
+      \arrow[rr, dashed, "\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x:\tau}_P"]
+      \ar{drr}[
+        , rotate = -16.5
+        , xshift = +1.5em
+        , yshift = -0.2em
+        , swap
+      ]{{\nu}_{\Gamma} {\pi}_{3}}
+      \ar{dl}[
+        , rotate = +11.5
+        , xshift = -2.75em
+      ]{\IMode{\Tuple{\pi_\MV{m}\,\pi_1, \phi\,\pi_2, \psi}}}
+      \arrow[l, dashed, swap, "\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash\MApp{\MV{m}}{\vec{u}}{\vec{M}}: \tau}_P"]
+    \&[-2.0em]
+    \&[+3.0em]{\IMode{P_\tau}}
+    \\
+    {\IMode{P_{\tau}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}\times S^{\{\vec{\sigma}\}} \times P^{[\vec{\tau}]}}}
+      \arrow[u, "\IMode{\varsigma^\tau_{\vec{\sigma}\parallel\vec{\tau}}}"]
+    \&{}
+    \&{}
+    \& {\IMode{P^{[\Gamma]}}}
+      \arrow[u, swap, "{\pi}_{x}"]
+    \\
+    {\IMode{P^{[\vec{\tau}]}}}
+    \& {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
+      \arrow[l, swap, "\Define{\psi\,}{\,\Tuple{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M:\tau}_P}_{(M,\tau)\in (\vec{M},\vec{\tau})}}"]
+    \& {\IMode{S^{\{\Upsilon\}}}}
+      \arrow[r, "\Define{\phi\,}{\,\Tuple{\pi_u}_{u\in\vec{u}}}"]
+    \& {\IMode{S^{\{\vec{\sigma}\}}}}
   \end{tikzcd}
-  \tag{var}
-\]
-
-Next interpret metavariable applications
-$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash\MApp{\MV{m}}{\vec{u}}{\vec{M}}
-: \tau}_P$ where
-$\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$, which
-proceeds by projecting the metavariable's interpretation from the environment
-and intantiating it via substitution:
-\[
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
-          \arrow[r, "\IMode{\Tuple{\pi_\MV{m}\,\pi_1, \phi\,\pi_2, \psi}}"]
-    \&[4.0em] {\IMode{P_{\tau}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}\times S^{\{\vec{\sigma}\}} \times P^{[\vec{\tau}]}}}
-          \arrow[r, "\IMode{\varsigma^\tau_{\vec{\sigma}\parallel\vec{\tau}}}"]
-    \&[1.0em] \IMode{P_\tau}
-  \end{tikzcd}
-  \tag{metavar}
-\]
-where
-\vspace{-0.5em}
-\begin{gather*}
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{S^{\{\Upsilon\}}}} \arrow[r, "\Define{\phi\,}{\,\Tuple{\pi_u}_{u\in\vec{u}}}"]
-    \&[3.0em] {\IMode{S^{\{\vec{\sigma}\}}}}
-  \end{tikzcd}
-  \qquad\begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
-      \arrow[r, "\Define{\psi\,}{\,\Tuple{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M:\tau}_P}_{(M,\tau)\in (\vec{M},\vec{\tau})}}"]
-    \&[11.0em] {\IMode{P^{[\vec{\tau}]}}}
-  \end{tikzcd}
-\end{gather*}
-
-The interpretation of operator applications is most complicated. Recall that,
+\]
+
+Interpretation of operator applications is the most complicated. Recall that,
 unlike in standard treatments of universal algebra, our operators are indexed by
@@
 \]
-Argument $\lambda\IMode{{\gamma}_{i}}$ are the exponential transposes (curried
+Arguments $\lambda\IMode{{\gamma}_{i}}$ are the exponential transposes (curried
 form) of the composites $\IMode{{\gamma}_{i}}$:

### 2015-11-12 paper.tex
[paper] write extension in pointful style

@@
     functor.} on the functor $\Of{V}{\widehat{\HypCat}^\Sorts}$ of variables.
-  Consequently, the monoid multiplication $\varsigma$ can be replaced with an
-  extension operation $\Of{{(\,-\,)}^{\circledast}}{{\IMode{{P(\,=\,)}^{V(\,-\,)}}}\to{\IMode{{P(\,=\,)}^{P(\,-\,)}}}}$.
+  Consequently, monoid multiplication $\varsigma$ can be replaced with extension
+$\Of{\circledast}{\int_{\Tuple{\tau, (\Upsilon\parallel\Gamma), (\Upsilon'\parallel\Gamma')}}{{{P}_{\tau}(\Upsilon'\parallel\Gamma')}^{{V}_{\tau}(\Upsilon\parallel\Gamma)} \longrightarrow {{P}_{\tau}(\Upsilon'\parallel\Gamma')}^{{P}_{\tau}(\Upsilon\parallel\Gamma)}}}$.
 \end{remark}

### 2015-11-12 paper.tex
[paper] use \mathbb{1} from bbmath

@@
 By the Grothendieck construction
-$\Iso{\Grothendieck{}{\Operators}}{\Comma{*}{\Operators}}$ we have a category of
+$\Iso{\Grothendieck{}{\Operators}}{\Comma{\mathbb{1}}{\Operators}}$ we have a category of
 elements of $\Operators$ where objects are triples
@@
   \qquad\begin{tikzcd}[cramped, sep = small, ampersand replacement = \&]
-    \& *
+    \& \mathbb{1}
       \arrow[dl, "\vartheta", swap]

### 2015-11-12 paper.tex
[paper] fix chktex warnings

@@
 where operators may express variable binding structure as part of their
-arities.  Originally formulated by Peter Aczel \cite{aczel:1978}, unisorted
+arities.  Originally formulated by Peter Aczel~\cite{aczel:1978}, unisorted
 abts have been deployed successfully as the uniform syntactic framework for
 several implementations of Constructive Type Theory, including Nuprl
-\cite{constable:1986}, MetaPRL \cite{hickey:2003} and JonPRL
+\cite{constable:1986}, MetaPRL~\cite{hickey:2003} and JonPRL
 \cite{jonprl:2015}.
 
-In \emph{Practical Foundations for Programming Languages} \cite{harper:2016},
+In \emph{Practical Foundations for Programming Languages}~\cite{harper:2016},
 Robert Harper develops the multi-sorted version of abstract binding trees and
@@
 developed the categorical semantics for abts formally in Constructive
-Type Theory using the Agda proof assistant \cite{norell:2007}.
+Type Theory using the Agda proof assistant~\cite{norell:2007}.
 
@@
 the abt logical framework, we can proceed to define proper wellformedness via
-\emph{inductive refinement} \cite{harper:2016}. The basic idea is to introduce a new form of
+\emph{inductive refinement}~\cite{harper:2016}. The basic idea is to introduce a new form of
 (meta)-judgment $\IsWf{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$ which expresses the

### 2015-11-12 paper.tex
[paper] fix typo

@@
 $\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$) which
-projects their interpretation from the environment and intantiates it via
+projects their interpretation from the environment and instantiates it via
 substitution:

### 2015-11-13 paper.tex
[paper] adjust modes and edge styles
Clean up packages
Use tikz for inline arrows

@@
 \usepackage{catchfilebetweentags}
-%\usepackage{fontspec}
 \usepackage{geometry}
@@
 \usepackage{ifthen}
-%\usepackage{lualatex-math}
 \usepackage{mathtools}
@@
 \usepackage{tikz-cd}
-%\usepackage{unicode-math}
+\usepackage{tikzpfeile}
 \usepackage{url}
-%\usepackage{xunicode}
 
@@
 \microtypecontext{spacing=nonfrench}
-% \setmainfont{Linux Libertine O}
-% \setmathfont{Asana Math}
 
@@
 \newcommand\Discrete[1]{#1_{\equiv}}
-\newcommand\Arr[1]{\mathbf{{#1}^{\rightarrow}}}
+\newcommand\Arr[1]{\mathbf{{#1}^{\ra}}}
 \newcommand\Sets{\mathbf{Set}}
@@
 category of contexts of symbols, whose objects are finite sets of symbols $U$
-and sort-assignments $\Of{\mathfrak{s}}{U\to\Sorts}$, and whose morphisms are
+and sort-assignments $\Of{\mathfrak{s}}{U\ra\Sorts}$, and whose morphisms are
 sort-preserving renamings; we will write $\Upsilon$ for a symbol context
@@
 Then, fix a covariant presheaf (copresheaf) of operators
-$\Of{\Operators}{\SCtx\times\Discrete{\Arities}\to\Sets}$ such that the arrows
+$\Of{\Operators}{\SCtx\times\Discrete{\Arities}\ra\Sets}$ such that the arrows
 in $\SCtx$ lift to renamings of operators' parameters. Together, $\Sorts$ and
@@
 specifying the fibers of
-$\Of{\pi_{\Operators}}{\Grothendieck{}{\Operators}\to\SCtx\times\Discrete{\Arities}}$
+$\Of{\pi_{\Operators}}{\Grothendieck{}{\Operators}\ra\SCtx\times\Discrete{\Arities}}$
 in which they reside:
 \begin{alignat*}{3}
-  &\IMode{{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}}&&\,\mathrel{\ {{:}}}\,&&\IMode{\SCtx\times\Discrete{\Arities}\to\Sets}\\
+  &\IMode{{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}}&&\,\mathrel{\ {{:}}}\,&&\IMode{\SCtx\times\Discrete{\Arities}\ra\Sets}\\
   &\IMode{{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}}&&\,=\,\ && %
@@
 Note that the functoriality of $\Operators$ guarantees for any renaming
-$\Of{\varrho}{\Upsilon\to\Upsilon'}$ and $\Of{f}{a\to a'}$, a lifted map
-$\Of{\FLift{\varrho}{f}}{\Operators\Pair{\Upsilon}{a}\to\Operators\Pair{\Upsilon'}{a'}}$
+$\Of{\varrho}{\Upsilon\ra\Upsilon'}$ and $\Of{f}{a\ra a'}$, a lifted map
+$\Of{\FLift{\varrho}{f}}{\Operators\Pair{\Upsilon}{a}\ra\Operators\Pair{\Upsilon'}{a'}}$
 such that $\IsOperator{\Upsilon'}{\FLift{\varrho}{f}(\vartheta)}{a'}$ when
@@
 $\IsRenaming{\varrho}{E}{F}$, presupposing
-$\Of{\varrho}{\Upsilon\to\Upsilon'}$, and $\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$
+$\Of{\varrho}{\Upsilon\ra\Upsilon'}$, and $\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$
 and $\IsAbt{\Omega}{\Upsilon}{\Gamma}{E}{v}$ respectively:
@@
   Consequently, monoid multiplication $\varsigma$ can be replaced with extension
-$\Of{\circledast}{\int_{\Tuple{\tau, (\Upsilon\parallel\Gamma), (\Upsilon'\parallel\Gamma')}}{{{P}_{\tau}(\Upsilon'\parallel\Gamma')}^{{V}_{\tau}(\Upsilon\parallel\Gamma)} \longrightarrow {{P}_{\tau}(\Upsilon'\parallel\Gamma')}^{{P}_{\tau}(\Upsilon\parallel\Gamma)}}}$.
+$\Of{\circledast}{\int_{\Tuple{\tau, (\Upsilon\parallel\Gamma), (\Upsilon'\parallel\Gamma')}}{{{P}_{\tau}(\Upsilon'\parallel\Gamma')}^{{V}_{\tau}(\Upsilon\parallel\Gamma)} \ra {{P}_{\tau}(\Upsilon'\parallel\Gamma')}^{{P}_{\tau}(\Upsilon\parallel\Gamma)}}}$.
 \end{remark}
@@
 endofunctor
-$\Of{\SigFun{\Sigma}}{\widehat{\HypCat}^\Sorts\to\widehat{\HypCat}^\Sorts}$,
+$\Of{\SigFun{\Sigma}}{\widehat{\HypCat}^\Sorts\ra\widehat{\HypCat}^\Sorts}$,
 which is defined as follows:
@@
 Then, a $\Sigma$-model is a $\Sigma$-monoid $P$ which is equipped with an initial
-algebra $\Of{\alpha}{\SigFun{\Sigma}(P)\to P}$, which shall interpret
+algebra $\Of{\alpha}{\SigFun{\Sigma}(P)\ra P}$, which shall interpret
 applications of each operator.
@@
   \Of{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash%
-      M:\tau}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P\longrightarrow P_\tau}%
+      M:\tau}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P\ra P_\tau}%
 \]
@@
     \&[8.0em] {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
-      \arrow[rr, dashed, "\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x:\tau}_P"]
+      \arrow[rr, dashed, "\OMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x:\tau}_P}"]
       \ar{drr}[
@@
       ]{\IMode{\Tuple{\pi_\MV{m}\,\pi_1, \phi\,\pi_2, \psi}}}
-      \arrow[l, dashed, swap, "\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash\MApp{\MV{m}}{\vec{u}}{\vec{M}}: \tau}_P"]
+      \arrow[l, dashed, swap, "\OMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash\MApp{\MV{m}}{\vec{u}}{\vec{M}}: \tau}_P}"]
     \&[-2.0em]
@@
     \& {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
-      \arrow[l, swap, "\Define{\psi\,}{\,\Tuple{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M:\tau}_P}_{(M,\tau)\in (\vec{M},\vec{\tau})}}"]
+      \arrow[l, dashed, swap, "\Define{\psi\,}{\,\Tuple{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M:\tau}_P}_{(M,\tau)\in (\vec{M},\vec{\tau})}}"]
     \& {\IMode{S^{\{\Upsilon\}}}}
-      \arrow[r, "\Define{\phi\,}{\,\Tuple{\pi_u}_{u\in\vec{u}}}"]
+      \arrow[r, dashed, "\Define{\phi\,}{\,\Tuple{\pi_u}_{u\in\vec{u}}}"]
     \& {\IMode{S^{\{\vec{\sigma}\}}}}
@@
     {\IMode{\Operators_a}}
-    \arrow[from = Src, to = Tgt, bend left, dashed, "\IMode{\SemBrackets{\vartheta}_{P}}"]
+    \arrow[from = Src, to = Tgt, bend left, dashed, "\OMode{\SemBrackets{\vartheta}_{P}}"]
   \end{tikzcd}
@@
     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
-       \arrow[r, "\Define{\beta\,}{\,\Tuple{\SemBrackets{\vartheta}_P, \lambda\gamma}}"]
+       \arrow[r, dashed, "\Define{\beta\,}{\,\Tuple{\SemBrackets{\vartheta}_P, \lambda\gamma}}"]
     \&[4.0em] {\IMode{%
@@
     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_{P}}}
-      \arrow[d, dashed, swap, "\IMode{\lambda{\gamma}_{i}}"]
+      \arrow[d, dashed, swap, "\OMode{\lambda{\gamma}_{i}}"]
     \&
@@
       \arrow[r, "\IMode{\Tuple{{\pi}_{[1,1]},{\phi}_{i},{\psi}_{i}}}"]
-      \arrow[d, dashed, swap, "\IMode{{\gamma}_{i}}"]
+      \arrow[d, dashed, swap, "\OMode{{\gamma}_{i}}"]
     \&[3.5em]
@@
     }}
-      \arrow[r, "\Define{\phi_i\,}{\,\Pair{\Tuple{\pi_u\circ\pi_1}_{u\in\Dom{\Upsilon}}}{\,\Tuple{\pi_u\circ\pi_{[2,1]}}_{u\in\Dom{\vec{\sigma}_i}}}}"]
+      \arrow[r, dashed, "\Define{\phi_i\,}{\,\Pair{\Tuple{\pi_u\circ\pi_1}_{u\in\Dom{\Upsilon}}}{\,\Tuple{\pi_u\circ\pi_{[2,1]}}_{u\in\Dom{\vec{\sigma}_i}}}}"]
     \&[12.0em] {\IMode{S^{\{\Upsilon,\vec{\sigma}_i\}}}}\\[-1.0em]
@@
     }}
-      \arrow[r, "\Define{\psi_i\,}{\,\Pair{\Tuple{\pi_x\circ\pi_1}_{x\in\Dom{\Gamma}}}{\,\Tuple{\pi_x\circ\pi_{[2,1]}}_{x\in\Dom{\vec{\tau}_i}}}}"]
+      \arrow[r, dashed, "\Define{\psi_i\,}{\,\Pair{\Tuple{\pi_x\circ\pi_1}_{x\in\Dom{\Gamma}}}{\,\Tuple{\pi_x\circ\pi_{[2,1]}}_{x\in\Dom{\vec{\tau}_i}}}}"]
     \&[12.0em] {\IMode{V^{[\Gamma,\vec{\tau}_i]}}}

### 2015-11-13 paper.tex
Use todonotes
Clean up

@@
 \usepackage{tikzpfeile}
+\usepackage{todonotes}
 \usepackage{url}
@@
 Abstract binding trees (abts) are a generalization of abstract syntax trees
-where operators may express variable binding structure as part of their
-arities.  Originally formulated by Peter Aczel~\cite{aczel:1978}, unisorted
-abts have been deployed successfully as the uniform syntactic framework for
-several implementations of Constructive Type Theory, including Nuprl
-\cite{constable:1986}, MetaPRL~\cite{hickey:2003} and JonPRL
-\cite{jonprl:2015}.
+where operators may express variable binding structure as part of their arities.
+Originally formulated by Peter Aczel~\cite{aczel:1978}, unisorted abts have been
+deployed successfully as the uniform syntactic framework for several
+implementations of Constructive Type Theory, including
+Nuprl~\cite{constable:1986}, MetaPRL~\cite{hickey:2003} and
+JonPRL~\cite{jonprl:2015}.
 
@@
 not appear in the syntax of abts, and are only introduced as parameters to
-operators. This extension to support symbols is essential for a
-correct treatment of programming languages with open sums (e.g. ML's
-\texttt{exn} type), as well as assignable references.
+operators. This extension to support symbols is essential for a correct
+treatment of programming languages with open sums (e.g. ML's \texttt{exn} type),
+as well as assignable references.
 
 In parallel, M.\ Fiore and his collaborators have developed the categorical
-semantics for several variations of second-order algebraic theories
-\cite{fiore:1999, fiore:2005, fiore-hur:2010, fiore-mamoud:2010}; of these, the
-simply sorted variants are equivalent to Harper's abstract binding trees
-augmented with a notion of second-order variable (metavariable).
-
-The contribution of this paper is the development of the syntax and semantics
-of multi-sorted abts, an extension of second order universal
-algebra to support symbol-indexed families of operators. Additionally, we have
-developed the categorical semantics for abts formally in Constructive
-Type Theory using the Agda proof assistant~\cite{norell:2007}.
+semantics for several variations of second-order algebraic
+theories~\cite{fiore:1999, fiore:2005, fiore-hur:2010, fiore-mamoud:2010}; of
+these, the simply sorted variants are equivalent to Harper's abstract binding
+trees augmented with a notion of second-order variable (metavariable).
+
+The contribution of this paper is the development of the syntax and semantics of
+multi-sorted abts, an extension of second order universal algebra to support
+symbol-indexed families of operators. Additionally, we have developed the
+categorical semantics for abts formally in Constructive Type Theory using the
+Agda proof assistant~\cite{norell:2007}.
 
@@
 Fix a set $\Sorts$ of \emph{sorts}. We will say $\IsSort{\tau}$ when
-$\Member{\tau}{\Sorts}$. A valence $\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}$ specifies an
-expression of sort $\tau$ which binds symbols in $\vec{\sigma}$ and variables in
-$\vec{\tau}$.
-
-\[
-  \infer{
-    \IsValence{\MkValence{\sigma_0,\dots,\sigma_m}{\tau_0,\dots,\tau_n}{\tau}}
-  }{
-    \IsSort{\tau} &
-    \HypJ{\IsSort{\sigma_i}}{\Leq{i}{m}} &
-    \HypJ{\IsSort{\tau_i}}{\Leq{i}{n}} &
+$\Member{\tau}{\Sorts}$. A valence $\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}$
+specifies an expression of sort $\tau$ which binds symbols in $\vec{\sigma}$ and
+variables in $\vec{\tau}$.
+
+\[
+  \infer{
+    \IsValence{
+      \MkValence
+        {\sigma_0,\dots,\sigma_m}
+        {\tau_0,\dots,\tau_n}
+        {\tau}
+    }
+  }{
+    \IsSort{\tau}
+&
+    \HypJ{\IsSort{\sigma_i}}{\Leq{i}{m}}
+&
+    \HypJ{\IsSort{\tau_i}}{\Leq{i}{n}}
   }
@@
   \infer{
-    \IsArity{\MkArity{v_0,\ldots,v_n}{\tau}}
-  }{
-    \IsSort{\tau} &
-    \HypJ{\IsValence{v_i}}{\Leq{i}{n}}
+    \IsArity{
+      \MkArity
+        {v_0,\ldots,v_n}
+        {\tau}
+    }
+  }{
+    \IsSort{\tau}
+&
+    \HypJ
+      {\IsValence{v_i}}
+      {\Leq{i}{n}}
   }
@@
 in $\SCtx$ lift to renamings of operators' parameters. Together, $\Sorts$ and
-$\Operators$ form a \emph{signature} $\Define{\Sigma}{\Pair{\Sorts}{\Operators}}$.
+$\Operators$ form a \emph{signature}
+$\Define{\Sigma}{\Pair{\Sorts}{\Operators}}$.
 
 By the Grothendieck construction
-$\Iso{\Grothendieck{}{\Operators}}{\Comma{\mathbb{1}}{\Operators}}$ we have a category of
-elements of $\Operators$ where objects are triples
+$\Iso{\Grothendieck{}{\Operators}}{\Comma{\mathbb{1}}{\Operators}}$ we have a
+category of elements of $\Operators$ where objects are triples
 $\Pair{\Pair{\Upsilon}{a}}{\vartheta}$, interpreted as
@@
     {\Pair{\Upsilon}{a}}
-      \arrow[r, "\Pair{\rho}{f}", swap]
-    \& {\Pair{\Upsilon'}{a'}}
+      \arrow[r, swap, "\Pair{\rho}{f}"]
+\&
+    {\Pair{\Upsilon'}{a'}}
   \end{tikzcd}
-  \qquad\begin{tikzcd}[cramped, sep = small, ampersand replacement = \&]
-    \& \mathbb{1}
-      \arrow[dl, "\vartheta", swap]
+\qquad
+  \begin{tikzcd}[cramped, sep = small, ampersand replacement = \&]
+    \null
+\&
+    \mathbb{1}
+      \arrow[dl, swap, "\vartheta"]
       \arrow[dr, "\vartheta'"]
 \&
+    \null
 \\
     {\Operators\Pair{\Upsilon}{a}}
-      \arrow[rr, "\Operators\Pair{\rho}{f}", swap]
-    \&
-    \& {\Operators\Pair{\Upsilon'}{a'}}
+      \arrow[rr, swap, "\Operators\Pair{\rho}{f}"]
+\&
+    \null
+\&
+    {\Operators\Pair{\Upsilon'}{a'}}
   \end{tikzcd}
@@
 \begin{alignat*}{3}
-  &\IMode{{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}}&&\,\mathrel{\ {{:}}}\,&&\IMode{\SCtx\times\Discrete{\Arities}\ra\Sets}\\
-  &\IMode{{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}}&&\,=\,\ && %
-    \OMode{%
+  \null
+&
+  \IMode{{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}}
+&&
+  \,\mathrel{\ {{:}}}\,
+&&
+  \IMode{\SCtx\times\Discrete{\Arities}\ra\Sets}
+\\
+  \null
+&
+  \IMode{{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}}
+&&
+  \,=\,\
+&&
+    \OMode{
       \MkSet{
         \vartheta\in\Pair{\Upsilon}{a}
-        \mid
-        {\pi_{\Operators}}{\Pair{\Pair{\Upsilon}{a}}{\vartheta}} =
+        \mid{\pi_{\Operators}}{\Pair{\Pair{\Upsilon}{a}}{\vartheta}} =
         \Pair{\Upsilon}{a}
@@
   \IsOperator{\Upsilon}{\mathtt{lam}}{
-    \MkArity{\MkValence{}{\SortExp}{\SortExp}}{\SortExp}
-  }\\
+    \MkArity{
+      \MkValence{}{\SortExp}{\SortExp}
+    }{\SortExp}
+  }
+\\
   \IsOperator{\Upsilon}{\mathtt{ap}}{
@@
 \end{gather*}
-These rules correspond to objects of $\Grothendieck{}{{\Operators}_{\lambda}}$ and we
-collect their fibers into sets of operators:
+These rules correspond to objects of $\Grothendieck{}{{\Operators}_{\lambda}}$
+and we collect their fibers into sets of operators:
 \begin{alignat*}{5}
-\IMode{{\pi_{{\Operators}_{\lambda}}}^{-1}}&\IMode{\Pair{\Upsilon}{{\MkArity{\MkValence{}{\SortExp}{\SortExp}}{\SortExp}}}}\quad&&=
-  &&\quad\OMode{\MkSet{\mathtt{lam}}}\\
-\IMode{{\pi_{{\Operators}_{\lambda}}}^{-1}}&\IMode{\Pair{\Upsilon}{\MkArity{\MkValence{}{}{\SortExp},\MkValence{}{}{\SortExp}}{\SortExp}}}\quad&&=
-  &&\quad\OMode{\MkSet{\mathtt{ap}}}\\
-\IMode{{\pi_{{\Sigma}_{\lambda}}}^{-1}}&\triangleq\IMode{\bigcup_{\Pair{\Upsilon}{a}}{{\pi_{{\Operators}_{\lambda}}}^{-1}{\Pair{\Upsilon}{a}}}}\quad&&=
-  &&\quad\OMode{\MkSet{\mathtt{lam},\ \mathtt{ap}}}\\
+  \IMode{{\pi_{{\Operators}_{\lambda}}}^{-1}}
+&
+  \IMode{
+    \Pair{
+      \Upsilon
+    }{
+      \MkArity{
+        \MkValence{}{\SortExp}{\SortExp}
+      }{\SortExp}
+    }
+  }
+\quad
+&&
+  =
+&&
+\quad
+  \OMode{\MkSet{\mathtt{lam}}}
+\\
+  \IMode{{\pi_{{\Operators}_{\lambda}}}^{-1}}
+&
+  \IMode{
+    \Pair{
+      \Upsilon
+    }{
+      \MkArity{
+        \MkValence{}{}{\SortExp},
+        \MkValence{}{}{\SortExp}
+      }{\SortExp}
+    }
+  }
+\quad
+&&
+  =
+&&
+\quad
+  \OMode{\MkSet{\mathtt{ap}}}
+\\
+  \IMode{{\pi_{{\Sigma}_{\lambda}}}^{-1}}
+&
+  \triangleq\IMode{
+    \bigcup_{\Pair{\Upsilon}{a}}{
+      {\pi_{{\Operators}_{\lambda}}}^{-1}
+      \Pair{\Upsilon}{a}
+    }
+  }
+\quad
+&&
+  =
+&&
+\quad
+  \OMode{
+    \MkSet{\mathtt{lam},\ \mathtt{ap}}
+  }
+\\
 \end{alignat*}
@@
     }
-  }\\
+  }
+\\
   \IsOperator{\Upsilon,u:\SortExp}{\mathtt{get}[u]}{
     \MkArity{}{\SortExp}
-  }\\
+  }
+\\
   \IsOperator{\Upsilon,u:\SortExp}{\mathtt{set}[u]}{
@@
 
-In general, we will have three kinds of context: metavariable contexts,
-variable contexts, and symbol (parameter) contexts. A metavariable context
-$\Omega$ consists of bindings of valences to metavariables; a variable context
-$\Gamma$ is a collection of bindings of sorts to variables, and a parameter
-context $\Upsilon$ is a collection of bindings of sorts to symbols.
+In general, we will have three kinds of context: metavariable contexts, variable
+contexts, and symbol (parameter) contexts. A metavariable context $\Omega$
+consists of bindings of valences to metavariables; a variable context $\Gamma$
+is a collection of bindings of sorts to variables, and a parameter context
+$\Upsilon$ is a collection of bindings of sorts to symbols.
 
@@
   }{
-  }\qquad
+  }
+\qquad
   \infer{
@@
   }{
-    \IsMetaCtx{\Omega} &
-    \IsValence{v} &
+    \IsMetaCtx{\Omega}
+&
+    \IsValence{v}
+&
     \NotIn{\MV{m}}{\Dom\Omega}
@@
   }{
-    \IsVarCtx{\Gamma} &
-    \IsSort{\tau} &
+    \IsVarCtx{\Gamma}
+&
+    \IsSort{\tau}
+&
     \NotIn{x}{\Dom\Gamma}
@@
   }{
-    \IsVarCtx{\Upsilon} &
-    \IsSort{\tau} &
+    \IsVarCtx{\Upsilon}
+&
+    \IsSort{\tau}
+&
     \NotIn{u}{\Dom\Upsilon}
@@
 Let the judgment $\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$ presuppose
-$\IsMetaCtx{\Omega}$, $\IsSymCtx\Upsilon$, $\IsVarCtx\Gamma$ and $\IsSort{\tau}$,
-meaning that $M$ is an abstract binding tree of sort $s$, with metavariables in
-$\Omega$, parameters in $\Upsilon$, and variables in $\Gamma$. Let the judgment
-$\IsBTm{\Omega}{\Upsilon}{\Gamma}{E}{v}$ presuppose $\IsValence{v}$.  Then, the
-syntax of abstract binding trees is inductively defined in four rules:
+$\IsMetaCtx{\Omega}$, $\IsSymCtx\Upsilon$, $\IsVarCtx\Gamma$ and
+$\IsSort{\tau}$, meaning that $M$ is an abstract binding tree of sort $s$, with
+metavariables in $\Omega$, parameters in $\Upsilon$, and variables in $\Gamma$.
+Let the judgment $\IsBTm{\Omega}{\Upsilon}{\Gamma}{E}{v}$ presuppose
+$\IsValence{v}$. Then, the syntax of abstract binding trees is inductively
+defined in four rules:
 
@@
     \Lookup{\Gamma}{x}{\tau}
-  }\\[6pt]
-  %
+  }
+\\[6pt]
   \infer[\textit{mvar}]{
@@
     \begin{array}{l}
-      \Lookup{\Omega}{\MV{m}}{\MkValence{\sigma_0,\dots,\sigma_m}{\tau_0,\dots,\tau_n}{\tau}}\\
-      \HypJ{\Lookup{\Upsilon}{u_i}{\sigma_i}}{\Leq{i}{m}}\\
+      \Lookup{\Omega}{\MV{m}}{
+        \MkValence{
+          \sigma_0,\dots,\sigma_m
+        }{
+          \tau_0,\dots,\tau_n
+        }{\tau}
+      }
+\\
+      \HypJ{
+        \Lookup{\Upsilon}{u_i}{\sigma_i}
+      }{\Leq{i}{m}}
+\\
       \HypJ{
         \IsAbt{\Omega}{\Upsilon}{\Gamma}{M_i}{\tau_i}
-      }{
-        \Leq{i}{n}
-      }
+      }{\Leq{i}{n}}
     \end{array}
-  }\\[6pt]
-  %
+  }
+\\[6pt]
   \infer[\textit{app}]{
@@
         \MkArity{v_1,\dots,v_n}{\tau}
-      }\\
+      }
+\\
       \HypJ{
         \IsBTm{\Omega}{\Upsilon}{\Gamma}{E_i}{v_i}
-      }{
-        \Leq{i}{n}
-      }
+      }{\Leq{i}{n}}
     \end{array}
-  }\\[6pt]
+  }
+\\[6pt]
   \infer[\textit{abs}]{
@@
   }{
-    \IsAbt{\Omega}{\Upsilon, \vec{u}:\vec{\sigma}}{\Gamma,\vec{x}:{\vec{\tau}}}{
+    \IsAbt{
+      \Omega
+    }{
+      \Upsilon, \vec{u}:\vec{\sigma}
+    }{
+      \Gamma,\vec{x}:{\vec{\tau}}
+    }{
       M
-    }{
-      \tau
-    }
+    }{\tau}
   }
@@
 of symbols free in $M$, and let $\FV{M}$ be the collection of variables free in
-$M$. %% TODO: actually define these
+$M$. \todo[inline]{actually define these}
 
@@
 
-The only place that symbols appear in our calculus is as parameters to
-operators (unlike variables, symbols are not terms). Therefore, the functorial
-action of the operator copresheaf can be lifted into terms by recursion on their
+The only place that symbols appear in our calculus is as parameters to operators
+(unlike variables, symbols are not terms). Therefore, the functorial action of
+the operator copresheaf can be lifted into terms by recursion on their
 structure, via a pair of judgments $\IsRenaming{\varrho}{M}{N}$ and
-$\IsRenaming{\varrho}{E}{F}$, presupposing
-$\Of{\varrho}{\Upsilon\ra\Upsilon'}$, and $\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$
-and $\IsAbt{\Omega}{\Upsilon}{\Gamma}{E}{v}$ respectively:
+$\IsRenaming{\varrho}{E}{F}$, presupposing $\Of{\varrho}{\Upsilon\ra\Upsilon'}$,
+and $\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$ and
+$\IsAbt{\Omega}{\Upsilon}{\Gamma}{E}{v}$ respectively:
 
@@
   }{
-  }\\[6pt]
+  }
+\\[6pt]
   \infer{
@@
   }{
-    \Match{\varrho(\vec{u})}{\vec{v}} &
+    \Match{\varrho(\vec{u})}{\vec{v}}
+&
     \HypJ{
@@
     }
-  }\\[6pt]
+  }
+\\[6pt]
   \infer{
@@
   }{
-    \Match{\FLift{\varrho}{}(\vartheta)}{\vartheta'} &
+    \Match{\FLift{\varrho}{}(\vartheta)}{\vartheta'}
+&
     \HypJ{
@@
 Above, the notation $\varrho\setminus\vec{u}$ means the omission of the
-variables $\vec{u}$ from the renaming $\varrho$. Because terms are identified
-up to $\alpha$-equivalence, the renaming judgment is functional in its input,
-and so we are justified in writing $\Rename{\varrho}{M}$ for $N$ when
+variables $\vec{u}$ from the renaming $\varrho$. Because terms are identified up
+to $\alpha$-equivalence, the renaming judgment is functional in its input, and
+so we are justified in writing $\Rename{\varrho}{M}$ for $N$ when
 $\IsRenaming{\varrho}{M}{N}$.
 
-%%% TODO: show that renaming preserves wellformedness
+\todo[inline]{show that renaming preserves wellformedness}
 
@@
 
-Variable substitution in abts is defined inductively by a
-pair of judgments, $\IsSubst{N}{x}{M}{M'}$ and $\IsSubst{N}{x}{E}{F}$:
+Variable substitution in abts is defined inductively by a pair of judgments,
+$\IsSubst{N}{x}{M}{M'}$ and $\IsSubst{N}{x}{E}{F}$:
 
@@
     \IsEq{x}{y}
-  }\qquad
+  }
+\qquad
   \infer{
@@
     \IsApart{x}{y}
-  }\\[6pt]
+  }
+\\[6pt]
   \infer{
@@
   }{
-    \NotIn{x}{\vec{y}} &
-    \IsApart{\vec{u}}{\FS{N}} &
-    \IsApart{\vec{y}}{\FV{N}} &
+    \NotIn{x}{\vec{y}}
+&
+    \IsApart{\vec{u}}{\FS{N}}
+&
+    \IsApart{\vec{y}}{\FV{N}}
+&
     \IsSubst{N}{x}{M}{M'}
-  }\qquad
+  }
+\qquad
   \infer{
@@
   }{
-    \Member{x}{\vec{y}} &
-    \IsApart{\vec{u}}{\FS{N}} &
+    \Member{x}{\vec{y}}
+&
+    \IsApart{\vec{u}}{\FS{N}}
+&
     \IsApart{\vec{y}}{\FV{N}}
@@
 Going forward, we will write $\Subst{N}{x}{M}$ for $M'$ when
-$\IsSubst{N}{x}{M}{M'}$, and $\Subst{\vec{N}}{\vec{x}}{M}$ for the
-simultaneous substitution of $\vec{N}$ for $\vec{x}$ in $M$.
+$\IsSubst{N}{x}{M}{M'}$, and $\Subst{\vec{N}}{\vec{x}}{M}$ for the simultaneous
+substitution of $\vec{N}$ for $\vec{x}$ in $M$.
 
@@
   }{
-    \IsApart{\MV{m}}{\MV{n}} &
+    \IsApart{\MV{m}}{\MV{n}}
+&
     \HypJ{
@@
   }{
-    \IsEq{\MV{m}}{\MV{n}} &
-    \IsSubst{\vec{M}}{\vec{x}}{N}{N'} &
+    \IsEq{\MV{m}}{\MV{n}}
+&
+    \IsSubst{\vec{M}}{\vec{x}}{N}{N'}
+&
     \IsRenaming{\MkSet{\vec{u}\mapsto\vec{v}}}{N'}{N''}
-  }\\[6pt]
+  }
+\end{gather*}
+\begin{gather*}
   \infer{
@@
   }{
-    \IsApart{\vec{u}}{\FS{E}} &
-    \IsApart{\vec{x}}{\FV{E}} &
+    \IsApart{\vec{u}}{\FS{E}}
+&
+    \IsApart{\vec{x}}{\FV{E}}
+&
     \IsSubst{E}{\MV{m}}{M}{N}
@@
 
-Let $\FinSets$ be the free cocartesian category on a single object (that is,
-the category of finite cardinals and functions between them); then the comma
+Let $\FinSets$ be the free cocartesian category on a single object (that is, the
+category of finite cardinals and functions between them); then the comma
 construction $\Define{\Ctx}{\Comma{\FinSets}{\Sorts}}$ is the category of
@@
 Let $\Define{V_\tau(\Upsilon\parallel\Gamma)}{\MkSet{x\in\Dom\Gamma\mid
-\Gamma(x)=\tau}}$ be called the presheaf of variables; additionally, we have a
-presheaf of symbols
+    \Gamma(x)=\tau}}$ be called the presheaf of variables; additionally, we have
+a presheaf of symbols
 $\Define{S_\tau(\Upsilon\parallel\Gamma)}{\MkSet{u\in\Dom\Upsilon\mid
-\Upsilon(x)=\tau}}$. Lastly, we have the presheaf of operators with arity $a$,
+    \Upsilon(x)=\tau}}$. Lastly, we have the presheaf of operators with arity
+$a$,
 $\Define{\Operators_a(\Upsilon\parallel\Gamma)}{\Operators\Pair{\Upsilon}{a}}$
@@
 $\Of{P}{\widehat{\HypCat}^\Sorts}$ is a $\Sigma$-monoid in case it is equipped
-with the following natural transformations where $\nu$ embeds variables into
-$P$ and $\varsigma$ equips $P$ with an operation for simultaneous substitutions
-of variables. Furthermore, $\nu$ and $\varsigma$ induce maps ${\nu}_{\Gamma}$ and
-${{\varsigma}^{\tau}}_{\Upsilon\parallel\Gamma}$:
-\vspace{-0.5em}
+with the following natural transformations where $\nu$ embeds variables into $P$
+and $\varsigma$ equips $P$ with an operation for simultaneous substitutions of
+variables. Furthermore, $\nu$ and $\varsigma$ induce maps ${\nu}_{\Gamma}$ and
+${{\varsigma}^{\tau}}_{\Upsilon\parallel\Gamma}$: \vspace{-0.5em}
 \[
@@
     {\IMode{V}} \arrow[r, "\IMode\nu"]
-    \& {\IMode{P}}
-    \& {\IMode{P\odot P}}
-    \arrow[l, "\IMode\varsigma", swap]
+\&
+    {\IMode{P}}
+\&
+    {\IMode{P\odot P}}
+      \arrow[l, swap, "\IMode\varsigma"]
   \end{tikzcd}
-  \qquad\begin{tikzcd}[cramped, ampersand replacement = \&]
+\qquad
+  \begin{tikzcd}[cramped, ampersand replacement = \&]
     {\IMode{{V}^{[\Gamma]}}}
       \arrow[r, "\IMode{{\nu}_{\Gamma}}"]
-    \& {\IMode{{P}^{[\Gamma]}}}
+\&
+    {\IMode{{P}^{[\Gamma]}}}
   \end{tikzcd}
-  \qquad\begin{tikzcd}[cramped, sep = large, ampersand replacement = \&]
+\qquad
+  \begin{tikzcd}[cramped, sep = large, ampersand replacement = \&]
     {\IMode{{P_\tau^{\Yoneda{\Upsilon\parallel\Gamma}}\times S^{\{\Upsilon\}}\times P^{[\Gamma]}}}}
       \arrow[r, "\IMode{\varsigma^\tau_{\Upsilon\parallel\Gamma}}"]
-    \& {\IMode{P_\tau}}
+\&
+    {\IMode{P_\tau}}
   \end{tikzcd}
 \]
-
-\begin{remark}
-  A presheaf $P$ with a substitution monoidal structure $\Pair{\nu}{\varsigma}$
-  also forms a \emph{relative monad}\footnote{Relative monads are skew-monoids
-    in the skew-monoidal structure induced by a functor category
-    $\mathbf{C}^\mathbf{J}$~\cite{altenkirch-chapman-uustalu:2015}. However,
-    relative monads can also be understood without reference to monoids by
-    viewing them as a generalization of \emph{extension systems} or
-    \emph{Kleisli structures}. This latter presentation of monads, popular in
-    computer science and logic, does not involve iteration of the monad
-    functor.} on the functor $\Of{V}{\widehat{\HypCat}^\Sorts}$ of variables.
-  Consequently, monoid multiplication $\varsigma$ can be replaced with extension
-$\Of{\circledast}{\int_{\Tuple{\tau, (\Upsilon\parallel\Gamma), (\Upsilon'\parallel\Gamma')}}{{{P}_{\tau}(\Upsilon'\parallel\Gamma')}^{{V}_{\tau}(\Upsilon\parallel\Gamma)} \ra {{P}_{\tau}(\Upsilon'\parallel\Gamma')}^{{P}_{\tau}(\Upsilon\parallel\Gamma)}}}$.
-\end{remark}
 
@@
   \IMode{{\SigFun{\Sigma}(X)}_{\tau}}
-    &\triangleq
+&
+    \triangleq
     \OMode{
-      \coprod_{\vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}}
-      \prod_{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau_i}\in\vec{v}}
+      \coprod_{
+        \vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}
+      }
+      \prod_{
+        \MkValence
+          {\vec{\sigma}}
+          {\vec{\tau}}
+          {\tau_i}
+        \in\vec{v}
+      }
       X_{\tau_i}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}
@@
 
-Then, a $\Sigma$-model is a $\Sigma$-monoid $P$ which is equipped with an initial
-algebra $\Of{\alpha}{\SigFun{\Sigma}(P)\ra P}$, which shall interpret
+Then, a $\Sigma$-model is a $\Sigma$-monoid $P$ which is equipped with an
+initial algebra $\Of{\alpha}{\SigFun{\Sigma}(P)\ra P}$, which shall interpret
 applications of each operator.
@@
 
-The metavariable, symbol and variable contexts are interpreted in a model $P$ as an
-environment presheaf in the following way:
+The metavariable, symbol and variable contexts are interpreted in a model $P$ as
+an environment presheaf in the following way:
 \[
@@
   }{
-    \left(\prod_{(m:\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau})\in\Omega}P_{\tau}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}\right)
+    \left(
+      \prod_{
+        (m:%
+        \MkValence
+          {\vec{\sigma}}
+          {\vec{\tau}}
+          {\tau})
+        \in\Omega
+      }{{P_{\tau}}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}}
+    \right)
     \times
@@
 to $P$:
-\[%
-  \Of{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash%
-      M:\tau}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P\ra P_\tau}%
+\[
+  \Of{
+    \SemBrackets{
+      \Omega\triangleright\Upsilon\parallel\Gamma\vdash M:\tau
+    }_{P}
+  }{
+    \SemBrackets{
+      \Omega\triangleright\Upsilon\parallel\Gamma
+    }_{P}\ra {P}_{\tau}
+  }
 \]
@@
 Variables are interpreted by the map
-$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x:\tau}_P$ which
+${\SemBrackets{
+  \Omega\triangleright\Upsilon\parallel\Gamma\vdash x:\tau
+}}_{P}$ which
 projects them from the environment and embeds them into the model. Metavariables
 are resolved by the map
-$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash\MApp{\MV{m}}{\vec{u}}{\vec{M}}
-  : \tau}_P$ (where
+${\SemBrackets{
+  \Omega\triangleright\Upsilon\parallel\Gamma\vdash\MApp{\MV{m}}{\vec{u}}{\vec{M}}
+  : \tau
+}}_{P}$ (where
 $\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$) which
@@
     {\IMode{P_\tau}}
-    \&[8.0em] {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
-      \arrow[rr, dashed, "\OMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x:\tau}_P}"]
+\&[8.0em]
+    {\IMode{
+      {\SemBrackets{
+        \Omega\triangleright\Upsilon\parallel\Gamma
+      }}_{P}
+    }}
+      \arrow[
+        , rr
+        , dashed
+        , "
+          \OMode{
+            {\SemBrackets{
+              \Omega\triangleright\Upsilon\parallel\Gamma
+              \vdash
+              x:\tau
+            }}_{P}
+          }
+          "
+      ]
       \ar{drr}[
@@
         , swap
-      ]{{\nu}_{\Gamma} {\pi}_{3}}
+      ]{
+        {\nu}_{\Gamma} {\pi}_{3}
+      }
       \ar{dl}[
@@
         , xshift = -2.75em
-      ]{\IMode{\Tuple{\pi_\MV{m}\,\pi_1, \phi\,\pi_2, \psi}}}
-      \arrow[l, dashed, swap, "\OMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash\MApp{\MV{m}}{\vec{u}}{\vec{M}}: \tau}_P}"]
+      ]{
+        \IMode{
+          \Tuple{
+            \pi_\MV{m}\,
+            \pi_1,
+            \phi\,
+            \pi_2,
+            \psi
+          }
+        }
+      }
+      \arrow[
+        , l
+        , dashed
+        , swap
+        , "
+          \OMode{
+            {\SemBrackets{
+              \Omega\triangleright\Upsilon\parallel\Gamma
+              \vdash
+              \MApp{\MV{m}}{\vec{u}}{\vec{M}}: \tau
+            }}_{P}
+          }"
+      ]
 \&[-2.0em]
-    \&[+3.0em]{\IMode{P_\tau}}
-    \\
-    {\IMode{P_{\tau}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}\times S^{\{\vec{\sigma}\}} \times P^{[\vec{\tau}]}}}
-      \arrow[u, "\IMode{\varsigma^\tau_{\vec{\sigma}\parallel\vec{\tau}}}"]
-    \&{}
-    \&{}
-    \& {\IMode{P^{[\Gamma]}}}
-      \arrow[u, swap, "{\pi}_{x}"]
+    \null
+\&[+3.0em]
+    {\IMode{P_\tau}}
+\\
+    {\IMode{
+      {P}_{\tau}^{
+        \Yoneda{\vec{\sigma}\parallel\vec{\tau}}
+      }
+      \times S^{\{\vec{\sigma}\}}
+      \times P^{[\vec{\tau}]}
+    }}
+      \arrow[
+        , u
+        , "
+          \IMode{
+            \varsigma^\tau_{\vec{\sigma}\parallel\vec{\tau}}
+          }
+          "
+        ]
+\&
+    \null
+\&
+    \null
+\&
+    {\IMode{P^{[\Gamma]}}}
+      \arrow[
+        , u
+        , swap
+        , "{\pi}_{x}"
+      ]
 \\
     {\IMode{P^{[\vec{\tau}]}}}
-    \& {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
-      \arrow[l, dashed, swap, "\Define{\psi\,}{\,\Tuple{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M:\tau}_P}_{(M,\tau)\in (\vec{M},\vec{\tau})}}"]
-    \& {\IMode{S^{\{\Upsilon\}}}}
-      \arrow[r, dashed, "\Define{\phi\,}{\,\Tuple{\pi_u}_{u\in\vec{u}}}"]
-    \& {\IMode{S^{\{\vec{\sigma}\}}}}
+\&
+    {\IMode{
+      {\SemBrackets{
+        \Omega\triangleright\Upsilon\parallel\Gamma
+      }}_{P}
+    }}
+      \arrow[
+        , dashed
+        , l
+        , swap
+        , "
+          \Define{
+            \psi\,
+          }{
+            \,\Tuple{
+              {\SemBrackets{
+                \Omega\triangleright\Upsilon\parallel\Gamma\vdash M:\tau
+              }}_{P}
+            }_{(M,\tau)\in (\vec{M},\vec{\tau})}
+          }
+          "]
+\&
+    {\IMode{S^{\{\Upsilon\}}}}
+      \arrow[
+        , dashed
+        , r
+        , "\Define{\phi\,}{\,\Tuple{\pi_u}_{u\in\vec{u}}}"
+      ]
+\&
+    {\IMode{S^{\{\vec{\sigma}\}}}}
   \end{tikzcd}
 \]
-
 Interpretation of operator applications is the most complicated. Recall that,
@@
     |[alias = Src]|
-    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
-      \arrow[r, swap, "\IMode{\pi_2}"]
+    {\IMode{
+      {\SemBrackets{
+        \Omega\triangleright\Upsilon\parallel\Gamma
+      }}_{P}
+    }}
+      \arrow[
+        , r
+        , swap
+        , "\IMode{\pi_2}"
+      ]
 \&
     {\IMode{S^{\{\Upsilon\}}}}
-      \arrow[r, swap, "\IMode{\FLift{\pi_{(-)}}{}}"]
+      \arrow[
+        , r
+        , swap
+        , "\IMode{\FLift{\pi_{(-)}}{}}"
+      ]
 \&[1.5em]
@@
     {\IMode{\Operators_a}}
-    \arrow[from = Src, to = Tgt, bend left, dashed, "\OMode{\SemBrackets{\vartheta}_{P}}"]
+      \arrow[
+        , from = Src
+        , to = Tgt
+        , bend left
+        , dashed
+        , "\OMode{\SemBrackets{\vartheta}_{P}}"
+      ]
   \end{tikzcd}
@@
 follows, by postcomposing it with a morphism $\beta$ from the environment into
-the signature endofunctor, which interprets the syntax of operator
-applications:
+the signature endofunctor, which interprets the syntax of operator applications:
 \[
   \begin{tikzcd}[cramped, ampersand replacement = \&]
-     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}} \arrow[r, "\IMode{\beta}"]
-     \& {\IMode{\SigFun{\Sigma}(P)_\tau}} \arrow[r, "\IMode{\alpha_\tau}"]
-     \& {\IMode{P_\tau}}
+    {\IMode{
+      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}}_{P}
+    }}
+      \arrow[r, "\IMode{\beta}"]
+\&
+    {\IMode{
+      {\SigFun{\Sigma}(P)}_{\tau}
+    }}
+      \arrow[r, "\IMode{\alpha_\tau}"]
+\&
+    {\IMode{P_\tau}}
   \end{tikzcd}
@@
   \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
-       \arrow[r, dashed, "\Define{\beta\,}{\,\Tuple{\SemBrackets{\vartheta}_P, \lambda\gamma}}"]
-    \&[4.0em] {\IMode{%
-       \coprod_{\vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}}%
-       \prod_{\MkValence{\vec{\sigma}_i}{\vec{\tau}_i}{\tau_i}\in\vec{v}}%
-       P_{\tau_i}^{\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}}%
+    {\IMode{
+      {\SemBrackets{
+        \Omega\triangleright\Upsilon\parallel\Gamma
+      }}_{P}
+    }}
+      \arrow[
+        , r
+        , dashed
+        , "
+          \Define{
+            \beta\,
+          }{
+            \,\Tuple{
+              \SemBrackets{\vartheta}_P, \lambda\gamma
+            }
+          }
+          "
+      ]
+\&[4.0em]
+    {\IMode{
+      \coprod_{
+        \vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}
+      }
+      \prod_{
+        \MkValence{\vec{\sigma}_i}{\vec{\tau}_i}{\tau_i}\in\vec{v}
+      }
+      P_{\tau_i}^{\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}}
     }}
@@
   \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_{P}}}
-      \arrow[d, dashed, swap, "\OMode{\lambda{\gamma}_{i}}"]
-    \&
-    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_{P}
-       \times\Yoneda{\vec{\sigma}_{i}\parallel\vec{\tau}_{i}}}
-    }
-      \arrow[r, "\IMode{\Tuple{{\pi}_{[1,1]},{\phi}_{i},{\psi}_{i}}}"]
-      \arrow[d, dashed, swap, "\OMode{{\gamma}_{i}}"]
+    {\IMode{
+      {\SemBrackets{
+        \Omega\triangleright\Upsilon\parallel\Gamma
+      }}_{P}
+    }}
+      \arrow[
+        , dashed
+        , d
+        , swap
+        , "\OMode{\lambda{\gamma}_{i}}"
+      ]
+\&
+    {\IMode{
+      {\SemBrackets{
+        \Omega\triangleright\Upsilon\parallel\Gamma
+      }}_{P}
+      \times
+      \Yoneda{
+        \vec{\sigma}_{i}\parallel\vec{\tau}_{i}}
+      }
+    }
+      \arrow[
+        , r
+        , "
+          \IMode{
+            \Tuple{
+              {\pi}_{[1,1]},
+              {\phi}_{i},
+              {\psi}_{i}
+            }
+          }
+          "
+        ]
+      \arrow[
+        , dashed
+        , d
+        , swap
+        , "\OMode{{\gamma}_{i}}"
+      ]
 \&[3.5em]
-    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon,\vec{\sigma}_i\parallel\Gamma,\vec{\tau}_{i}}_{P}}}
+    {\IMode{
+      \SemBrackets{
+        \Omega\triangleright\Upsilon,
+        \vec{\sigma}_i\parallel\Gamma,
+        \vec{\tau}_{i}
+      }_{P}
+    }}
       \ar{dl}[
@@
     {\IMode{{P}_{{\tau}_{i}}^{\Yoneda{\vec{\sigma}_{i}\parallel\vec{\tau}_{i}}}}}
-    \& {\IMode{P_{{\tau}_{i}}}}
+\&
+    {\IMode{P_{{\tau}_{i}}}}
 \&
@@
   \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
-       \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}
+    {\IMode{
+      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}}_{P}
+      \times
+      \Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}
     }}
-      \arrow[r, dashed, "\Define{\phi_i\,}{\,\Pair{\Tuple{\pi_u\circ\pi_1}_{u\in\Dom{\Upsilon}}}{\,\Tuple{\pi_u\circ\pi_{[2,1]}}_{u\in\Dom{\vec{\sigma}_i}}}}"]
-    \&[12.0em] {\IMode{S^{\{\Upsilon,\vec{\sigma}_i\}}}}\\[-1.0em]
-    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
-       \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}
+      \arrow[
+        , r
+        , dashed
+        , "
+          \Define
+            {
+              \phi_i\,
+            }
+            {
+              \,\Pair{
+                \Tuple{\pi_u\circ\pi_1}_{u\in\Dom{\Upsilon}}
+              }{
+                \,\Tuple{\pi_u\circ\pi_{[2,1]}}_{u\in\Dom{\vec{\sigma}_i}}
+              }
+            }
+          "
+      ]
+\&[12.0em]
+    {\IMode{S^{\{\Upsilon,\vec{\sigma}_i\}}}}
+\\[-1.0em]
+    {\IMode{
+      {\SemBrackets{
+        \Omega\triangleright\Upsilon\parallel\Gamma
+      }}_{P}
+      \times
+      \Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}
     }}
-      \arrow[r, dashed, "\Define{\psi_i\,}{\,\Pair{\Tuple{\pi_x\circ\pi_1}_{x\in\Dom{\Gamma}}}{\,\Tuple{\pi_x\circ\pi_{[2,1]}}_{x\in\Dom{\vec{\tau}_i}}}}"]
-    \&[12.0em] {\IMode{V^{[\Gamma,\vec{\tau}_i]}}}
+      \arrow[
+        , r
+        , dashed
+        , "
+          \Define{
+            \psi_i\,
+          }{
+            \,\Pair{
+              \Tuple{\pi_x\circ\pi_1}_{x\in\Dom{\Gamma}}
+            }{
+              \,\Tuple{\pi_x\circ\pi_{[2,1]}}_{x\in\Dom{\vec{\tau}_i}}
+            }
+          }
+          "
+      ]
+\&[12.0em]
+    {\IMode{V^{[\Gamma,\vec{\tau}_i]}}}
   \end{tikzcd}
@@
 The representation of telescopes and sequents in a logical framework is
-notoriously difficult; whilst it is possible to use higher-order abstract
-syntax or abts to encode the binding-structure of telescopes
-and sequents, the encoding is sufficiently laborious and obscure that it is not
-used in practice.
+notoriously difficult; whilst it is possible to use higher-order abstract syntax
+or abts to encode the binding-structure of telescopes and sequents, the encoding
+is sufficiently laborious and obscure that it is not used in practice.
 
 Crary has demonstrated a first-order encoding of contexts in the logical
-framework in bijection with actual LF-contexts~\cite{crary:2009}, which has
-been successfully used in large-scale mechanization efforts, including that of
+framework in bijection with actual LF-contexts~\cite{crary:2009}, which has been
+successfully used in large-scale mechanization efforts, including that of
 Standard ML~\cite{lee-crary-harper:2007} and the Edinburgh Logical Framework
@@
 
-%%% TODO: talk about context encodings in Abella
+\todo[inline]{talk about context encodings in Abella}
 
@@
 \emph{refinements} perspective, where a conservative approximation of the
-grammar is first given using the abt logical framework, and then the
-correctness of a code is expressed separately in a judgment that refines the
-existing specification.
-
-Because we have not committed to using the built-in binding machinery to
-express the well-scopedness of telescopes and sequents, we are free to use
+grammar is first given using the abt logical framework, and then the correctness
+of a code is expressed separately in a judgment that refines the existing
+specification.
+
+Because we have not committed to using the built-in binding machinery to express
+the well-scopedness of telescopes and sequents, we are free to use
 \emph{symbols} in order to model the variables in the context. This is in fact
 quite sensible if we are actually trying to faithfully represent the syntax of
-telescopes and sequents, rather than replace them with their counterparts on
-the meta-level.
-
-This insight leads the way to a simple abt signature for the theory of telescopes
-and sequents.
+telescopes and sequents, rather than replace them with their counterparts on the
+meta-level.
+
+This insight leads the way to a simple abt signature for the theory of
+telescopes and sequents.
 %
@@
 \newcommand\OpSequent{\mathtt{sequent}}
-%
+% %
 \[
@@
       \MkArity{}{\SortType}
-    }\\
+    }
+\\
     \IsOperator{\Upsilon}{\bot}{
       \MkArity{}{\SortType}
-    }\\
+    }
+\\
     \IsOperator{\Upsilon}{\OpBool}{
       \MkArity{}{\SortType}
-    }\\
+    }
+\\
     \IsOperator{\Upsilon}{\OpSo}{
@@
       }{\SortType}
-    }\\
+    }
+\\
     \IsOperator{\Upsilon}{\OpPi}{
@@
       }{\SortType}
-    }\\
+    }
+\\
     \IsOperator{\Upsilon}{\OpSg}{
@@
       }{\SortType}
-    }\\
+    }
+\\
   \end{array}
@@
 \begin{align*}
-  \IMode{\diamond} &\triangleq \OMode{\OpNil}\\
-  \IMode{H, u:P} &\triangleq \OMode{
+  \IMode{\diamond}
+&
+  \triangleq \OMode{\OpNil}
+\\
+  \IMode{H, u:P}
+&
+  \triangleq \OMode{
     \App{\OpSnoc{u}}{H, P}
-  }\\
-  \IMode{H\gg A} &\triangleq \OMode{
+  }
+\\
+  \IMode{H\gg A}
+&
+  \triangleq \OMode{
     \App{\OpSequent}{
@@
     }
-  }\\
-  \IMode{`u} &\triangleq \OMode{\OpVar{u}}
+  }
+\\
+  \IMode{`u}
+&
+  \triangleq \OMode{\OpVar{u}}
 \end{align*}
@@
   }{\cdot}{
-    \diamond,u:\OpBool,v:\App{\OpSo}{`u}%
-    \gg%
+    \diamond,u:\OpBool,v:\App{\OpSo}{`u}
+    \gg
     \App{\OpSo}{`u}
@@
       \MkBTm{u,v}{}{
-        \diamond,u:\OpBool,v:\App{\OpSo}{`u}%
-        \gg%
+        \diamond,u:\OpBool,v:\App{\OpSo}{`u}
+        \gg
         \App{\OpSo}{`u}
@@
 the abt logical framework, we can proceed to define proper wellformedness via
-\emph{inductive refinement}~\cite{harper:2016}. The basic idea is to introduce a new form of
-(meta)-judgment $\IsWf{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$ which expresses the
-extrinsic wellformedness properties we wish to verify, presupposing
-$\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$. Additionally, we introduce an
-analogous judgment on bound terms, $\IsWf{\Omega}{\Upsilon}{\Gamma}{E}{v}$
-presupposing $\IsAbt{\Omega}{\Upsilon}{\Gamma}{E}{v}$, defined uniformly as
-follows:
+\emph{inductive refinement}~\cite{harper:2016}. The basic idea is to introduce a
+new form of (meta)-judgment $\IsWf{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$ which
+expresses the extrinsic wellformedness properties we wish to verify,
+presupposing $\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$. Additionally, we
+introduce an analogous judgment on bound terms,
+$\IsWf{\Omega}{\Upsilon}{\Gamma}{E}{v}$ presupposing
+$\IsAbt{\Omega}{\Upsilon}{\Gamma}{E}{v}$, defined uniformly as follows:
 \[
@@
     \begin{array}{l}
-      \Lookup{\Omega}{\MV{m}}{\MkValence{\vec{\sigma}}{\tau_0,\dots,\tau_n}{\tau}}\\
+      \Lookup{\Omega}{\MV{m}}{\MkValence{\vec{\sigma}}{\tau_0,\dots,\tau_n}{\tau}}
+\\
       \HypJ{
@@
 
-The remainder of the definition of refinement proceeds by induction on sorts
-and operators. For the sake of this example, we will just stipulate that
-anything of sort $\SortExp$ or $\SortType$ is grammatical if its subterms are
-grammatical:
+The remainder of the definition of refinement proceeds by induction on sorts and
+operators. For the sake of this example, we will just stipulate that anything of
+sort $\SortExp$ or $\SortType$ is grammatical if its subterms are grammatical:
 
@@
         }{\tau}
-      }\\
+      }
+\\
       \HypJ{
@@
     \IsWf{\Omega}{\Upsilon,\vec{u}:\vec{\sigma}}{\Gamma}{J}{\SortJdg}
-  }\\[6pt]
+  }
+\\[6pt]
   \infer{
@@
   }{
-    \IsWf{\Omega}{\Upsilon}{\Gamma}{H}{\SortTele} &
+    \IsWf{\Omega}{\Upsilon}{\Gamma}{H}{\SortTele}
+&
     \IsWf{\Omega}{\Upsilon}{\Gamma}{A}{\SortType}
@@
     \begin{array}{l}
-      \IsWf{\Omega}{\Upsilon\setminus\MkSet{u}}{\Gamma}{H}{\SortTele}\\
+      \IsWf{\Omega}{\Upsilon\setminus\MkSet{u}}{\Gamma}{H}{\SortTele}
+\\
       \IsWf{\Omega}{\Upsilon\setminus\MkSet{u}}{\Gamma}{A}{\SortType}

### 2015-11-20 paper.tex
Switch from \Omega to \Theta for metavariable contexts; fix up modes

@@
 
-\newcommand\Iso[2]{#1\mathrel{{{\cong}}}#2}
+\newcommand\Iso[2]{\IMode{#1}\mathrel{{{\cong}}}\IMode{#2}}
 \newcommand\Cats{\mathbf{Cat}}
@@
   \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\Pair{\Upsilon}{a}}
-      \arrow[r, swap, "\Pair{\rho}{f}"]
-\&
-    {\Pair{\Upsilon'}{a'}}
+    \IMode{\Pair{\Upsilon}{a}}
+      \arrow[r, swap, "\IMode{\Pair{\rho}{f}}"]
+\&
+    \IMode{\Pair{\Upsilon'}{a'}}
   \end{tikzcd}
@@
 \&
-    \mathbb{1}
-      \arrow[dl, swap, "\vartheta"]
-      \arrow[dr, "\vartheta'"]
+    \IMode{\mathbb{1}}
+      \arrow[dl, swap, "\IMode{\vartheta}"]
+      \arrow[dr, "\IMode{\vartheta'}"]
 \&
@@
 \\
-    {\Operators\Pair{\Upsilon}{a}}
-      \arrow[rr, swap, "\Operators\Pair{\rho}{f}"]
+    \IMode{\Operators\Pair{\Upsilon}{a}}
+      \arrow[rr, swap, "\IMode{\Operators\Pair{\rho}{f}}"]
 \&
@@
 \&
-    {\Operators\Pair{\Upsilon'}{a'}}
+    \IMode{\Operators\Pair{\Upsilon'}{a'}}
   \end{tikzcd}
@@
 In general, we will have three kinds of context: metavariable contexts, variable
-contexts, and symbol (parameter) contexts. A metavariable context $\Omega$
+contexts, and symbol (parameter) contexts. A metavariable context $\Theta$
 consists of bindings of valences to metavariables; a variable context $\Gamma$
@@
   \infer{
-    \IsMetaCtx{\Omega,\MV{m}:v}
-  }{
-    \IsMetaCtx{\Omega}
+    \IsMetaCtx{\Theta,\MV{m}:v}
+  }{
+    \IsMetaCtx{\Theta}
 &
@@
 &
-    \NotIn{\MV{m}}{\Dom\Omega}
+    \NotIn{\MV{m}}{\Dom\Theta}
   }
@@
 
-Let the judgment $\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$ presuppose
-$\IsMetaCtx{\Omega}$, $\IsSymCtx\Upsilon$, $\IsVarCtx\Gamma$ and
+Let the judgment $\IsAbt{\Theta}{\Upsilon}{\Gamma}{M}{\tau}$ presuppose
+$\IsMetaCtx{\Theta}$, $\IsSymCtx\Upsilon$, $\IsVarCtx\Gamma$ and
 $\IsSort{\tau}$, meaning that $M$ is an abstract binding tree of sort $s$, with
-metavariables in $\Omega$, parameters in $\Upsilon$, and variables in $\Gamma$.
-Let the judgment $\IsBTm{\Omega}{\Upsilon}{\Gamma}{E}{v}$ presuppose
+metavariables in $\Theta$, parameters in $\Upsilon$, and variables in $\Gamma$.
+Let the judgment $\IsBTm{\Theta}{\Upsilon}{\Gamma}{E}{v}$ presuppose
 $\IsValence{v}$. Then, the syntax of abstract binding trees is inductively
@@
   \infer[\textit{var}]{
-    \IsAbt{\Omega}{\Upsilon}{\Gamma}{x}{\tau}
+    \IsAbt{\Theta}{\Upsilon}{\Gamma}{x}{\tau}
   }{
@@
   \infer[\textit{mvar}]{
-    \IsAbt{\Omega}{\Upsilon}{\Gamma}{
+    \IsAbt{\Theta}{\Upsilon}{\Gamma}{
       \MApp{\MV{m}}{u_0,\dots,u_m}{M_0,\dots,M_n}
@@
     \begin{array}{l}
-      \Lookup{\Omega}{\MV{m}}{
+      \Lookup{\Theta}{\MV{m}}{
         \MkValence{
@@
       \HypJ{
-        \IsAbt{\Omega}{\Upsilon}{\Gamma}{M_i}{\tau_i}
+        \IsAbt{\Theta}{\Upsilon}{\Gamma}{M_i}{\tau_i}
       }{\Leq{i}{n}}
@@
   \infer[\textit{app}]{
-    \IsAbt{\Omega}{\Upsilon}{\Gamma}{
+    \IsAbt{\Theta}{\Upsilon}{\Gamma}{
       \App{\vartheta}{E_0,\dots,E_n}
@@
       \HypJ{
-        \IsBTm{\Omega}{\Upsilon}{\Gamma}{E_i}{v_i}
+        \IsBTm{\Theta}{\Upsilon}{\Gamma}{E_i}{v_i}
       }{\Leq{i}{n}}
@@
   \infer[\textit{abs}]{
-    \IsBTm{\Omega}{\Upsilon}{\Gamma}{
+    \IsBTm{\Theta}{\Upsilon}{\Gamma}{
       \MkBTm{\vec{u}}{\vec{x}}{M}
@@
     \IsAbt{
-      \Omega
+      \Theta
     }{
@@
 $\IsRenaming{\varrho}{E}{F}$, presupposing $\Of{\varrho}{\Upsilon\ra\Upsilon'}$,
-and $\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$ and
-$\IsAbt{\Omega}{\Upsilon}{\Gamma}{E}{v}$ respectively:
+and $\IsAbt{\Theta}{\Upsilon}{\Gamma}{M}{\tau}$ and
+$\IsAbt{\Theta}{\Upsilon}{\Gamma}{E}{v}$ respectively:
 
@@
     \SemBrackets{
-      \Omega\triangleright\Upsilon\parallel\Gamma
+      \Theta\triangleright\Upsilon\parallel\Gamma
     }_P
@@
           {\tau})
-        \in\Omega
+        \in\Theta
       }{{P_{\tau}}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}}
@@
     \SemBrackets{
-      \Omega\triangleright\Upsilon\parallel\Gamma\vdash M:\tau
+      \Theta\triangleright\Upsilon\parallel\Gamma\vdash M:\tau
     }_{P}
@@
     \SemBrackets{
-      \Omega\triangleright\Upsilon\parallel\Gamma
+      \Theta\triangleright\Upsilon\parallel\Gamma
     }_{P}\ra {P}_{\tau}
@@
 ${\SemBrackets{
-  \Omega\triangleright\Upsilon\parallel\Gamma\vdash x:\tau
+  \Theta\triangleright\Upsilon\parallel\Gamma\vdash x:\tau
 }}_{P}$ which
@@
 ${\SemBrackets{
-  \Omega\triangleright\Upsilon\parallel\Gamma\vdash\MApp{\MV{m}}{\vec{u}}{\vec{M}}
+  \Theta\triangleright\Upsilon\parallel\Gamma\vdash\MApp{\MV{m}}{\vec{u}}{\vec{M}}
   : \tau
 }}_{P}$ (where
-$\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$) which
+$\Lookup{\Theta}{\MV{m}}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$) which
 projects their interpretation from the environment and instantiates it via
@@
       {\SemBrackets{
-        \Omega\triangleright\Upsilon\parallel\Gamma
+        \Theta\triangleright\Upsilon\parallel\Gamma
       }}_{P}
@@
             {\SemBrackets{
-              \Omega\triangleright\Upsilon\parallel\Gamma
+              \Theta\triangleright\Upsilon\parallel\Gamma
               \vdash
@@
               {\SemBrackets{
-                \Omega\triangleright\Upsilon\parallel\Gamma\vdash M:\tau
+                \Theta\triangleright\Upsilon\parallel\Gamma\vdash M:\tau
               }}_{P}
@@
     {\IMode{
-      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}}_{P}
+      {\SemBrackets{\Theta\triangleright\Upsilon\parallel\Gamma}}_{P}
     }}
@@
       \SemBrackets{
-        \Omega\triangleright\Upsilon,
+        \Theta\triangleright\Upsilon,
         \vec{\sigma}_i\parallel\Gamma,
@@
         , yshift = -0.2em
-      ]{\IMode{\SemBrackets{\Omega\triangleright\Upsilon,\vec{\sigma}_{i}\parallel\Gamma,\vec{\tau}_{i}\vdash M_i:\tau_i}_P}}
+      ]{\IMode{\SemBrackets{\Theta\triangleright\Upsilon,\vec{\sigma}_{i}\parallel\Gamma,\vec{\tau}_{i}\vdash M_i:\tau_i}_P}}
 \\[2.0em]
@@
     {\IMode{
-      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}}_{P}
+      {\SemBrackets{\Theta\triangleright\Upsilon\parallel\Gamma}}_{P}
       \times
@@
 \emph{inductive refinement}~\cite{harper:2016}. The basic idea is to introduce a
-new form of (meta)-judgment $\IsWf{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$ which
+new form of (meta)-judgment $\IsWf{\Theta}{\Upsilon}{\Gamma}{M}{\tau}$ which
 expresses the extrinsic wellformedness properties we wish to verify,
-presupposing $\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{\tau}$. Additionally, we
+presupposing $\IsAbt{\Theta}{\Upsilon}{\Gamma}{M}{\tau}$. Additionally, we
 introduce an analogous judgment on bound terms,
-$\IsWf{\Omega}{\Upsilon}{\Gamma}{E}{v}$ presupposing
-$\IsAbt{\Omega}{\Upsilon}{\Gamma}{E}{v}$, defined uniformly as follows:
-\[
-  \infer{
-    \IsWf{\Omega}{\Upsilon}{\Gamma}{
+$\IsWf{\Theta}{\Upsilon}{\Gamma}{E}{v}$ presupposing
+$\IsAbt{\Theta}{\Upsilon}{\Gamma}{E}{v}$, defined uniformly as follows:
+\[
+  \infer{
+    \IsWf{\Theta}{\Upsilon}{\Gamma}{
       \MkBTm{\vec{u}}{\vec{x}}{M}
@@
   }{
-    \IsWf{\Omega}{\Upsilon,\vec{u}:\vec{\sigma}}{\Gamma,\vec{x}:\vec{\tau}}{M}{\tau}
+    \IsWf{\Theta}{\Upsilon,\vec{u}:\vec{\sigma}}{\Gamma,\vec{x}:\vec{\tau}}{M}{\tau}
   }
@@
   \infer{
-    \IsWf{\Omega}{\Upsilon}{\Gamma}{x}{\tau}
+    \IsWf{\Theta}{\Upsilon}{\Gamma}{x}{\tau}
   }{
@@
   \infer{
-    \IsWf{\Omega}{\Upsilon}{\Gamma}{
+    \IsWf{\Theta}{\Upsilon}{\Gamma}{
       \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
@@
     \begin{array}{l}
-      \Lookup{\Omega}{\MV{m}}{\MkValence{\vec{\sigma}}{\tau_0,\dots,\tau_n}{\tau}}
+      \Lookup{\Theta}{\MV{m}}{\MkValence{\vec{\sigma}}{\tau_0,\dots,\tau_n}{\tau}}
 \\
       \HypJ{
-        \IsWf{\Omega}{\Upsilon}{\Gamma}{M_i}{\tau_i}
+        \IsWf{\Theta}{\Upsilon}{\Gamma}{M_i}{\tau_i}
       }{
@@
   Note that the refinement for variables $x$ is not trivial, since it is only
-  defined in case the presupposition $\IsAbt{\Omega}{\Upsilon}{\Gamma}{x}{\tau}$
+  defined in case the presupposition $\IsAbt{\Theta}{\Upsilon}{\Gamma}{x}{\tau}$
   is satisfied.
@@
   ]{
-    \IsWf{\Omega}{\Upsilon}{\Gamma}{
+    \IsWf{\Theta}{\Upsilon}{\Gamma}{
       \App{\vartheta}{E_0,\dots,E_n}
@@
       \HypJ{
-        \IsWf{\Omega}{\Upsilon}{\Gamma}{E_i}{v_i}
+        \IsWf{\Theta}{\Upsilon}{\Gamma}{E_i}{v_i}
       }{
@@
   \infer{
-    \IsWf{\Omega}{\Upsilon}{\Gamma}{
+    \IsWf{\Theta}{\Upsilon}{\Gamma}{
       \App{\OpNabla{\vec{\sigma}}}{
@@
   }{
-    \IsWf{\Omega}{\Upsilon,\vec{u}:\vec{\sigma}}{\Gamma}{J}{\SortJdg}
+    \IsWf{\Theta}{\Upsilon,\vec{u}:\vec{\sigma}}{\Gamma}{J}{\SortJdg}
   }
@@
   \infer{
-    \IsWf{\Omega}{\Upsilon}{\Gamma}{
+    \IsWf{\Theta}{\Upsilon}{\Gamma}{
       H\gg A
@@
   }{
-    \IsWf{\Omega}{\Upsilon}{\Gamma}{H}{\SortTele}
-&
-    \IsWf{\Omega}{\Upsilon}{\Gamma}{A}{\SortType}
+    \IsWf{\Theta}{\Upsilon}{\Gamma}{H}{\SortTele}
+&
+    \IsWf{\Theta}{\Upsilon}{\Gamma}{A}{\SortType}
   }
@@
   \infer{
-    \IsWf{\Omega}{\Upsilon}{\Gamma}{\diamond}{\SortTele}
+    \IsWf{\Theta}{\Upsilon}{\Gamma}{\diamond}{\SortTele}
   }{
@@
   \infer{
-    \IsWf{\Omega}{\Upsilon}{\Gamma}{H,u:A}{\SortTele}
+    \IsWf{\Theta}{\Upsilon}{\Gamma}{H,u:A}{\SortTele}
   }{
     \begin{array}{l}
-      \IsWf{\Omega}{\Upsilon\setminus\MkSet{u}}{\Gamma}{H}{\SortTele}
-\\
-      \IsWf{\Omega}{\Upsilon\setminus\MkSet{u}}{\Gamma}{A}{\SortType}
+      \IsWf{\Theta}{\Upsilon\setminus\MkSet{u}}{\Gamma}{H}{\SortTele}
+\\
+      \IsWf{\Theta}{\Upsilon\setminus\MkSet{u}}{\Gamma}{A}{\SortType}
     \end{array}

### 2015-11-23 paper.tex
Layout and notation improvements

@@
 \newcommand\Rename[2]{%
-  #2\upharpoonright#1%
+  #2\{#1\}%
 }
@@
 category of contexts of symbols, whose objects are finite sets of symbols $U$
-and sort-assignments $\Of{\mathfrak{s}}{U\ra\Sorts}$, and whose morphisms are
+and sort-assignments
+\begin{tikzcd}[cramped]
+  \IMode{U}\arrow[r, "\IMode{\mathfrak{s}}"] &\IMode{\Sorts}
+\end{tikzcd}, and whose morphisms are
 sort-preserving renamings; we will write $\Upsilon$ for a symbol context
@@
 \]
+
 We will write $\IsOperator{\Upsilon}{\vartheta}{a}$ in case
@@
 \end{alignat*}
+
 So far, we have made no use of symbols and parameters; however, consider the
@@
 Note that the functoriality of $\Operators$ guarantees for any renaming
-$\Of{\varrho}{\Upsilon\ra\Upsilon'}$ and $\Of{f}{a\ra a'}$, a lifted map
-$\Of{\FLift{\varrho}{f}}{\Operators\Pair{\Upsilon}{a}\ra\Operators\Pair{\Upsilon'}{a'}}$
+$\Of{\varrho}{\Upsilon\inj\Upsilon'}$ and $\Of{f}{a\ra a'}$, a lifted map
+$\Of{\FLift{\varrho}{f}}{\Operators\Pair{\Upsilon}{a}\inj\Operators\Pair{\Upsilon'}{a'}}$
 such that $\IsOperator{\Upsilon'}{\FLift{\varrho}{f}(\vartheta)}{a'}$ when
@@
 structure, via a pair of judgments $\IsRenaming{\varrho}{M}{N}$ and
-$\IsRenaming{\varrho}{E}{F}$, presupposing $\Of{\varrho}{\Upsilon\ra\Upsilon'}$,
+$\IsRenaming{\varrho}{E}{F}$, presupposing $\Of{\varrho}{\Upsilon\inj\Upsilon'}$,
 and $\IsAbt{\Theta}{\Upsilon}{\Gamma}{M}{\tau}$ and
@@
 $P^{[\Gamma]}$ to mean $\prod_{x\in\Dom\Gamma}P_{\Gamma(x)}$; likewise,
-$S^{\{\Upsilon\}}$ shall mean $\prod_{u\in\Dom\Upsilon}S_{\Upsilon(u)}$.
-
-For a presheaf $\Of{A}{\widehat{\HypCat}}$ and a sort-indexed family of
-presheaves $\Of{P}{\widehat{\HypCat}^\Sorts}$, we have an operation $A\bullet
-P$, defined as a coend in the following way:
+$S^{\{\Upsilon\}}$ shall mean $\prod_{u\in\Dom\Upsilon}S_{\Upsilon(u)}$. For a
+presheaf $\Of{A}{\widehat{\HypCat}}$ and a sort-indexed family of presheaves
+$\Of{P}{\widehat{\HypCat}^\Sorts}$, we have an operation $A\bullet P$, defined
+as a coend in the following way:
 \[
@@
 variables. Furthermore, $\nu$ and $\varsigma$ induce maps ${\nu}_{\Gamma}$ and
-${{\varsigma}^{\tau}}_{\Upsilon\parallel\Gamma}$: \vspace{-0.5em}
-\[
+${\varsigma}^{\tau}_{\Upsilon\parallel\Gamma}$:
+\begin{gather*}
   \begin{tikzcd}[cramped, ampersand replacement = \&]
@@
       \arrow[l, swap, "\IMode\varsigma"]
-  \end{tikzcd}
-\qquad
+  \end{tikzcd}\\[6pt]
+  %
   \begin{tikzcd}[cramped, ampersand replacement = \&]
@@
   \end{tikzcd}
-\]
+\end{gather*}
 
@@
 \]
+
 Interpretation of operator applications is the most complicated. Recall that,

### 2015-11-24 paper.tex
factor out macros into packages
Remove dead code
I don't like the word copresheaf
It always confuses me, and since we're only dealing with covariant
presheaves, I'm inclined to just say presheaf when it is clear.

@@
 
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
+\usepackage{notation/modes}
+\usepackage{notation/judgments}
+\usepackage{notation/abts}
+\usepackage{notation/common}
+
 \usepackage{amsmath}
@@
 
-% make lualatex play nice with agda
-\makeatletter
-\ifluatex%
-   \newwrite\CATCHPATCHOUT%
-\fi
-\def\CatchFBT@Fin@l#1[#2]{%
-   \begingroup%
-      \endlinechar\m@ne\makeatletter #2%
-      \ifluatex%
-         \immediate\openout\CATCHPATCHOUT=\jobname.catchpatch%
-         \immediate\write\CATCHPATCHOUT{\CatchFBT@tok{\the\CatchFBT@tok}}%
-         \immediate\closeout\CATCHPATCHOUT%
-         \@@input\jobname.catchpatch%
-      \else%
-      \scantokens\expandafter{%
-         \expandafter\CatchFBT@tok\expandafter{\the\CatchFBT@tok}}%
-      \fi%
-      \CatchFBT@IsAToken{#1}%
-         {\global#1\expandafter{\the\CatchFBT@tok}}%
-         {\xdef#1{\the\CatchFBT@tok}}%
-      \ifx\CatchFBT@tok#1\else\global\CatchFBT@tok{}\fi%
-   \endgroup%
-}% \CatchFBT@Final
-\makeatother
-
 \newtheorem{thm}{Theorem}[section]
@@
 \numberwithin{equation}{section}
-
-\def\IModeColorName{MidnightBlue}
-\def\OModeColorName{Maroon}
-
 %% Euler doesn't have \varsigma and \varrho
@@
 
-\newcommand\IMode[1]{{\color{\IModeColorName}{#1}}}
-\newcommand\OMode[1]{{\color{\OModeColorName}{#1}}}
-
-\newcommand\HypJ[2]{#1\ \ (#2)}
-\newcommand\GenJ[2]{\vert_{\IMode{#1}}\; #2}
-
-\newcommand\IsTac[1]{\IMode{#1}\ \textit{tactic}}
-
-\newcommand\MkSet[1]{\left\{\,#1\,\right\}}
-\newcommand\Member[2]{\IMode{#1}\in\IMode{#2}}
-\newcommand\EqMember[3]{\IMode{#1}=\IMode{#2}\in\IMode{#3}}
-\newcommand\Of[2]{\IMode{#1}: \IMode{#2}}
-
-\newcommand\MkValence[3]{%
-  \ifthenelse{\equal{#1}{}}{}{%
-    \{#1\}%
-  }%
-  \ifthenelse{\equal{#2}{}}{}{%
-    [#2]%
-  }.\, #3%
-}
-\newcommand\MkBTm[3]{\mathbb{\lambda}\{#1\}[#2].\, #3}
-\newcommand\MkArity[2]{(#1)\, #2}
-\newcommand\IsArity[1]{\IMode{#1}\ \textit{arity}}
-\newcommand\IsValence[1]{\IMode{#1}\ \textit{valence}}
-\newcommand\IsSort[1]{\IMode{#1}\ \textit{sort}}
-\newcommand\Leq[2]{\IMode{#1}\leq\IMode{#2}}
-
-\newcommand\ArrId[1]{1_{#1}}
-\newcommand\Shriek[1]{\,{{{\mathexclam}}}_{#1}}
-\newcommand\Match[2]{\IMode{#1}\equiv\OMode{#2}}
-
 \newcommand\Symbols{\mathbf{I}}
@@
 \newcommand\FinSets{\mathbb{F}}
-\newcommand\Sorts{\mathcal{S}}
-\newcommand\Valences{\mathcal{V}}
-\newcommand\Arities{\mathcal{A}}
-\newcommand\Operators{\mathcal{O}}
-
-\newcommand\Iso[2]{\IMode{#1}\mathrel{{{\cong}}}\IMode{#2}}
-\newcommand\Cats{\mathbf{Cat}}
-\newcommand\Discrete[1]{#1_{\equiv}}
-\newcommand\Arr[1]{\mathbf{{#1}^{\ra}}}
-\newcommand\Sets{\mathbf{Set}}
-\newcommand\SCtx{\mathbf{SCtx}}
-\newcommand\Ctx{\mathbf{Ctx}}
-\newcommand\Seq{\mathbf{Seq}}
 \newcommand\HypCat{\mathbf{H}}
 
-\newcommand\OpCat[1]{#1^{\mathtt{op}}}
-
-\newcommand\IsOperator[3]{
-  \IMode{#1}\,{\Vdash}\,\IMode{#2}: \OMode{#3}
-}
-
-\newcommand\IsMetaCtx[1]{\IMode{#1}\ \textit{mctx}}
-\newcommand\IsVarCtx[1]{\IMode{#1}\ \textit{vctx}}
-\newcommand\IsSymCtx[1]{\IMode{#1}\ \textit{sctx}}
-
-\newcommand\Lookup[3]{\IMode{#1}\ni\IMode{#2}:\OMode{#3}}
-
-\newcommand\Grothendieck[2]{\oint^{#1} #2#1}
-\newcommand\Pair[2]{\left\langle\,#1,#2\,\right\rangle}
-\newcommand\Tuple[1]{\left\langle#1 \right\rangle}
 \newcommand\FLift[2]{
@@
 }
-\newcommand\PairTimes[2]{\langle#1\, {\scriptstyle{{\times}}}\, #2 \rangle}
-\newcommand\SSingleton{\bullet}
-\newcommand\Forget[1]{|\, #1\, |}
-
-\newcommand\Comma[2]{#1 \downarrow#2}
-\newcommand\Coslice[2]{#1\, /\, #2}
-
-\newcommand\IsAbt[5]{%
-  \IMode{#1}\triangleright%
-  \IMode{#2}\parallel%
-  \,{\IMode{#3}\vdash}\,%
-  \IMode{#4}:\OMode{#5}%
-}
+
+\newcommand\SortExp{\mathtt{exp}}
+
 \newcommand\IsWf[5]{%
@@
 }
-
-\newcommand\IsBTm[5]{%
-  \IMode{#1}\triangleright%
-  \IMode{#2}\parallel%
-  \,{\IMode{#3}\vdash}\,%
-  \IMode{#4}:\IMode{#5}%
-}
-
-\newcommand\MV[1]{\mathfrak{#1}}
-\newcommand\MApp[3]{#1\{#2\}(#3)}
-\newcommand\App[2]{#1(#2)}
-\newcommand\Dom[1]{\left\vert#1\right\vert}
-\newcommand\FV[1]{\mathbf{FV}\left(#1\right)}
-\newcommand\FS[1]{\mathbf{FS}\left(#1\right)}
-\newcommand\NotIn[2]{\IMode{#1}\notin\IMode{#2}}
-\newcommand\SortExp{\mathtt{exp}}
-
-\newcommand\Hom[3]{#1\left[#2,#3\right]}
-
-\newcommand\Rename[2]{%
-  #2\{#1\}%
-}
-
-\newcommand\IsRenaming[3]{%
-  \Rename{\IMode{#1}}{\IMode{#2}}%
-  \leadsto\OMode{#3}%
-}
-
-\newcommand\Subst[3]{%
-  \left[#1\, /\, #2\right]#3%
-}
-
-\newcommand\IsSubst[4]{%
-  \Subst{\IMode{#1}}{\IMode{#2}}{\IMode{#3}}%
-  \leadsto\OMode{#4}%
-}
-
-\newcommand\IsEq[2]{\IMode{#1} = \IMode{#2}}
-\newcommand\IsApart[2]{\IMode{#1} \mathrel{\#} \IMode{#2}}
-
-\newcommand\Define[2]{\IMode{#1}\triangleq\OMode{#2}}
-\newcommand\ADefine[2]{\IMode{#1}&\triangleq\OMode{#2}}
-
-\newcommand\SigFun[1]{\mathcal{F}_{#1}}
-
-\newcommand\SemBrackets[1]{\left\llbracket#1\right\rrbracket}
-\newcommand\Yoneda[1]{\boldsymbol{y} (#1)}
 
@@
 
-Then, fix a covariant presheaf (copresheaf) of operators
+Then, fix a covariant presheaf of operators
 $\Of{\Operators}{\SCtx\times\Discrete{\Arities}\ra\Sets}$ such that the arrows
@@
 (unlike variables, symbols are not terms). Therefore, the functorial action of
-the operator copresheaf can be lifted into terms by recursion on their
+the operator presheaf can be lifted into terms by recursion on their
 structure, via a pair of judgments $\IsRenaming{\varrho}{M}{N}$ and

### 2015-11-24 paper.tex
Begin talking about sheaves & support

@@
 
-\newcommand\Symbols{\mathbf{I}}
-\newcommand\SymSets{\mathbb{F}_\Symbols}
+\newcommand\SymSets{\mathbb{I}}
 \newcommand\FinSets{\mathbb{F}}
@@
 }
+
+\newcommand\Sheaves[1]{\mathbf{Sh}(#1)}
 
@@
 
-Let $\Symbols$ be an infinite set of symbols. Let $\SymSets$ be the category of
-finite subsets of $\Symbols$ and their injective maps; then the comma
-construction $\Define{\SCtx}{\Comma{\SymSets}{\Discrete{\Sorts}}}$, with
-$\Discrete{\Sorts}$ the discrete category on the set $\mathcal{S}$, is the
-category of contexts of symbols, whose objects are finite sets of symbols $U$
-and sort-assignments
+\subsection{Symbols, contexts and their sheaves}
+
+Let $\SymSets$ be the category of finite cardinals and their injective maps;
+the the comma construction
+$\Define{\SCtx}{\Comma{\SymSets}{\Discrete{\Sorts}}}$, with $\Discrete{\Sorts}$
+the discrete category on the set $\Sorts$, is the category of contexts of
+symbols whose objects are finite sets of symbols $U$ and sort-assignments
 \begin{tikzcd}[cramped]
@@
 
-Then, fix a covariant presheaf of operators
-$\Of{\Operators}{\SCtx\times\Discrete{\Arities}\ra\Sets}$ such that the arrows
+Now, a covariant presheaf $X$ on $\SCtx$ is an intensional set which is subject to
+renamings
+\begin{tikzcd}[cramped]
+  \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
+\end{tikzcd};
+that is, each element $\Member{m}{X(\Upsilon)}$ can be mapped to a unique
+element $\Member{m\varrho}{X(\Upsilon')}$.
+
+\begin{definition}[Support]
+  For a presheaf $\Of{X}{\Sets^\SCtx}$, when
+  \begin{tikzcd}[cramped]
+    \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
+  \end{tikzcd},
+  we say that $\Upsilon$ supports $\Member{m}{\Upsilon'}$ when, for all
+  \begin{tikzcd}[cramped]
+    \IMode{\Upsilon'}\arrow[r, hook, "\IMode{\varrho_1}{,} \IMode{\varrho_2}"] & \IMode{\Upsilon''}
+  \end{tikzcd},
+  if $\IsEq{\varrho_1\circ\varrho}{\varrho_2\circ\varrho}$ then $\IsEq{m\varrho_1}{m\varrho_2}$.
+\end{definition}
+
+Intuitively, when $\Upsilon$ supports $\Member{m}{X(\Upsilon')}$, we would
+expect that we can work backward to a unique $\Of{m'}{\Upsilon}$ such that
+$\IsEq{m'\varrho}{m}$; whilst this is not in general the case for presheaves
+$X$, it is precisely the glueing condition for covariant sheaves on $\SCtx$
+under the atomic topology. The sheaf condition, then, ensures that the notion
+of support is well-behaved.
+
+\begin{remark}
+  The presentation we have given above is related both to the Schanuel topos
+  and, equivalently, Pitts' category of nominal sets; we differ only in
+  discussing sheaves on the category of symbol contexts $\SCtx$, as opposed to
+  sheaves on finite sets of symbols $\SymSets$ (i.e.\ unisorted contexts).
+\end{remark}
+
+\subsection{Operators and signatures}
+
+Fix a sheaf of operators
+$\Of{\Operators}{\Sheaves{\OpCat{\SCtx}\times\Discrete{\Arities}}}$ such that the arrows
 in $\SCtx$ lift to renamings of operators' parameters. Together, $\Sorts$ and

### 2015-11-24 paper.tex
require satisfaction of glueing condition for operators

@@
 $\Of{\pi_{\Operators}}{\Grothendieck{}{\Operators}\ra\SCtx\times\Discrete{\Arities}}$
-in which they reside:
+in which they reside, such that the glueing condition is satisfied:
 \begin{alignat*}{3}

### 2015-11-24 paper.tex
Define the least support

@@
 \newcommand\HypCat{\mathbf{H}}
+\newcommand\Supp[1]{\mathbf{supp}(#1)}
+\newcommand\SupportsUnmoded[3]{#1 \blacktriangleright_{#2} #3}
+\newcommand\Supports[3]{\SupportsUnmoded{\IMode{#1}}{\IMode{#2}}{\IMode{#3}}}
 
@@
   \end{tikzcd},
-  we say that $\Upsilon$ supports $\Member{m}{\Upsilon'}$ when, for all
+  we say that $\Upsilon$ supports $\Member{m}{\Upsilon'}$ (written
+  $\Supports{\Upsilon}{\varrho}{m}$) when, for all
   \begin{tikzcd}[cramped]
@@
 of support is well-behaved.
+
+Let $\Supp{m}$ be the \emph{least support} of $\Member{m}{X(\Upsilon')}$:
+\[
+  \Define{\Supp{m}}{
+    \bigcap_{
+      \begin{tikzcd}[cramped, ampersand replacement = \&]
+        \Upsilon\arrow[r, hook, "\varrho"] \& \Upsilon'
+      \end{tikzcd}
+    }
+    \SupportsUnmoded{\Upsilon}{\varrho}{m}
+  }
+\]
+Intuitively, $\Supp{m}$ is the exact symbol context that $m$ depends on.
 
@@
 
-Abts are identified up to $\alpha$-equivalence. Let $\FS{M}$ be the collection
-of symbols free in $M$, and let $\FV{M}$ be the collection of variables free in
-$M$. \todo[inline]{actually define these}
+Abts are identified up to $\alpha$-equivalence. Let $\FV{M}$ be the collection
+of variables free in $M$, which is easily calculated by recursion on the
+structure of $M$. We also intend to calculate the collection $\FS{M}$ of
+symbols free in $M$, but this is more involved.
 
@@
 (unlike variables, symbols are not terms). Therefore, the functorial action of
-the operator presheaf can be lifted into terms by recursion on their
+the operator sheaf can be lifted into terms by recursion on their
 structure, via a pair of judgments $\IsRenaming{\varrho}{M}{N}$ and

### 2015-11-24 paper.tex
Define calculation of free variables

@@
 
+\newcommand\CalcFV[2]{
+  \FV{\IMode{#1}}\leadsto \OMode{#2}
+}
+
 \newcommand\FLift[2]{
@@
 
-Abts are identified up to $\alpha$-equivalence. Let $\FV{M}$ be the collection
-of variables free in $M$, which is easily calculated by recursion on the
-structure of $M$. We also intend to calculate the collection $\FS{M}$ of
-symbols free in $M$, but this is more involved.
-
-\subsection{Renaming of Symbols}
+Abts are identified up to $\alpha$-equivalence.
+
+\subsection{Calculating free variables}
+
+Let $\FV{M}$ be the collection of variables free in $M$, which is easily
+calculated by recursion on the structure of $M$, as follows:
+
+\begin{gather*}
+  \infer[\textit{var}]{
+    \CalcFV{x}{\MkSet{x}}
+  }{
+  }\\[6pt]
+  %
+  \infer[\textit{mvar}]{
+    \CalcFV{
+      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
+    }{
+      \bigcup_{i\leq n} \vec{x}_i
+    }
+  }{
+    \HypJ{
+      \CalcFV{M_i}{\vec{x}_i}
+    }{
+      \Leq{i}{n}
+    }
+  }\\[6pt]
+  %
+  \infer[\textit{app}]{
+    \CalcFV{
+      \App{\vartheta}{E_0,\dots,E_n}
+    }{
+      \bigcup_{i\leq n} \vec{x}_i
+    }
+  }{
+    \HypJ{
+      \CalcFV{E_i}{\vec{x}_i}
+    }{
+      \Leq{i}{n}
+    }
+  }\\[6pt]
+  %
+  \infer[\textit{abs}]{
+    \CalcFV{
+      \MkBTm{\vec{u}}{\vec{y}}{M}
+    }{
+      \vec{x}\setminus\vec{y}
+    }
+  }{
+    \CalcFV{M}{\vec{x}}
+  }
+\end{gather*}
+
+Because this is a total relation, henceforth we will write $\FV{M}$ for
+$\vec{x}$ when $\CalcFV{M}{x}$.
+
+\subsection{Renaming of symbols}
 
@@
 
-\subsection{Substitution of Variables}
+\subsection{Substitution of variables}
 
@@
 
-\subsection{Substitution of Metavariables}
+\subsection{Substitution of metavariables}
 

### 2015-11-24 paper.tex
Define calculation of free symbols

@@
 
+\newcommand\CalcFS[2]{
+  \FS{\IMode{#1}}\leadsto \OMode{#2}
+}
+
 \newcommand\FLift[2]{
@@
 
-Let $\FV{M}$ be the collection of variables free in $M$, which is easily
-calculated by recursion on the structure of $M$, as follows:
+We can easily calculate the variables free in an expression by recursion on its
+structure:
 
@@
 $\vec{x}$ when $\CalcFV{M}{x}$.
+
+\subsection{Calculating free symbols}
+
+Whereas the calculation of free variables pivoted on the \emph{var} rule, the
+calculation of free symbols will pivot on the \emph{app} rule, because the only
+place a symbol can be introduced is as a parameter to an operator.
+
+\begin{gather*}
+  \infer[\textit{var}]{
+    \CalcFS{x}{\MkSet{}}
+  }{
+  }\\[6pt]
+  %
+  \infer[\textit{mvar}]{
+    \CalcFS{
+      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
+    }{
+      \bigcup_{i\leq n} \vec{u}_i
+    }
+  }{
+    \HypJ{
+      \CalcFS{M_i}{\vec{u}_i}
+    }{
+      \Leq{i}{n}
+    }
+  }\\[6pt]
+  %
+  \infer[\textit{app}]{
+    \CalcFS{
+      \App{\vartheta}{E_0,\dots,E_n}
+    }{
+      \Dom{\Supp{\vartheta}} \cup \bigcup_{i\leq n} \vec{u}_i
+    }
+  }{
+    \HypJ{
+      \CalcFS{E_i}{\vec{u}_i}
+    }{
+      \Leq{i}{n}
+    }
+  }\\[6pt]
+  %
+  \infer[\textit{abs}]{
+    \CalcFS{
+      \MkBTm{\vec{v}}{\vec{x}}{M}
+    }{
+      \vec{u}\setminus\vec{v}
+    }
+  }{
+    \CalcFS{M}{\vec{u}}
+  }
+\end{gather*}
+
+Because this is a total relation, henceforth we will write $\FS{M}$ for
+$\vec{u}$ when $\CalcFS{M}{u}$.
 

### 2015-11-26 paper.tex
Just write out the comma constructions instead of "SCtx", "Ctx"

@@
 
-\newcommand\SymSets{\mathbb{I}}
-\newcommand\FinSets{\mathbb{F}}
 \newcommand\HypCat{\mathbf{H}}
@@
 Let $\SymSets$ be the category of finite cardinals and their injective maps;
-the the comma construction
-$\Define{\SCtx}{\Comma{\SymSets}{\Discrete{\Sorts}}}$, with $\Discrete{\Sorts}$
+then the comma construction $\SCtx$, with $\Discrete{\Sorts}$
 the discrete category on the set $\Sorts$, is the category of contexts of
@@
   and, equivalently, Pitts' category of nominal sets; we differ only in
-  discussing sheaves on the category of symbol contexts $\SCtx$, as opposed to
-  sheaves on finite sets of symbols $\SymSets$ (i.e.\ unisorted contexts).
+  discussing covariant sheaves on the category of symbol contexts $\SCtx$, as
+  opposed to covariant sheaves on finite sets of symbols $\SymSets$ (i.e.\
+  unisorted contexts).
 \end{remark}
@@
 Fix a sheaf of operators
-$\Of{\Operators}{\Sheaves{\OpCat{\SCtx}\times\Discrete{\Arities}}}$ such that the arrows
-in $\SCtx$ lift to renamings of operators' parameters. Together, $\Sorts$ and
-$\Operators$ form a \emph{signature}
+$\Of{\Operators}{\Sheaves{\OpCat{\left(\SCtx\right)}\times\Discrete{\Arities}}}$
+such that the arrows in $\SCtx$ lift to renamings of operators' parameters.
+Together, $\Sorts$ and $\Operators$ form a \emph{signature}
 $\Define{\Sigma}{\Pair{\Sorts}{\Operators}}$.
@@
 
-Let $\FinSets$ be the free cocartesian category on a single object (that is, the
-category of finite cardinals and functions between them); then the comma
-construction $\Define{\Ctx}{\Comma{\FinSets}{\Sorts}}$ is the category of
-variable contexts. Let $\Define{\HypCat}{\OpCat{(\SCtx\times\Ctx)}}$; then we
-fix the functor category $\widehat{\HypCat}^\Sorts$ as our semantic universe.
+Let $\FinSets$ be category of finite cardinals and \emph{any} functions between
+them; then the comma construction $\Ctx$ is the category of variable contexts.
+Let $\Define{\HypCat}{\OpCat{(\SCtx\times\Ctx)}}$; then we fix the functor
+category $\widehat{\HypCat}^\Sorts$ as our semantic universe.
 

### 2015-11-26 paper.tex
Fix typo
Rework section on contexts

@@
 
-In general, we will have three kinds of context: metavariable contexts, variable
-contexts, and symbol (parameter) contexts. A metavariable context $\Theta$
-consists of bindings of valences to metavariables; a variable context $\Gamma$
-is a collection of bindings of sorts to variables, and a parameter context
-$\Upsilon$ is a collection of bindings of sorts to symbols.
-
+In general, we will have three kinds of context: symbolic (parameter) contexts,
+variable contexts, and metavaraible contexts The symbol contexts have already
+been defined via the comma construction $\SCtx$, but they also admit a
+syntactic characterization,
 \begin{gather*}
   \infer{
+    \IsSymCtx{\cdot}
+  }{
+  }
+\qquad
+  \infer{
+    \IsSymCtx{\Upsilon,u:\tau}
+  }{
+    \IsSymCtx{\Upsilon}
+&
+    \IsSort{\tau}
+&
+    \NotIn{u}{\Dom\Upsilon}
+  }
+\end{gather*}
+%
+Because, modulo notation, we have $\Member{\Upsilon}{\SCtx}$ just when
+$\IsSymCtx{\Upsilon}$, we will use the syntactic view when it is convenient.
+
+Contexts of variables are similar to contexts of symbols, except that they
+admit \emph{any} renamings, not just the injective ones. As such, when
+$\FinSets$ is the category of finite cardinals and all functions between them, the
+comma construction $\Ctx$ is the category of variable contexts. As above, we
+can give them an equivalent syntactic treatment:
+\begin{gather*}
+  \infer{
+    \IsVarCtx{\cdot}
+  }{
+  }
+\qquad
+  \infer{
+    \IsVarCtx{\Gamma,x:\tau}
+  }{
+    \IsVarCtx{\Gamma}
+&
+    \IsSort{\tau}
+&
+    \NotIn{x}{\Dom\Gamma}
+  }
+\end{gather*}
+
+A metavariable context consists of bindings of \emph{valences} to
+metavariables; let $\Define{\Valences}{\MkSet{v\mid \IsValenceUnmoded{v}}}$ be the set
+of valences. Then, the category of metavariable contexts is the comma
+construction $\MCtx$, which likewise admits an equivalent inductive definition:
+\begin{gather*}
+  \infer{
     \IsMetaCtx{\cdot}
@@
     \NotIn{\MV{m}}{\Dom\Theta}
-  }
-\end{gather*}
-
-\begin{gather*}
-  \infer{
-    \IsVarCtx{\cdot}
-  }{
-  }
-\qquad
-  \infer{
-    \IsVarCtx{\Gamma,x:\tau}
-  }{
-    \IsVarCtx{\Gamma}
-&
-    \IsSort{\tau}
-&
-    \NotIn{x}{\Dom\Gamma}
-  }
-\end{gather*}
-
-\begin{gather*}
-  \infer{
-    \IsSymCtx{\cdot}
-  }{
-  }
-\qquad
-  \infer{
-    \IsSymCtx{\Upsilon,u:\tau}
-  }{
-    \IsVarCtx{\Upsilon}
-&
-    \IsSort{\tau}
-&
-    \NotIn{u}{\Dom\Upsilon}
   }
@@
 
-\todo[inline]{show that renaming preserves wellformedness}
-
 \subsection{Substitution of variables}
@@
 &
-    \IsRenaming{\MkSet{\vec{u}\mapsto\vec{v}}}{N'}{N''}
-  }
-\end{gather*}
-\begin{gather*}
+    \IsRenaming{\vec{u}\mapsto\vec{v}}{N'}{N''}
+  }\\[6pt]
+  %
   \infer{
@@
 
-Let $\FinSets$ be category of finite cardinals and \emph{any} functions between
-them; then the comma construction $\Ctx$ is the category of variable contexts.
 Let $\Define{\HypCat}{\OpCat{(\SCtx\times\Ctx)}}$; then we fix the functor

### 2015-11-26 paper.tex
Add citations for schanuel / nominal sets stuff

@@
 $\IsEq{m'\varrho}{m}$; whilst this is not in general the case for presheaves
-$X$, it is precisely the glueing condition for covariant sheaves on $\SCtx$
-under the atomic topology. The sheaf condition, then, ensures that the notion
+$X$, it is precisely the sheaf condition for covariant presheafs on $\SCtx$
+under the atomic topology \cite[p.~126]{mac-lane-moerdijk:1992}. The sheaf condition, then, ensures that the notion
 of support is well-behaved.
@@
   The presentation we have given above is related both to the Schanuel topos
-  and, equivalently, Pitts' category of nominal sets; we differ only in
+  and, equivalently, Pitts' category of nominal sets \cite{fiore-staton:2006}; we differ only in
   discussing covariant sheaves on the category of symbol contexts $\SCtx$, as

### 2015-11-26 paper.tex
Add macros for \SymPsh and \VarPsh; bold

@@
 \subsection{Substitution monoidal structures}
+\newcommand\SymPsh{\mathbf{S}}
+\newcommand\VarPsh{\mathbf{V}}
 
@@
 $P^{[\Gamma]}$ to mean $\prod_{x\in\Dom\Gamma}P_{\Gamma(x)}$; likewise,
-$S^{\{\Upsilon\}}$ shall mean $\prod_{u\in\Dom\Upsilon}S_{\Upsilon(u)}$. For a
+$\SymPsh^{\{\Upsilon\}}$ shall mean $\prod_{u\in\Dom\Upsilon}S_{\Upsilon(u)}$. For a
 presheaf $\Of{A}{\widehat{\HypCat}}$ and a sort-indexed family of presheaves
@@
       A(\Upsilon'\parallel\Delta)
-      \times S^{\{\Upsilon'\}}(\Upsilon\parallel\Gamma)
+      \times \SymPsh^{\{\Upsilon'\}}(\Upsilon\parallel\Gamma)
       \times P^{[\Delta]}(\Upsilon\parallel\Gamma)
@@
   \begin{tikzcd}[cramped, sep = large, ampersand replacement = \&]
-    {\IMode{{P_\tau^{\Yoneda{\Upsilon\parallel\Gamma}}\times S^{\{\Upsilon\}}\times P^{[\Gamma]}}}}
+    {\IMode{{P_\tau^{\Yoneda{\Upsilon\parallel\Gamma}}\times \SymPsh^{\{\Upsilon\}}\times P^{[\Gamma]}}}}
       \arrow[r, "\IMode{\varsigma^\tau_{\Upsilon\parallel\Gamma}}"]
@@
     \times
-    S^{\{\Upsilon\}}
+    \SymPsh^{\{\Upsilon\}}
     \times
-    V^{[\Gamma]}
+    \VarPsh^{[\Gamma]}
   }
@@
       }
-      \times S^{\{\vec{\sigma}\}}
+      \times \SymPsh^{\{\vec{\sigma}\}}
       \times P^{[\vec{\tau}]}
@@
 \&
-    {\IMode{S^{\{\Upsilon\}}}}
+    {\IMode{\SymPsh^{\{\Upsilon\}}}}
       \arrow[
@@
 \&
-    {\IMode{S^{\{\vec{\sigma}\}}}}
+    {\IMode{\SymPsh^{\{\vec{\sigma}\}}}}
   \end{tikzcd}
@@
 \&[12.0em]
-    {\IMode{S^{\{\Upsilon,\vec{\sigma}_i\}}}}
+    {\IMode{\SymPsh^{\{\Upsilon,\vec{\sigma}_i\}}}}
 \\[-1.0em]
@@
 \&[12.0em]
-    {\IMode{V^{[\Gamma,\vec{\tau}_i]}}}
+    {\IMode{\VarPsh^{[\Gamma,\vec{\tau}_i]}}}
   \end{tikzcd}

### 2015-11-27 paper.tex
adjust notation for renaming operators

@@
 such that $\IsOperator{\Upsilon'}{\FLift{\varrho}{f}(\vartheta)}{a'}$ when
-$\IsOperator{\Upsilon}{\vartheta}{a}$. Because the category
-$\Discrete{\Arities}$ is discrete and has only identity arrows, through an abuse
-of notation we will often write $\FLift{\varrho}{}$ to mean
-$\FLift{\varrho}{\ArrId{}}$. In particular, the renaming
-$\Upsilon,u\mapsto\Upsilon,v$ shall take $\mathtt{get}[u]$ to $\mathtt{get}[v]$.
+$\IsOperator{\Upsilon}{\vartheta}{a}$. In particular, the renaming
+$\Upsilon,u\mapsto\Upsilon,v$ shall take $\mathtt{get}[u]$ to
+$\mathtt{get}[v]$. Because the category $\Discrete{\Arities}$ is discrete and
+has only identity arrows, through an abuse of notation we will often write
+$\vartheta\varrho$ to mean $\FLift{\varrho}{\ArrId{}}(\vartheta)$.
 
@@
     }{
-      \App{\vartheta'}{F_0,\dots,F_n}
-    }
-  }{
-    \Match{\FLift{\varrho}{}(\vartheta)}{\vartheta'}
-&
+      \App{\vartheta\varrho}{F_0,\dots,F_n}
+    }
+  }{
     \HypJ{
@@
         , swap
-        , "\IMode{\FLift{\pi_{(-)}}{}}"
+        , "\IMode{\FLift{\pi_{(-)}}{\ArrId{}}}"
       ]

### 2015-12-02 paper.tex
Rework the treatment of operators
@freebroccolo What do you think of this?

@@
 
-\newcommand\FLift[2]{
-  \Operators\langle#1%
-  \ifthenelse{\equal{#2}{}}{}{, #2}\rangle%
-}
-
 \newcommand\SortExp{\mathtt{exp}}
@@
 
-\newcommand\Sheaves[1]{\mathbf{Sh}(#1)}
+\newcommand\Sheaves[1]{\mathbf{Sh}\left(#1\right)}
+\newcommand\Presheaves[1]{\mathbf{Psh}\left(#1\right)}
+\newcommand\Sheafify[1]{#1^{\#}}
+\newcommand\ConstPsh[1]{\Delta[#1]}
+\newcommand\ConstSh[1]{\Delta^\#[#1]}
+\newcommand\SCtxSite{\SymSets[\Sorts]}
 
@@
 
-Now, a covariant presheaf $X$ on $\SCtx$ is an intensional set which is subject to
-renamings
+Now, a covariant presheaf $X$ on $\SCtx$ is an intensional set which is subject
+to renamings
 \begin{tikzcd}[cramped]
@@
 $\IsEq{m'\varrho}{m}$; whilst this is not in general the case for presheaves
-$X$, it is precisely the sheaf condition for covariant presheafs on $\SCtx$
-under the atomic topology \cite[p.~126]{mac-lane-moerdijk:1992}. The sheaf condition, then, ensures that the notion
-of support is well-behaved.
+$X$, it is precisely the sheaf condition for covariant presheaves on $\SCtx$
+under the atomic topology \cite[p.~126]{mac-lane-moerdijk:1992}. The sheaf
+condition, then, ensures that the notion of support is well-behaved.
 
@@
 
+Going forward, we will write $\SCtxSite$ for the Grothendieck site
+$\Pair{\OpCat{\left(\SCtx\right)}}{A}$, where $A$ is the atomic coverage (sc.\
+every non-empty family covers).
+
 \begin{remark}
@@
   and, equivalently, Pitts' category of nominal sets \cite{fiore-staton:2006}; we differ only in
-  discussing covariant sheaves on the category of symbol contexts $\SCtx$, as
-  opposed to covariant sheaves on finite sets of symbols $\SymSets$ (i.e.\
-  unisorted contexts).
+  discussing sheaves on $\SCtxSite$, as
+  opposed to sheaves on $\OpCat{\SymSets}$ (i.e.\ unisorted contexts).
 \end{remark}
 
+\subsubsection{Constructions on sheaves}
+
+For any Grothendieck site $\Match{\mathfrak{S}}{\Pair{\mathcal{C}}{J}}$, the
+inclusion $\Of{\iota}{\Sheaves{\mathfrak{S}}\ra\Presheaves{\mathcal{C}}}$ of
+sheaves into presheaves has a left adjoint
+$\IMode{\Sheafify{(-)}}\dashv\IMode{\iota}$, which takes any presheaf
+$\Of{X}{\Presheaves{\mathcal{C}}}$ to a unique sheaf
+$\Of{\Sheafify{X}}{\Sheaves{\mathfrak{S}}}$, called the \emph{sheafification}
+of $X$.
+
+Every category of sheaves on a site gives rise to a topos, which equips us with
+a number of standard constructions, including (among other things) the disjoint
+union of sheaves $X\oplus Y$, the product of sheaves $X\otimes Y$, and the terminal sheaf $\mathbb{1}$.
+
+Returning to the site $\SCtxSite$, we will write $\mathbf{S}_\tau$ for the sheaf
+of symbols of sort $\tau$, a subobject of the Yoneda embedding of the empty
+symbol context, $\Yoneda{\cdot}$.
+
 \subsection{Operators and signatures}
 
-Fix a sheaf of operators
-$\Of{\Operators}{\Sheaves{\OpCat{\left(\SCtx\right)}\times\Discrete{\Arities}}}$
-such that the arrows in $\SCtx$ lift to renamings of operators' parameters.
-Together, $\Sorts$ and $\Operators$ form a \emph{signature}
+Fix a family of sheaves of operators
+$\Member{\Operators_a}{\Sheaves{\SCtxSite}}$, indexed by arities
+$\Member{a}{\Arities}$; for each $\Operators_a$, arrows in $\SCtx$ will lift to
+renamings in operators' symbolic parameters. Together, $\Sorts$ and $\Operators$ are said to form a \emph{signature}
 $\Define{\Sigma}{\Pair{\Sorts}{\Operators}}$.
 
-By the Grothendieck construction
-$\Iso{\Grothendieck{}{\Operators}}{\Comma{\mathbb{1}}{\Operators}}$ we have a
-category of elements of $\Operators$ where objects are triples
-$\Pair{\Pair{\Upsilon}{a}}{\vartheta}$, interpreted as
-$\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$, and morphisms are paired
-diagrams:
-\[
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    \IMode{\Pair{\Upsilon}{a}}
-      \arrow[r, swap, "\IMode{\Pair{\rho}{f}}"]
-\&
-    \IMode{\Pair{\Upsilon'}{a'}}
-  \end{tikzcd}
-\qquad
-  \begin{tikzcd}[cramped, sep = small, ampersand replacement = \&]
-    \null
-\&
-    \IMode{\mathbb{1}}
-      \arrow[dl, swap, "\IMode{\vartheta}"]
-      \arrow[dr, "\IMode{\vartheta'}"]
-\&
-    \null
-\\
-    \IMode{\Operators\Pair{\Upsilon}{a}}
-      \arrow[rr, swap, "\IMode{\Operators\Pair{\rho}{f}}"]
-\&
-    \null
-\&
-    \IMode{\Operators\Pair{\Upsilon'}{a'}}
-  \end{tikzcd}
-\]
-
 We will write $\IsOperator{\Upsilon}{\vartheta}{a}$ in case
-$\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$ and note that this judgment
-enjoys the structural properties of weakening and exchange via functoriality of
-$\Operators$. Operators $\vartheta_{0}, \ldots, \vartheta_{n}$ are defined by
-specifying the fibers of
-$\Of{\pi_{\Operators}}{\Grothendieck{}{\Operators}\ra\SCtx\times\Discrete{\Arities}}$
-in which they reside, such that the glueing condition is satisfied:
-\begin{alignat*}{3}
-  \null
-&
-  \IMode{{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}}
-&&
-  \,\mathrel{\ {{:}}}\,
-&&
-  \IMode{\SCtx\times\Discrete{\Arities}\ra\Sets}
-\\
-  \null
-&
-  \IMode{{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}}
-&&
-  \,=\,\
-&&
-    \OMode{
-      \MkSet{
-        \vartheta\in\Pair{\Upsilon}{a}
-        \mid{\pi_{\Operators}}{\Pair{\Pair{\Upsilon}{a}}{\vartheta}} =
-        \Pair{\Upsilon}{a}
-      }
-    }
-\end{alignat*}
+$\Member{\vartheta}{\Operators_a(\Upsilon)}$; note that this judgment enjoys
+the structural properties of weakening and exchange via functoriality of
+$\Operators$. An operator signature is defined by specifying, for each arity
+$a$, the sheaf $\Operators_a$, whose ``elements'' are the operators of arity
+$a$.
 
@@
 
-For instance, consider the $\lambda$-calculus with a single sort, $\SortExp$; we
+For instance, consider a $\lambda$-calculus with a single sort, $\SortExp$; we
 give its signature $\Sigma_{\lambda}$ by asserting the following about its
@@
     }{\SortExp}
-  }
-\\
+  }\\
+  \IsOperator{\Upsilon}{\mathtt{fix}}{
+    \MkArity{
+      \MkValence{}{\SortExp}{\SortExp}
+    }{\SortExp}
+  }\\
   \IsOperator{\Upsilon}{\mathtt{ap}}{
@@
 \end{gather*}
-These rules correspond to objects of $\Grothendieck{}{{\Operators}_{\lambda}}$
-and we collect their fibers into sets of operators:
-\begin{alignat*}{5}
-  \IMode{{\pi_{{\Operators}_{\lambda}}}^{-1}}
-&
-  \IMode{
-    \Pair{
-      \Upsilon
-    }{
+
+These rules correspond to the following definition of $\Operators$:
+\begin{align*}
+  \ADefine{
+    \Operators_{
        \MkArity{
@@
     }
-  }
-\quad
-&&
-  =
-&&
-\quad
-  \OMode{\MkSet{\mathtt{lam}}}
-\\
-  \IMode{{\pi_{{\Operators}_{\lambda}}}^{-1}}
-&
-  \IMode{
-    \Pair{
-      \Upsilon
-    }{
+  }{
+    \mathbb{1}\oplus\mathbb{1}
+  }
+  \\
+  %
+  \ADefine{
+    \Operators_{
       \MkArity{
@@
     }
-  }
-\quad
-&&
-  =
-&&
-\quad
-  \OMode{\MkSet{\mathtt{ap}}}
-\\
-  \IMode{{\pi_{{\Sigma}_{\lambda}}}^{-1}}
-&
-  \triangleq\IMode{
-    \bigcup_{\Pair{\Upsilon}{a}}{
-      {\pi_{{\Operators}_{\lambda}}}^{-1}
-      \Pair{\Upsilon}{a}
-    }
-  }
-\quad
-&&
-  =
-&&
-\quad
-  \OMode{
-    \MkSet{\mathtt{lam},\ \mathtt{ap}}
-  }
-\\
-\end{alignat*}
+  }{
+    \mathbb{1}
+  }
+\end{align*}
 
@@
 extension of the calculus with assignables (references):
-
+%
 \begin{gather*}
@@
 
+These rules correspond to the following family of sheaves:
+\begin{align*}
+  \ADefine{
+    \Operators_{
+      \MkArity{
+        \MkValence{}{}{\SortExp},
+        \MkValence{\SortExp}{}{\SortExp}
+      }{
+        \SortExp
+      }
+    }
+  }{
+    \mathbb{1}
+  }\\
+  \ADefine{
+    \Operators_{
+      \MkArity{}{\SortExp}
+    }
+  }{
+    \mathbf{S}_\SortExp
+  }
+\end{align*}
+
 Declaring a new assignable consists in providing an initial value, and an
@@
 Note that the functoriality of $\Operators$ guarantees for any renaming
-$\Of{\varrho}{\Upsilon\inj\Upsilon'}$ and $\Of{f}{a\ra a'}$, a lifted map
-$\Of{\FLift{\varrho}{f}}{\Operators\Pair{\Upsilon}{a}\inj\Operators\Pair{\Upsilon'}{a'}}$
-such that $\IsOperator{\Upsilon'}{\FLift{\varrho}{f}(\vartheta)}{a'}$ when
+$\Of{\varrho}{\Upsilon\inj\Upsilon'}$, a family of lifted maps
+$\Of{\Operators_a(\varrho)}{\Operators_a(\Upsilon)\inj\Operators_a(\Upsilon')}$
+natural in $a$, such that
+$\IsOperator{\Upsilon'}{\Operators_a(\varrho)(\vartheta)}{a}$ when
 $\IsOperator{\Upsilon}{\vartheta}{a}$. In particular, the renaming
 $\Upsilon,u\mapsto\Upsilon,v$ shall take $\mathtt{get}[u]$ to
-$\mathtt{get}[v]$. Because the category $\Discrete{\Arities}$ is discrete and
-has only identity arrows, through an abuse of notation we will often write
-$\vartheta\varrho$ to mean $\FLift{\varrho}{\ArrId{}}(\vartheta)$.
+$\mathtt{get}[v]$. Going forward, we will write $\vartheta\varrho$ for
+$\Operators_a(\varrho)$.
 
@@
 \[
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
+  \begin{tikzcd}[cramped, ampersand replacement = \&, sep = huge]
     |[alias = Src]|
@@
         , swap
-        , "\IMode{\FLift{\pi_{(-)}}{\ArrId{}}}"
+        , "\IMode{\Operators_a(\pi_{(-)})(\vartheta)}"
       ]

### 2015-12-02 paper.tex
finish up operator interpretation example
Define sheafification

@@
 Going forward, we will write $\SCtxSite$ for the Grothendieck site
-$\Pair{\OpCat{\left(\SCtx\right)}}{A}$, where $A$ is the atomic coverage (sc.\
-every non-empty family covers).
+$\Pair{\OpCat{\left(\SCtx\right)}}{J_\mathsf{atm}}$, where $J_\mathsf{atm}$
+is the atomic coverage (sc.\ every non-empty family covers).
 
@@
 $\Of{\Sheafify{X}}{\Sheaves{\mathfrak{S}}}$, called the \emph{sheafification}
-of $X$.
+of $X$. Sheafification is the twice-iterated application of the
+\emph{+-construction} $X^+$, which turns a presheaf into a separated presheaf,
+and a separated presheaf into a sheaf:
+\begin{align*}
+  \ADefine{
+    X^+(C)
+  }{
+    \int^{S\in J(C)}
+    \Hom{\Presheaves{\mathcal{C}}}{S}{X}
+  }\\
+  \ADefine{
+    \Sheafify{X}
+  }{
+    X^{++}
+  }
+\end{align*}
+
+We also have a canonical map $\Of{\eta}{X\ra X^+}$, defined
+using the maximal sieve $\Define{t_C}{\Yoneda{C}}$:
+\[
+  \HypJ{
+    \Define{\eta_C(m)}{
+      \Pair{t_C}{
+        X(-)(m)
+      }
+    }
+  }{
+    \Member{m}{X(C)}
+  }
+\]
 
@@
       \MkArity{}{\SortExp}
+    }
+  }{
+    \mathbf{S}_\SortExp
+  }\\
+  \ADefine{
+    \Operators_{
+      \MkArity{
+        \MkValence{}{}{\SortExp}
+      }{
+        \SortExp
+      }
     }

### 2015-12-02 paper.tex
Add more citations

@@
   The presentation we have given above is related both to the Schanuel topos
-  and, equivalently, Pitts' category of nominal sets \cite{fiore-staton:2006}; we differ only in
-  discussing sheaves on $\SCtxSite$, as
-  opposed to sheaves on $\OpCat{\SymSets}$ (i.e.\ unisorted contexts).
+  and, equivalently, Pitts' category of nominal sets \cite{fiore-staton:2006,
+  pitts:2013}; we differ only in discussing sheaves on $\SCtxSite$, as opposed
+  to sheaves on $\OpCat{\SymSets}$ (i.e.\ unisorted contexts).
 \end{remark}
@@
 \emph{+-construction} $X^+$, which turns a presheaf into a separated presheaf,
-and a separated presheaf into a sheaf:
+and a separated presheaf into a sheaf \cite{mac-lane-moerdijk:1992}:
 \begin{align*}

### 2015-12-02 paper.tex
Minor cleanup of case study
[case study] rename OpVar -> OpHyp
Discuss Abella; call them *nominal abstract binding trees*

@@
 
-The contribution of this paper is the development of the syntax and semantics of
-multi-sorted abts, an extension of second order universal algebra to support
-symbol-indexed families of operators. Additionally, we have developed the
-categorical semantics for abts formally in Constructive Type Theory using the
-Agda proof assistant~\cite{norell:2007}.
+The contribution of this paper is the development of the syntax and semantics
+of multi-sorted nominal abts, an extension of second order universal algebra to
+support symbol-indexed families of operators. Additionally, we have developed
+the categorical semantics for abts formally in Constructive Type Theory using
+the Agda proof assistant~\cite{norell:2007}.
 
@@
 
-\section{Abstract Binding Trees}
+\section{Nominal Abstract Binding Trees}
 
@@
 The representation of telescopes and sequents in a logical framework is
-notoriously difficult; whilst it is possible to use higher-order abstract syntax
-or abts to encode the binding-structure of telescopes and sequents, the encoding
-is sufficiently laborious and obscure that it is not used in practice.
-
+notoriously difficult; whilst it is possible to use higher-order abstract
+syntax or abts to encode the binding-structure of telescopes and sequents, the
+encoding is sufficiently laborious and obscure that it is not used in practice.
 Crary has demonstrated a first-order encoding of contexts in the logical
-framework in bijection with actual LF-contexts~\cite{crary:2009}, which has been
-successfully used in large-scale mechanization efforts, including that of
+framework in bijection with actual LF-contexts~\cite{crary:2009}, which has
+been successfully used in large-scale mechanization efforts, including that of
 Standard ML~\cite{lee-crary-harper:2007} and the Edinburgh Logical Framework
@@
 
-\todo[inline]{talk about context encodings in Abella}
-
 We will approach the problem of encoding telescopes and sequents from the
 \emph{refinements} perspective, where a conservative approximation of the
-grammar is first given using the abt logical framework, and then the correctness
-of a code is expressed separately in a judgment that refines the existing
-specification.
-
-Because we have not committed to using the built-in binding machinery to express
-the well-scopedness of telescopes and sequents, we are free to use
-\emph{symbols} in order to model the variables in the context. This is in fact
-quite sensible if we are actually trying to faithfully represent the syntax of
-telescopes and sequents, rather than replace them with their counterparts on the
-meta-level.
+grammar is first given using the abt logical framework, and then the
+correctness of a code is expressed separately in a judgment that refines the
+existing specification.  Because we have not committed to using the built-in
+binding machinery to express the well-scopedness of telescopes and sequents, we
+are free to use \emph{symbols} in order to model the variables in the context.
+This is in fact quite sensible if we are actually trying to faithfully
+represent the syntax of telescopes and sequents, rather than replace them with
+their counterparts on the meta-level.
+
+The approach outlined above is actually quite similar to the standard practice
+in the Abella proof assistant \cite{gacek:2008}; in Abella, in addition to the
+canonical forms, each type is inhabited by an infinite number of nominal
+constants or symbols, which can be used in the syntax of contexts. In order to
+specify where such symbols are allowed to occur, a wellformedness predicate (a
+``scheme'') is separately defined. Our setting is similar in that we may also
+use symbols to encode the syntax of contexts, except that here, we have better
+control over the proliferation of symbols because, contrary to the state of
+affairs in Abella, symbols are \emph{not} terms.
 
@@
 \newcommand\OpSnoc[1]{\mathtt{snoc}[#1]}
-\newcommand\OpVar[1]{\mathtt{var}[#1]}
+\newcommand\OpHyp[1]{\mathtt{hyp}[#1]}
 \newcommand\OpSequent{\mathtt{sequent}}
@@
     \IsSort{\SortJdg}\\\\
-    \IsOperator{\Upsilon, u:\SortExpr}{\OpVar{u}}{
+    \IsOperator{\Upsilon, u:\SortExpr}{\OpHyp{u}}{
       \MkArity{}{\SortExpr}
@@
 
-\newcommand\OpPi{\mathtt{pi}}
-\newcommand\OpSg{\mathtt{sg}}
-\newcommand\OpBool{\mathtt{bool}}
-\newcommand\OpSo{\mathtt{isTrue}}
-
 Suppose we have encoded a fragment of type theory as well:
@@
   \begin{array}{l}
-    \IsOperator{\Upsilon}{\top}{
+    \IsOperator{\Upsilon}{\mathtt{A}}{
       \MkArity{}{\SortType}
@@
 \\
-    \IsOperator{\Upsilon}{\bot}{
-      \MkArity{}{\SortType}
-    }
-\\
-    \IsOperator{\Upsilon}{\OpBool}{
-      \MkArity{}{\SortType}
-    }
-\\
-    \IsOperator{\Upsilon}{\OpSo}{
+    \IsOperator{\Upsilon}{\mathtt{pred}}{
       \MkArity{
         \MkValence{}{}{\SortExp}
-      }{\SortType}
-    }
-\\
-    \IsOperator{\Upsilon}{\OpPi}{
-      \MkArity{
-        \MkValence{}{}{\SortType},
-        \MkValence{}{\SortExp}{\SortType}
-      }{\SortType}
-    }
-\\
-    \IsOperator{\Upsilon}{\OpSg}{
-      \MkArity{
-        \MkValence{}{}{\SortType},
-        \MkValence{}{\SortExp}{\SortType}
       }{\SortType}
@@
 &
-  \triangleq \OMode{\OpVar{u}}
+  \triangleq \OMode{\OpHyp{u}}
 \end{align*}
@@
   }{\cdot}{
-    \diamond,u:\OpBool,v:\App{\OpSo}{`u}
+    \diamond,u:\mathtt{A},v:\App{\mathtt{pred}}{`u}
     \gg
-    \App{\OpSo}{`u}
+    \App{\mathtt{pred}}{`u}
   }{\SortJdg}
@@
       \MkBTm{u,v}{}{
-        \diamond,u:\OpBool,v:\App{\OpSo}{`u}
+        \diamond,u:\mathtt{A},v:\App{\mathtt{pred}}{`u}
         \gg
-        \App{\OpSo}{`u}
+        \App{\mathtt{pred}}{`u}
       }

### 2015-12-03 paper.tex
[case study] Add "comments" to signature block
[case study] notation
[case study] wordsmithing

@@
 in the Abella proof assistant \cite{gacek:2008}; in Abella, in addition to the
-canonical forms, each type is inhabited by an infinite number of nominal
+canonical forms, each type is inhabited by an infinite collection of nominal
 constants or symbols, which can be used in the syntax of contexts. In order to
@@
 use symbols to encode the syntax of contexts, except that here, we have better
-control over the proliferation of symbols because, contrary to the state of
+control over their proliferation because, contrary to the state of
 affairs in Abella, symbols are \emph{not} terms.
@@
 \newcommand\SortJdg{\mathtt{jdg}}
-\newcommand\SortType{\mathtt{type}}
+\newcommand\SortProp{\mathtt{prop}}
 \newcommand\SortExpr{\mathtt{exp}}
@@
 \[
-  \begin{array}{l}
-    \IsSort{\SortTele}\\
-    \IsSort{\SortExpr}\\
-    \IsSort{\SortType}\\
-    \IsSort{\SortJdg}\\\\
+  \begin{array}{lr}
+    \IsSort{\SortTele}&\textit{telescopes}\\
+    \IsSort{\SortExpr}&\textit{expressions}\\
+    \IsSort{\SortProp}&\textit{propositions}\\
+    \IsSort{\SortJdg}&\textit{judgments}\\\\
     \IsOperator{\Upsilon, u:\SortExpr}{\OpHyp{u}}{
@@
         \MkValence{}{}{\SortTele},
-        \MkValence{}{}{\SortType}
+        \MkValence{}{}{\SortProp}
       }{\SortTele}
@@
         \MkValence{}{}{\SortTele},
-        \MkValence{}{}{\SortType}
+        \MkValence{}{}{\SortProp}
       }{\SortJdg}
@@
   \begin{array}{l}
-    \IsOperator{\Upsilon}{\mathtt{A}}{
-      \MkArity{}{\SortType}
+    \IsOperator{\Upsilon}{\mathtt{P}}{
+      \MkArity{}{\SortProp}
     }
@@
         \MkValence{}{}{\SortExp}
-      }{\SortType}
+      }{\SortProp}
     }
@@
   }{\cdot}{
-    \diamond,u:\mathtt{A},v:\App{\mathtt{pred}}{`u}
+    \diamond,u:\mathtt{P},v:\App{\mathtt{pred}}{`u}
     \gg
@@
       \MkBTm{u,v}{}{
-        \diamond,u:\mathtt{A},v:\App{\mathtt{pred}}{`u}
+        \diamond,u:\mathtt{P},v:\App{\mathtt{pred}}{`u}
         \gg
@@
 operators. For the sake of this example, we will just stipulate that anything of
-sort $\SortExp$ or $\SortType$ is grammatical if its subterms are grammatical:
+sort $\SortExp$ or $\SortProp$ is grammatical if its subterms are grammatical:
 
@@
   \infer[
-    \text{for $\Member{\tau}{\MkSet{\SortExp,\SortType}}$}
+    \text{for $\Member{\tau}{\MkSet{\SortExp,\SortProp}}$}
   ]{
@@
 &
-    \IsWf{\Theta}{\Upsilon}{\Gamma}{A}{\SortType}
+    \IsWf{\Theta}{\Upsilon}{\Gamma}{A}{\SortProp}
   }
@@
 \\
-      \IsWf{\Theta}{\Upsilon\setminus\MkSet{u}}{\Gamma}{A}{\SortType}
+      \IsWf{\Theta}{\Upsilon\setminus\MkSet{u}}{\Gamma}{A}{\SortProp}
     \end{array}

### 2015-12-04 paper.tex
Start adding summary of nominal sets, schanuel topos, proofs

@@
 
+\DeclareMathOperator{\Perm}{\mathbf{Perm}}
+
 \begin{document}
@@
 
-Going forward, we will write $\SCtxSite$ for the Grothendieck site
+In the future, we will write $\SCtxSite$ for the Grothendieck site
 $\Pair{\OpCat{\left(\SCtx\right)}}{J_\mathsf{atm}}$, where $J_\mathsf{atm}$
@@
 
-\begin{remark}
-  The presentation we have given above is related both to the Schanuel topos
-  and, equivalently, Pitts' category of nominal sets \cite{fiore-staton:2006,
-  pitts:2013}; we differ only in discussing sheaves on $\SCtxSite$, as opposed
-  to sheaves on $\OpCat{\SymSets}$ (i.e.\ unisorted contexts).
-\end{remark}
+\subsubsection{Relation to Nominal Sets and the Schanuel Topos}
+
+Pitts defines a category $\mathbf{Nom}$ of nominal sets in \cite{pitts:2013}
+which function equivalently to the sheaves that we considered above, in the
+case of $\IMode{\Sorts}\cong\IMode{\mathbb{1}}$. Fix a countably infinite set
+of atoms $\mathbb{A}$; then let $\Perm\mathbb{A}$ be the group of permutations
+(i.e.\ autoequivalences) on $\mathbb{A}$.  Considered as a category,
+$\Perm\mathbb{A}$ has a single object $\bullet$ with morphisms
+$\Of{\pi}{\bullet\ra\bullet}$ for each $\Member{\pi}{\Dom{\Perm\mathbb{A}}}$.
+Then, the category of nominal sets $\mathbf{Nom}$ is the subcategory of
+$\Sets^{\Perm\mathbb{A}}$ containing just the presheaves which satisfy a
+\emph{finite support} condition analogous to the one that we have formulated
+above.\todo[inline]{write it out}
+
+The category $\mathbf{Nom}$ is equivalent to the category of covariant sheaves
+on $\SymSets$ under the atomic coverage, called the Schanuel Topos
+$\Define{\mathbb{S}}{\Sheaves{\OpCat{\SymSets}}}$ \cite{pitts:2013,
+fiore-staton:2006}; equivalently, the Schanuel topos is the subcategory of
+$\Sets^\SymSets$ containing only pullbacks-preserving functors.
+
+In addition to the unisorted nominal sets framework, Pitts also briefly
+discusses a multi-sorted version of the apparatus in which the countably
+infinite set of atoms $\mathbb{A}$ is equipped with a sort assignment
+$\Of{\mathfrak{s}}{\mathbb{A}\to\Sorts}$ such that all the fibres of
+$\mathfrak{s}$ are countably infinite; then, he defines for any such assignment
+the category $\mathbf{Nom}_\mathfrak{s}$ of $\mathfrak{s}$-sorted nominal sets
+as the category of presheaves on the group of $\mathfrak{s}$-respecting
+permutations on $\mathbb{A}$, $\Perm_\mathfrak{s}\mathbb{A}$.
+
+In the same way as the Schanuel topos is equivalent to the unisorted nominal
+sets, we expect to find an equivalence between $\Sheaves{\SCtxSite}$ and the
+limit $\int_{\mathfrak{s}}\mathbf{Nom}_\mathfrak{s}$. At the very least,
+the sheaf topos $\Sheaves{\SCtxSite}$ may serve as a multi-sorted
+generalization of the Schanuel topos, for which we may carry over certain
+useful results, including its characterization as the subcategory of
+$\Sets^\SCtx$ which contains just the pullbacks-preserving functors.
+
+\begin{thm}
+  A presheaf $\Of{X}{\Sets^{\SCtx}}$ is a sheaf on $\SCtxSite$ just when it
+  preserves pullbacks.
+\end{thm}
+
+\begin{proof}
+
+  The proof is essentially the same as that of the analogous lemma for the
+  Schanuel topos as presented in \cite[A.2.1.11.h]{johnstone:2002}, but we will
+  give a slightly more detailed version here. The presheaf $X$ is a sheaf on the
+  atomic site $\SCtxSite$ just when, for any renaming
+  $\Of{\varrho}{C\inj D}$ and any $\Member{n}{X(D)}$ if
+  $\IsEq{n\varrho_0}{n\varrho_1}$ for all diagrams
+  \[
+    \begin{tikzcd}[sep=large]
+      \IMode{C} \arrow[r, hook, "\IMode{\varrho}"] &
+      \IMode{D}
+        \arrow[r, hook, shift left, "\IMode{\varrho_0}"]
+        \arrow[r, hook, swap, shift right, "\IMode{\varrho_1}"]
+      &
+      \IMode{E}
+    \end{tikzcd}
+  \]
+  such that $\IsEq{\varrho_0\circ\varrho}{\varrho_1\circ\varrho}$, then there
+  exists a unique $\Member{m}{X(C)}$ such that $\IsEq{m\varrho}{n}$. In other
+  words, the arrow $X(\varrho)$ is an equalizer, as in the following:
+  \[
+    \begin{tikzcd}[sep=large]
+      \IMode{X(C)} \arrow[r, "\IMode{X(\varrho)}"] &
+      \IMode{X(D)}
+        \arrow[r, shift left, "\IMode{X(\varrho_0)}"]
+        \arrow[r, swap, shift right, "\IMode{X(\varrho_1)}"]&
+      \IMode{X(E)}
+      \\
+      \IMode{\mathbb{1}}
+        \arrow[u, densely dotted, "\OMode{m}"]
+        \arrow[ur, "\IMode{n}"]
+    \end{tikzcd}
+  \]
+
+  Now, because $\Sets$ is a regular category, $X(\varrho)$ is an equalizer
+  precisely when it is a monomorphism. Therefore, if we have assumed that $X$
+  preserves pullbacks, in order to demonstrate that $X$ is a sheaf, it suffices
+  to show that $X(\varrho)$ is monic. If $X$ preserves pullbacks, then it also
+  preserves monomorphisms, because in any category with pullbacks, $\Of{f}{A\ra
+  B}$ is monic just when the square
+  \[
+    \begin{tikzcd}[sep=large]
+      \IMode{A}
+        \arrow[r, "\IMode{\ArrId{A}}"]
+        \arrow[d, "\IMode{\ArrId{A}}"] &
+      \IMode{A} \arrow[d, "\IMode{f}"]\\
+      \IMode{A} \arrow[r, "\IMode{f}"] &
+      \IMode{B}
+    \end{tikzcd}
+  \]
+  is a pullback. Because $X$ preserves monomorphisms, and all arrows in $\SCtx$
+  are monic, then $X(\varrho)$ is monic. Hence, $X$ is a sheaf.
+
+  \todo[inline]{prove the other direction}
+
+\end{proof}
 

### 2015-12-04 paper.tex
slight improvement of notation

@@
 \numberwithin{equation}{section}
-%% Euler doesn't have \varsigma and \varrho
+%% Euler doesn't have \varsigma
 \DeclareSymbolFont{greekletters}{OML}{cmr}{m}{it}
-\DeclareMathSymbol{\varrho}{\mathalpha}{greekletters}{"25}
 \DeclareMathSymbol{\varsigma}{\mathalpha}{greekletters}{"26}
@@
   $\Of{\varrho}{C\inj D}$ and any $\Member{n}{X(D)}$ if
-  $\IsEq{n\varrho_0}{n\varrho_1}$ for all diagrams
+  $\IsEq{X(\varrho_0)(n)}{X(\varrho_1)(n)}$ for all diagrams
   \[
@@
   such that $\IsEq{\varrho_0\circ\varrho}{\varrho_1\circ\varrho}$, then there
-  exists a unique $\Member{m}{X(C)}$ such that $\IsEq{m\varrho}{n}$. In other
+  exists a unique $\Member{m}{X(C)}$ such that $\IsEq{X(\varrho)(n)}{n}$. In other
   words, the arrow $X(\varrho)$ is an equalizer, as in the following:

### 2015-12-04 paper.tex
typo

@@
 In general, we will have three kinds of context: symbolic (parameter) contexts,
-variable contexts, and metavaraible contexts The symbol contexts have already
+variable contexts, and metavariable contexts The symbol contexts have already
 been defined via the comma construction $\SCtx$, but they also admit a

### 2015-12-04 paper.tex
typo

@@
   such that $\IsEq{\varrho_0\circ\varrho}{\varrho_1\circ\varrho}$, then there
-  exists a unique $\Member{m}{X(C)}$ such that $\IsEq{X(\varrho)(n)}{n}$. In other
+  exists a unique $\Member{m}{X(C)}$ such that $\IsEq{X(\varrho)(m)}{n}$. In other
   words, the arrow $X(\varrho)$ is an equalizer, as in the following:

### 2015-12-04 paper.tex
correction to free symbols rules

@@
 Whereas the calculation of free variables pivoted on the \emph{var} rule, the
-calculation of free symbols will pivot on the \emph{app} rule, because the only
-place a symbol can be introduced is as a parameter to an operator.
+calculation of free symbols will pivot on the \emph{app} and \emph{mvar} rule,
+because the only way a symbol can be introduced is as a parameter to an
+operator or as a parameter to a metavariable.
 
@@
     }{
-      \bigcup_{i\leq n} \vec{u}_i
+      \vec{u} \cup \bigcup_{i\leq n} \vec{u}_i
     }

### 2015-12-04 paper.tex
Finish up the preserves pullbacks <=> atomic sheaf proof

@@
   \]
-  is a pullback. Because $X$ preserves monomorphisms, and all arrows in $\SCtx$
-  are monic, then $X(\varrho)$ is monic. Hence, $X$ is a sheaf.
-
-  \todo[inline]{prove the other direction}
+  is a pullback \cite[p.\ 16]{mac-lane-moerdijk:1992}. Because $X$ preserves
+  monomorphisms, and all arrows in $\SCtx$ are monic, then $X(\varrho)$ is
+  monic. Hence, $X$ is a sheaf.
+
+  Next, we must show that if $X$ is a sheaf, then it preserves pullbacks; we
+  will loosely track the proof of Proposition~2.6.15 in \cite{biering:2004}.
+  Fix the following diagram such that it is a pullback square in $\SCtx$:
+  \[
+    \begin{tikzcd}[sep=large]
+      \IMode{C}
+        \arrow[r, "\IMode{f_j}"]
+        \arrow[d, "\IMode{f_i}"] &
+      \IMode{C_j} \arrow[d, "\IMode{s_j}"]\\
+      \IMode{C_i} \arrow[r, "\IMode{s_i}"] &
+      \IMode{D}
+    \end{tikzcd}
+  \]
+
+  Then it suffices to show that the following diagram is also pullback square
+  in $\Sets$:
+  \[
+    \begin{tikzcd}[sep=large]
+      \IMode{X(C)}
+        \arrow[r, "\IMode{X(f_j)}"]
+        \arrow[d, "\IMode{X(f_i)}"] &
+      \IMode{X(C_j)} \arrow[d, "\IMode{X(s_j)}"]\\
+      \IMode{X(C_i)} \arrow[r, "\IMode{X(s_i)}"] &
+      \IMode{X(D)}
+    \end{tikzcd}
+  \]
+
+  To prove that the above diagram is a pullback, it suffices to show that for
+  any $\Member{m_i}{X(C_i)}$ and $\Member{m_j}{X(C_j)}$ such that
+  $\IsEq{X(s_i)(m_i)}{X(s_j)(m_j)}$, we have a unique $\Member{m}{X(C)}$; the
+  universal property of the pullback follows immediately from this.
+
+  Consider the singleton cover on $C$ that contains $\Define{h}{s_i\circ
+  f_i}=\OMode{s_j\circ f_j}$; a matching family $\phi$ for the cover
+  $\MkSet{h}$ is a map that takes $h$ to an object in $X(D)$, and an
+  amalgamation for $\phi$ consists in an object $\Member{m}{X(C)}$ such that
+  $\IsEq{\phi(h)}{X(h)(m)}$. Now recall that $X$ is a sheaf on $\SCtxSite$ if
+  and only if every matching family has a unique amalgamation; for the matching
+  family $\Define{\phi(h)}{X(s_i)(m_i)}$, then, we must have a unique amalgamation
+  $\Member{m}{X(C)}$ since $X$ is a sheaf. Therefore, $X$ preserves pullbacks.
 

### 2015-12-04 paper.tex
typo fixes; minor improvements

@@
 expect that we can work backward to a unique $\Of{m'}{\Upsilon}$ such that
-$\IsEq{m'\varrho}{m}$; whilst this is not in general the case for presheaves
-$X$, it is precisely the sheaf condition for covariant presheaves on $\SCtx$
-under the atomic topology \cite[p.~126]{mac-lane-moerdijk:1992}. The sheaf
-condition, then, ensures that the notion of support is well-behaved.
+$\IsEq{m'\varrho}{m}$; whilst it is not in general the case for presheaves $X$,
+this is precisely the sheaf condition for covariant presheaves on $\SCtx$ under
+the atomic topology \cite[p.~126]{mac-lane-moerdijk:1992}. The sheaf condition,
+then, ensures that the notion of support is well-behaved.
 
@@
     }
+    \MkSet{
+      \Upsilon
+      \mid
       \SupportsUnmoded{\Upsilon}{\varrho}{m}
+    }
   }
@@
 the category $\mathbf{Nom}_\mathfrak{s}$ of $\mathfrak{s}$-sorted nominal sets
-as the category of presheaves on the group of $\mathfrak{s}$-respecting
-permutations on $\mathbb{A}$, $\Perm_\mathfrak{s}\mathbb{A}$.
+as a subcategory of the category of presheaves on the group of
+$\mathfrak{s}$-respecting permutations on $\mathbb{A}$,
+$\Perm_\mathfrak{s}\mathbb{A}$.
 
@@
 Because this is a total relation, henceforth we will write $\FV{M}$ for
-$\vec{x}$ when $\CalcFV{M}{x}$.
+$\vec{x}$ when $\CalcFV{M}{\vec{x}}$.
 
@@
 Because this is a total relation, henceforth we will write $\FS{M}$ for
-$\vec{u}$ when $\CalcFS{M}{u}$.
+$\vec{u}$ when $\CalcFS{M}{\vec{u}}$.
 
@@
 \begin{gather*}
-  \infer{
+  \infer[\textit{var}]{
     \IsRenaming{\varrho}{x}{x}
@@
 \\[6pt]
-  \infer{
+  \infer[\textit{mvar}]{
     \IsRenaming{\varrho}{
@@
 \\[6pt]
-  \infer{
+  \infer[\textit{app}]{
     \IsRenaming{\varrho}{
@@
 \\[6pt]
-  \infer{
+  \infer[\textit{abs}]{
     \IsRenaming{\varrho}{
@@
 \begin{gather*}
-  \infer{
+  \infer[\textit{var}_1]{
     \IsSubst{N}{x}{y}{N}
@@
 \qquad
-  \infer{
+  \infer[\textit{var}_2]{
     \IsSubst{N}{x}{y}{y}
@@
 \\[6pt]
-  \infer{
+  \infer[\textit{mvar}]{
     \IsSubst{N}{x}{
@@
 \\[6pt]
-  \infer{
+  \infer[\textit{app}]{
     \IsSubst{N}{x}{
@@
 \begin{gather*}
-  \infer{
+  \infer[\textit{abs}_1]{
     \IsSubst{N}{x}{
@@
   }
-\qquad
-  \infer{
+\\[6pt]
+  \infer[\textit{abs}_2]{
     \IsSubst{N}{x}{
@@
 \begin{gather*}
-  \infer{
+  \infer[\textit{var}]{
     \IsSubst{E}{\MV{m}}{x}{x}
@@
 \\[6pt]
-  \infer{
+  \infer[\textit{mvar}_1]{
     \IsSubst{E}{\MV{m}}{
@@
 \\[6pt]
-  \infer{
+  \infer[\textit{mvar}_2]{
     \IsSubst{
@@
   %
-  \infer{
+  \infer[\textit{app}]{
     \IsSubst{E}{\MV{m}}{
@@
 \\[6pt]
-  \infer{
+  \infer[\textit{abs}]{
     \IsSubst{E}{\MV{m}}{

### 2015-12-05 paper.tex
typo
get rid of some brackets (this is concatenation, not pairing)
in => for

@@
 $\mathtt{get}[v]$. Going forward, we will write $\vartheta\varrho$ for
-$\Operators_a(\varrho)$.
+$\Operators_a(\varrho)(\vartheta)$.
 
@@
 
-The metavariable, symbol and variable contexts are interpreted in a model $P$ as
+The metavariable, symbol and variable contexts are interpreted for a model $P$ as
 an environment presheaf in the following way:
@@
             {
-              \,\Pair{
+              \,
                 \Tuple{\pi_u\circ\pi_1}_{u\in\Dom{\Upsilon}}
-              }{
+              ,
                 \,\Tuple{\pi_u\circ\pi_{[2,1]}}_{u\in\Dom{\vec{\sigma}_i}}
-              }
             }
@@
           }{
-            \,\Pair{
+            \,
               \Tuple{\pi_x\circ\pi_1}_{x\in\Dom{\Gamma}}
-            }{
+            ,
               \,\Tuple{\pi_x\circ\pi_{[2,1]}}_{x\in\Dom{\vec{\tau}_i}}
-            }
           }

### 2015-12-05 paper.tex
fix up outdated notation
fix typo

@@
 natural in $a$, such that
-$\IsOperator{\Upsilon'}{\Operators_a(\varrho)(\vartheta)}{a}$ when
+$\IsOperator{\Upsilon'}{\Operators(\varrho)(\vartheta)}{a}$ when
 $\IsOperator{\Upsilon}{\vartheta}{a}$. In particular, the renaming
@@
 $a$,
-$\Define{\Operators_a(\Upsilon\parallel\Gamma)}{\Operators\Pair{\Upsilon}{a}}$
+$\Define{\Operators_a(\Upsilon\parallel\Gamma)}{\Operators_a(\Upsilon)}$
 

### 2015-12-05 paper.tex
begin rearranging categorical machinery; explain stuff in detail
- moved most of the categorical machinery toward the beginning
- explained the comma construction notation in detail
- defined sites, sieves, matching families, amalgamations, separated
  presheaves and sheaves

@@
 
-\section{Preliminaries}
-
-Fix a set $\Sorts$ of \emph{sorts}. We will say $\IsSort{\tau}$ when
-$\Member{\tau}{\Sorts}$. A valence $\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}$
-specifies an expression of sort $\tau$ which binds symbols in $\vec{\sigma}$ and
-variables in $\vec{\tau}$.
-
-\[
-  \infer{
-    \IsValence{
-      \MkValence
-        {\sigma_0,\dots,\sigma_m}
-        {\tau_0,\dots,\tau_n}
-        {\tau}
-    }
-  }{
-    \IsSort{\tau}
-&
-    \HypJ{\IsSort{\sigma_i}}{\Leq{i}{m}}
-&
-    \HypJ{\IsSort{\tau_i}}{\Leq{i}{n}}
-  }
-\]
-
-An arity $\MkArity{\vec{v}}{\tau}$ specifies an operator of sort $\tau$ with
-arguments of valences $\vec{v}$. We will call the set of valences $\Valences$,
-and the set of arities $\Arities$.
-
-\[
-  \infer{
-    \IsArity{
-      \MkArity
-        {v_0,\ldots,v_n}
-        {\tau}
-    }
-  }{
-    \IsSort{\tau}
-&
-    \HypJ
-      {\IsValence{v_i}}
-      {\Leq{i}{n}}
-  }
-\]
-
-\subsection{Symbols, contexts and their sheaves}
+\section{Categorical Preliminaries}
+
+Fix a set $\Sorts$ of \emph{sorts}; we will say $\IsSort{\tau}$ when
+$\Member{\tau}{\Sorts}$. In this section, we will develop a theory of sets
+varying over collections of $\Sorts$-sorted symbols.
 
 Let $\SymSets$ be the category of finite cardinals and their injective maps;
-then the comma construction $\SCtx$, with $\Discrete{\Sorts}$
-the discrete category on the set $\Sorts$, is the category of contexts of
-symbols whose objects are finite sets of symbols $U$ and sort-assignments
+then the comma construction $\SCtx$, is the category of contexts of symbols
+whose objects are finite sets of symbols $U$ and sort-assignments
 \begin{tikzcd}[cramped]
@@
 
-Now, a covariant presheaf $X$ on $\SCtx$ is an intensional set which is subject
-to renamings
+\begin{remark}
+
+  To be precise, $\SCtx$ is an abuse of notation for the comma category
+  $\Dom{-}_\SymSets\downarrow \Delta[\Sorts]$, with $\Dom{-}_\SymSets$ the
+  forgetful functor from $\SymSets$ to $\Sets$, and
+  $\Define{\Delta[\Sorts](*)}{\Sorts}$ a constant functor from $\mathbb{1}$ to $\Sets$.
+  Formally, an object of $\SCtx$ is, then, a triple $\Tuple{U, *,
+  \mathfrak{s}}$ with $U$ an object in $\SymSets$, $*$ the unique object in
+  $\mathbb{1}$, and $\mathfrak{s}$ a function $\Dom{U}_\SymSets\ra
+  \Sorts$; an arrow $\Of{\varrho}{\Tuple{U,*,\mathfrak{s}}\ra
+  \Tuple{V,*,\mathfrak{t}}}$ is a commuting triangle like the following:
+  \[
+    \begin{tikzcd}[cramped]
+      \IMode{\Dom{U}_\SymSets}
+        \arrow[rr, "\IMode{\Dom{\rho}_\SymSets}"]
+        \arrow[dr, swap, "\IMode{\mathfrak{s}}"] &&
+      \IMode{\Dom{V}_\SymSets}
+      \arrow[dl, "\IMode{\mathfrak{t}}"] \\
+      &\IMode{\Sorts}
+    \end{tikzcd}
+  \]
+
+  Because the object $\Of{*}{\mathbb{1}}$ is unique, we will always omit it
+  from the data of an object in $\SCtx$.
+\end{remark}
+
+\paragraph{Presheaves on $\SCtx$}
+
+A presheaf on a category $\mathcal{C}$ is a contravariant functor
+$\OpCat{\mathcal{C}}\ra\Sets$; when $\Match{\mathcal{C}}{\OpCat{\mathcal{D}}}$,
+we call this a \emph{copresheaf} (covariant presheaf) on $\mathcal{D}$.
+Therefore, a copresheaf on $\SCtx$ is a functor $\SCtx\ra\Sets$; colloquially,
+such a copresheaf can be thought of as a set that \emph{varies over} symbol
+contexts.
+
+This leads to a very intuitive setting in which to consider sets
+that vary over collections of $\Sorts$-sorted symbols. In particular, for any
+morphism
+%
 \begin{tikzcd}[cramped]
   \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
-\end{tikzcd};
-that is, each element $\Member{m}{X(\Upsilon)}$ can be mapped to a unique
-element $\Member{m\varrho}{X(\Upsilon')}$.
+\end{tikzcd}
+%
+in $\SCtx$, we have a corresponding functorial renaming action on any
+copresheaf $\Of{X}{\Sets^\SCtx}$, namely $\Of{X(\varrho)}{X(\Upsilon)\ra
+X(\Upsilon')}$. When the copresheaf $X$ is clear from context, we will write
+$m\cdot\varrho$ for $X(\varrho)(m)$.
+
+Now, when $\Member{m}{X(\Upsilon)}$ for a copresheaf $\Of{X}{\Sets^\SCtx}$, we
+can define the notion that $m$ depends on \emph{at most} a subcontext
+$\Upsilon'$ of $\Upsilon$; this is called \emph{support}.
 
 \begin{definition}[Support]
-  For a presheaf $\Of{X}{\Sets^\SCtx}$, when
+  \label{def:support}
+  When
   \begin{tikzcd}[cramped]
@@
   \end{tikzcd},
-  if $\IsEq{\varrho_1\circ\varrho}{\varrho_2\circ\varrho}$ then $\IsEq{m\varrho_1}{m\varrho_2}$.
+  if $\IsEq{\varrho_1\circ\varrho}{\varrho_2\circ\varrho}$ then $\IsEq{m\cdot\varrho_1}{m\cdot\varrho_2}$.
 \end{definition}
 
-Intuitively, when $\Upsilon$ supports $\Member{m}{X(\Upsilon')}$, we would
-expect that we can work backward to a unique $\Of{m'}{\Upsilon}$ such that
-$\IsEq{m'\varrho}{m}$; whilst it is not in general the case for presheaves $X$,
-this is precisely the sheaf condition for covariant presheaves on $\SCtx$ under
-the atomic topology \cite[p.~126]{mac-lane-moerdijk:1992}. The sheaf condition,
-then, ensures that the notion of support is well-behaved.
-
-Let $\Supp{m}$ be the \emph{least support} of $\Member{m}{X(\Upsilon')}$:
+We can also always calculate the \emph{least support} of a term
+$\Member{m}{X(\Upsilon')}$, written $\Supp{m}$:
 \[
@@
 
+\subsection{Sheaves on the atomic site}
+
+Intuitively, when $\Upsilon$ supports $\Member{m}{X(\Upsilon')}$, we would
+expect that we can work backward to a unique $\Of{m'}{\Upsilon}$ such that
+$\IsEq{m'\cdot\varrho}{m}$; whilst it is not in general the case for presheaves
+$X$, this is precisely the sheaf condition for covariant presheaves on $\SCtx$
+under the atomic topology \cite[p.~126]{mac-lane-moerdijk:1992}. The sheaf
+condition, then, ensures that the notion of support is well-behaved.
+
+\paragraph{Sites and coverages}
+
+We will briefly discuss the basics of sheaves on a site. A \emph{site} is a
+pair $\Pair{\mathcal{C}}{J}$, with $\mathcal{C}$ a category of ``worlds'' and
+$J$ a Grothendieck topology (or ``coverage''). A coverage is a predicate that
+picks out for each object $\Of{C}{\mathcal{C}}$, the \emph{covering sieves} of
+$C$, subject to several axioms (see \cite{mac-lane-moerdijk:1992} for details).
+
+\paragraph{Covering sieves and matching families}
+
+A \emph{sieve} on an object $C$ is a subfunctor of the Yoneda embedding
+$\Define{\Yoneda{C}}{\Hom{\mathcal{C}}{-}{C}}$. We say that such a sieve $S$
+covers $C$ just when $\Member{S}{J(C)}$.  For a presheaf $X$ on the site
+$\Pair{\mathcal{C}}{J}$, a \emph{matching family} for the sieve $S$ is a
+natural transformation $\Of{\phi}{S\ra X}$. An \emph{amalgamation} for $\phi$
+is an object $\Member{m}{X(C)}$ such that $\IsEq{m\cdot f}{\phi_D(f)}$ for each
+$\Member{f}{S(D)}$, for all objects $D$.
+
+\paragraph{Separation and sheafhood}
+Then, we say that the presheaf $X$ is \emph{separated} in case for every sieve
+$\Member{S}{J(C)}$ and matching family $\Of{\phi}{S\ra X}$, there is at most
+one amalgamation $\Member{m}{X(C)}$ for $\phi$. Additionally, $X$ is called a
+\emph{sheaf} when there is \emph{exactly one} such amalgamation.
+
+\paragraph{The atomic coverage}
+
+The atomic coverage $J_\mathsf{atm}$ is the one in which all non-empty sieves
+cover. A sieve $S\sqsubseteq\Yoneda{C}$ is non-empty when the collection
+$\bigcup_D S(D)$ of all its arrows is non-empty.
+
 In the future, we will write $\SCtxSite$ for the Grothendieck site
-$\Pair{\OpCat{\left(\SCtx\right)}}{J_\mathsf{atm}}$, where $J_\mathsf{atm}$
-is the atomic coverage (sc.\ every non-empty family covers).
-
-\subsubsection{Relation to Nominal Sets and the Schanuel Topos}
-
-Pitts defines a category $\mathbf{Nom}$ of nominal sets in \cite{pitts:2013}
-which function equivalently to the sheaves that we considered above, in the
-case of $\IMode{\Sorts}\cong\IMode{\mathbb{1}}$. Fix a countably infinite set
-of atoms $\mathbb{A}$; then let $\Perm\mathbb{A}$ be the group of permutations
-(i.e.\ autoequivalences) on $\mathbb{A}$.  Considered as a category,
-$\Perm\mathbb{A}$ has a single object $\bullet$ with morphisms
-$\Of{\pi}{\bullet\ra\bullet}$ for each $\Member{\pi}{\Dom{\Perm\mathbb{A}}}$.
-Then, the category of nominal sets $\mathbf{Nom}$ is the subcategory of
-$\Sets^{\Perm\mathbb{A}}$ containing just the presheaves which satisfy a
-\emph{finite support} condition analogous to the one that we have formulated
-above.\todo[inline]{write it out}
-
-The category $\mathbf{Nom}$ is equivalent to the category of covariant sheaves
-on $\SymSets$ under the atomic coverage, called the Schanuel Topos
-$\Define{\mathbb{S}}{\Sheaves{\OpCat{\SymSets}}}$ \cite{pitts:2013,
-fiore-staton:2006}; equivalently, the Schanuel topos is the subcategory of
-$\Sets^\SymSets$ containing only pullbacks-preserving functors.
-
-In addition to the unisorted nominal sets framework, Pitts also briefly
-discusses a multi-sorted version of the apparatus in which the countably
-infinite set of atoms $\mathbb{A}$ is equipped with a sort assignment
-$\Of{\mathfrak{s}}{\mathbb{A}\to\Sorts}$ such that all the fibres of
-$\mathfrak{s}$ are countably infinite; then, he defines for any such assignment
-the category $\mathbf{Nom}_\mathfrak{s}$ of $\mathfrak{s}$-sorted nominal sets
-as a subcategory of the category of presheaves on the group of
-$\mathfrak{s}$-respecting permutations on $\mathbb{A}$,
-$\Perm_\mathfrak{s}\mathbb{A}$.
-
-In the same way as the Schanuel topos is equivalent to the unisorted nominal
-sets, we expect to find an equivalence between $\Sheaves{\SCtxSite}$ and the
-limit $\int_{\mathfrak{s}}\mathbf{Nom}_\mathfrak{s}$. At the very least,
-the sheaf topos $\Sheaves{\SCtxSite}$ may serve as a multi-sorted
-generalization of the Schanuel topos, for which we may carry over certain
-useful results, including its characterization as the subcategory of
-$\Sets^\SCtx$ which contains just the pullbacks-preserving functors.
-
-\begin{thm}
-  A presheaf $\Of{X}{\Sets^{\SCtx}}$ is a sheaf on $\SCtxSite$ just when it
-  preserves pullbacks.
-\end{thm}
-
-\begin{proof}
-
-  The proof is essentially the same as that of the analogous lemma for the
-  Schanuel topos as presented in \cite[A.2.1.11.h]{johnstone:2002}, but we will
-  give a slightly more detailed version here. The presheaf $X$ is a sheaf on the
-  atomic site $\SCtxSite$ just when, for any renaming
-  $\Of{\varrho}{C\inj D}$ and any $\Member{n}{X(D)}$ if
-  $\IsEq{X(\varrho_0)(n)}{X(\varrho_1)(n)}$ for all diagrams
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{C} \arrow[r, hook, "\IMode{\varrho}"] &
-      \IMode{D}
-        \arrow[r, hook, shift left, "\IMode{\varrho_0}"]
-        \arrow[r, hook, swap, shift right, "\IMode{\varrho_1}"]
-      &
-      \IMode{E}
+$\Pair{\OpCat{\left(\SCtx\right)}}{J_\mathsf{atm}}$, where $J_\mathsf{atm}$ is
+the atomic coverage. Sheafhood on the atomic site $\SCtxSite$ amounts to the
+following condition, using Definition~\ref{def:support}:
+\begin{quote}
+  A presheaf $X:\Sets^\SCtx$ is a sheaf on $\SCtxSite$ just when, for all $\Member{n}{X(\Upsilon')}$, if we have
+  \begin{tikzcd}[cramped]
+    \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
   \end{tikzcd}
-  \]
-  such that $\IsEq{\varrho_0\circ\varrho}{\varrho_1\circ\varrho}$, then there
-  exists a unique $\Member{m}{X(C)}$ such that $\IsEq{X(\varrho)(m)}{n}$. In other
-  words, the arrow $X(\varrho)$ is an equalizer, as in the following:
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{X(C)} \arrow[r, "\IMode{X(\varrho)}"] &
-      \IMode{X(D)}
-        \arrow[r, shift left, "\IMode{X(\varrho_0)}"]
-        \arrow[r, swap, shift right, "\IMode{X(\varrho_1)}"]&
-      \IMode{X(E)}
-      \\
-      \IMode{\mathbb{1}}
-        \arrow[u, densely dotted, "\OMode{m}"]
-        \arrow[ur, "\IMode{n}"]
-    \end{tikzcd}
-  \]
-
-  Now, because $\Sets$ is a regular category, $X(\varrho)$ is an equalizer
-  precisely when it is a monomorphism. Therefore, if we have assumed that $X$
-  preserves pullbacks, in order to demonstrate that $X$ is a sheaf, it suffices
-  to show that $X(\varrho)$ is monic. If $X$ preserves pullbacks, then it also
-  preserves monomorphisms, because in any category with pullbacks, $\Of{f}{A\ra
-  B}$ is monic just when the square
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{A}
-        \arrow[r, "\IMode{\ArrId{A}}"]
-        \arrow[d, "\IMode{\ArrId{A}}"] &
-      \IMode{A} \arrow[d, "\IMode{f}"]\\
-      \IMode{A} \arrow[r, "\IMode{f}"] &
-      \IMode{B}
-    \end{tikzcd}
-  \]
-  is a pullback \cite[p.\ 16]{mac-lane-moerdijk:1992}. Because $X$ preserves
-  monomorphisms, and all arrows in $\SCtx$ are monic, then $X(\varrho)$ is
-  monic. Hence, $X$ is a sheaf.
-
-  Next, we must show that if $X$ is a sheaf, then it preserves pullbacks; we
-  will loosely track the proof of Proposition~2.6.15 in \cite{biering:2004}.
-  Fix the following diagram such that it is a pullback square in $\SCtx$:
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{C}
-        \arrow[r, "\IMode{f_j}"]
-        \arrow[d, "\IMode{f_i}"] &
-      \IMode{C_j} \arrow[d, "\IMode{s_j}"]\\
-      \IMode{C_i} \arrow[r, "\IMode{s_i}"] &
-      \IMode{D}
-    \end{tikzcd}
-  \]
-
-  Then it suffices to show that the following diagram is also pullback square
-  in $\Sets$:
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{X(C)}
-        \arrow[r, "\IMode{X(f_j)}"]
-        \arrow[d, "\IMode{X(f_i)}"] &
-      \IMode{X(C_j)} \arrow[d, "\IMode{X(s_j)}"]\\
-      \IMode{X(C_i)} \arrow[r, "\IMode{X(s_i)}"] &
-      \IMode{X(D)}
-    \end{tikzcd}
-  \]
-
-  To prove that the above diagram is a pullback, it suffices to show that for
-  any $\Member{m_i}{X(C_i)}$ and $\Member{m_j}{X(C_j)}$ such that
-  $\IsEq{X(s_i)(m_i)}{X(s_j)(m_j)}$, we have a unique $\Member{m}{X(C)}$; the
-  universal property of the pullback follows immediately from this.
-
-  Consider the singleton cover on $C$ that contains $\Define{h}{s_i\circ
-  f_i}=\OMode{s_j\circ f_j}$; a matching family $\phi$ for the cover
-  $\MkSet{h}$ is a map that takes $h$ to an object in $X(D)$, and an
-  amalgamation for $\phi$ consists in an object $\Member{m}{X(C)}$ such that
-  $\IsEq{\phi(h)}{X(h)(m)}$. Now recall that $X$ is a sheaf on $\SCtxSite$ if
-  and only if every matching family has a unique amalgamation; for the matching
-  family $\Define{\phi(h)}{X(s_i)(m_i)}$, then, we must have a unique amalgamation
-  $\Member{m}{X(C)}$ since $X$ is a sheaf. Therefore, $X$ preserves pullbacks.
-
-\end{proof}
+  and $\Supports{\Upsilon}{\varrho}{n}$, then there is a unique
+  $\Member{m}{X(\Upsilon)}$ such that $\IsEq{m\cdot\varrho}{n}$.
+\end{quote}
+
+Therefore, a sheaf on $\SCtxSite$ is a set that varies over $\Sorts$-sorted
+symbols, which has a well-behaved notion of support. Another way to put it is,
+if $\Member{m}{X(\Upsilon)}$ only ``uses'' some of the symbols available at
+world $\Upsilon$, there's a canonical way to restrict $m$ to a term in the
+world which contains \emph{only} the symbols that it uses.
 
@@
 
-\subsection{Operators and signatures}
+\subsection{Relation to Nominal Sets and the Schanuel Topos}
+
+Pitts defines a category $\mathbf{Nom}$ of nominal sets in \cite{pitts:2013}
+which function equivalently to the sheaves that we considered above, in the
+case of $\IMode{\Sorts}\cong\IMode{\mathbb{1}}$. Fix a countably infinite set
+of atoms $\mathbb{A}$; then let $\Perm\mathbb{A}$ be the group of permutations
+(i.e.\ autoequivalences) on $\mathbb{A}$.  Considered as a category,
+$\Perm\mathbb{A}$ has a single object $\bullet$ with morphisms
+$\Of{\pi}{\bullet\ra\bullet}$ for each $\Member{\pi}{\Dom{\Perm\mathbb{A}}}$.
+Then, the category of nominal sets $\mathbf{Nom}$ is the subcategory of
+$\Sets^{\Perm\mathbb{A}}$ containing just the presheaves which satisfy a
+\emph{finite support} condition analogous to the one that we have formulated
+above.\todo[inline]{write it out}
+
+The category $\mathbf{Nom}$ is equivalent to the category of covariant sheaves
+on $\SymSets$ under the atomic coverage, called the Schanuel Topos
+$\Define{\mathbb{S}}{\Sheaves{\OpCat{\SymSets}}}$ \cite{pitts:2013,
+fiore-staton:2006}; equivalently, the Schanuel topos is the subcategory of
+$\Sets^\SymSets$ containing only pullbacks-preserving functors.
+
+In addition to the unisorted nominal sets framework, Pitts also briefly
+discusses a multi-sorted version of the apparatus in which the countably
+infinite set of atoms $\mathbb{A}$ is equipped with a sort assignment
+$\Of{\mathfrak{s}}{\mathbb{A}\to\Sorts}$ such that all the fibres of
+$\mathfrak{s}$ are countably infinite; then, he defines for any such assignment
+the category $\mathbf{Nom}_\mathfrak{s}$ of $\mathfrak{s}$-sorted nominal sets
+as a subcategory of the category of presheaves on the group of
+$\mathfrak{s}$-respecting permutations on $\mathbb{A}$,
+$\Perm_\mathfrak{s}\mathbb{A}$.
+
+In the same way as the Schanuel topos is equivalent to the unisorted nominal
+sets, we expect to find an equivalence between $\Sheaves{\SCtxSite}$ and the
+limit $\int_{\mathfrak{s}}\mathbf{Nom}_\mathfrak{s}$. At the very least,
+the sheaf topos $\Sheaves{\SCtxSite}$ may serve as a multi-sorted
+generalization of the Schanuel topos, for which we may carry over certain
+useful results, including its characterization as the subcategory of
+$\Sets^\SCtx$ which contains just the pullbacks-preserving functors.
+
+\begin{thm}
+  A presheaf $\Of{X}{\Sets^{\SCtx}}$ is a sheaf on $\SCtxSite$ just when it
+  preserves pullbacks.
+\end{thm}
+
+\begin{proof}
+
+  The proof is essentially the same as that of the analogous lemma for the
+  Schanuel topos as presented in \cite[A.2.1.11.h]{johnstone:2002}, but we will
+  give a slightly more detailed version here. The presheaf $X$ is a sheaf on the
+  atomic site $\SCtxSite$ just when, for any renaming
+  $\Of{\varrho}{C\inj D}$ and any $\Member{n}{X(D)}$ if
+  $\IsEq{X(\varrho_0)(n)}{X(\varrho_1)(n)}$ for all diagrams
+  \[
+    \begin{tikzcd}[sep=large]
+      \IMode{C} \arrow[r, hook, "\IMode{\varrho}"] &
+      \IMode{D}
+        \arrow[r, hook, shift left, "\IMode{\varrho_0}"]
+        \arrow[r, hook, swap, shift right, "\IMode{\varrho_1}"]
+      &
+      \IMode{E}
+    \end{tikzcd}
+  \]
+  such that $\IsEq{\varrho_0\circ\varrho}{\varrho_1\circ\varrho}$, then there
+  exists a unique $\Member{m}{X(C)}$ such that $\IsEq{X(\varrho)(m)}{n}$. In other
+  words, the arrow $X(\varrho)$ is an equalizer, as in the following:
+  \[
+    \begin{tikzcd}[sep=large]
+      \IMode{X(C)} \arrow[r, "\IMode{X(\varrho)}"] &
+      \IMode{X(D)}
+        \arrow[r, shift left, "\IMode{X(\varrho_0)}"]
+        \arrow[r, swap, shift right, "\IMode{X(\varrho_1)}"]&
+      \IMode{X(E)}
+      \\
+      \IMode{\mathbb{1}}
+        \arrow[u, densely dotted, "\OMode{m}"]
+        \arrow[ur, "\IMode{n}"]
+    \end{tikzcd}
+  \]
+
+  Now, because $\Sets$ is a regular category, $X(\varrho)$ is an equalizer
+  precisely when it is a monomorphism. Therefore, if we have assumed that $X$
+  preserves pullbacks, in order to demonstrate that $X$ is a sheaf, it suffices
+  to show that $X(\varrho)$ is monic. If $X$ preserves pullbacks, then it also
+  preserves monomorphisms, because in any category with pullbacks, $\Of{f}{A\ra
+  B}$ is monic just when the square
+  \[
+    \begin{tikzcd}[sep=large]
+      \IMode{A}
+        \arrow[r, "\IMode{\ArrId{A}}"]
+        \arrow[d, "\IMode{\ArrId{A}}"] &
+      \IMode{A} \arrow[d, "\IMode{f}"]\\
+      \IMode{A} \arrow[r, "\IMode{f}"] &
+      \IMode{B}
+    \end{tikzcd}
+  \]
+  is a pullback \cite[p.\ 16]{mac-lane-moerdijk:1992}. Because $X$ preserves
+  monomorphisms, and all arrows in $\SCtx$ are monic, then $X(\varrho)$ is
+  monic. Hence, $X$ is a sheaf.
+
+  Next, we must show that if $X$ is a sheaf, then it preserves pullbacks; we
+  will loosely track the proof of Proposition~2.6.15 in \cite{biering:2004}.
+  Fix the following diagram such that it is a pullback square in $\SCtx$:
+  \[
+    \begin{tikzcd}[sep=large]
+      \IMode{C}
+        \arrow[r, "\IMode{f_j}"]
+        \arrow[d, "\IMode{f_i}"] &
+      \IMode{C_j} \arrow[d, "\IMode{s_j}"]\\
+      \IMode{C_i} \arrow[r, "\IMode{s_i}"] &
+      \IMode{D}
+    \end{tikzcd}
+  \]
+
+  Then it suffices to show that the following diagram is also pullback square
+  in $\Sets$:
+  \[
+    \begin{tikzcd}[sep=large]
+      \IMode{X(C)}
+        \arrow[r, "\IMode{X(f_j)}"]
+        \arrow[d, "\IMode{X(f_i)}"] &
+      \IMode{X(C_j)} \arrow[d, "\IMode{X(s_j)}"]\\
+      \IMode{X(C_i)} \arrow[r, "\IMode{X(s_i)}"] &
+      \IMode{X(D)}
+    \end{tikzcd}
+  \]
+
+  To prove that the above diagram is a pullback, it suffices to show that for
+  any $\Member{m_i}{X(C_i)}$ and $\Member{m_j}{X(C_j)}$ such that
+  $\IsEq{X(s_i)(m_i)}{X(s_j)(m_j)}$, we have a unique $\Member{m}{X(C)}$; the
+  universal property of the pullback follows immediately from this.
+
+  Consider the singleton cover on $C$ that contains $\Define{h}{s_i\circ
+  f_i}=\OMode{s_j\circ f_j}$; a matching family $\phi$ for the cover
+  $\MkSet{h}$ is a map that takes $h$ to an object in $X(D)$, and an
+  amalgamation for $\phi$ consists in an object $\Member{m}{X(C)}$ such that
+  $\IsEq{\phi(h)}{X(h)(m)}$. Now recall that $X$ is a sheaf on $\SCtxSite$ if
+  and only if every matching family has a unique amalgamation; for the matching
+  family $\Define{\phi(h)}{X(s_i)(m_i)}$, then, we must have a unique amalgamation
+  $\Member{m}{X(C)}$ since $X$ is a sheaf. Therefore, $X$ preserves pullbacks.
+
+\end{proof}
+
+\section{Operators and signatures}
+
+A valence $\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}$ specifies an expression
+of sort $\tau$ which binds symbols in $\vec{\sigma}$ and variables in
+$\vec{\tau}$.
+
+\[
+  \infer{
+    \IsValence{
+      \MkValence
+        {\sigma_0,\dots,\sigma_m}
+        {\tau_0,\dots,\tau_n}
+        {\tau}
+    }
+  }{
+    \IsSort{\tau}
+&
+    \HypJ{\IsSort{\sigma_i}}{\Leq{i}{m}}
+&
+    \HypJ{\IsSort{\tau_i}}{\Leq{i}{n}}
+  }
+\]
+
+An arity $\MkArity{\vec{v}}{\tau}$ specifies an operator of sort $\tau$ with
+arguments of valences $\vec{v}$. We will call the set of valences $\Valences$,
+and the set of arities $\Arities$.
+
+\[
+  \infer{
+    \IsArity{
+      \MkArity
+        {v_0,\ldots,v_n}
+        {\tau}
+    }
+  }{
+    \IsSort{\tau}
+&
+    \HypJ
+      {\IsValence{v_i}}
+      {\Leq{i}{n}}
+  }
+\]
 

### 2015-12-05 paper.tex
fix mode annotations

@@
 The atomic coverage $J_\mathsf{atm}$ is the one in which all non-empty sieves
-cover. A sieve $S\sqsubseteq\Yoneda{C}$ is non-empty when the collection
-$\bigcup_D S(D)$ of all its arrows is non-empty.
+cover. A sieve $\IMode{S}\sqsubseteq\IMode{\Yoneda{C}}$ is non-empty when the
+collection $\bigcup_D S(D)$ of all its arrows is non-empty.
 
@@
 \begin{quote}
-  A presheaf $X:\Sets^\SCtx$ is a sheaf on $\SCtxSite$ just when, for all $\Member{n}{X(\Upsilon')}$, if we have
+  A presheaf $\Of{X}{\Sets^\SCtx}$ is a sheaf on $\SCtxSite$ just when, for all $\Member{n}{X(\Upsilon')}$, if we have
   \begin{tikzcd}[cramped]

### 2015-12-05 paper.tex
change section level
notation; remove TODO

@@
 
-\subsubsection{Constructions on sheaves}
+\subsection{Constructions on sheaves}
 
@@
 
-We also have a canonical map $\Of{\eta}{X\ra X^+}$, defined
+We also have a canonical natural transformation $\Of{\eta}{X\ra X^+}$, defined
 using the maximal sieve $\Define{t_C}{\Yoneda{C}}$:
@@
 $\Sets^{\Perm\mathbb{A}}$ containing just the presheaves which satisfy a
-\emph{finite support} condition analogous to the one that we have formulated
-above.\todo[inline]{write it out}
+\emph{finite support} condition.
 
@@
   $\Of{\varrho}{C\inj D}$ and any $\Member{n}{X(D)}$ if
-  $\IsEq{X(\varrho_0)(n)}{X(\varrho_1)(n)}$ for all diagrams
+  $\IsEq{n\cdot \varrho_0}{n\cdot \varrho_1}$ for all diagrams
   \[
@@
   such that $\IsEq{\varrho_0\circ\varrho}{\varrho_1\circ\varrho}$, then there
-  exists a unique $\Member{m}{X(C)}$ such that $\IsEq{X(\varrho)(m)}{n}$. In other
+  exists a unique $\Member{m}{X(C)}$ such that $\IsEq{m\cdot\varrho}{n}$. In other
   words, the arrow $X(\varrho)$ is an equalizer, as in the following:
@@
 In general, we will have three kinds of context: symbolic (parameter) contexts,
-variable contexts, and metavariable contexts The symbol contexts have already
+variable contexts, and metavariable contexts. The symbol contexts have already
 been defined via the comma construction $\SCtx$, but they also admit a

### 2015-12-06 paper.tex
move a proof into the appendix

@@
 
+\usepackage{appendix}
 \usepackage{amsmath}
@@
 sets, we expect to find an equivalence between $\Sheaves{\SCtxSite}$ and the
-limit $\int_{\mathfrak{s}}\mathbf{Nom}_\mathfrak{s}$. At the very least,
-the sheaf topos $\Sheaves{\SCtxSite}$ may serve as a multi-sorted
-generalization of the Schanuel topos, for which we may carry over certain
-useful results, including its characterization as the subcategory of
-$\Sets^\SCtx$ which contains just the pullbacks-preserving functors.
-
-\begin{thm}
-  A presheaf $\Of{X}{\Sets^{\SCtx}}$ is a sheaf on $\SCtxSite$ just when it
-  preserves pullbacks.
-\end{thm}
-
-\begin{proof}
-
-  The proof is essentially the same as that of the analogous lemma for the
-  Schanuel topos as presented in \cite[A.2.1.11.h]{johnstone:2002}, but we will
-  give a slightly more detailed version here. The presheaf $X$ is a sheaf on the
-  atomic site $\SCtxSite$ just when, for any renaming
-  $\Of{\varrho}{C\inj D}$ and any $\Member{n}{X(D)}$ if
-  $\IsEq{n\cdot \varrho_0}{n\cdot \varrho_1}$ for all diagrams
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{C} \arrow[r, hook, "\IMode{\varrho}"] &
-      \IMode{D}
-        \arrow[r, hook, shift left, "\IMode{\varrho_0}"]
-        \arrow[r, hook, swap, shift right, "\IMode{\varrho_1}"]
-      &
-      \IMode{E}
-    \end{tikzcd}
-  \]
-  such that $\IsEq{\varrho_0\circ\varrho}{\varrho_1\circ\varrho}$, then there
-  exists a unique $\Member{m}{X(C)}$ such that $\IsEq{m\cdot\varrho}{n}$. In other
-  words, the arrow $X(\varrho)$ is an equalizer, as in the following:
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{X(C)} \arrow[r, "\IMode{X(\varrho)}"] &
-      \IMode{X(D)}
-        \arrow[r, shift left, "\IMode{X(\varrho_0)}"]
-        \arrow[r, swap, shift right, "\IMode{X(\varrho_1)}"]&
-      \IMode{X(E)}
-      \\
-      \IMode{\mathbb{1}}
-        \arrow[u, densely dotted, "\OMode{m}"]
-        \arrow[ur, "\IMode{n}"]
-    \end{tikzcd}
-  \]
-
-  Now, because $\Sets$ is a regular category, $X(\varrho)$ is an equalizer
-  precisely when it is a monomorphism. Therefore, if we have assumed that $X$
-  preserves pullbacks, in order to demonstrate that $X$ is a sheaf, it suffices
-  to show that $X(\varrho)$ is monic. If $X$ preserves pullbacks, then it also
-  preserves monomorphisms, because in any category with pullbacks, $\Of{f}{A\ra
-  B}$ is monic just when the square
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{A}
-        \arrow[r, "\IMode{\ArrId{A}}"]
-        \arrow[d, "\IMode{\ArrId{A}}"] &
-      \IMode{A} \arrow[d, "\IMode{f}"]\\
-      \IMode{A} \arrow[r, "\IMode{f}"] &
-      \IMode{B}
-    \end{tikzcd}
-  \]
-  is a pullback \cite[p.\ 16]{mac-lane-moerdijk:1992}. Because $X$ preserves
-  monomorphisms, and all arrows in $\SCtx$ are monic, then $X(\varrho)$ is
-  monic. Hence, $X$ is a sheaf.
-
-  Next, we must show that if $X$ is a sheaf, then it preserves pullbacks; we
-  will loosely track the proof of Proposition~2.6.15 in \cite{biering:2004}.
-  Fix the following diagram such that it is a pullback square in $\SCtx$:
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{C}
-        \arrow[r, "\IMode{f_j}"]
-        \arrow[d, "\IMode{f_i}"] &
-      \IMode{C_j} \arrow[d, "\IMode{s_j}"]\\
-      \IMode{C_i} \arrow[r, "\IMode{s_i}"] &
-      \IMode{D}
-    \end{tikzcd}
-  \]
-
-  Then it suffices to show that the following diagram is also pullback square
-  in $\Sets$:
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{X(C)}
-        \arrow[r, "\IMode{X(f_j)}"]
-        \arrow[d, "\IMode{X(f_i)}"] &
-      \IMode{X(C_j)} \arrow[d, "\IMode{X(s_j)}"]\\
-      \IMode{X(C_i)} \arrow[r, "\IMode{X(s_i)}"] &
-      \IMode{X(D)}
-    \end{tikzcd}
-  \]
-
-  To prove that the above diagram is a pullback, it suffices to show that for
-  any $\Member{m_i}{X(C_i)}$ and $\Member{m_j}{X(C_j)}$ such that
-  $\IsEq{X(s_i)(m_i)}{X(s_j)(m_j)}$, we have a unique $\Member{m}{X(C)}$; the
-  universal property of the pullback follows immediately from this.
-
-  Consider the singleton cover on $C$ that contains $\Define{h}{s_i\circ
-  f_i}=\OMode{s_j\circ f_j}$; a matching family $\phi$ for the cover
-  $\MkSet{h}$ is a map that takes $h$ to an object in $X(D)$, and an
-  amalgamation for $\phi$ consists in an object $\Member{m}{X(C)}$ such that
-  $\IsEq{\phi(h)}{X(h)(m)}$. Now recall that $X$ is a sheaf on $\SCtxSite$ if
-  and only if every matching family has a unique amalgamation; for the matching
-  family $\Define{\phi(h)}{X(s_i)(m_i)}$, then, we must have a unique amalgamation
-  $\Member{m}{X(C)}$ since $X$ is a sheaf. Therefore, $X$ preserves pullbacks.
-
-\end{proof}
+limit $\int_{\mathfrak{s}}\mathbf{Nom}_\mathfrak{s}$. At the very least, the
+sheaf topos $\Sheaves{\SCtxSite}$ may serve as a multi-sorted generalization of
+the Schanuel topos, for which we may carry over certain useful results,
+including its characterization as the subcategory of $\Sets^\SCtx$ which
+contains just the pullbacks-preserving functors (see
+Theorem~\ref{thm:pullback-preservation} for the proof).
 
@@
 
+\newpage
+\appendix
+\appendixpage
+
+\section{Sheafhood on $\SCtxSite$}
+
+\begin{thm}
+  \label{thm:pullback-preservation}
+  A presheaf $\Of{X}{\Sets^{\SCtx}}$ is a sheaf on $\SCtxSite$ just when it
+  preserves pullbacks.
+\end{thm}
+
+\begin{proof}
+
+  The proof is essentially the same as that of the analogous lemma for the
+  Schanuel topos as presented in \cite[A.2.1.11.h]{johnstone:2002}, but we will
+  give a slightly more detailed version here. The presheaf $X$ is a sheaf on the
+  atomic site $\SCtxSite$ just when, for any renaming
+  $\Of{\varrho}{C\inj D}$ and any $\Member{n}{X(D)}$ if
+  $\IsEq{n\cdot \varrho_0}{n\cdot \varrho_1}$ for all diagrams
+  \[
+    \begin{tikzcd}[sep=large]
+      \IMode{C} \arrow[r, hook, "\IMode{\varrho}"] &
+      \IMode{D}
+        \arrow[r, hook, shift left, "\IMode{\varrho_0}"]
+        \arrow[r, hook, swap, shift right, "\IMode{\varrho_1}"]
+      &
+      \IMode{E}
+    \end{tikzcd}
+  \]
+  such that $\IsEq{\varrho_0\circ\varrho}{\varrho_1\circ\varrho}$, then there
+  exists a unique $\Member{m}{X(C)}$ such that $\IsEq{m\cdot\varrho}{n}$. In other
+  words, the arrow $X(\varrho)$ is an equalizer, as in the following:
+  \[
+    \begin{tikzcd}[sep=large]
+      \IMode{X(C)} \arrow[r, "\IMode{X(\varrho)}"] &
+      \IMode{X(D)}
+        \arrow[r, shift left, "\IMode{X(\varrho_0)}"]
+        \arrow[r, swap, shift right, "\IMode{X(\varrho_1)}"]&
+      \IMode{X(E)}
+      \\
+      \IMode{\mathbb{1}}
+        \arrow[u, densely dotted, "\OMode{m}"]
+        \arrow[ur, "\IMode{n}"]
+    \end{tikzcd}
+  \]
+
+  Now, because $\Sets$ is a regular category, $X(\varrho)$ is an equalizer
+  precisely when it is a monomorphism. Therefore, if we have assumed that $X$
+  preserves pullbacks, in order to demonstrate that $X$ is a sheaf, it suffices
+  to show that $X(\varrho)$ is monic. If $X$ preserves pullbacks, then it also
+  preserves monomorphisms, because in any category with pullbacks, $\Of{f}{A\ra
+  B}$ is monic just when the square
+  \[
+    \begin{tikzcd}[sep=large]
+      \IMode{A}
+        \arrow[r, "\IMode{\ArrId{A}}"]
+        \arrow[d, "\IMode{\ArrId{A}}"] &
+      \IMode{A} \arrow[d, "\IMode{f}"]\\
+      \IMode{A} \arrow[r, "\IMode{f}"] &
+      \IMode{B}
+    \end{tikzcd}
+  \]
+  is a pullback \cite[p.\ 16]{mac-lane-moerdijk:1992}. Because $X$ preserves
+  monomorphisms, and all arrows in $\SCtx$ are monic, then $X(\varrho)$ is
+  monic. Hence, $X$ is a sheaf.
+
+  Next, we must show that if $X$ is a sheaf, then it preserves pullbacks; we
+  will loosely track the proof of Proposition~2.6.15 in \cite{biering:2004}.
+  Fix the following diagram such that it is a pullback square in $\SCtx$:
+  \[
+    \begin{tikzcd}[sep=large]
+      \IMode{C}
+        \arrow[r, "\IMode{f_j}"]
+        \arrow[d, "\IMode{f_i}"] &
+      \IMode{C_j} \arrow[d, "\IMode{s_j}"]\\
+      \IMode{C_i} \arrow[r, "\IMode{s_i}"] &
+      \IMode{D}
+    \end{tikzcd}
+  \]
+
+  Then it suffices to show that the following diagram is also pullback square
+  in $\Sets$:
+  \[
+    \begin{tikzcd}[sep=large]
+      \IMode{X(C)}
+        \arrow[r, "\IMode{X(f_j)}"]
+        \arrow[d, "\IMode{X(f_i)}"] &
+      \IMode{X(C_j)} \arrow[d, "\IMode{X(s_j)}"]\\
+      \IMode{X(C_i)} \arrow[r, "\IMode{X(s_i)}"] &
+      \IMode{X(D)}
+    \end{tikzcd}
+  \]
+
+  To prove that the above diagram is a pullback, it suffices to show that for
+  any $\Member{m_i}{X(C_i)}$ and $\Member{m_j}{X(C_j)}$ such that
+  $\IsEq{X(s_i)(m_i)}{X(s_j)(m_j)}$, we have a unique $\Member{m}{X(C)}$; the
+  universal property of the pullback follows immediately from this.
+
+  Consider the singleton cover on $C$ that contains $\Define{h}{s_i\circ
+  f_i}=\OMode{s_j\circ f_j}$; a matching family $\phi$ for the cover
+  $\MkSet{h}$ is a map that takes $h$ to an object in $X(D)$, and an
+  amalgamation for $\phi$ consists in an object $\Member{m}{X(C)}$ such that
+  $\IsEq{\phi(h)}{X(h)(m)}$. Now recall that $X$ is a sheaf on $\SCtxSite$ if
+  and only if every matching family has a unique amalgamation; for the matching
+  family $\Define{\phi(h)}{X(s_i)(m_i)}$, then, we must have a unique amalgamation
+  $\Member{m}{X(C)}$ since $X$ is a sheaf. Therefore, $X$ preserves pullbacks.
+
+\end{proof}
+
 \ifdraft{}{
-  \nocite{*}
+  \newpage
+  % \nocite{*}
   \bibliographystyle{abbrv}

### 2015-12-06 paper.tex
layout
use copresheaf
ehh, I changed my mind.

@@
 $\IsEq{m'\cdot\varrho}{m}$; whilst it is not in general the case for presheaves
-$X$, this is precisely the sheaf condition for covariant presheaves on $\SCtx$
+$X$, this is precisely the sheaf condition for copresheaves on $\SCtx$
 under the atomic topology \cite[p.~126]{mac-lane-moerdijk:1992}. The sheaf
@@
 
-\begin{gather*}
+\[
   \infer[\textit{var}]{
@@
   }
-\\[6pt]
+\]
+\[
   \infer[\textit{mvar}]{
@@
   }
-\\[6pt]
+\]
+\[
   \infer[\textit{app}]{
@@
   }
-\\[6pt]
+\]
+\[
   \infer[\textit{abs}]{
@@
   }
-\end{gather*}
+\]
 

### 2015-12-06 paper.tex
Explain "presupposition"

@@
 Let the judgment $\IsAbt{\Theta}{\Upsilon}{\Gamma}{M}{\tau}$ presuppose
+\footnote{
+  In the \emph{judgmental method}, as pioneered by Per Martin-L\"of, a form of
+  judgment is propounded first by specifying its range of significance (i.e.\
+  the circumstances under which it shall have a meaning), and then, a meaning
+  explanation (definition) for the judgment is given. The range of significance
+  of a judgment is called its \emph{presupposition}, and the meaning of the
+  judgment may proceed by induction on the evidence that the presupposition
+  obtains. See \cite{schroeder-heister:1987} for a more detailed explanation of
+  presuppositions and the judgmental method.
+}
 $\IsMetaCtx{\Theta}$, $\IsSymCtx\Upsilon$, $\IsVarCtx\Gamma$ and

### 2015-12-09 paper.tex
mention sml implementation

@@
 the categorical semantics for abts formally in Constructive Type Theory using
-the Agda proof assistant~\cite{norell:2007}.
+the Agda proof assistant~\cite{norell:2007}; we have also developed an
+implementation for abstract binding trees in Standard ML, which we intend to
+integrate into the JonPRL proof assistant \cite{jonprl:2015}.
 

### 2015-12-12 paper.tex
Add brief acknowledgement
Fix up facile proof

@@
 
+\section*{Acknowledgements}
+
+The first author wishes to thank Robert Harper for numerous conversations about
+abstract binding trees and symbolic parameters, and Andy Pitts for his help in
+understanding multi-sorted nominal sets and atomic sheaves.
+
 \newpage
@@
 
-  To prove that the above diagram is a pullback, it suffices to show that for
-  any $\Member{m_i}{X(C_i)}$ and $\Member{m_j}{X(C_j)}$ such that
-  $\IsEq{X(s_i)(m_i)}{X(s_j)(m_j)}$, we have a unique $\Member{m}{X(C)}$; the
-  universal property of the pullback follows immediately from this.
-
-  Consider the singleton cover on $C$ that contains $\Define{h}{s_i\circ
-  f_i}=\OMode{s_j\circ f_j}$; a matching family $\phi$ for the cover
-  $\MkSet{h}$ is a map that takes $h$ to an object in $X(D)$, and an
-  amalgamation for $\phi$ consists in an object $\Member{m}{X(C)}$ such that
-  $\IsEq{\phi(h)}{X(h)(m)}$. Now recall that $X$ is a sheaf on $\SCtxSite$ if
-  and only if every matching family has a unique amalgamation; for the matching
-  family $\Define{\phi(h)}{X(s_i)(m_i)}$, then, we must have a unique amalgamation
+  We have to show that $X(C)$ is, up to equivalence, the pullback of $X(s_j)$
+  along $X(s_j)$ in $\Sets$, namely:
+  \[
+    \IMode{X(C_i)\times_{X(D)}X(C_j)}
+      \cong
+    \OMode{
+      \MkSet{
+        (m_i, m_j)\in X(C_i)\times X(C_j)
+          \mid
+        m_i \cdot s_i = m_j \cdot s_j
+      }
+    }
+  \]
+  The first direction of the equivalence is trivial: for each
+  $\Member{m}{X(C)}$, we can easily exhibit $\Member{(m\cdot f_i,m\cdot
+  f_j)}{X(C_i)\times_{X(D)}X(C_j)}$.
+
+  For the other direction, fix $\Member{(m_i,
+  m_j)}{X(C_i)\times_{X(D)}X(C_j)}$.  Consider the singleton cover on $C$ that
+  contains $\Define{h}{s_i\circ f_i}=\OMode{s_j\circ f_j}$; a matching family
+  $\phi$ for the cover $\MkSet{h}$ is a map that takes $h$ to an object in
+  $X(D)$, and an amalgamation for $\phi$ consists in an object
+  $\Member{m}{X(C)}$ such that $\IsEq{\phi(h)}{X(h)(m)}$. Now recall that $X$
+  is a sheaf on $\SCtxSite$ if and only if every matching family has a unique
+  amalgamation; for the matching family $\Define{\phi(h)}{m_i\cdot
+  s_i}=\OMode{m_j\cdot s_j}$, then, we must have a unique amalgamation
   $\Member{m}{X(C)}$ since $X$ is a sheaf. Therefore, $X$ preserves pullbacks.

### 2015-12-12 paper.tex
typo
notation

@@
 
-  We have to show that $X(C)$ is, up to equivalence, the pullback of $X(s_j)$
+  We have to show that $X(C)$ is, up to equivalence, the pullback of $X(s_i)$
   along $X(s_j)$ in $\Sets$, namely:
@@
   $X(D)$, and an amalgamation for $\phi$ consists in an object
-  $\Member{m}{X(C)}$ such that $\IsEq{\phi(h)}{X(h)(m)}$. Now recall that $X$
+  $\Member{m}{X(C)}$ such that $\IsEq{\phi(h)}{m\cdot h}$. Now recall that $X$
   is a sheaf on $\SCtxSite$ if and only if every matching family has a unique

### 2015-12-12 paper.tex
Add further motivation for sequent case study

@@
 
+There are two reasons that one might want to model sequents in a (meta)-logical
+framework. Perhaps the most obvious reason is to prove things about a logic
+with sequent judgments; this is what motivated Crary's encoding of contexts in
+LF, for instance. However, another common use-case is the practical development
+of a proof assistant in which users can define new sequent rules; such
+user-defined rules may only be compiled into well-behaved tactics in the proof
+system in case they obey certain wellformedness properties including (among
+other things) correct scoping of names. MetaPRL is a salient example of such a
+proof assistant, whose implementation might have benefited from the more
+principled treatment of symbols and metavariables which we present here.
+
 This insight leads the way to a simple abt signature for the theory of

### 2015-12-12 paper.tex
Clean up refinements section
get rid of refinements for metavariables

@@
 
-\newcommand\IsWf[5]{%
-  \IMode{#1}\triangleright%
-  \IMode{#2}\parallel%
-  \IMode{#3}\vdash%
-  \IMode{#4}\in_{\mathbf{wf}}\IMode{#5}%
+\newcommand\IsWf[4]{%
+  \IMode{#1}\parallel%
+  \IMode{#2}\vdash%
+  \IMode{#3}\in_{\mathbf{wf}}\IMode{#4}%
 }
@@
 
-Now, we have the following well-formed sequent:
+Now, we can write the following sequent:
 \[
@@
 \emph{inductive refinement}~\cite{harper:2016}. The basic idea is to introduce a
-new form of (meta)-judgment $\IsWf{\Theta}{\Upsilon}{\Gamma}{M}{\tau}$ which
+new form of (meta)-judgment $\IsWf{\Upsilon}{\Gamma}{M}{\tau}$ which
 expresses the extrinsic wellformedness properties we wish to verify,
-presupposing $\IsAbt{\Theta}{\Upsilon}{\Gamma}{M}{\tau}$. Additionally, we
+presupposing $\IsAbt{\cdot}{\Upsilon}{\Gamma}{M}{\tau}$. Additionally, we
 introduce an analogous judgment on bound terms,
-$\IsWf{\Theta}{\Upsilon}{\Gamma}{E}{v}$ presupposing
-$\IsAbt{\Theta}{\Upsilon}{\Gamma}{E}{v}$, defined uniformly as follows:
+$\IsWf{\Upsilon}{\Gamma}{E}{v}$ presupposing
+$\IsAbt{\cdot}{\Upsilon}{\Gamma}{E}{v}$, defined uniformly as follows:
 \[
   \infer{
-    \IsWf{\Theta}{\Upsilon}{\Gamma}{
+    \IsWf{\Upsilon}{\Gamma}{
       \MkBTm{\vec{u}}{\vec{x}}{M}
@@
   }{
-    \IsWf{\Theta}{\Upsilon,\vec{u}:\vec{\sigma}}{\Gamma,\vec{x}:\vec{\tau}}{M}{\tau}
-  }
-\]
-
-Likewise, wellformedness for variables and metavariables is defined uniformly:
-\begin{gather*}
+    \IsWf{\Upsilon,\vec{u}:\vec{\sigma}}{\Gamma,\vec{x}:\vec{\tau}}{M}{\tau}
+  }
+\]
+
+Likewise, wellformedness for variables is defined uniformly:
+\[
   \infer{
-    \IsWf{\Theta}{\Upsilon}{\Gamma}{x}{\tau}
-  }{
-  }
-\qquad
-  \infer{
-    \IsWf{\Theta}{\Upsilon}{\Gamma}{
-      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
-    }{\tau}
-  }{
-    \begin{array}{l}
-      \Lookup{\Theta}{\MV{m}}{\MkValence{\vec{\sigma}}{\tau_0,\dots,\tau_n}{\tau}}
-\\
-      \HypJ{
-        \IsWf{\Theta}{\Upsilon}{\Gamma}{M_i}{\tau_i}
-      }{
-        \Leq{i}{n}
-      }
-    \end{array}
-  }
-\end{gather*}
-
-\begin{remark}
+    \IsWf{\Upsilon}{\Gamma}{x}{\tau}
+  }{
+  }
+\]
 Note that the refinement for variables $x$ is not trivial, since it is only
-  defined in case the presupposition $\IsAbt{\Theta}{\Upsilon}{\Gamma}{x}{\tau}$
+defined in case the presupposition $\IsAbt{\cdot}{\Upsilon}{\Gamma}{x}{\tau}$
 is satisfied.
-\end{remark}
 
@@
   ]{
-    \IsWf{\Theta}{\Upsilon}{\Gamma}{
+    \IsWf{\Upsilon}{\Gamma}{
       \App{\vartheta}{E_0,\dots,E_n}
@@
       \HypJ{
-        \IsWf{\Theta}{\Upsilon}{\Gamma}{E_i}{v_i}
+        \IsWf{\Upsilon}{\Gamma}{E_i}{v_i}
       }{
@@
   \infer{
-    \IsWf{\Theta}{\Upsilon}{\Gamma}{
+    \IsWf{\Upsilon}{\Gamma}{
       \App{\OpNabla{\vec{\sigma}}}{
@@
   }{
-    \IsWf{\Theta}{\Upsilon,\vec{u}:\vec{\sigma}}{\Gamma}{J}{\SortJdg}
+    \IsWf{\Upsilon,\vec{u}:\vec{\sigma}}{\Gamma}{J}{\SortJdg}
   }
@@
   \infer{
-    \IsWf{\Theta}{\Upsilon}{\Gamma}{
+    \IsWf{\Upsilon}{\Gamma}{
       H\gg A
@@
   }{
-    \IsWf{\Theta}{\Upsilon}{\Gamma}{H}{\SortTele}
-&
-    \IsWf{\Theta}{\Upsilon}{\Gamma}{A}{\SortProp}
+    \IsWf{\Upsilon}{\Gamma}{H}{\SortTele}
+&
+    \IsWf{\Upsilon}{\Gamma}{A}{\SortProp}
   }
@@
   \infer{
-    \IsWf{\Theta}{\Upsilon}{\Gamma}{\diamond}{\SortTele}
+    \IsWf{\Upsilon}{\Gamma}{\diamond}{\SortTele}
   }{
@@
   \infer{
-    \IsWf{\Theta}{\Upsilon}{\Gamma}{H,u:A}{\SortTele}
+    \IsWf{\Upsilon}{\Gamma}{H,u:A}{\SortTele}
   }{
     \begin{array}{l}
-      \IsWf{\Theta}{\Upsilon\setminus\MkSet{u}}{\Gamma}{H}{\SortTele}
+      \IsWf{\Upsilon\setminus\MkSet{u}}{\Gamma}{H}{\SortTele}
 \\
-      \IsWf{\Theta}{\Upsilon\setminus\MkSet{u}}{\Gamma}{A}{\SortProp}
+      \IsWf{\Upsilon\setminus\MkSet{u}}{\Gamma}{A}{\SortProp}
     \end{array}
@@
 \end{gather*}
+
+Based on these rules, it is easy to show that the example sequent from the
+previous section is wellformed:
+\[
+  \IsWf{\cdot}{\cdot}{
+    \App{\OpNabla{\SortExp,\SortExp}}{
+      \MkBTm{u,v}{}{
+        \diamond,u:\mathtt{P},v:\App{\mathtt{pred}}{`u}
+        \gg
+        \App{\mathtt{pred}}{`u}
+      }
+    }
+  }{\SortJdg}
+\]
 

### 2015-12-12 paper.tex
add remark about a general thy of behavioral refinements

@@
 
+\begin{remark}
+  In the full development of a theory of behavioral refinements, the ad-hoc
+  judgment $\IsWf{\Upsilon}{\Gamma}{M}{\tau}$ would be replaced with a general
+  notion of refinement $\IMode{\phi}\sqsubset\tau$ ($\phi$ refines sort
+  $\tau$), and a new judgment $\IMode{\Upsilon}\parallel\IMode{\Phi}\vDash
+  \IMode{M} \in \IMode{\phi}$, such that $\phi$ refines $\tau$ the
+  refinement context $\Phi$ refines some sort context $\Gamma$. Furthermore, in
+  addition to restricting membership, refinements might also be designed to
+  coarsen equivalence.
+\end{remark}
+
 \section*{Acknowledgements}

### 2015-12-12 paper.tex
typo

@@
 encoding is sufficiently laborious and obscure that it is not used in practice.
-Crary has demonstrated a first-order encoding of contexts in the logical
-framework in bijection with actual LF-contexts~\cite{crary:2009}, which has
-been successfully used in large-scale mechanization efforts, including that of
-Standard ML~\cite{lee-crary-harper:2007} and the Edinburgh Logical Framework
+Crary has demonstrated a first-order encoding of contexts in the Edinburgh
+Logical Framework (ELF) in bijection with actual LF-contexts~\cite{crary:2009},
+which has been successfully used in large-scale mechanization efforts,
+including that of Standard ML~\cite{lee-crary-harper:2007} and the ELF
 itself~\cite{martens-crary:2012}.
@@
   $\tau$), and a new judgment $\IMode{\Upsilon}\parallel\IMode{\Phi}\vDash
-  \IMode{M} \in \IMode{\phi}$, such that $\phi$ refines $\tau$ the
+  \IMode{M} \in \IMode{\phi}$, such that $\phi$ refines $\tau$ and the
   refinement context $\Phi$ refines some sort context $\Gamma$. Furthermore, in

### 2015-12-18 paper.tex
presheaf -> copresheaf; remove sheafification for now
we weren't using sheafification anywhere, so no point in defining it.

@@
 
-\paragraph{Presheaves on $\SCtx$}
+\paragraph{Copresheaves on $\SCtx$}
 
@@
 \begin{quote}
-  A presheaf $\Of{X}{\Sets^\SCtx}$ is a sheaf on $\SCtxSite$ just when, for all $\Member{n}{X(\Upsilon')}$, if we have
+  A copresheaf $\Of{X}{\Sets^\SCtx}$ is a sheaf on $\SCtxSite$ just when, for all $\Member{n}{X(\Upsilon')}$, if we have
   \begin{tikzcd}[cramped]
@@
 Therefore, a sheaf on $\SCtxSite$ is a set that varies over $\Sorts$-sorted
-symbols, which has a well-behaved notion of support. Another way to put it is,
-if $\Member{m}{X(\Upsilon)}$ only ``uses'' some of the symbols available at
-world $\Upsilon$, there's a canonical way to restrict $m$ to a term in the
+symbol contexts, which has a well-behaved notion of support. Another way to put
+it is, if $\Member{m}{X(\Upsilon)}$ only ``uses'' some of the symbols available
+at world $\Upsilon$, there's a canonical way to restrict $m$ to a term in the
 world which contains \emph{only} the symbols that it uses.
 
-\subsection{Constructions on sheaves}
-
-For any Grothendieck site $\Match{\mathfrak{S}}{\Pair{\mathcal{C}}{J}}$, the
-inclusion $\Of{\iota}{\Sheaves{\mathfrak{S}}\ra\Presheaves{\mathcal{C}}}$ of
-sheaves into presheaves has a left adjoint
-$\IMode{\Sheafify{(-)}}\dashv\IMode{\iota}$, which takes any presheaf
-$\Of{X}{\Presheaves{\mathcal{C}}}$ to a unique sheaf
-$\Of{\Sheafify{X}}{\Sheaves{\mathfrak{S}}}$, called the \emph{sheafification}
-of $X$. Sheafification is the twice-iterated application of the
-\emph{+-construction} $X^+$, which turns a presheaf into a separated presheaf,
-and a separated presheaf into a sheaf \cite{mac-lane-moerdijk:1992}:
-\begin{align*}
-  \ADefine{
-    X^+(C)
-  }{
-    \int^{S\in J(C)}
-    \Hom{\Presheaves{\mathcal{C}}}{S}{X}
-  }\\
-  \ADefine{
-    \Sheafify{X}
-  }{
-    X^{++}
-  }
-\end{align*}
-
-We also have a canonical natural transformation $\Of{\eta}{X\ra X^+}$, defined
-using the maximal sieve $\Define{t_C}{\Yoneda{C}}$:
-\[
-  \HypJ{
-    \Define{\eta_C(m)}{
-      \Pair{t_C}{
-        X(-)(m)
-      }
-    }
-  }{
-    \Member{m}{X(C)}
-  }
-\]
+\paragraph{Constructions on sheaves}
+
+%% For any Grothendieck site $\Match{\mathfrak{S}}{\Pair{\mathcal{C}}{J}}$, the
+%% inclusion $\Of{\iota}{\Sheaves{\mathfrak{S}}\ra\Presheaves{\mathcal{C}}}$ of
+%% sheaves into presheaves has a left adjoint
+%% $\IMode{\Sheafify{(-)}}\dashv\IMode{\iota}$, which takes any presheaf
+%% $\Of{X}{\Presheaves{\mathcal{C}}}$ to a unique sheaf
+%% $\Of{\Sheafify{X}}{\Sheaves{\mathfrak{S}}}$, called the \emph{sheafification}
+%% of $X$. Sheafification is the twice-iterated application of the
+%% \emph{+-construction} $X^+$, which turns a presheaf into a separated presheaf,
+%% and a separated presheaf into a sheaf \cite{mac-lane-moerdijk:1992}:
+%% \begin{align*}
+%%   \ADefine{
+%%     X^+(C)
+%%   }{
+%%     \int^{S\in J(C)}
+%%     \Hom{\Presheaves{\mathcal{C}}}{S}{X}
+%%   }\\
+%%   \ADefine{
+%%     \Sheafify{X}
+%%   }{
+%%     X^{++}
+%%   }
+%% \end{align*}
+%%
+%% We also have a canonical natural transformation $\Of{\eta}{X\ra X^+}$, defined
+%% using the maximal sieve $\Define{t_C}{\Yoneda{C}}$:
+%% \[
+%%   \HypJ{
+%%     \Define{\eta_C(m)}{
+%%       \Pair{t_C}{
+%%         X(-)(m)
+%%       }
+%%     }
+%%   }{
+%%     \Member{m}{X(C)}
+%%   }
+%% \]
 

### 2015-12-22 paper.tex
Shorten arrows from tikz-pfeile
fix renaming notation
remove bureaucratic corporatespeak

@@
 \usepackage{tikz-cd}
-\usepackage{tikzpfeile}
+\usepackage[length=.5]{tikzpfeile}
 \usepackage{todonotes}
@@
 $\Upsilon,u\mapsto\Upsilon,v$ shall take $\mathtt{get}[u]$ to
-$\mathtt{get}[v]$. Going forward, we will write $\vartheta\varrho$ for
-$\Operators_a(\varrho)(\vartheta)$.
+$\mathtt{get}[v]$. In the future, we will write $\Rename{\varrho}{\vartheta}$
+for $\Operators_a(\varrho)(\vartheta)$.
 
@@
     }{
-      \App{\vartheta\varrho}{F_0,\dots,F_n}
+      \App{\Rename{\varrho}{\vartheta}}{F_0,\dots,F_n}
     }
@@
 
-Going forward, we will write $\Subst{N}{x}{M}$ for $M'$ when
+In the remainder, we will write $\Subst{N}{x}{M}$ for $M'$ when
 $\IsSubst{N}{x}{M}{M'}$, and $\Subst{\vec{N}}{\vec{x}}{M}$ for the simultaneous

### 2015-12-29 paper.tex
fix freudian slip

@@
 
-\subsection{The signature endofunctor and its initial algebras}
+\subsection{The signature endofunctor and its algebras}
 
@@
 Then, a $\Sigma$-model is a $\Sigma$-monoid $P$ which is equipped with an
-initial algebra $\Of{\alpha}{\SigFun{\Sigma}(P)\ra P}$, which shall interpret
+algebra $\Of{\alpha}{\SigFun{\Sigma}(P)\ra P}$, which shall interpret
 applications of each operator.
@@
 
-We will proceed using the initial $\SigFun{\Sigma}$-algebra $\alpha$, as
+We will proceed using the $\SigFun{\Sigma}$-algebra $\alpha$, as
 follows, by postcomposing it with a morphism $\beta$ from the environment into

### 2015-12-29 paper.tex
move comparison w/ nominal sets to appendix

@@
 
-\subsection{Relation to Nominal Sets and the Schanuel Topos}
-
-Pitts defines a category $\mathbf{Nom}$ of nominal sets in \cite{pitts:2013}
-which function equivalently to the sheaves that we considered above, in the
-case of $\IMode{\Sorts}\cong\IMode{\mathbb{1}}$. Fix a countably infinite set
-of atoms $\mathbb{A}$; then let $\Perm\mathbb{A}$ be the group of permutations
-(i.e.\ autoequivalences) on $\mathbb{A}$.  Considered as a category,
-$\Perm\mathbb{A}$ has a single object $\bullet$ with morphisms
-$\Of{\pi}{\bullet\ra\bullet}$ for each $\Member{\pi}{\Dom{\Perm\mathbb{A}}}$.
-Then, the category of nominal sets $\mathbf{Nom}$ is the subcategory of
-$\Sets^{\Perm\mathbb{A}}$ containing just the presheaves which satisfy a
-\emph{finite support} condition.
-
-The category $\mathbf{Nom}$ is equivalent to the category of covariant sheaves
-on $\SymSets$ under the atomic coverage, called the Schanuel Topos
-$\Define{\mathbb{S}}{\Sheaves{\OpCat{\SymSets}}}$ \cite{pitts:2013,
-fiore-staton:2006}; equivalently, the Schanuel topos is the subcategory of
-$\Sets^\SymSets$ containing only pullbacks-preserving functors.
-
-In addition to the unisorted nominal sets framework, Pitts also briefly
-discusses a multi-sorted version of the apparatus in which the countably
-infinite set of atoms $\mathbb{A}$ is equipped with a sort assignment
-$\Of{\mathfrak{s}}{\mathbb{A}\to\Sorts}$ such that all the fibres of
-$\mathfrak{s}$ are countably infinite; then, he defines for any such assignment
-the category $\mathbf{Nom}_\mathfrak{s}$ of $\mathfrak{s}$-sorted nominal sets
-as a subcategory of the category of presheaves on the group of
-$\mathfrak{s}$-respecting permutations on $\mathbb{A}$,
-$\Perm_\mathfrak{s}\mathbb{A}$.
-
-In the same way as the Schanuel topos is equivalent to the unisorted nominal
-sets, we expect to find an equivalence between $\Sheaves{\SCtxSite}$ and the
-limit $\int_{\mathfrak{s}}\mathbf{Nom}_\mathfrak{s}$. At the very least, the
-sheaf topos $\Sheaves{\SCtxSite}$ may serve as a multi-sorted generalization of
-the Schanuel topos, for which we may carry over certain useful results,
-including its characterization as the subcategory of $\Sets^\SCtx$ which
-contains just the pullbacks-preserving functors (see
-Theorem~\ref{thm:pullback-preservation} for the proof).
-
 \section{Operators and signatures}
@@
 
-\section{Sheafhood on $\SCtxSite$}
+\section{$\SCtxSite$, Nominal Sets, and the Schanuel Topos}
+
+Pitts defines a category $\mathbf{Nom}$ of nominal sets in \cite{pitts:2013}
+which function equivalently to the sheaves that we considered above, in the
+case of $\IMode{\Sorts}\cong\IMode{\mathbb{1}}$. Fix a countably infinite set
+of atoms $\mathbb{A}$; then let $\Perm\mathbb{A}$ be the group of permutations
+(i.e.\ autoequivalences) on $\mathbb{A}$.  Considered as a category,
+$\Perm\mathbb{A}$ has a single object $\bullet$ with morphisms
+$\Of{\pi}{\bullet\ra\bullet}$ for each $\Member{\pi}{\Dom{\Perm\mathbb{A}}}$.
+Then, the category of nominal sets $\mathbf{Nom}$ is the subcategory of
+$\Sets^{\Perm\mathbb{A}}$ containing just the presheaves which satisfy a
+\emph{finite support} condition.
+
+The category $\mathbf{Nom}$ is equivalent to the category of covariant sheaves
+on $\SymSets$ under the atomic coverage, called the Schanuel Topos
+$\Define{\mathbb{S}}{\Sheaves{\OpCat{\SymSets}}}$ \cite{pitts:2013,
+fiore-staton:2006}; equivalently, the Schanuel topos is the subcategory of
+$\Sets^\SymSets$ containing only pullbacks-preserving functors.
+
+In addition to the unisorted nominal sets framework, Pitts also briefly
+discusses a multi-sorted version of the apparatus in which the countably
+infinite set of atoms $\mathbb{A}$ is equipped with a sort assignment
+$\Of{\mathfrak{s}}{\mathbb{A}\to\Sorts}$ such that all the fibres of
+$\mathfrak{s}$ are countably infinite; then, he defines for any such assignment
+the category $\mathbf{Nom}_\mathfrak{s}$ of $\mathfrak{s}$-sorted nominal sets
+as a subcategory of the category of presheaves on the group of
+$\mathfrak{s}$-respecting permutations on $\mathbb{A}$,
+$\Perm_\mathfrak{s}\mathbb{A}$.
+
+In the same way as the Schanuel topos is equivalent to the unisorted nominal
+sets, we expect to find an equivalence between $\Sheaves{\SCtxSite}$ and the
+limit $\int_{\mathfrak{s}}\mathbf{Nom}_\mathfrak{s}$. At the very least, the
+sheaf topos $\Sheaves{\SCtxSite}$ may serve as a multi-sorted generalization of
+the Schanuel topos, for which we may carry over certain useful results,
+including its characterization as the subcategory of $\Sets^\SCtx$ which
+contains just the pullbacks-preserving functors.
 

### 2015-12-29 paper.tex
"bound term" ⇒ "abstraction" (#42)

@@
 
-Metavariables are substituted by bound terms; since a metavariable may only
+Metavariables are substituted by abstractions; since a metavariable may only
 appear in an application expression $\MApp{\MV{m}}{\cdots}{\cdots}$, we will
-instantiate the bound term at the supplied parameters and arguments.
+instantiate the abstraction at the supplied parameters and arguments.
 Substitution for metavariables is defined inductively by the judgments
@@
 presupposing $\IsAbt{\cdot}{\Upsilon}{\Gamma}{M}{\tau}$. Additionally, we
-introduce an analogous judgment on bound terms,
+introduce an analogous judgment on abstractions,
 $\IsWf{\Upsilon}{\Gamma}{E}{v}$ presupposing

### 2015-12-29 paper.tex
Add rule names

@@
 \DeclareMathOperator{\Perm}{\mathbf{Perm}}
+
+\newcommand\RuleNameVar[2]{{#1}_{\textit{var}_{#2}}}
+\newcommand\RuleNameMVar[2]{{#1}_{\textit{mvar}_{#2}}}
+\newcommand\RuleNameApp[1]{{#1}_\textit{app}}
+\newcommand\RuleNameAbs[2]{{#1}_{\textit{abs}_{#2}}}
 
@@
 
-\[
-  \infer[\textit{var}]{
+\newcommand\RuleIsAbtVar{\RuleNameVar{\vdash}{}}
+\newcommand\RuleIsAbtMVar{\RuleNameMVar{\vdash}{}}
+\newcommand\RuleIsAbtApp{\RuleNameApp{\vdash}}
+\newcommand\RuleIsAbtAbs{\RuleNameAbs{\vdash}{}}
+
+\[
+  \infer[\RuleIsAbtVar]{
     \IsAbt{\Theta}{\Upsilon}{\Gamma}{x}{\tau}
@@
 \[
-  \infer[\textit{mvar}]{
+  \infer[\RuleIsAbtMVar]{
     \IsAbt{\Theta}{\Upsilon}{\Gamma}{
@@
 \[
-  \infer[\textit{app}]{
+  \infer[\RuleIsAbtApp]{
     \IsAbt{\Theta}{\Upsilon}{\Gamma}{
@@
 \[
-  \infer[\textit{abs}]{
+  \infer[\RuleIsAbtAbs]{
     \IsBTm{\Theta}{\Upsilon}{\Gamma}{
@@
 
+\newcommand\RuleFVVar{\RuleNameVar{\mathbf{FV}}{}}
+\newcommand\RuleFVMVar{\RuleNameMVar{\mathbf{FV}}{}}
+\newcommand\RuleFVApp{\RuleNameApp{\mathbf{FV}}}
+\newcommand\RuleFVAbs{\RuleNameAbs{\mathbf{FV}}{}}
+
 We can easily calculate the variables free in an expression by recursion on its
@@
 \begin{gather*}
-  \infer[\textit{var}]{
+  \infer[\RuleFVVar]{
     \CalcFV{x}{\MkSet{x}}
@@
   %
-  \infer[\textit{mvar}]{
+  \infer[\RuleFVMVar]{
     \CalcFV{
@@
   %
-  \infer[\textit{app}]{
+  \infer[\RuleFVApp]{
     \CalcFV{
@@
   %
-  \infer[\textit{abs}]{
+  \infer[\RuleFVAbs]{
     \CalcFV{
@@
 
-Whereas the calculation of free variables pivoted on the \emph{var} rule, the
-calculation of free symbols will pivot on the \emph{app} and \emph{mvar} rule,
+\newcommand\RuleFSVar{\RuleNameVar{\mathbf{FS}}{}}
+\newcommand\RuleFSMVar{\RuleNameMVar{\mathbf{FS}}{}}
+\newcommand\RuleFSApp{\RuleNameApp{\mathbf{FS}}}
+\newcommand\RuleFSAbs{\RuleNameAbs{\mathbf{FS}}{}}
+
+Whereas the calculation of free variables pivoted on the $\RuleFVVar$ rule, the
+calculation of free symbols will pivot on the $\RuleFSApp$ and $\RuleFSMVar$ rules,
 because the only way a symbol can be introduced is as a parameter to an
@@
 \begin{gather*}
-  \infer[\textit{var}]{
+  \infer[\RuleFSVar]{
     \CalcFS{x}{\MkSet{}}
@@
   %
-  \infer[\textit{mvar}]{
+  \infer[\RuleFSMVar]{
     \CalcFS{
@@
   %
-  \infer[\textit{app}]{
+  \infer[\RuleFSApp]{
     \CalcFS{
@@
   %
-  \infer[\textit{abs}]{
+  \infer[\RuleFSAbs]{
     \CalcFS{
@@
 \subsection{Renaming of symbols}
+
+\newcommand\RuleRenPrefix{\bullet}
+\newcommand\RuleRenVar{\RuleNameVar{\RuleRenPrefix}{}}
+\newcommand\RuleRenMVar{\RuleNameMVar{\RuleRenPrefix}{}}
+\newcommand\RuleRenApp{\RuleNameApp{\RuleRenPrefix}}
+\newcommand\RuleRenAbs{\RuleNameAbs{\RuleRenPrefix}{}}
 
@@
 \begin{gather*}
-  \infer[\textit{var}]{
+  \infer[\RuleRenVar]{
     \IsRenaming{\varrho}{x}{x}
@@
 \\[6pt]
-  \infer[\textit{mvar}]{
+  \infer[\RuleRenMVar]{
     \IsRenaming{\varrho}{
@@
 \\[6pt]
-  \infer[\textit{app}]{
+  \infer[\RuleRenApp]{
     \IsRenaming{\varrho}{
@@
 \\[6pt]
-  \infer[\textit{abs}]{
+  \infer[\RuleRenAbs]{
     \IsRenaming{\varrho}{
@@
 
+\newcommand\RuleSubstPrefix{/}
+\newcommand\RuleSubstVar[1]{\RuleNameVar{\RuleSubstPrefix}{#1}}
+\newcommand\RuleSubstMVar{\RuleNameMVar{\RuleSubstPrefix}{}}
+\newcommand\RuleSubstApp{\RuleNameApp{\RuleSubstPrefix}}
+\newcommand\RuleSubstAbs[1]{\RuleNameAbs{\RuleSubstPrefix}{#1}}
+
 Variable substitution in abts is defined inductively by a pair of judgments,
@@
 \begin{gather*}
-  \infer[\textit{var}_1]{
+  \infer[\RuleSubstVar{1}]{
     \IsSubst{N}{x}{y}{N}
@@
   \qquad
-  \infer[\textit{var}_2]{
+  \infer[\RuleSubstVar{2}]{
       \IsSubst{N}{x}{y}{y}
@@
 \\[6pt]
-  \infer[\textit{mvar}]{
+  \infer[\RuleSubstMVar]{
     \IsSubst{N}{x}{
@@
 \\[6pt]
-  \infer[\textit{app}]{
+  \infer[\RuleSubstApp]{
     \IsSubst{N}{x}{
@@
 \begin{gather*}
-  \infer[\textit{abs}_1]{
+  \infer[\RuleSubstAbs{1}]{
     \IsSubst{N}{x}{
@@
 \\[6pt]
-  \infer[\textit{abs}_2]{
+  \infer[\RuleSubstAbs{2}]{
     \IsSubst{N}{x}{
@@
 \subsection{Substitution of metavariables}
+
+\newcommand\RuleMSubstPrefix{/}
+\newcommand\RuleMSubstSuffix{\texttt{meta}}
+\newcommand\RuleMSubstVar{\RuleNameVar{\RuleMSubstPrefix}{}^\RuleMSubstSuffix}
+\newcommand\RuleMSubstMVar[1]{\RuleNameMVar{\RuleMSubstPrefix}{#1}^\RuleMSubstSuffix}
+\newcommand\RuleMSubstApp{\RuleNameApp{\RuleMSubstPrefix}^\RuleMSubstSuffix}
+\newcommand\RuleMSubstAbs{\RuleNameAbs{\RuleMSubstPrefix}{}^\RuleMSubstSuffix}
 
@@
 \begin{gather*}
-  \infer[\textit{var}]{
+  \infer[\RuleMSubstVar]{
     \IsSubst{E}{\MV{m}}{x}{x}
@@
 \\[6pt]
-  \infer[\textit{mvar}_1]{
+  \infer[\RuleMSubstMVar{1}]{
     \IsSubst{E}{\MV{m}}{
@@
 \\[6pt]
-  \infer[\textit{mvar}_2]{
+  \infer[\RuleMSubstMVar{2}]{
     \IsSubst{
@@
   %
-  \infer[\textit{app}]{
+  \infer[\RuleMSubstApp]{
     \IsSubst{E}{\MV{m}}{
@@
 \\[6pt]
-  \infer[\textit{abs}]{
+  \infer[\RuleMSubstAbs]{
     \IsSubst{E}{\MV{m}}{

### 2015-12-31 paper.tex
update notation package
fix some wording and add clarifications (#42)

@@
 %
-in $\SCtx$, we have a corresponding functorial renaming action on any
-copresheaf $\Of{X}{\Sets^\SCtx}$, namely $\Of{X(\varrho)}{X(\Upsilon)\ra
-X(\Upsilon')}$. When the copresheaf $X$ is clear from context, we will write
-$m\cdot\varrho$ for $X(\varrho)(m)$.
+in $\SCtx$, any copresheaf $\Of{X}{\Sets^\SCtx}$ has a symbol-renaming action
+$\Of{X(\varrho)}{X(\Upsilon)\ra X(\Upsilon')}$. When the copresheaf $X$ is
+clear from context, we will write $m\cdot\varrho$ for $X(\varrho)(m)$.
 
@@
 
-Intuitively, when $\Upsilon$ supports $\Member{m}{X(\Upsilon')}$, we would
+When
+  \begin{tikzcd}[cramped]
+    \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
+  \end{tikzcd}
+such that $\Supports{\Upsilon}{\varrho}{m}$ for some $\Member{m}{X(\Upsilon')}$
+, we would
 expect that we can work backward to a unique $\Of{m'}{\Upsilon}$ such that
@@
 $X$, this is precisely the sheaf condition for copresheaves on $\SCtx$
-under the atomic topology \cite[p.~126]{mac-lane-moerdijk:1992}. The sheaf
-condition, then, ensures that the notion of support is well-behaved.
+under the atomic topology \cite[p.~126]{mac-lane-moerdijk:1992}.
+%%% TODO: add definition of atomic coverage
+The sheaf condition, then, ensures that the notion of support is well-behaved in the
+following sense: if some term lies in two different fibers, then it may be
+\emph{strengthened} uniquely into the fibre of its support.
 
@@
 metavariables in $\Theta$, parameters in $\Upsilon$, and variables in $\Gamma$.
-Let the judgment $\IsBTm{\Theta}{\Upsilon}{\Gamma}{E}{v}$ presuppose
+Let the judgment $\IsAbs{\Theta}{\Upsilon}{\Gamma}{E}{v}$ presuppose
 $\IsValence{v}$. Then, the syntax of abstract binding trees is inductively
@@
       \HypJ{
-        \IsBTm{\Theta}{\Upsilon}{\Gamma}{E_i}{v_i}
+        \IsAbs{\Theta}{\Upsilon}{\Gamma}{E_i}{v_i}
       }{\Leq{i}{n}}
@@
   \infer[\RuleIsAbtAbs]{
-    \IsBTm{\Theta}{\Upsilon}{\Gamma}{
-      \MkBTm{\vec{u}}{\vec{x}}{M}
+    \IsAbs{\Theta}{\Upsilon}{\Gamma}{
+      \MkAbs{\vec{u}}{\vec{x}}{M}
     }{
@@
     \CalcFV{
-      \MkBTm{\vec{u}}{\vec{y}}{M}
+      \MkAbs{\vec{u}}{\vec{y}}{M}
     }{
@@
     \CalcFS{
-      \MkBTm{\vec{v}}{\vec{x}}{M}
+      \MkAbs{\vec{v}}{\vec{x}}{M}
     }{
@@
     \IsRenaming{\varrho}{
-      \MkBTm{\vec{u}}{\vec{x}}{M}
-    }{
-      \MkBTm{\vec{u}}{\vec{x}}{N}
+      \MkAbs{\vec{u}}{\vec{x}}{M}
+    }{
+      \MkAbs{\vec{u}}{\vec{x}}{N}
     }
@@
     \IsSubst{N}{x}{
-      \MkBTm{\vec{u}}{\vec{y}}{M}
-    }{
-      \MkBTm{\vec{u}}{\vec{y}}{M'}
+      \MkAbs{\vec{u}}{\vec{y}}{M}
+    }{
+      \MkAbs{\vec{u}}{\vec{y}}{M'}
     }
@@
     \IsSubst{N}{x}{
-      \MkBTm{\vec{u}}{\vec{y}}{M}
-    }{
-      \MkBTm{\vec{u}}{\vec{y}}{M}
+      \MkAbs{\vec{u}}{\vec{y}}{M}
+    }{
+      \MkAbs{\vec{u}}{\vec{y}}{M}
     }
@@
     \IsSubst{
-      \MkBTm{\vec{u}}{\vec{x}}{N}
+      \MkAbs{\vec{u}}{\vec{x}}{N}
     }{\MV{m}}{
@@
     \IsSubst{E}{\MV{m}}{
-      \MkBTm{\vec{u}}{\vec{x}}{M}
-    }{
-      \MkBTm{\vec{u}}{\vec{x}}{N}
+      \MkAbs{\vec{u}}{\vec{x}}{M}
+    }{
+      \MkAbs{\vec{u}}{\vec{x}}{N}
     }
@@
     \App{\OpNabla{\SortExp,\SortExp}}{
-      \MkBTm{u,v}{}{
+      \MkAbs{u,v}{}{
         \diamond,u:\mathtt{P},v:\App{\mathtt{pred}}{`u}
@@
     \IsWf{\Upsilon}{\Gamma}{
-      \MkBTm{\vec{u}}{\vec{x}}{M}
+      \MkAbs{\vec{u}}{\vec{x}}{M}
     }{
@@
       \App{\OpNabla{\vec{\sigma}}}{
-        \MkBTm{\vec{u}}{}{J}
+        \MkAbs{\vec{u}}{}{J}
       }

### 2015-12-31 paper.tex
rework sheaf theory preliminaries a bit (#42)

@@
 \newcommand\SCtxSite{\SymSets[\Sorts]}
+\newcommand\AtomicCoverage{J_{\mathtt{atm}}}
 
@@
   \end{tikzcd}
-such that $\Supports{\Upsilon}{\varrho}{m}$ for some $\Member{m}{X(\Upsilon')}$
-, we would
-expect that we can work backward to a unique $\Of{m'}{\Upsilon}$ such that
-$\IsEq{m'\cdot\varrho}{m}$; whilst it is not in general the case for presheaves
-$X$, this is precisely the sheaf condition for copresheaves on $\SCtx$
-under the atomic topology \cite[p.~126]{mac-lane-moerdijk:1992}.
-%%% TODO: add definition of atomic coverage
-The sheaf condition, then, ensures that the notion of support is well-behaved in the
+such that $\Supports{\Upsilon}{\varrho}{m}$ for some $\Member{m}{X(\Upsilon')}$,
+we would expect that we can work backward to a unique $\Of{m'}{\Upsilon}$
+such that $\IsEq{m'\cdot\varrho}{m}$; whilst it is not in general the case for
+presheaves $X$, this is precisely the sheaf condition for copresheaves on
+$\SCtx$ under the atomic topology \cite[p.~126]{mac-lane-moerdijk:1992}, which
+we will define in Definition~$\ref{def:atomic-topology}$.
+
+The sheaf condition ensures that the notion of support is well-behaved in the
 following sense: if some term lies in two different fibers, then it may be
-\emph{strengthened} uniquely into the fibre of its support.
-
-\paragraph{Sites and coverages}
-
-We will briefly discuss the basics of sheaves on a site. A \emph{site} is a
-pair $\Pair{\mathcal{C}}{J}$, with $\mathcal{C}$ a category of ``worlds'' and
-$J$ a Grothendieck topology (or ``coverage''). A coverage is a predicate that
-picks out for each object $\Of{C}{\mathcal{C}}$, the \emph{covering sieves} of
-$C$, subject to several axioms (see \cite{mac-lane-moerdijk:1992} for details).
-
-\paragraph{Covering sieves and matching families}
-
-A \emph{sieve} on an object $C$ is a subfunctor of the Yoneda embedding
-$\Define{\Yoneda{C}}{\Hom{\mathcal{C}}{-}{C}}$. We say that such a sieve $S$
-covers $C$ just when $\Member{S}{J(C)}$.  For a presheaf $X$ on the site
-$\Pair{\mathcal{C}}{J}$, a \emph{matching family} for the sieve $S$ is a
-natural transformation $\Of{\phi}{S\ra X}$. An \emph{amalgamation} for $\phi$
-is an object $\Member{m}{X(C)}$ such that $\IsEq{m\cdot f}{\phi_D(f)}$ for each
+\emph{strengthened} uniquely into the fiber of its support.
+
+\begin{definition}[Sieves and matching families]
+  A \emph{sieve} $S$ on an object $C$ is a subfunctor of the Yoneda embedding
+  $\Define{\Yoneda{C}}{\Hom{\mathcal{C}}{-}{C}}$, i.e.\
+  $\IMode{S}\sqsubseteq\IMode{\Yoneda{C}}$. For a presheaf $X$ on the category
+  $\mathcal{C}$, a \emph{matching family} for a sieve $S$ is a natural
+  transformation $\Of{\phi}{S\ra X}$. An \emph{amalgamation} for $\phi$ is an
+  object $\Member{m}{X(C)}$ such that $\IsEq{m\cdot f}{\phi_D(f)}$ for each
   $\Member{f}{S(D)}$, for all objects $D$.
-
-\paragraph{Separation and sheafhood}
-Then, we say that the presheaf $X$ is \emph{separated} in case for every sieve
+\end{definition}
+
+\begin{definition}[Sites and coverages]
+  A \emph{site} is a pair $\Pair{\mathcal{C}}{J}$, with $\mathcal{C}$ a category
+  of ``worlds'' and $J$ a Grothendieck topology (or ``coverage''). A Grothendieck
+  topology $J$ (or \emph{coverage}) is a family of sets of sieves, indexed by
+  $\mathcal{C}$-objects; that is, for each object $\Of{C}{\mathcal{C}}$, $J(C)$
+  is the set of $C$-sieves which will be said to $\emph{cover}$ $C$. A topology
+  $J$ must satisfy several axioms (see \cite{mac-lane-moerdijk:1992} for
+  details).
+\end{definition}
+
+\begin{definition}[Separation and sheafhood]
+  A presheaf on a site $\Pair{\mathcal{C}}{J}$ is a presheaf on $\mathcal{C}$; we
+  say that such a presheaf $X$ is \emph{separated} in case for every sieve
   $\Member{S}{J(C)}$ and matching family $\Of{\phi}{S\ra X}$, there is at most
-one amalgamation $\Member{m}{X(C)}$ for $\phi$. Additionally, $X$ is called a
+  one amalgamation $\Member{m}{X(C)}$ for $\phi$. Additionally, $X$ is a
   \emph{sheaf} when there is \emph{exactly one} such amalgamation.
-
-\paragraph{The atomic coverage}
-
-The atomic coverage $J_\mathsf{atm}$ is the one in which all non-empty sieves
-cover. A sieve $\IMode{S}\sqsubseteq\IMode{\Yoneda{C}}$ is non-empty when the
-collection $\bigcup_D S(D)$ of all its arrows is non-empty.
+\end{definition}
+
+\begin{definition}[The atomic topology]
+  \label{def:atomic-topology}
+  The atomic coverage $\AtomicCoverage$ is the one in which all non-empty sieves
+  cover; in other words:
+  \[
+    \Define{
+      \AtomicCoverage(C)
+    }{
+      \MkSet{
+        S\sqsubseteq\Yoneda{C}
+        \mid
+        \textstyle
+        \bigcup_D S(D)\ \textit{non-empty}
+      }
+    }
+  \]
+\end{definition}
 
 In the future, we will write $\SCtxSite$ for the Grothendieck site
-$\Pair{\OpCat{\left(\SCtx\right)}}{J_\mathsf{atm}}$, where $J_\mathsf{atm}$ is
-the atomic coverage. Sheafhood on the atomic site $\SCtxSite$ amounts to the
+$\Pair{\OpCat{\left(\SCtx\right)}}{\AtomicCoverage}$.
+Sheafhood on the atomic site $\SCtxSite$ amounts to the
 following condition, using Definition~\ref{def:support}:
@@
 infinite set of atoms $\mathbb{A}$ is equipped with a sort assignment
-$\Of{\mathfrak{s}}{\mathbb{A}\to\Sorts}$ such that all the fibres of
+$\Of{\mathfrak{s}}{\mathbb{A}\to\Sorts}$ such that all the fibers of
 $\mathfrak{s}$ are countably infinite; then, he defines for any such assignment

### 2015-12-31 paper.tex
mention strengthening; move explanation of symbols sheaf (#42)

@@
 
-Returning to the site $\SCtxSite$, we will write $\mathbf{S}_\tau$ for the sheaf
-of symbols of sort $\tau$, a subobject of the Yoneda embedding of the empty
-symbol context, $\Yoneda{\cdot}$.
-
 \section{Operators and signatures}
@@
 the structural properties of weakening and exchange via functoriality of
-$\Operators$. An operator signature is defined by specifying, for each arity
-$a$, the sheaf $\Operators_a$, whose ``elements'' are the operators of arity
-$a$.
+$\Operators$, and strengthening via the sheaf condition. An operator signature
+is defined by specifying, for each arity $a$, the sheaf $\Operators_a$, whose
+``elements'' are the operators of arity $a$.
 
@@
 
-These rules correspond to the following family of sheaves:
+Let $\mathbf{S}_\tau$ be the sheaf of symbols of sort $\tau$, a subobject of
+the Yoneda embedding of the empty symbol context, $\Yoneda{\cdot}$:
+\[
+  \Define{\mathbf{S}_\tau(\Upsilon)}{
+    \MkSet{
+      u\in\Dom{\Upsilon}
+      \mid
+      \Upsilon(u) \equiv \tau
+    }
+  }
+\]
+%
+Then the above rules correspond to the following family of sheaves:
 \begin{align*}

### 2015-12-31 paper.tex
clarify valences over which sorts (#42)

@@
 
-A metavariable context consists of bindings of \emph{valences} to
+A metavariable context consists of bindings of $\Sorts$-\emph{valences} to
 metavariables; let $\Define{\Valences}{\MkSet{v\mid \IsValenceUnmoded{v}}}$ be the set

### 2015-12-31 paper.tex
Reference Watkins / hereditary substitutions (#42)
Explain \widehat notation (#42)

@@
 
+The rules for metavariable substitution are reminiscent of the ``hereditary
+substitutions'' first devised for the Concurrent Logical Framework by Watkins
+in \cite{watkins:2004}.
+%
 As usual, we will write $\Subst{E}{\MV{m}}{M}$ for $N$ when
@@
 Let $\Define{\HypCat}{\OpCat{(\SCtx\times\Ctx)}}$; then we fix the functor
-category $\widehat{\HypCat}^\Sorts$ as our semantic universe.
-
+category $\widehat{\HypCat}^\Sorts$ as our semantic universe, using the
+notation of the French school,
+$\Define{\widehat{\mathcal{C}}}{\Sets^{\OpCat{\mathcal{C}}}}$.
+%
 Let $\Define{V_\tau(\Upsilon\parallel\Gamma)}{\MkSet{x\in\Dom\Gamma\mid
-    \Gamma(x)=\tau}}$ be called the presheaf of variables; additionally, we have
+    \Gamma(x)\equiv\tau}}$ be called the presheaf of variables; additionally, we have
 a presheaf of symbols
 $\Define{S_\tau(\Upsilon\parallel\Gamma)}{\MkSet{u\in\Dom\Upsilon\mid
-    \Upsilon(x)=\tau}}$. Lastly, we have the presheaf of operators with arity
+    \Upsilon(u)\equiv\tau}}$. Lastly, we have the presheaf of operators with arity
 $a$,

### 2015-12-31 paper.tex
use \SymPsh, \VarPsh macros appropriately
Explain the idea behind the substitution tensor (#42)
just say "compose" instead of "postcompose" (#42)
because I can't figure out which is right ;-)

@@
 
+\newcommand\SymPsh{\mathbf{S}}
+\newcommand\VarPsh{\mathbf{V}}
+
 Let $\Define{\HypCat}{\OpCat{(\SCtx\times\Ctx)}}$; then we fix the functor
@@
 %
-Let $\Define{V_\tau(\Upsilon\parallel\Gamma)}{\MkSet{x\in\Dom\Gamma\mid
+Let $\Define{\VarPsh_\tau(\Upsilon\parallel\Gamma)}{\MkSet{x\in\Dom\Gamma\mid
     \Gamma(x)\equiv\tau}}$ be called the presheaf of variables; additionally, we have
 a presheaf of symbols
-$\Define{S_\tau(\Upsilon\parallel\Gamma)}{\MkSet{u\in\Dom\Upsilon\mid
+$\Define{\SymPsh_\tau(\Upsilon\parallel\Gamma)}{\MkSet{u\in\Dom\Upsilon\mid
     \Upsilon(u)\equiv\tau}}$. Lastly, we have the presheaf of operators with arity
@@
 \subsection{Substitution monoidal structures}
-\newcommand\SymPsh{\mathbf{S}}
-\newcommand\VarPsh{\mathbf{V}}
 
@@
 
+Intuitively, the coend construction $A\bullet P$ provides the data of a
+suspended substitution of an $A$-term's variables by $P$-terms.
+%
 Using this, we can define a tensor $P\odot Q$ for
@@
 We will proceed using the $\SigFun{\Sigma}$-algebra $\alpha$, as
-follows, by postcomposing it with a morphism $\beta$ from the environment into
+follows, by composing it with a morphism $\beta$ from the environment into
 the signature endofunctor, which interprets the syntax of operator applications:

### 2015-12-31 paper.tex
add a few more citations of PFPL ;-) (#42)

@@
 use symbols to encode the syntax of contexts, except that here, we have better
-control over their proliferation because, contrary to the state of
-affairs in Abella, symbols are \emph{not} terms.
+control over their proliferation because, contrary to the state of affairs in
+Abella, symbols are \emph{not} expressions; following
+Harper~\cite{harper:2016}, symbols are meant only to serve as an open-ended
+indexing scheme for operators, not as actual expressions.
 
@@
 The above sequent has free symbols, but we can close over them by adding a form
-of parametric higher-order judgment to our object language, indexed by a
-collection of sorts $\vec{\sigma}$:
+of parametric higher-order judgment (as in \cite{harper:2016}) to our object
+language, indexed by a collection of sorts $\vec{\sigma}$:
 \[

### 2015-12-31 paper.tex
Acknowledge bob's feedback & help

@@
 The first author wishes to thank Robert Harper for numerous conversations about
-abstract binding trees and symbolic parameters, and Andy Pitts for his help in
-understanding multi-sorted nominal sets and atomic sheaves.
+abstract binding trees and symbolic parameters, and for his feedback on a draft
+of this paper; and Andy Pitts for his help in understanding multi-sorted
+nominal sets and atomic sheaves.
 

### 2016-01-21 ijcar-abts.tex
Add IJCAR edition

--- 2015-12-31 paper.tex
@@
-\documentclass[11pt]{article}
-
-\usepackage{notation/modes}
+\documentclass{llncs}
+
+\usepackage[bw]{notation/modes}
 \usepackage{notation/judgments}
@@
 \usepackage{amssymb}
-\usepackage{amsthm}
-%\usepackage{agda} % this should come after ams*
+\usepackage{bbold}
+% \usepackage{amsthm}
 \usepackage{catchfilebetweentags}
-\usepackage{geometry}
 \usepackage{graphicx}
@@
 \usepackage{url}
+\usepackage[h]{esvect}
 
 \definecolor{dark-gray}{gray}{0.20}
-
-% fonts
-\usepackage{pxfonts}
-\usepackage{eulervm}
-\usepackage{bbold}
-\usepackage[
-  activate={true,nocompatibility},
-  kerning=true,
-  spacing=true,
-  tracking=true
-]{microtype}
-\microtypecontext{spacing=nonfrench}
 
@@
 \newtheorem{cor}[thm]{Corollary}
-\theoremstyle{definition}
-\newtheorem{definition}[thm]{Definition}
-\newtheorem{example}[thm]{Example}
-\theoremstyle{remark}
-\newtheorem{remark}[thm]{Remark}
 \numberwithin{equation}{section}
-%% Euler doesn't have \varsigma
-\DeclareSymbolFont{greekletters}{OML}{cmr}{m}{it}
-\DeclareMathSymbol{\varsigma}{\mathalpha}{greekletters}{"26}
+\renewcommand\vec[1]{\protect\vv{#1}}
 
@@
 
-\newcommand\RuleNameVar[2]{{#1}_{\textit{var}_{#2}}}
-\newcommand\RuleNameMVar[2]{{#1}_{\textit{mvar}_{#2}}}
-\newcommand\RuleNameApp[1]{{#1}_\textit{app}}
-\newcommand\RuleNameAbs[2]{{#1}_{\textit{abs}_{#2}}}
-
 \begin{document}
@@
 \author{Jon Sterling and Darin Morrison}
+\institute{}
 \maketitle
@@
 semantics for several variations of second-order algebraic
-theories~\cite{fiore:1999, fiore:2005, fiore-hur:2010, fiore-mamoud:2010}; of
+theories~\cite{fiore:1999,fiore:2005,fiore-hur:2010,fiore-mamoud:2010}; of
 these, the simply sorted variants are equivalent to Harper's abstract binding
@@
 
-\newcommand\RuleIsAbtVar{\RuleNameVar{\vdash}{}}
-\newcommand\RuleIsAbtMVar{\RuleNameMVar{\vdash}{}}
-\newcommand\RuleIsAbtApp{\RuleNameApp{\vdash}}
-\newcommand\RuleIsAbtAbs{\RuleNameAbs{\vdash}{}}
-
 \[
@@
   addition to restricting membership, refinements might also be designed to
-  coarsen equivalence.
+  coarsen equivalence as in \cite{denney:1998}.
 \end{remark}
@@
 
-\newpage
 \appendix
@@
 on $\SymSets$ under the atomic coverage, called the Schanuel Topos
-$\Define{\mathbb{S}}{\Sheaves{\OpCat{\SymSets}}}$ \cite{pitts:2013,
-fiore-staton:2006}; equivalently, the Schanuel topos is the subcategory of
-$\Sets^\SymSets$ containing only pullbacks-preserving functors.
+$\Define{\mathbb{S}}{\Sheaves{\OpCat{\SymSets}}}$
+\cite{pitts:2013,fiore-staton:2006}; equivalently, the Schanuel topos is the
+subcategory of $\Sets^\SymSets$ containing only pullbacks-preserving functors.
 

### 2016-01-23 appendices.tex
Split up paper

--- 2016-01-21 ijcar-abts.tex
@@
-\documentclass{llncs}
-
-\usepackage[bw]{notation/modes}
-\usepackage{notation/judgments}
-\usepackage{notation/abts}
-\usepackage{notation/common}
-
-\usepackage{appendix}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{bbold}
-% \usepackage{amsthm}
-\usepackage{catchfilebetweentags}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{ifdraft}
-\usepackage{ifthen}
-\usepackage{mathtools}
-\usepackage{proof}
-\usepackage{scalerel}
-\usepackage{setspace}
-\usepackage{stmaryrd}
-\usepackage{supertabular}
-\usepackage{tikz-cd}
-\usepackage[length=.5]{tikzpfeile}
-\usepackage{todonotes}
-\usepackage{url}
-\usepackage[h]{esvect}
-
-\definecolor{dark-gray}{gray}{0.20}
-
-\newtheorem{thm}{Theorem}[section]
-\newtheorem{prop}[thm]{Proposition}
-\newtheorem{lem}[thm]{Lemma}
-\newtheorem{cor}[thm]{Corollary}
-\numberwithin{equation}{section}
-\renewcommand\vec[1]{\protect\vv{#1}}
-
-\newcommand\HypCat{\mathbf{H}}
-\newcommand\Supp[1]{\mathbf{supp}(#1)}
-\newcommand\SupportsUnmoded[3]{#1 \blacktriangleright_{#2} #3}
-\newcommand\Supports[3]{\SupportsUnmoded{\IMode{#1}}{\IMode{#2}}{\IMode{#3}}}
-
-\newcommand\CalcFV[2]{
-  \FV{\IMode{#1}}\leadsto \OMode{#2}
-}
-
-\newcommand\CalcFS[2]{
-  \FS{\IMode{#1}}\leadsto \OMode{#2}
-}
-
-\newcommand\SortExp{\mathtt{exp}}
-
-\newcommand\IsWf[4]{%
-  \IMode{#1}\parallel%
-  \IMode{#2}\vdash%
-  \IMode{#3}\in_{\mathbf{wf}}\IMode{#4}%
-}
-
-\newcommand\Sheaves[1]{\mathbf{Sh}\left(#1\right)}
-\newcommand\Presheaves[1]{\mathbf{Psh}\left(#1\right)}
-\newcommand\Sheafify[1]{#1^{\#}}
-\newcommand\ConstPsh[1]{\Delta[#1]}
-\newcommand\ConstSh[1]{\Delta^\#[#1]}
-\newcommand\SCtxSite{\SymSets[\Sorts]}
-\newcommand\AtomicCoverage{J_{\mathtt{atm}}}
-
-\DeclareMathOperator{\Perm}{\mathbf{Perm}}
-
-\begin{document}
-
-\title{Syntax and Semantics of Abstract Binding Trees}
-\date{}
-\author{Jon Sterling and Darin Morrison}
-\institute{}
-\maketitle
-
-Abstract binding trees (abts) are a generalization of abstract syntax trees
-where operators may express variable binding structure as part of their arities.
-Originally formulated by Peter Aczel~\cite{aczel:1978}, unisorted abts have been
-deployed successfully as the uniform syntactic framework for several
-implementations of Constructive Type Theory, including
-Nuprl~\cite{constable:1986}, MetaPRL~\cite{hickey:2003} and
-JonPRL~\cite{jonprl:2015}.
-
-In \emph{Practical Foundations for Programming Languages}~\cite{harper:2016},
-Robert Harper develops the multi-sorted version of abstract binding trees and
-proposes an extension to include families of operators indexed by
-\emph{symbols}, which are, unlike variables, subject to only
-distinctness-preserving renaming and not substitution; furthermore, symbols do
-not appear in the syntax of abts, and are only introduced as parameters to
-operators. This extension to support symbols is essential for a correct
-treatment of programming languages with open sums (e.g. ML's \texttt{exn} type),
-as well as assignable references.
-
-In parallel, M.\ Fiore and his collaborators have developed the categorical
-semantics for several variations of second-order algebraic
-theories~\cite{fiore:1999,fiore:2005,fiore-hur:2010,fiore-mamoud:2010}; of
-these, the simply sorted variants are equivalent to Harper's abstract binding
-trees augmented with a notion of second-order variable (metavariable).
-
-The contribution of this paper is the development of the syntax and semantics
-of multi-sorted nominal abts, an extension of second order universal algebra to
-support symbol-indexed families of operators. Additionally, we have developed
-the categorical semantics for abts formally in Constructive Type Theory using
-the Agda proof assistant~\cite{norell:2007}; we have also developed an
-implementation for abstract binding trees in Standard ML, which we intend to
-integrate into the JonPRL proof assistant \cite{jonprl:2015}.
-
-\section{Categorical Preliminaries}
-
-Fix a set $\Sorts$ of \emph{sorts}; we will say $\IsSort{\tau}$ when
-$\Member{\tau}{\Sorts}$. In this section, we will develop a theory of sets
-varying over collections of $\Sorts$-sorted symbols.
-
-Let $\SymSets$ be the category of finite cardinals and their injective maps;
-then the comma construction $\SCtx$, is the category of contexts of symbols
-whose objects are finite sets of symbols $U$ and sort-assignments
-\begin{tikzcd}[cramped]
-  \IMode{U}\arrow[r, "\IMode{\mathfrak{s}}"] &\IMode{\Sorts}
-\end{tikzcd}, and whose morphisms are
-sort-preserving renamings; we will write $\Upsilon$ for a symbol context
-$(U,\mathfrak{s})$.
-
-\begin{remark}
-
-  To be precise, $\SCtx$ is an abuse of notation for the comma category
-  $\Dom{-}_\SymSets\downarrow \Delta[\Sorts]$, with $\Dom{-}_\SymSets$ the
-  forgetful functor from $\SymSets$ to $\Sets$, and
-  $\Define{\Delta[\Sorts](*)}{\Sorts}$ a constant functor from $\mathbb{1}$ to $\Sets$.
-  Formally, an object of $\SCtx$ is, then, a triple $\Tuple{U, *,
-  \mathfrak{s}}$ with $U$ an object in $\SymSets$, $*$ the unique object in
-  $\mathbb{1}$, and $\mathfrak{s}$ a function $\Dom{U}_\SymSets\ra
-  \Sorts$; an arrow $\Of{\varrho}{\Tuple{U,*,\mathfrak{s}}\ra
-  \Tuple{V,*,\mathfrak{t}}}$ is a commuting triangle like the following:
-  \[
-    \begin{tikzcd}[cramped]
-      \IMode{\Dom{U}_\SymSets}
-        \arrow[rr, "\IMode{\Dom{\rho}_\SymSets}"]
-        \arrow[dr, swap, "\IMode{\mathfrak{s}}"] &&
-      \IMode{\Dom{V}_\SymSets}
-      \arrow[dl, "\IMode{\mathfrak{t}}"] \\
-      &\IMode{\Sorts}
-    \end{tikzcd}
-  \]
-
-  Because the object $\Of{*}{\mathbb{1}}$ is unique, we will always omit it
-  from the data of an object in $\SCtx$.
-\end{remark}
-
-\paragraph{Copresheaves on $\SCtx$}
-
-A presheaf on a category $\mathcal{C}$ is a contravariant functor
-$\OpCat{\mathcal{C}}\ra\Sets$; when $\Match{\mathcal{C}}{\OpCat{\mathcal{D}}}$,
-we call this a \emph{copresheaf} (covariant presheaf) on $\mathcal{D}$.
-Therefore, a copresheaf on $\SCtx$ is a functor $\SCtx\ra\Sets$; colloquially,
-such a copresheaf can be thought of as a set that \emph{varies over} symbol
-contexts.
-
-This leads to a very intuitive setting in which to consider sets
-that vary over collections of $\Sorts$-sorted symbols. In particular, for any
-morphism
-%
-\begin{tikzcd}[cramped]
-  \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
-\end{tikzcd}
-%
-in $\SCtx$, any copresheaf $\Of{X}{\Sets^\SCtx}$ has a symbol-renaming action
-$\Of{X(\varrho)}{X(\Upsilon)\ra X(\Upsilon')}$. When the copresheaf $X$ is
-clear from context, we will write $m\cdot\varrho$ for $X(\varrho)(m)$.
-
-Now, when $\Member{m}{X(\Upsilon)}$ for a copresheaf $\Of{X}{\Sets^\SCtx}$, we
-can define the notion that $m$ depends on \emph{at most} a subcontext
-$\Upsilon'$ of $\Upsilon$; this is called \emph{support}.
-
-\begin{definition}[Support]
-  \label{def:support}
-  When
-  \begin{tikzcd}[cramped]
-    \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
-  \end{tikzcd},
-  we say that $\Upsilon$ supports $\Member{m}{\Upsilon'}$ (written
-  $\Supports{\Upsilon}{\varrho}{m}$) when, for all
-  \begin{tikzcd}[cramped]
-    \IMode{\Upsilon'}\arrow[r, hook, "\IMode{\varrho_1}{,} \IMode{\varrho_2}"] & \IMode{\Upsilon''}
-  \end{tikzcd},
-  if $\IsEq{\varrho_1\circ\varrho}{\varrho_2\circ\varrho}$ then $\IsEq{m\cdot\varrho_1}{m\cdot\varrho_2}$.
-\end{definition}
-
-We can also always calculate the \emph{least support} of a term
-$\Member{m}{X(\Upsilon')}$, written $\Supp{m}$:
-\[
-  \Define{\Supp{m}}{
-    \bigcap_{
-      \begin{tikzcd}[cramped, ampersand replacement = \&]
-        \Upsilon\arrow[r, hook, "\varrho"] \& \Upsilon'
-      \end{tikzcd}
-    }
-    \MkSet{
-      \Upsilon
-      \mid
-      \SupportsUnmoded{\Upsilon}{\varrho}{m}
-    }
-  }
-\]
-Intuitively, $\Supp{m}$ is the exact symbol context that $m$ depends on.
-
-\subsection{Sheaves on the atomic site}
-
-When
-  \begin{tikzcd}[cramped]
-    \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
-  \end{tikzcd}
-such that $\Supports{\Upsilon}{\varrho}{m}$ for some $\Member{m}{X(\Upsilon')}$,
-we would expect that we can work backward to a unique $\Of{m'}{\Upsilon}$
-such that $\IsEq{m'\cdot\varrho}{m}$; whilst it is not in general the case for
-presheaves $X$, this is precisely the sheaf condition for copresheaves on
-$\SCtx$ under the atomic topology \cite[p.~126]{mac-lane-moerdijk:1992}, which
-we will define in Definition~$\ref{def:atomic-topology}$.
-
-The sheaf condition ensures that the notion of support is well-behaved in the
-following sense: if some term lies in two different fibers, then it may be
-\emph{strengthened} uniquely into the fiber of its support.
-
-\begin{definition}[Sieves and matching families]
-  A \emph{sieve} $S$ on an object $C$ is a subfunctor of the Yoneda embedding
-  $\Define{\Yoneda{C}}{\Hom{\mathcal{C}}{-}{C}}$, i.e.\
-  $\IMode{S}\sqsubseteq\IMode{\Yoneda{C}}$. For a presheaf $X$ on the category
-  $\mathcal{C}$, a \emph{matching family} for a sieve $S$ is a natural
-  transformation $\Of{\phi}{S\ra X}$. An \emph{amalgamation} for $\phi$ is an
-  object $\Member{m}{X(C)}$ such that $\IsEq{m\cdot f}{\phi_D(f)}$ for each
-  $\Member{f}{S(D)}$, for all objects $D$.
-\end{definition}
-
-\begin{definition}[Sites and coverages]
-  A \emph{site} is a pair $\Pair{\mathcal{C}}{J}$, with $\mathcal{C}$ a category
-  of ``worlds'' and $J$ a Grothendieck topology (or ``coverage''). A Grothendieck
-  topology $J$ (or \emph{coverage}) is a family of sets of sieves, indexed by
-  $\mathcal{C}$-objects; that is, for each object $\Of{C}{\mathcal{C}}$, $J(C)$
-  is the set of $C$-sieves which will be said to $\emph{cover}$ $C$. A topology
-  $J$ must satisfy several axioms (see \cite{mac-lane-moerdijk:1992} for
-  details).
-\end{definition}
-
-\begin{definition}[Separation and sheafhood]
-  A presheaf on a site $\Pair{\mathcal{C}}{J}$ is a presheaf on $\mathcal{C}$; we
-  say that such a presheaf $X$ is \emph{separated} in case for every sieve
-  $\Member{S}{J(C)}$ and matching family $\Of{\phi}{S\ra X}$, there is at most
-  one amalgamation $\Member{m}{X(C)}$ for $\phi$. Additionally, $X$ is a
-  \emph{sheaf} when there is \emph{exactly one} such amalgamation.
-\end{definition}
-
-\begin{definition}[The atomic topology]
-  \label{def:atomic-topology}
-  The atomic coverage $\AtomicCoverage$ is the one in which all non-empty sieves
-  cover; in other words:
-  \[
-    \Define{
-      \AtomicCoverage(C)
-    }{
-      \MkSet{
-        S\sqsubseteq\Yoneda{C}
-        \mid
-        \textstyle
-        \bigcup_D S(D)\ \textit{non-empty}
-      }
-    }
-  \]
-\end{definition}
-
-In the future, we will write $\SCtxSite$ for the Grothendieck site
-$\Pair{\OpCat{\left(\SCtx\right)}}{\AtomicCoverage}$.
-Sheafhood on the atomic site $\SCtxSite$ amounts to the
-following condition, using Definition~\ref{def:support}:
-\begin{quote}
-  A copresheaf $\Of{X}{\Sets^\SCtx}$ is a sheaf on $\SCtxSite$ just when, for all $\Member{n}{X(\Upsilon')}$, if we have
-  \begin{tikzcd}[cramped]
-    \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
-  \end{tikzcd}
-  and $\Supports{\Upsilon}{\varrho}{n}$, then there is a unique
-  $\Member{m}{X(\Upsilon)}$ such that $\IsEq{m\cdot\varrho}{n}$.
-\end{quote}
-
-Therefore, a sheaf on $\SCtxSite$ is a set that varies over $\Sorts$-sorted
-symbol contexts, which has a well-behaved notion of support. Another way to put
-it is, if $\Member{m}{X(\Upsilon)}$ only ``uses'' some of the symbols available
-at world $\Upsilon$, there's a canonical way to restrict $m$ to a term in the
-world which contains \emph{only} the symbols that it uses.
-
-\paragraph{Constructions on sheaves}
-
-%% For any Grothendieck site $\Match{\mathfrak{S}}{\Pair{\mathcal{C}}{J}}$, the
-%% inclusion $\Of{\iota}{\Sheaves{\mathfrak{S}}\ra\Presheaves{\mathcal{C}}}$ of
-%% sheaves into presheaves has a left adjoint
-%% $\IMode{\Sheafify{(-)}}\dashv\IMode{\iota}$, which takes any presheaf
-%% $\Of{X}{\Presheaves{\mathcal{C}}}$ to a unique sheaf
-%% $\Of{\Sheafify{X}}{\Sheaves{\mathfrak{S}}}$, called the \emph{sheafification}
-%% of $X$. Sheafification is the twice-iterated application of the
-%% \emph{+-construction} $X^+$, which turns a presheaf into a separated presheaf,
-%% and a separated presheaf into a sheaf \cite{mac-lane-moerdijk:1992}:
-%% \begin{align*}
-%%   \ADefine{
-%%     X^+(C)
-%%   }{
-%%     \int^{S\in J(C)}
-%%     \Hom{\Presheaves{\mathcal{C}}}{S}{X}
-%%   }\\
-%%   \ADefine{
-%%     \Sheafify{X}
-%%   }{
-%%     X^{++}
-%%   }
-%% \end{align*}
-%%
-%% We also have a canonical natural transformation $\Of{\eta}{X\ra X^+}$, defined
-%% using the maximal sieve $\Define{t_C}{\Yoneda{C}}$:
-%% \[
-%%   \HypJ{
-%%     \Define{\eta_C(m)}{
-%%       \Pair{t_C}{
-%%         X(-)(m)
-%%       }
-%%     }
-%%   }{
-%%     \Member{m}{X(C)}
-%%   }
-%% \]
-
-Every category of sheaves on a site gives rise to a topos, which equips us with
-a number of standard constructions, including (among other things) the disjoint
-union of sheaves $X\oplus Y$, the product of sheaves $X\otimes Y$, and the terminal sheaf $\mathbb{1}$.
-
-\section{Operators and signatures}
-
-A valence $\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}$ specifies an expression
-of sort $\tau$ which binds symbols in $\vec{\sigma}$ and variables in
-$\vec{\tau}$.
-
-\[
-  \infer{
-    \IsValence{
-      \MkValence
-        {\sigma_0,\dots,\sigma_m}
-        {\tau_0,\dots,\tau_n}
-        {\tau}
-    }
-  }{
-    \IsSort{\tau}
-&
-    \HypJ{\IsSort{\sigma_i}}{\Leq{i}{m}}
-&
-    \HypJ{\IsSort{\tau_i}}{\Leq{i}{n}}
-  }
-\]
-
-An arity $\MkArity{\vec{v}}{\tau}$ specifies an operator of sort $\tau$ with
-arguments of valences $\vec{v}$. We will call the set of valences $\Valences$,
-and the set of arities $\Arities$.
-
-\[
-  \infer{
-    \IsArity{
-      \MkArity
-        {v_0,\ldots,v_n}
-        {\tau}
-    }
-  }{
-    \IsSort{\tau}
-&
-    \HypJ
-      {\IsValence{v_i}}
-      {\Leq{i}{n}}
-  }
-\]
-
-Fix a family of sheaves of operators
-$\Member{\Operators_a}{\Sheaves{\SCtxSite}}$, indexed by arities
-$\Member{a}{\Arities}$; for each $\Operators_a$, arrows in $\SCtx$ will lift to
-renamings in operators' symbolic parameters. Together, $\Sorts$ and $\Operators$ are said to form a \emph{signature}
-$\Define{\Sigma}{\Pair{\Sorts}{\Operators}}$.
-
-We will write $\IsOperator{\Upsilon}{\vartheta}{a}$ in case
-$\Member{\vartheta}{\Operators_a(\Upsilon)}$; note that this judgment enjoys
-the structural properties of weakening and exchange via functoriality of
-$\Operators$, and strengthening via the sheaf condition. An operator signature
-is defined by specifying, for each arity $a$, the sheaf $\Operators_a$, whose
-``elements'' are the operators of arity $a$.
-
-\paragraph{Examples}
-
-For instance, consider a $\lambda$-calculus with a single sort, $\SortExp$; we
-give its signature $\Sigma_{\lambda}$ by asserting the following about its
-operators:
-\begin{gather*}
-  \IsOperator{\Upsilon}{\mathtt{lam}}{
-    \MkArity{
-      \MkValence{}{\SortExp}{\SortExp}
-    }{\SortExp}
-  }\\
-  \IsOperator{\Upsilon}{\mathtt{fix}}{
-    \MkArity{
-      \MkValence{}{\SortExp}{\SortExp}
-    }{\SortExp}
-  }\\
-  \IsOperator{\Upsilon}{\mathtt{ap}}{
-    \MkArity{
-      \MkValence{}{}{\SortExp},
-      \MkValence{}{}{\SortExp}
-    }{\SortExp}
-  }
-\end{gather*}
-
-These rules correspond to the following definition of $\Operators$:
-\begin{align*}
-  \ADefine{
-    \Operators_{
-       \MkArity{
-        \MkValence{}{\SortExp}{\SortExp}
-      }{\SortExp}
-    }
-  }{
-    \mathbb{1}\oplus\mathbb{1}
-  }
-  \\
-  %
-  \ADefine{
-    \Operators_{
-      \MkArity{
-        \MkValence{}{}{\SortExp},
-        \MkValence{}{}{\SortExp}
-      }{\SortExp}
-    }
-  }{
-    \mathbb{1}
-  }
-\end{align*}
-
-So far, we have made no use of symbols and parameters; however, consider the
-extension of the calculus with assignables (references):
-%
-\begin{gather*}
-  \IsOperator{\Upsilon}{\mathtt{decl}}{
-    \MkArity{
-      \MkValence{}{}{\SortExp},
-      \MkValence{\SortExp}{}{\SortExp}
-    }{
-      \SortExp
-    }
-  }
-\\
-  \IsOperator{\Upsilon,u:\SortExp}{\mathtt{get}[u]}{
-    \MkArity{}{\SortExp}
-  }
-\\
-  \IsOperator{\Upsilon,u:\SortExp}{\mathtt{set}[u]}{
-    \MkArity{
-      \MkValence{}{}{\SortExp}
-    }{
-      \SortExp
-    }
-  }
-\end{gather*}
-
-Let $\mathbf{S}_\tau$ be the sheaf of symbols of sort $\tau$, a subobject of
-the Yoneda embedding of the empty symbol context, $\Yoneda{\cdot}$:
-\[
-  \Define{\mathbf{S}_\tau(\Upsilon)}{
-    \MkSet{
-      u\in\Dom{\Upsilon}
-      \mid
-      \Upsilon(u) \equiv \tau
-    }
-  }
-\]
-%
-Then the above rules correspond to the following family of sheaves:
-\begin{align*}
-  \ADefine{
-    \Operators_{
-      \MkArity{
-        \MkValence{}{}{\SortExp},
-        \MkValence{\SortExp}{}{\SortExp}
-      }{
-        \SortExp
-      }
-    }
-  }{
-    \mathbb{1}
-  }\\
-  \ADefine{
-    \Operators_{
-      \MkArity{}{\SortExp}
-    }
-  }{
-    \mathbf{S}_\SortExp
-  }\\
-  \ADefine{
-    \Operators_{
-      \MkArity{
-        \MkValence{}{}{\SortExp}
-      }{
-        \SortExp
-      }
-    }
-  }{
-    \mathbf{S}_\SortExp
-  }
-\end{align*}
-
-Declaring a new assignable consists in providing an initial value, and an
-expression binding a symbol (which shall represent the assignable in scope).
-Note that the functoriality of $\Operators$ guarantees for any renaming
-$\Of{\varrho}{\Upsilon\inj\Upsilon'}$, a family of lifted maps
-$\Of{\Operators_a(\varrho)}{\Operators_a(\Upsilon)\inj\Operators_a(\Upsilon')}$
-natural in $a$, such that
-$\IsOperator{\Upsilon'}{\Operators(\varrho)(\vartheta)}{a}$ when
-$\IsOperator{\Upsilon}{\vartheta}{a}$. In particular, the renaming
-$\Upsilon,u\mapsto\Upsilon,v$ shall take $\mathtt{get}[u]$ to
-$\mathtt{get}[v]$. In the future, we will write $\Rename{\varrho}{\vartheta}$
-for $\Operators_a(\varrho)(\vartheta)$.
-
-\section{Contexts}
-
-In general, we will have three kinds of context: symbolic (parameter) contexts,
-variable contexts, and metavariable contexts. The symbol contexts have already
-been defined via the comma construction $\SCtx$, but they also admit a
-syntactic characterization,
-\begin{gather*}
-  \infer{
-    \IsSymCtx{\cdot}
-  }{
-  }
-\qquad
-  \infer{
-    \IsSymCtx{\Upsilon,u:\tau}
-  }{
-    \IsSymCtx{\Upsilon}
-&
-    \IsSort{\tau}
-&
-    \NotIn{u}{\Dom\Upsilon}
-  }
-\end{gather*}
-%
-Because, modulo notation, we have $\Member{\Upsilon}{\SCtx}$ just when
-$\IsSymCtx{\Upsilon}$, we will use the syntactic view when it is convenient.
-
-Contexts of variables are similar to contexts of symbols, except that they
-admit \emph{any} renamings, not just the injective ones. As such, when
-$\FinSets$ is the category of finite cardinals and all functions between them, the
-comma construction $\Ctx$ is the category of variable contexts. As above, we
-can give them an equivalent syntactic treatment:
-\begin{gather*}
-  \infer{
-    \IsVarCtx{\cdot}
-  }{
-  }
-\qquad
-  \infer{
-    \IsVarCtx{\Gamma,x:\tau}
-  }{
-    \IsVarCtx{\Gamma}
-&
-    \IsSort{\tau}
-&
-    \NotIn{x}{\Dom\Gamma}
-  }
-\end{gather*}
-
-A metavariable context consists of bindings of $\Sorts$-\emph{valences} to
-metavariables; let $\Define{\Valences}{\MkSet{v\mid \IsValenceUnmoded{v}}}$ be the set
-of valences. Then, the category of metavariable contexts is the comma
-construction $\MCtx$, which likewise admits an equivalent inductive definition:
-\begin{gather*}
-  \infer{
-    \IsMetaCtx{\cdot}
-  }{
-  }
-\qquad
-  \infer{
-    \IsMetaCtx{\Theta,\MV{m}:v}
-  }{
-    \IsMetaCtx{\Theta}
-&
-    \IsValence{v}
-&
-    \NotIn{\MV{m}}{\Dom\Theta}
-  }
-\end{gather*}
-
-\section{Nominal Abstract Binding Trees}
-
-Let the judgment $\IsAbt{\Theta}{\Upsilon}{\Gamma}{M}{\tau}$ presuppose
-\footnote{
-  In the \emph{judgmental method}, as pioneered by Per Martin-L\"of, a form of
-  judgment is propounded first by specifying its range of significance (i.e.\
-  the circumstances under which it shall have a meaning), and then, a meaning
-  explanation (definition) for the judgment is given. The range of significance
-  of a judgment is called its \emph{presupposition}, and the meaning of the
-  judgment may proceed by induction on the evidence that the presupposition
-  obtains. See \cite{schroeder-heister:1987} for a more detailed explanation of
-  presuppositions and the judgmental method.
-}
-$\IsMetaCtx{\Theta}$, $\IsSymCtx\Upsilon$, $\IsVarCtx\Gamma$ and
-$\IsSort{\tau}$, meaning that $M$ is an abstract binding tree of sort $s$, with
-metavariables in $\Theta$, parameters in $\Upsilon$, and variables in $\Gamma$.
-Let the judgment $\IsAbs{\Theta}{\Upsilon}{\Gamma}{E}{v}$ presuppose
-$\IsValence{v}$. Then, the syntax of abstract binding trees is inductively
-defined in four rules:
-
-\[
-  \infer[\RuleIsAbtVar]{
-    \IsAbt{\Theta}{\Upsilon}{\Gamma}{x}{\tau}
-  }{
-    \Lookup{\Gamma}{x}{\tau}
-  }
-\]
-\[
-  \infer[\RuleIsAbtMVar]{
-    \IsAbt{\Theta}{\Upsilon}{\Gamma}{
-      \MApp{\MV{m}}{u_0,\dots,u_m}{M_0,\dots,M_n}
-    }{\tau}
-  }{
-    \begin{array}{l}
-      \Lookup{\Theta}{\MV{m}}{
-        \MkValence{
-          \sigma_0,\dots,\sigma_m
-        }{
-          \tau_0,\dots,\tau_n
-        }{\tau}
-      }
-\\
-      \HypJ{
-        \Lookup{\Upsilon}{u_i}{\sigma_i}
-      }{\Leq{i}{m}}
-\\
-      \HypJ{
-        \IsAbt{\Theta}{\Upsilon}{\Gamma}{M_i}{\tau_i}
-      }{\Leq{i}{n}}
-    \end{array}
-  }
-\]
-\[
-  \infer[\RuleIsAbtApp]{
-    \IsAbt{\Theta}{\Upsilon}{\Gamma}{
-      \App{\vartheta}{E_0,\dots,E_n}
-    }{\tau}
-  }{
-    \begin{array}{l}
-      \IsOperator{\Upsilon}{\vartheta}{
-        \MkArity{v_1,\dots,v_n}{\tau}
-      }
-\\
-      \HypJ{
-        \IsAbs{\Theta}{\Upsilon}{\Gamma}{E_i}{v_i}
-      }{\Leq{i}{n}}
-    \end{array}
-  }
-\]
-\[
-  \infer[\RuleIsAbtAbs]{
-    \IsAbs{\Theta}{\Upsilon}{\Gamma}{
-      \MkAbs{\vec{u}}{\vec{x}}{M}
-    }{
-      \MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}
-    }
-  }{
-    \IsAbt{
-      \Theta
-    }{
-      \Upsilon, \vec{u}:\vec{\sigma}
-    }{
-      \Gamma,\vec{x}:{\vec{\tau}}
-    }{
-      M
-    }{\tau}
-  }
-\]
-
-Abts are identified up to $\alpha$-equivalence.
-
-\subsection{Calculating free variables}
-
-\newcommand\RuleFVVar{\RuleNameVar{\mathbf{FV}}{}}
-\newcommand\RuleFVMVar{\RuleNameMVar{\mathbf{FV}}{}}
-\newcommand\RuleFVApp{\RuleNameApp{\mathbf{FV}}}
-\newcommand\RuleFVAbs{\RuleNameAbs{\mathbf{FV}}{}}
-
-We can easily calculate the variables free in an expression by recursion on its
-structure:
-
-\begin{gather*}
-  \infer[\RuleFVVar]{
-    \CalcFV{x}{\MkSet{x}}
-  }{
-  }\\[6pt]
-  %
-  \infer[\RuleFVMVar]{
-    \CalcFV{
-      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
-    }{
-      \bigcup_{i\leq n} \vec{x}_i
-    }
-  }{
-    \HypJ{
-      \CalcFV{M_i}{\vec{x}_i}
-    }{
-      \Leq{i}{n}
-    }
-  }\\[6pt]
-  %
-  \infer[\RuleFVApp]{
-    \CalcFV{
-      \App{\vartheta}{E_0,\dots,E_n}
-    }{
-      \bigcup_{i\leq n} \vec{x}_i
-    }
-  }{
-    \HypJ{
-      \CalcFV{E_i}{\vec{x}_i}
-    }{
-      \Leq{i}{n}
-    }
-  }\\[6pt]
-  %
-  \infer[\RuleFVAbs]{
-    \CalcFV{
-      \MkAbs{\vec{u}}{\vec{y}}{M}
-    }{
-      \vec{x}\setminus\vec{y}
-    }
-  }{
-    \CalcFV{M}{\vec{x}}
-  }
-\end{gather*}
-
-Because this is a total relation, henceforth we will write $\FV{M}$ for
-$\vec{x}$ when $\CalcFV{M}{\vec{x}}$.
-
-\subsection{Calculating free symbols}
-
-\newcommand\RuleFSVar{\RuleNameVar{\mathbf{FS}}{}}
-\newcommand\RuleFSMVar{\RuleNameMVar{\mathbf{FS}}{}}
-\newcommand\RuleFSApp{\RuleNameApp{\mathbf{FS}}}
-\newcommand\RuleFSAbs{\RuleNameAbs{\mathbf{FS}}{}}
-
-Whereas the calculation of free variables pivoted on the $\RuleFVVar$ rule, the
-calculation of free symbols will pivot on the $\RuleFSApp$ and $\RuleFSMVar$ rules,
-because the only way a symbol can be introduced is as a parameter to an
-operator or as a parameter to a metavariable.
-
-\begin{gather*}
-  \infer[\RuleFSVar]{
-    \CalcFS{x}{\MkSet{}}
-  }{
-  }\\[6pt]
-  %
-  \infer[\RuleFSMVar]{
-    \CalcFS{
-      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
-    }{
-      \vec{u} \cup \bigcup_{i\leq n} \vec{u}_i
-    }
-  }{
-    \HypJ{
-      \CalcFS{M_i}{\vec{u}_i}
-    }{
-      \Leq{i}{n}
-    }
-  }\\[6pt]
-  %
-  \infer[\RuleFSApp]{
-    \CalcFS{
-      \App{\vartheta}{E_0,\dots,E_n}
-    }{
-      \Dom{\Supp{\vartheta}} \cup \bigcup_{i\leq n} \vec{u}_i
-    }
-  }{
-    \HypJ{
-      \CalcFS{E_i}{\vec{u}_i}
-    }{
-      \Leq{i}{n}
-    }
-  }\\[6pt]
-  %
-  \infer[\RuleFSAbs]{
-    \CalcFS{
-      \MkAbs{\vec{v}}{\vec{x}}{M}
-    }{
-      \vec{u}\setminus\vec{v}
-    }
-  }{
-    \CalcFS{M}{\vec{u}}
-  }
-\end{gather*}
-
-Because this is a total relation, henceforth we will write $\FS{M}$ for
-$\vec{u}$ when $\CalcFS{M}{\vec{u}}$.
-
-\subsection{Renaming of symbols}
-
-\newcommand\RuleRenPrefix{\bullet}
-\newcommand\RuleRenVar{\RuleNameVar{\RuleRenPrefix}{}}
-\newcommand\RuleRenMVar{\RuleNameMVar{\RuleRenPrefix}{}}
-\newcommand\RuleRenApp{\RuleNameApp{\RuleRenPrefix}}
-\newcommand\RuleRenAbs{\RuleNameAbs{\RuleRenPrefix}{}}
-
-The only place that symbols appear in our calculus is as parameters to operators
-(unlike variables, symbols are not terms). Therefore, the functorial action of
-the operator sheaf can be lifted into terms by recursion on their
-structure, via a pair of judgments $\IsRenaming{\varrho}{M}{N}$ and
-$\IsRenaming{\varrho}{E}{F}$, presupposing $\Of{\varrho}{\Upsilon\inj\Upsilon'}$,
-and $\IsAbt{\Theta}{\Upsilon}{\Gamma}{M}{\tau}$ and
-$\IsAbt{\Theta}{\Upsilon}{\Gamma}{E}{v}$ respectively:
-
-\begin{gather*}
-  \infer[\RuleRenVar]{
-    \IsRenaming{\varrho}{x}{x}
-  }{
-  }
-\\[6pt]
-  \infer[\RuleRenMVar]{
-    \IsRenaming{\varrho}{
-      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
-    }{
-      \MApp{\MV{m}}{\vec{v}}{N_0,\dots,N_n}
-    }
-  }{
-    \Match{\varrho(\vec{u})}{\vec{v}}
-&
-    \HypJ{
-      \IsRenaming{\varrho}{M_i}{N_i}
-    }{
-      \Leq{i}{n}
-    }
-  }
-\\[6pt]
-  \infer[\RuleRenApp]{
-    \IsRenaming{\varrho}{
-      \App{\vartheta}{E_0,\dots,E_n}
-    }{
-      \App{\Rename{\varrho}{\vartheta}}{F_0,\dots,F_n}
-    }
-  }{
-    \HypJ{
-      \IsRenaming{\varrho}{E_i}{F_i}
-    }{
-      \Leq{i}{n}
-    }
-  }
-\\[6pt]
-  \infer[\RuleRenAbs]{
-    \IsRenaming{\varrho}{
-      \MkAbs{\vec{u}}{\vec{x}}{M}
-    }{
-      \MkAbs{\vec{u}}{\vec{x}}{N}
-    }
-  }{
-    \IsRenaming{\varrho\setminus\vec{u}}{M}{N}
-  }
-\end{gather*}
-
-Above, the notation $\varrho\setminus\vec{u}$ means the omission of the
-variables $\vec{u}$ from the renaming $\varrho$. Because terms are identified up
-to $\alpha$-equivalence, the renaming judgment is functional in its input, and
-so we are justified in writing $\Rename{\varrho}{M}$ for $N$ when
-$\IsRenaming{\varrho}{M}{N}$.
-
-\subsection{Substitution of variables}
-
-\newcommand\RuleSubstPrefix{/}
-\newcommand\RuleSubstVar[1]{\RuleNameVar{\RuleSubstPrefix}{#1}}
-\newcommand\RuleSubstMVar{\RuleNameMVar{\RuleSubstPrefix}{}}
-\newcommand\RuleSubstApp{\RuleNameApp{\RuleSubstPrefix}}
-\newcommand\RuleSubstAbs[1]{\RuleNameAbs{\RuleSubstPrefix}{#1}}
-
-Variable substitution in abts is defined inductively by a pair of judgments,
-$\IsSubst{N}{x}{M}{M'}$ and $\IsSubst{N}{x}{E}{F}$:
-
-\begin{gather*}
-  \infer[\RuleSubstVar{1}]{
-    \IsSubst{N}{x}{y}{N}
-  }{
-    \IsEq{x}{y}
-  }
-  \qquad
-  \infer[\RuleSubstVar{2}]{
-      \IsSubst{N}{x}{y}{y}
-    }{
-      \IsApart{x}{y}
-    }
-\\[6pt]
-  \infer[\RuleSubstMVar]{
-    \IsSubst{N}{x}{
-      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
-    }{
-      \MApp{\MV{m}}{\vec{u}}{M'_0,\dots,M'_n}
-    }
-  }{
-    \HypJ{
-      \IsSubst{N}{x}{M_i}{M'_i}
-    }{
-      \Leq{i}{n}
-    }
-  }
-\\[6pt]
-  \infer[\RuleSubstApp]{
-    \IsSubst{N}{x}{
-      \App{\vartheta}{E_0,\dots,E_n}
-    }{
-      \App{\vartheta}{F_0,\dots,F_n}
-    }
-  }{
-    \HypJ{
-      \IsSubst{N}{x}{E_i}{F_i}
-    }{
-      \Leq{i}{n}
-    }
-  }
-\end{gather*}
-\begin{gather*}
-  \infer[\RuleSubstAbs{1}]{
-    \IsSubst{N}{x}{
-      \MkAbs{\vec{u}}{\vec{y}}{M}
-    }{
-      \MkAbs{\vec{u}}{\vec{y}}{M'}
-    }
-  }{
-    \NotIn{x}{\vec{y}}
-&
-    \IsApart{\vec{u}}{\FS{N}}
-&
-    \IsApart{\vec{y}}{\FV{N}}
-&
-    \IsSubst{N}{x}{M}{M'}
-  }
-\\[6pt]
-  \infer[\RuleSubstAbs{2}]{
-    \IsSubst{N}{x}{
-      \MkAbs{\vec{u}}{\vec{y}}{M}
-    }{
-      \MkAbs{\vec{u}}{\vec{y}}{M}
-    }
-  }{
-    \Member{x}{\vec{y}}
-&
-    \IsApart{\vec{u}}{\FS{N}}
-&
-    \IsApart{\vec{y}}{\FV{N}}
-  }
-\end{gather*}
-
-In the remainder, we will write $\Subst{N}{x}{M}$ for $M'$ when
-$\IsSubst{N}{x}{M}{M'}$, and $\Subst{\vec{N}}{\vec{x}}{M}$ for the simultaneous
-substitution of $\vec{N}$ for $\vec{x}$ in $M$.
-
-\subsection{Substitution of metavariables}
-
-\newcommand\RuleMSubstPrefix{/}
-\newcommand\RuleMSubstSuffix{\texttt{meta}}
-\newcommand\RuleMSubstVar{\RuleNameVar{\RuleMSubstPrefix}{}^\RuleMSubstSuffix}
-\newcommand\RuleMSubstMVar[1]{\RuleNameMVar{\RuleMSubstPrefix}{#1}^\RuleMSubstSuffix}
-\newcommand\RuleMSubstApp{\RuleNameApp{\RuleMSubstPrefix}^\RuleMSubstSuffix}
-\newcommand\RuleMSubstAbs{\RuleNameAbs{\RuleMSubstPrefix}{}^\RuleMSubstSuffix}
-
-Metavariables are substituted by abstractions; since a metavariable may only
-appear in an application expression $\MApp{\MV{m}}{\cdots}{\cdots}$, we will
-instantiate the abstraction at the supplied parameters and arguments.
-Substitution for metavariables is defined inductively by the judgments
-$\IsSubst{E}{\MV{m}}{M}{N}$ and $\IsSubst{E}{\MV{m}}{F}{F'}$:
-
-\begin{gather*}
-  \infer[\RuleMSubstVar]{
-    \IsSubst{E}{\MV{m}}{x}{x}
-  }{
-  }
-\\[6pt]
-  \infer[\RuleMSubstMVar{1}]{
-    \IsSubst{E}{\MV{m}}{
-      \MApp{\MV{n}}{\vec{u}}{M_0,\dots,M_n}
-    }{
-      \MApp{\MV{n}}{\vec{u}}{N_0,\dots,N_n}
-    }
-  }{
-    \IsApart{\MV{m}}{\MV{n}}
-&
-    \HypJ{
-      \IsSubst{E}{\MV{n}}{M_i}{N_i}
-    }{
-      \Leq{i}{n}
-    }
-  }
-\\[6pt]
-  \infer[\RuleMSubstMVar{2}]{
-    \IsSubst{
-      \MkAbs{\vec{u}}{\vec{x}}{N}
-    }{\MV{m}}{
-      \MApp{\MV{n}}{\vec{v}}{\vec{M}}
-    }{
-      N''
-    }
-  }{
-    \IsEq{\MV{m}}{\MV{n}}
-&
-    \IsSubst{\vec{M}}{\vec{x}}{N}{N'}
-&
-    \IsRenaming{\vec{u}\mapsto\vec{v}}{N'}{N''}
-  }\\[6pt]
-  %
-  \infer[\RuleMSubstApp]{
-    \IsSubst{E}{\MV{m}}{
-      \App{\vartheta}{F_0,\dots,F_n}
-    }{
-      \App{\vartheta}{F'_0,\dots,F'_n}
-    }
-  }{
-    \HypJ{
-      \IsSubst{E}{\MV{m}}{F_i}{F'_i}
-    }{
-      \Leq{i}{n}
-    }
-  }
-\\[6pt]
-  \infer[\RuleMSubstAbs]{
-    \IsSubst{E}{\MV{m}}{
-      \MkAbs{\vec{u}}{\vec{x}}{M}
-    }{
-      \MkAbs{\vec{u}}{\vec{x}}{N}
-    }
-  }{
-    \IsApart{\vec{u}}{\FS{E}}
-&
-    \IsApart{\vec{x}}{\FV{E}}
-&
-    \IsSubst{E}{\MV{m}}{M}{N}
-  }
-\end{gather*}
-
-The rules for metavariable substitution are reminiscent of the ``hereditary
-substitutions'' first devised for the Concurrent Logical Framework by Watkins
-in \cite{watkins:2004}.
-%
-As usual, we will write $\Subst{E}{\MV{m}}{M}$ for $N$ when
-$\IsSubst{E}{\MV{m}}{M}{N}$.
-
-\section{Model Theory}
-
-\newcommand\SymPsh{\mathbf{S}}
-\newcommand\VarPsh{\mathbf{V}}
-
-Let $\Define{\HypCat}{\OpCat{(\SCtx\times\Ctx)}}$; then we fix the functor
-category $\widehat{\HypCat}^\Sorts$ as our semantic universe, using the
-notation of the French school,
-$\Define{\widehat{\mathcal{C}}}{\Sets^{\OpCat{\mathcal{C}}}}$.
-%
-Let $\Define{\VarPsh_\tau(\Upsilon\parallel\Gamma)}{\MkSet{x\in\Dom\Gamma\mid
-    \Gamma(x)\equiv\tau}}$ be called the presheaf of variables; additionally, we have
-a presheaf of symbols
-$\Define{\SymPsh_\tau(\Upsilon\parallel\Gamma)}{\MkSet{u\in\Dom\Upsilon\mid
-    \Upsilon(u)\equiv\tau}}$. Lastly, we have the presheaf of operators with arity
-$a$,
-$\Define{\Operators_a(\Upsilon\parallel\Gamma)}{\Operators_a(\Upsilon)}$
-
-\subsection{Substitution monoidal structures}
-
-For an object $\Of{P}{\widehat{\HypCat}^\Sorts}$, we will use the notation
-$P^{[\Gamma]}$ to mean $\prod_{x\in\Dom\Gamma}P_{\Gamma(x)}$; likewise,
-$\SymPsh^{\{\Upsilon\}}$ shall mean $\prod_{u\in\Dom\Upsilon}S_{\Upsilon(u)}$. For a
-presheaf $\Of{A}{\widehat{\HypCat}}$ and a sort-indexed family of presheaves
-$\Of{P}{\widehat{\HypCat}^\Sorts}$, we have an operation $A\bullet P$, defined
-as a coend in the following way:
-\[
-  \Define{(A\bullet P)(\Upsilon\parallel\Gamma)}{
-    \int^{(\Upsilon'\parallel\Delta)\in\HypCat}
-      A(\Upsilon'\parallel\Delta)
-      \times \SymPsh^{\{\Upsilon'\}}(\Upsilon\parallel\Gamma)
-      \times P^{[\Delta]}(\Upsilon\parallel\Gamma)
-  }
-\]
-
-Intuitively, the coend construction $A\bullet P$ provides the data of a
-suspended substitution of an $A$-term's variables by $P$-terms.
-%
-Using this, we can define a tensor $P\odot Q$ for
-$\Of{P,Q}{\widehat{\HypCat}^\Sorts}$ as follows:
-\[
-  \HypJ{
-    \Define{{(P\odot Q)}_{\tau}}{
-      P_\tau\bullet Q
-    }
-\quad
-  }{
-    \Member{\tau}{\Sorts}
-  }
-\]
-
-Then, $V$ is the unit to this tensor. We will say that an object
-$\Of{P}{\widehat{\HypCat}^\Sorts}$ is a $\Sigma$-monoid in case it is equipped
-with the following natural transformations where $\nu$ embeds variables into $P$
-and $\varsigma$ equips $P$ with an operation for simultaneous substitutions of
-variables. Furthermore, $\nu$ and $\varsigma$ induce maps ${\nu}_{\Gamma}$ and
-${\varsigma}^{\tau}_{\Upsilon\parallel\Gamma}$:
-\begin{gather*}
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{V}} \arrow[r, "\IMode\nu"]
-    \&
-    {\IMode{P}}
-    \&
-    {\IMode{P\odot P}}
-      \arrow[l, swap, "\IMode\varsigma"]
-  \end{tikzcd}\\[6pt]
-  %
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{{V}^{[\Gamma]}}}
-      \arrow[r, "\IMode{{\nu}_{\Gamma}}"]
-    \&
-    {\IMode{{P}^{[\Gamma]}}}
-  \end{tikzcd}
-  \qquad
-  \begin{tikzcd}[cramped, sep = large, ampersand replacement = \&]
-    {\IMode{{P_\tau^{\Yoneda{\Upsilon\parallel\Gamma}}\times \SymPsh^{\{\Upsilon\}}\times P^{[\Gamma]}}}}
-      \arrow[r, "\IMode{\varsigma^\tau_{\Upsilon\parallel\Gamma}}"]
-    \&
-    {\IMode{P_\tau}}
-  \end{tikzcd}
-\end{gather*}
-
-\subsection{The signature endofunctor and its algebras}
-
-For each signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we have an
-endofunctor
-$\Of{\SigFun{\Sigma}}{\widehat{\HypCat}^\Sorts\ra\widehat{\HypCat}^\Sorts}$,
-which is defined as follows:
-\begin{align*}
-  \IMode{{\SigFun{\Sigma}(X)}_{\tau}}
-&
-    \triangleq
-    \OMode{
-      \coprod_{
-        \vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}
-      }
-      \prod_{
-        \MkValence
-          {\vec{\sigma}}
-          {\vec{\tau}}
-          {\tau_i}
-        \in\vec{v}
-      }
-      X_{\tau_i}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}
-    }
-\end{align*}
-
-Then, a $\Sigma$-model is a $\Sigma$-monoid $P$ which is equipped with an
-algebra $\Of{\alpha}{\SigFun{\Sigma}(P)\ra P}$, which shall interpret
-applications of each operator.
-
-\subsection{Interpretation of terms}
-
-The metavariable, symbol and variable contexts are interpreted for a model $P$ as
-an environment presheaf in the following way:
-\[
-  \Define{
-    \SemBrackets{
-      \Theta\triangleright\Upsilon\parallel\Gamma
-    }_P
-  }{
-    \left(
-      \prod_{
-        (m:%
-        \MkValence
-          {\vec{\sigma}}
-          {\vec{\tau}}
-          {\tau})
-        \in\Theta
-      }{{P_{\tau}}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}}
-    \right)
-    \times
-    \SymPsh^{\{\Upsilon\}}
-    \times
-    \VarPsh^{[\Gamma]}
-  }
-\]
-
-Then, the interpretation of a term in a model $P$ is a map from its environment
-to $P$:
-\[
-  \Of{
-    \SemBrackets{
-      \Theta\triangleright\Upsilon\parallel\Gamma\vdash M:\tau
-    }_{P}
-  }{
-    \SemBrackets{
-      \Theta\triangleright\Upsilon\parallel\Gamma
-    }_{P}\ra {P}_{\tau}
-  }
-\]
-
-Variables are interpreted by the map
-${\SemBrackets{
-  \Theta\triangleright\Upsilon\parallel\Gamma\vdash x:\tau
-}}_{P}$ which
-projects them from the environment and embeds them into the model. Metavariables
-are resolved by the map
-${\SemBrackets{
-  \Theta\triangleright\Upsilon\parallel\Gamma\vdash\MApp{\MV{m}}{\vec{u}}{\vec{M}}
-  : \tau
-}}_{P}$ (where
-$\Lookup{\Theta}{\MV{m}}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$) which
-projects their interpretation from the environment and instantiates it via
-substitution:
-\[
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{P_\tau}}
-\&[8.0em]
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-    }}
-      \arrow[
-        , rr
-        , dashed
-        , "
-          \OMode{
-            {\SemBrackets{
-              \Theta\triangleright\Upsilon\parallel\Gamma
-              \vdash
-              x:\tau
-            }}_{P}
-          }
-          "
-      ]
-      \ar{drr}[
-        , rotate = -16.5
-        , xshift = +1.5em
-        , yshift = -0.2em
-        , swap
-      ]{
-        {\nu}_{\Gamma} {\pi}_{3}
-      }
-      \ar{dl}[
-        , rotate = +11.5
-        , xshift = -2.75em
-      ]{
-        \IMode{
-          \Tuple{
-            \pi_\MV{m}\,
-            \pi_1,
-            \phi\,
-            \pi_2,
-            \psi
-          }
-        }
-      }
-      \arrow[
-        , l
-        , dashed
-        , swap
-        , "
-          \OMode{
-            {\SemBrackets{
-              \Theta\triangleright\Upsilon\parallel\Gamma
-              \vdash
-              \MApp{\MV{m}}{\vec{u}}{\vec{M}}: \tau
-            }}_{P}
-          }"
-      ]
-\&[-2.0em]
-    \null
-\&[+3.0em]
-    {\IMode{P_\tau}}
-\\
-    {\IMode{
-      {P}_{\tau}^{
-        \Yoneda{\vec{\sigma}\parallel\vec{\tau}}
-      }
-      \times \SymPsh^{\{\vec{\sigma}\}}
-      \times P^{[\vec{\tau}]}
-    }}
-      \arrow[
-        , u
-        , "
-          \IMode{
-            \varsigma^\tau_{\vec{\sigma}\parallel\vec{\tau}}
-          }
-          "
-        ]
-\&
-    \null
-\&
-    \null
-\&
-    {\IMode{P^{[\Gamma]}}}
-      \arrow[
-        , u
-        , swap
-        , "{\pi}_{x}"
-      ]
-\\
-    {\IMode{P^{[\vec{\tau}]}}}
-\&
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-    }}
-      \arrow[
-        , dashed
-        , l
-        , swap
-        , "
-          \Define{
-            \psi\,
-          }{
-            \,\Tuple{
-              {\SemBrackets{
-                \Theta\triangleright\Upsilon\parallel\Gamma\vdash M:\tau
-              }}_{P}
-            }_{(M,\tau)\in (\vec{M},\vec{\tau})}
-          }
-          "]
-\&
-    {\IMode{\SymPsh^{\{\Upsilon\}}}}
-      \arrow[
-        , dashed
-        , r
-        , "\Define{\phi\,}{\,\Tuple{\pi_u}_{u\in\vec{u}}}"
-      ]
-\&
-    {\IMode{\SymPsh^{\{\vec{\sigma}\}}}}
-  \end{tikzcd}
-\]
-
-Interpretation of operator applications is the most complicated. Recall that,
-unlike in standard treatments of universal algebra, our operators are indexed by
-symbol collections; therefore, operators must pass through suitable renamings in
-order to be used in the interpretation. Let us begin by constructing for each
-operator $\IsOperator{\Upsilon}{\vartheta}{a}$ the morphism
-$\SemBrackets{\vartheta}_P$ which shall rename the parameters of the operator
-using the environment:
-\[
-  \begin{tikzcd}[cramped, ampersand replacement = \&, sep = huge]
-    |[alias = Src]|
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-    }}
-      \arrow[
-        , r
-        , swap
-        , "\IMode{\pi_2}"
-      ]
-\&
-    {\IMode{\SymPsh^{\{\Upsilon\}}}}
-      \arrow[
-        , r
-        , swap
-        , "\IMode{\Operators_a(\pi_{(-)})(\vartheta)}"
-      ]
-\&[1.5em]
-    |[alias = Tgt]|
-    {\IMode{\Operators_a}}
-      \arrow[
-        , from = Src
-        , to = Tgt
-        , bend left
-        , dashed
-        , "\OMode{\SemBrackets{\vartheta}_{P}}"
-      ]
-  \end{tikzcd}
-\]
-
-We will proceed using the $\SigFun{\Sigma}$-algebra $\alpha$, as
-follows, by composing it with a morphism $\beta$ from the environment into
-the signature endofunctor, which interprets the syntax of operator applications:
-\[
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{
-      {\SemBrackets{\Theta\triangleright\Upsilon\parallel\Gamma}}_{P}
-    }}
-      \arrow[r, "\IMode{\beta}"]
-\&
-    {\IMode{
-      {\SigFun{\Sigma}(P)}_{\tau}
-    }}
-      \arrow[r, "\IMode{\alpha_\tau}"]
-\&
-    {\IMode{P_\tau}}
-  \end{tikzcd}
-\]
-
-The construction of $\beta$ proceeds by renaming the parameters of the operator
-$\vartheta$ and constructing the (bound) exponentiated arguments $\gamma$ of the
-operator.
-\[
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-    }}
-      \arrow[
-        , r
-        , dashed
-        , "
-          \Define{
-            \beta\,
-          }{
-            \,\Tuple{
-              \SemBrackets{\vartheta}_P, \lambda\gamma
-            }
-          }
-          "
-      ]
-\&[4.0em]
-    {\IMode{
-      \coprod_{
-        \vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}
-      }
-      \prod_{
-        \MkValence{\vec{\sigma}_i}{\vec{\tau}_i}{\tau_i}\in\vec{v}
-      }
-      P_{\tau_i}^{\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}}
-    }}
-  \end{tikzcd}
-\]
-Arguments $\lambda\IMode{{\gamma}_{i}}$ are the exponential transposes (curried
-form) of the composites $\IMode{{\gamma}_{i}}$:
-\[
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-    }}
-      \arrow[
-        , dashed
-        , d
-        , swap
-        , "\OMode{\lambda{\gamma}_{i}}"
-      ]
-\&
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-      \times
-      \Yoneda{
-        \vec{\sigma}_{i}\parallel\vec{\tau}_{i}}
-      }
-    }
-      \arrow[
-        , r
-        , "
-          \IMode{
-            \Tuple{
-              {\pi}_{[1,1]},
-              {\phi}_{i},
-              {\psi}_{i}
-            }
-          }
-          "
-        ]
-      \arrow[
-        , dashed
-        , d
-        , swap
-        , "\OMode{{\gamma}_{i}}"
-      ]
-\&[3.5em]
-    {\IMode{
-      \SemBrackets{
-        \Theta\triangleright\Upsilon,
-        \vec{\sigma}_i\parallel\Gamma,
-        \vec{\tau}_{i}
-      }_{P}
-    }}
-      \ar{dl}[
-        , rotate = +19.25
-        , xshift = -4.5em
-        , yshift = -0.2em
-      ]{\IMode{\SemBrackets{\Theta\triangleright\Upsilon,\vec{\sigma}_{i}\parallel\Gamma,\vec{\tau}_{i}\vdash M_i:\tau_i}_P}}
-\\[2.0em]
-    {\IMode{{P}_{{\tau}_{i}}^{\Yoneda{\vec{\sigma}_{i}\parallel\vec{\tau}_{i}}}}}
-\&
-    {\IMode{P_{{\tau}_{i}}}}
-\&
-  \end{tikzcd}
-\]
-where $\phi_i,\psi_i$ are defined as follows:
-\begin{gather*}
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{
-      {\SemBrackets{\Theta\triangleright\Upsilon\parallel\Gamma}}_{P}
-      \times
-      \Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}
-    }}
-      \arrow[
-        , r
-        , dashed
-        , "
-          \Define
-            {
-              \phi_i\,
-            }
-            {
-              \,
-                \Tuple{\pi_u\circ\pi_1}_{u\in\Dom{\Upsilon}}
-              ,
-                \,\Tuple{\pi_u\circ\pi_{[2,1]}}_{u\in\Dom{\vec{\sigma}_i}}
-            }
-          "
-      ]
-\&[12.0em]
-    {\IMode{\SymPsh^{\{\Upsilon,\vec{\sigma}_i\}}}}
-\\[-1.0em]
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-      \times
-      \Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}
-    }}
-      \arrow[
-        , r
-        , dashed
-        , "
-          \Define{
-            \psi_i\,
-          }{
-            \,
-              \Tuple{\pi_x\circ\pi_1}_{x\in\Dom{\Gamma}}
-            ,
-              \,\Tuple{\pi_x\circ\pi_{[2,1]}}_{x\in\Dom{\vec{\tau}_i}}
-          }
-          "
-      ]
-\&[12.0em]
-    {\IMode{\VarPsh^{[\Gamma,\vec{\tau}_i]}}}
-  \end{tikzcd}
-\end{gather*}
-
-This concludes the interpretation of well-sorted terms into any $\Sigma$-model.
-
-\section{Case Study: Wellformed Sequents}
-
-The representation of telescopes and sequents in a logical framework is
-notoriously difficult; whilst it is possible to use higher-order abstract
-syntax or abts to encode the binding-structure of telescopes and sequents, the
-encoding is sufficiently laborious and obscure that it is not used in practice.
-Crary has demonstrated a first-order encoding of contexts in the Edinburgh
-Logical Framework (ELF) in bijection with actual LF-contexts~\cite{crary:2009},
-which has been successfully used in large-scale mechanization efforts,
-including that of Standard ML~\cite{lee-crary-harper:2007} and the ELF
-itself~\cite{martens-crary:2012}.
-
-We will approach the problem of encoding telescopes and sequents from the
-\emph{refinements} perspective, where a conservative approximation of the
-grammar is first given using the abt logical framework, and then the
-correctness of a code is expressed separately in a judgment that refines the
-existing specification.  Because we have not committed to using the built-in
-binding machinery to express the well-scopedness of telescopes and sequents, we
-are free to use \emph{symbols} in order to model the variables in the context.
-This is in fact quite sensible if we are actually trying to faithfully
-represent the syntax of telescopes and sequents, rather than replace them with
-their counterparts on the meta-level.
-
-The approach outlined above is actually quite similar to the standard practice
-in the Abella proof assistant \cite{gacek:2008}; in Abella, in addition to the
-canonical forms, each type is inhabited by an infinite collection of nominal
-constants or symbols, which can be used in the syntax of contexts. In order to
-specify where such symbols are allowed to occur, a wellformedness predicate (a
-``scheme'') is separately defined. Our setting is similar in that we may also
-use symbols to encode the syntax of contexts, except that here, we have better
-control over their proliferation because, contrary to the state of affairs in
-Abella, symbols are \emph{not} expressions; following
-Harper~\cite{harper:2016}, symbols are meant only to serve as an open-ended
-indexing scheme for operators, not as actual expressions.
-
-There are two reasons that one might want to model sequents in a (meta)-logical
-framework. Perhaps the most obvious reason is to prove things about a logic
-with sequent judgments; this is what motivated Crary's encoding of contexts in
-LF, for instance. However, another common use-case is the practical development
-of a proof assistant in which users can define new sequent rules; such
-user-defined rules may only be compiled into well-behaved tactics in the proof
-system in case they obey certain wellformedness properties including (among
-other things) correct scoping of names. MetaPRL is a salient example of such a
-proof assistant, whose implementation might have benefited from the more
-principled treatment of symbols and metavariables which we present here.
-
-This insight leads the way to a simple abt signature for the theory of
-telescopes and sequents.
-%
-\newcommand\SortTele{\mathtt{tele}}
-\newcommand\SortJdg{\mathtt{jdg}}
-\newcommand\SortProp{\mathtt{prop}}
-\newcommand\SortExpr{\mathtt{exp}}
-\newcommand\OpNil{\mathtt{nil}}
-\newcommand\OpSnoc[1]{\mathtt{snoc}[#1]}
-\newcommand\OpHyp[1]{\mathtt{hyp}[#1]}
-\newcommand\OpSequent{\mathtt{sequent}}
-% %
-\[
-  \begin{array}{lr}
-    \IsSort{\SortTele}&\textit{telescopes}\\
-    \IsSort{\SortExpr}&\textit{expressions}\\
-    \IsSort{\SortProp}&\textit{propositions}\\
-    \IsSort{\SortJdg}&\textit{judgments}\\\\
-    \IsOperator{\Upsilon, u:\SortExpr}{\OpHyp{u}}{
-      \MkArity{}{\SortExpr}
-    }\\\\
-    \IsOperator{\Upsilon}{\OpNil}{
-      \MkArity{}{\SortTele}
-    }\\
-    \IsOperator{\Upsilon, u:\SortExpr}{\OpSnoc{u}}{
-      \MkArity{
-        \MkValence{}{}{\SortTele},
-        \MkValence{}{}{\SortProp}
-      }{\SortTele}
-    }\\\\
-    \IsOperator{\Upsilon}{\OpSequent}{
-      \MkArity{
-        \MkValence{}{}{\SortTele},
-        \MkValence{}{}{\SortProp}
-      }{\SortJdg}
-    }
-  \end{array}
-\]
-
-Suppose we have encoded a fragment of type theory as well:
-\[
-  \begin{array}{l}
-    \IsOperator{\Upsilon}{\mathtt{P}}{
-      \MkArity{}{\SortProp}
-    }
-\\
-    \IsOperator{\Upsilon}{\mathtt{pred}}{
-      \MkArity{
-        \MkValence{}{}{\SortExp}
-      }{\SortProp}
-    }
-\\
-  \end{array}
-\]
-
-Terms written using the abstract syntax will be difficult to read, so let us
-define some notation:
-\begin{align*}
-  \IMode{\diamond}
-&
-  \triangleq \OMode{\OpNil}
-\\
-  \IMode{H, u:P}
-&
-  \triangleq \OMode{
-    \App{\OpSnoc{u}}{H, P}
-  }
-\\
-  \IMode{H\gg A}
-&
-  \triangleq \OMode{
-    \App{\OpSequent}{
-      H, A
-    }
-  }
-\\
-  \IMode{`u}
-&
-  \triangleq \OMode{\OpHyp{u}}
-\end{align*}
-
-Now, we can write the following sequent:
-\[
-  \IsAbt{\cdot}{
-    u:\SortExp,
-    v:\SortExp
-  }{\cdot}{
-    \diamond,u:\mathtt{P},v:\App{\mathtt{pred}}{`u}
-    \gg
-    \App{\mathtt{pred}}{`u}
-  }{\SortJdg}
-\]
-
-\newcommand\OpNabla[1]{\nabla[#1]}
-
-The above sequent has free symbols, but we can close over them by adding a form
-of parametric higher-order judgment (as in \cite{harper:2016}) to our object
-language, indexed by a collection of sorts $\vec{\sigma}$:
-\[
-  \IsOperator{\Upsilon}{\OpNabla{\vec{\sigma}}}{
-    \MkArity{
-      \MkValence{\vec{\sigma}}{}{\SortJdg}
-    }{\SortJdg}
-  }
-\]
-
-Then, we may write a closed sequent judgment as follows:
-\[
-  \IsAbt{\cdot}{\cdot}{\cdot}{
-    \App{\OpNabla{\SortExp,\SortExp}}{
-      \MkAbs{u,v}{}{
-        \diamond,u:\mathtt{P},v:\App{\mathtt{pred}}{`u}
-        \gg
-        \App{\mathtt{pred}}{`u}
-      }
-    }
-  }{\SortJdg}
-\]
-
-\subsection{Refinements for wellformedness}
-
-Having specified an approximation of the grammar of telescopes and sequents in
-the abt logical framework, we can proceed to define proper wellformedness via
-\emph{inductive refinement}~\cite{harper:2016}. The basic idea is to introduce a
-new form of (meta)-judgment $\IsWf{\Upsilon}{\Gamma}{M}{\tau}$ which
-expresses the extrinsic wellformedness properties we wish to verify,
-presupposing $\IsAbt{\cdot}{\Upsilon}{\Gamma}{M}{\tau}$. Additionally, we
-introduce an analogous judgment on abstractions,
-$\IsWf{\Upsilon}{\Gamma}{E}{v}$ presupposing
-$\IsAbt{\cdot}{\Upsilon}{\Gamma}{E}{v}$, defined uniformly as follows:
-\[
-  \infer{
-    \IsWf{\Upsilon}{\Gamma}{
-      \MkAbs{\vec{u}}{\vec{x}}{M}
-    }{
-      \MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}
-    }
-  }{
-    \IsWf{\Upsilon,\vec{u}:\vec{\sigma}}{\Gamma,\vec{x}:\vec{\tau}}{M}{\tau}
-  }
-\]
-
-Likewise, wellformedness for variables is defined uniformly:
-\[
-  \infer{
-    \IsWf{\Upsilon}{\Gamma}{x}{\tau}
-  }{
-  }
-\]
-Note that the refinement for variables $x$ is not trivial, since it is only
-defined in case the presupposition $\IsAbt{\cdot}{\Upsilon}{\Gamma}{x}{\tau}$
-is satisfied.
-
-The remainder of the definition of refinement proceeds by induction on sorts and
-operators. For the sake of this example, we will just stipulate that anything of
-sort $\SortExp$ or $\SortProp$ is grammatical if its subterms are grammatical:
-
-\[
-  \infer[
-    \text{for $\Member{\tau}{\MkSet{\SortExp,\SortProp}}$}
-  ]{
-    \IsWf{\Upsilon}{\Gamma}{
-      \App{\vartheta}{E_0,\dots,E_n}
-    }{\tau}
-  }{
-    \begin{array}{l}
-      \IsOperator{\Upsilon}{\vartheta}{
-        \MkArity{
-          v_0,\dots,v_n
-        }{\tau}
-      }
-\\
-      \HypJ{
-        \IsWf{\Upsilon}{\Gamma}{E_i}{v_i}
-      }{
-        \Leq{i}{n}
-      }
-    \end{array}
-  }
-\]
-
-The refinements for parametric judgment and sequents simply delegate to their
-subterms as well:
-\begin{gather*}
-  \infer{
-    \IsWf{\Upsilon}{\Gamma}{
-      \App{\OpNabla{\vec{\sigma}}}{
-        \MkAbs{\vec{u}}{}{J}
-      }
-    }{\SortJdg}
-  }{
-    \IsWf{\Upsilon,\vec{u}:\vec{\sigma}}{\Gamma}{J}{\SortJdg}
-  }
-\\[6pt]
-  \infer{
-    \IsWf{\Upsilon}{\Gamma}{
-      H\gg A
-    }{\SortJdg}
-  }{
-    \IsWf{\Upsilon}{\Gamma}{H}{\SortTele}
-&
-    \IsWf{\Upsilon}{\Gamma}{A}{\SortProp}
-  }
-\end{gather*}
-
-The refinement for telescopes proceeds by induction:
-\begin{gather*}
-  \infer{
-    \IsWf{\Upsilon}{\Gamma}{\diamond}{\SortTele}
-  }{
-  }
-\qquad
-  \infer{
-    \IsWf{\Upsilon}{\Gamma}{H,u:A}{\SortTele}
-  }{
-    \begin{array}{l}
-      \IsWf{\Upsilon\setminus\MkSet{u}}{\Gamma}{H}{\SortTele}
-\\
-      \IsWf{\Upsilon\setminus\MkSet{u}}{\Gamma}{A}{\SortProp}
-    \end{array}
-  }
-\end{gather*}
-
-Based on these rules, it is easy to show that the example sequent from the
-previous section is wellformed:
-\[
-  \IsWf{\cdot}{\cdot}{
-    \App{\OpNabla{\SortExp,\SortExp}}{
-      \MkAbs{u,v}{}{
-        \diamond,u:\mathtt{P},v:\App{\mathtt{pred}}{`u}
-        \gg
-        \App{\mathtt{pred}}{`u}
-      }
-    }
-  }{\SortJdg}
-\]
-
-\begin{remark}
-  In the full development of a theory of behavioral refinements, the ad-hoc
-  judgment $\IsWf{\Upsilon}{\Gamma}{M}{\tau}$ would be replaced with a general
-  notion of refinement $\IMode{\phi}\sqsubset\tau$ ($\phi$ refines sort
-  $\tau$), and a new judgment $\IMode{\Upsilon}\parallel\IMode{\Phi}\vDash
-  \IMode{M} \in \IMode{\phi}$, such that $\phi$ refines $\tau$ and the
-  refinement context $\Phi$ refines some sort context $\Gamma$. Furthermore, in
-  addition to restricting membership, refinements might also be designed to
-  coarsen equivalence as in \cite{denney:1998}.
-\end{remark}
-
-\section*{Acknowledgements}
-
-The first author wishes to thank Robert Harper for numerous conversations about
-abstract binding trees and symbolic parameters, and for his feedback on a draft
-of this paper; and Andy Pitts for his help in understanding multi-sorted
-nominal sets and atomic sheaves.
-
-\appendix
-\appendixpage
-
 \section{$\SCtxSite$, Nominal Sets, and the Schanuel Topos}
@@
 \end{proof}
-
-\ifdraft{}{
-  \newpage
-  % \nocite{*}
-  \bibliographystyle{abbrv}
-  \bibliography{refs}
-}
-
-\end{document}

### 2016-01-23 body.tex
Split up paper

--- 2016-01-21 ijcar-abts.tex
@@
-\documentclass{llncs}
-
-\usepackage[bw]{notation/modes}
-\usepackage{notation/judgments}
-\usepackage{notation/abts}
-\usepackage{notation/common}
-
-\usepackage{appendix}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{bbold}
-% \usepackage{amsthm}
-\usepackage{catchfilebetweentags}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{ifdraft}
-\usepackage{ifthen}
-\usepackage{mathtools}
-\usepackage{proof}
-\usepackage{scalerel}
-\usepackage{setspace}
-\usepackage{stmaryrd}
-\usepackage{supertabular}
-\usepackage{tikz-cd}
-\usepackage[length=.5]{tikzpfeile}
-\usepackage{todonotes}
-\usepackage{url}
-\usepackage[h]{esvect}
-
-\definecolor{dark-gray}{gray}{0.20}
-
-\newtheorem{thm}{Theorem}[section]
-\newtheorem{prop}[thm]{Proposition}
-\newtheorem{lem}[thm]{Lemma}
-\newtheorem{cor}[thm]{Corollary}
-\numberwithin{equation}{section}
-\renewcommand\vec[1]{\protect\vv{#1}}
-
-\newcommand\HypCat{\mathbf{H}}
-\newcommand\Supp[1]{\mathbf{supp}(#1)}
-\newcommand\SupportsUnmoded[3]{#1 \blacktriangleright_{#2} #3}
-\newcommand\Supports[3]{\SupportsUnmoded{\IMode{#1}}{\IMode{#2}}{\IMode{#3}}}
-
-\newcommand\CalcFV[2]{
-  \FV{\IMode{#1}}\leadsto \OMode{#2}
-}
-
-\newcommand\CalcFS[2]{
-  \FS{\IMode{#1}}\leadsto \OMode{#2}
-}
-
-\newcommand\SortExp{\mathtt{exp}}
-
-\newcommand\IsWf[4]{%
-  \IMode{#1}\parallel%
-  \IMode{#2}\vdash%
-  \IMode{#3}\in_{\mathbf{wf}}\IMode{#4}%
-}
-
-\newcommand\Sheaves[1]{\mathbf{Sh}\left(#1\right)}
-\newcommand\Presheaves[1]{\mathbf{Psh}\left(#1\right)}
-\newcommand\Sheafify[1]{#1^{\#}}
-\newcommand\ConstPsh[1]{\Delta[#1]}
-\newcommand\ConstSh[1]{\Delta^\#[#1]}
-\newcommand\SCtxSite{\SymSets[\Sorts]}
-\newcommand\AtomicCoverage{J_{\mathtt{atm}}}
-
-\DeclareMathOperator{\Perm}{\mathbf{Perm}}
-
-\begin{document}
-
-\title{Syntax and Semantics of Abstract Binding Trees}
-\date{}
-\author{Jon Sterling and Darin Morrison}
-\institute{}
-\maketitle
-
 Abstract binding trees (abts) are a generalization of abstract syntax trees
@@
 \begin{align*}
-  \ADefine{
+  \ADefineEquiv{
     \Operators_{
@@
   %
-  \ADefine{
+  \ADefineEquiv{
     \Operators_{
@@
   }\\
-  \ADefine{
+  \ADefineEquiv{
     \Operators_{
@@
 nominal sets and atomic sheaves.
-
-\appendix
-\appendixpage
-
-\section{$\SCtxSite$, Nominal Sets, and the Schanuel Topos}
-
-Pitts defines a category $\mathbf{Nom}$ of nominal sets in \cite{pitts:2013}
-which function equivalently to the sheaves that we considered above, in the
-case of $\IMode{\Sorts}\cong\IMode{\mathbb{1}}$. Fix a countably infinite set
-of atoms $\mathbb{A}$; then let $\Perm\mathbb{A}$ be the group of permutations
-(i.e.\ autoequivalences) on $\mathbb{A}$.  Considered as a category,
-$\Perm\mathbb{A}$ has a single object $\bullet$ with morphisms
-$\Of{\pi}{\bullet\ra\bullet}$ for each $\Member{\pi}{\Dom{\Perm\mathbb{A}}}$.
-Then, the category of nominal sets $\mathbf{Nom}$ is the subcategory of
-$\Sets^{\Perm\mathbb{A}}$ containing just the presheaves which satisfy a
-\emph{finite support} condition.
-
-The category $\mathbf{Nom}$ is equivalent to the category of covariant sheaves
-on $\SymSets$ under the atomic coverage, called the Schanuel Topos
-$\Define{\mathbb{S}}{\Sheaves{\OpCat{\SymSets}}}$
-\cite{pitts:2013,fiore-staton:2006}; equivalently, the Schanuel topos is the
-subcategory of $\Sets^\SymSets$ containing only pullbacks-preserving functors.
-
-In addition to the unisorted nominal sets framework, Pitts also briefly
-discusses a multi-sorted version of the apparatus in which the countably
-infinite set of atoms $\mathbb{A}$ is equipped with a sort assignment
-$\Of{\mathfrak{s}}{\mathbb{A}\to\Sorts}$ such that all the fibers of
-$\mathfrak{s}$ are countably infinite; then, he defines for any such assignment
-the category $\mathbf{Nom}_\mathfrak{s}$ of $\mathfrak{s}$-sorted nominal sets
-as a subcategory of the category of presheaves on the group of
-$\mathfrak{s}$-respecting permutations on $\mathbb{A}$,
-$\Perm_\mathfrak{s}\mathbb{A}$.
-
-In the same way as the Schanuel topos is equivalent to the unisorted nominal
-sets, we expect to find an equivalence between $\Sheaves{\SCtxSite}$ and the
-limit $\int_{\mathfrak{s}}\mathbf{Nom}_\mathfrak{s}$. At the very least, the
-sheaf topos $\Sheaves{\SCtxSite}$ may serve as a multi-sorted generalization of
-the Schanuel topos, for which we may carry over certain useful results,
-including its characterization as the subcategory of $\Sets^\SCtx$ which
-contains just the pullbacks-preserving functors.
-
-\begin{thm}
-  \label{thm:pullback-preservation}
-  A presheaf $\Of{X}{\Sets^{\SCtx}}$ is a sheaf on $\SCtxSite$ just when it
-  preserves pullbacks.
-\end{thm}
-
-\begin{proof}
-
-  The proof is essentially the same as that of the analogous lemma for the
-  Schanuel topos as presented in \cite[A.2.1.11.h]{johnstone:2002}, but we will
-  give a slightly more detailed version here. The presheaf $X$ is a sheaf on the
-  atomic site $\SCtxSite$ just when, for any renaming
-  $\Of{\varrho}{C\inj D}$ and any $\Member{n}{X(D)}$ if
-  $\IsEq{n\cdot \varrho_0}{n\cdot \varrho_1}$ for all diagrams
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{C} \arrow[r, hook, "\IMode{\varrho}"] &
-      \IMode{D}
-        \arrow[r, hook, shift left, "\IMode{\varrho_0}"]
-        \arrow[r, hook, swap, shift right, "\IMode{\varrho_1}"]
-      &
-      \IMode{E}
-    \end{tikzcd}
-  \]
-  such that $\IsEq{\varrho_0\circ\varrho}{\varrho_1\circ\varrho}$, then there
-  exists a unique $\Member{m}{X(C)}$ such that $\IsEq{m\cdot\varrho}{n}$. In other
-  words, the arrow $X(\varrho)$ is an equalizer, as in the following:
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{X(C)} \arrow[r, "\IMode{X(\varrho)}"] &
-      \IMode{X(D)}
-        \arrow[r, shift left, "\IMode{X(\varrho_0)}"]
-        \arrow[r, swap, shift right, "\IMode{X(\varrho_1)}"]&
-      \IMode{X(E)}
-      \\
-      \IMode{\mathbb{1}}
-        \arrow[u, densely dotted, "\OMode{m}"]
-        \arrow[ur, "\IMode{n}"]
-    \end{tikzcd}
-  \]
-
-  Now, because $\Sets$ is a regular category, $X(\varrho)$ is an equalizer
-  precisely when it is a monomorphism. Therefore, if we have assumed that $X$
-  preserves pullbacks, in order to demonstrate that $X$ is a sheaf, it suffices
-  to show that $X(\varrho)$ is monic. If $X$ preserves pullbacks, then it also
-  preserves monomorphisms, because in any category with pullbacks, $\Of{f}{A\ra
-  B}$ is monic just when the square
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{A}
-        \arrow[r, "\IMode{\ArrId{A}}"]
-        \arrow[d, "\IMode{\ArrId{A}}"] &
-      \IMode{A} \arrow[d, "\IMode{f}"]\\
-      \IMode{A} \arrow[r, "\IMode{f}"] &
-      \IMode{B}
-    \end{tikzcd}
-  \]
-  is a pullback \cite[p.\ 16]{mac-lane-moerdijk:1992}. Because $X$ preserves
-  monomorphisms, and all arrows in $\SCtx$ are monic, then $X(\varrho)$ is
-  monic. Hence, $X$ is a sheaf.
-
-  Next, we must show that if $X$ is a sheaf, then it preserves pullbacks; we
-  will loosely track the proof of Proposition~2.6.15 in \cite{biering:2004}.
-  Fix the following diagram such that it is a pullback square in $\SCtx$:
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{C}
-        \arrow[r, "\IMode{f_j}"]
-        \arrow[d, "\IMode{f_i}"] &
-      \IMode{C_j} \arrow[d, "\IMode{s_j}"]\\
-      \IMode{C_i} \arrow[r, "\IMode{s_i}"] &
-      \IMode{D}
-    \end{tikzcd}
-  \]
-
-  Then it suffices to show that the following diagram is also pullback square
-  in $\Sets$:
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{X(C)}
-        \arrow[r, "\IMode{X(f_j)}"]
-        \arrow[d, "\IMode{X(f_i)}"] &
-      \IMode{X(C_j)} \arrow[d, "\IMode{X(s_j)}"]\\
-      \IMode{X(C_i)} \arrow[r, "\IMode{X(s_i)}"] &
-      \IMode{X(D)}
-    \end{tikzcd}
-  \]
-
-  We have to show that $X(C)$ is, up to equivalence, the pullback of $X(s_i)$
-  along $X(s_j)$ in $\Sets$, namely:
-  \[
-    \IMode{X(C_i)\times_{X(D)}X(C_j)}
-      \cong
-    \OMode{
-      \MkSet{
-        (m_i, m_j)\in X(C_i)\times X(C_j)
-          \mid
-        m_i \cdot s_i = m_j \cdot s_j
-      }
-    }
-  \]
-  The first direction of the equivalence is trivial: for each
-  $\Member{m}{X(C)}$, we can easily exhibit $\Member{(m\cdot f_i,m\cdot
-  f_j)}{X(C_i)\times_{X(D)}X(C_j)}$.
-
-  For the other direction, fix $\Member{(m_i,
-  m_j)}{X(C_i)\times_{X(D)}X(C_j)}$.  Consider the singleton cover on $C$ that
-  contains $\Define{h}{s_i\circ f_i}=\OMode{s_j\circ f_j}$; a matching family
-  $\phi$ for the cover $\MkSet{h}$ is a map that takes $h$ to an object in
-  $X(D)$, and an amalgamation for $\phi$ consists in an object
-  $\Member{m}{X(C)}$ such that $\IsEq{\phi(h)}{m\cdot h}$. Now recall that $X$
-  is a sheaf on $\SCtxSite$ if and only if every matching family has a unique
-  amalgamation; for the matching family $\Define{\phi(h)}{m_i\cdot
-  s_i}=\OMode{m_j\cdot s_j}$, then, we must have a unique amalgamation
-  $\Member{m}{X(C)}$ since $X$ is a sheaf. Therefore, $X$ preserves pullbacks.
-
-\end{proof}
-
-\ifdraft{}{
-  \newpage
-  % \nocite{*}
-  \bibliographystyle{abbrv}
-  \bibliography{refs}
-}
-
-\end{document}

### 2016-01-23 paper.tex
cite Denney
update notation
notation
Split up paper

--- 2015-12-31 paper.tex
@@
 \usepackage{amsthm}
-%\usepackage{agda} % this should come after ams*
 \usepackage{catchfilebetweentags}
@@
 \usepackage{url}
-
-\definecolor{dark-gray}{gray}{0.20}
+\usepackage{ssabt}
 
@@
 
-\newcommand\HypCat{\mathbf{H}}
-\newcommand\Supp[1]{\mathbf{supp}(#1)}
-\newcommand\SupportsUnmoded[3]{#1 \blacktriangleright_{#2} #3}
-\newcommand\Supports[3]{\SupportsUnmoded{\IMode{#1}}{\IMode{#2}}{\IMode{#3}}}
-
-\newcommand\CalcFV[2]{
-  \FV{\IMode{#1}}\leadsto \OMode{#2}
-}
-
-\newcommand\CalcFS[2]{
-  \FS{\IMode{#1}}\leadsto \OMode{#2}
-}
-
-\newcommand\SortExp{\mathtt{exp}}
-
-\newcommand\IsWf[4]{%
-  \IMode{#1}\parallel%
-  \IMode{#2}\vdash%
-  \IMode{#3}\in_{\mathbf{wf}}\IMode{#4}%
-}
-
-\newcommand\Sheaves[1]{\mathbf{Sh}\left(#1\right)}
-\newcommand\Presheaves[1]{\mathbf{Psh}\left(#1\right)}
-\newcommand\Sheafify[1]{#1^{\#}}
-\newcommand\ConstPsh[1]{\Delta[#1]}
-\newcommand\ConstSh[1]{\Delta^\#[#1]}
-\newcommand\SCtxSite{\SymSets[\Sorts]}
-\newcommand\AtomicCoverage{J_{\mathtt{atm}}}
-
-\DeclareMathOperator{\Perm}{\mathbf{Perm}}
-
-\newcommand\RuleNameVar[2]{{#1}_{\textit{var}_{#2}}}
-\newcommand\RuleNameMVar[2]{{#1}_{\textit{mvar}_{#2}}}
-\newcommand\RuleNameApp[1]{{#1}_\textit{app}}
-\newcommand\RuleNameAbs[2]{{#1}_{\textit{abs}_{#2}}}
-
 \begin{document}
@@
 
-Abstract binding trees (abts) are a generalization of abstract syntax trees
-where operators may express variable binding structure as part of their arities.
-Originally formulated by Peter Aczel~\cite{aczel:1978}, unisorted abts have been
-deployed successfully as the uniform syntactic framework for several
-implementations of Constructive Type Theory, including
-Nuprl~\cite{constable:1986}, MetaPRL~\cite{hickey:2003} and
-JonPRL~\cite{jonprl:2015}.
-
-In \emph{Practical Foundations for Programming Languages}~\cite{harper:2016},
-Robert Harper develops the multi-sorted version of abstract binding trees and
-proposes an extension to include families of operators indexed by
-\emph{symbols}, which are, unlike variables, subject to only
-distinctness-preserving renaming and not substitution; furthermore, symbols do
-not appear in the syntax of abts, and are only introduced as parameters to
-operators. This extension to support symbols is essential for a correct
-treatment of programming languages with open sums (e.g. ML's \texttt{exn} type),
-as well as assignable references.
-
-In parallel, M.\ Fiore and his collaborators have developed the categorical
-semantics for several variations of second-order algebraic
-theories~\cite{fiore:1999, fiore:2005, fiore-hur:2010, fiore-mamoud:2010}; of
-these, the simply sorted variants are equivalent to Harper's abstract binding
-trees augmented with a notion of second-order variable (metavariable).
-
-The contribution of this paper is the development of the syntax and semantics
-of multi-sorted nominal abts, an extension of second order universal algebra to
-support symbol-indexed families of operators. Additionally, we have developed
-the categorical semantics for abts formally in Constructive Type Theory using
-the Agda proof assistant~\cite{norell:2007}; we have also developed an
-implementation for abstract binding trees in Standard ML, which we intend to
-integrate into the JonPRL proof assistant \cite{jonprl:2015}.
-
-\section{Categorical Preliminaries}
-
-Fix a set $\Sorts$ of \emph{sorts}; we will say $\IsSort{\tau}$ when
-$\Member{\tau}{\Sorts}$. In this section, we will develop a theory of sets
-varying over collections of $\Sorts$-sorted symbols.
-
-Let $\SymSets$ be the category of finite cardinals and their injective maps;
-then the comma construction $\SCtx$, is the category of contexts of symbols
-whose objects are finite sets of symbols $U$ and sort-assignments
-\begin{tikzcd}[cramped]
-  \IMode{U}\arrow[r, "\IMode{\mathfrak{s}}"] &\IMode{\Sorts}
-\end{tikzcd}, and whose morphisms are
-sort-preserving renamings; we will write $\Upsilon$ for a symbol context
-$(U,\mathfrak{s})$.
-
-\begin{remark}
-
-  To be precise, $\SCtx$ is an abuse of notation for the comma category
-  $\Dom{-}_\SymSets\downarrow \Delta[\Sorts]$, with $\Dom{-}_\SymSets$ the
-  forgetful functor from $\SymSets$ to $\Sets$, and
-  $\Define{\Delta[\Sorts](*)}{\Sorts}$ a constant functor from $\mathbb{1}$ to $\Sets$.
-  Formally, an object of $\SCtx$ is, then, a triple $\Tuple{U, *,
-  \mathfrak{s}}$ with $U$ an object in $\SymSets$, $*$ the unique object in
-  $\mathbb{1}$, and $\mathfrak{s}$ a function $\Dom{U}_\SymSets\ra
-  \Sorts$; an arrow $\Of{\varrho}{\Tuple{U,*,\mathfrak{s}}\ra
-  \Tuple{V,*,\mathfrak{t}}}$ is a commuting triangle like the following:
-  \[
-    \begin{tikzcd}[cramped]
-      \IMode{\Dom{U}_\SymSets}
-        \arrow[rr, "\IMode{\Dom{\rho}_\SymSets}"]
-        \arrow[dr, swap, "\IMode{\mathfrak{s}}"] &&
-      \IMode{\Dom{V}_\SymSets}
-      \arrow[dl, "\IMode{\mathfrak{t}}"] \\
-      &\IMode{\Sorts}
-    \end{tikzcd}
-  \]
-
-  Because the object $\Of{*}{\mathbb{1}}$ is unique, we will always omit it
-  from the data of an object in $\SCtx$.
-\end{remark}
-
-\paragraph{Copresheaves on $\SCtx$}
-
-A presheaf on a category $\mathcal{C}$ is a contravariant functor
-$\OpCat{\mathcal{C}}\ra\Sets$; when $\Match{\mathcal{C}}{\OpCat{\mathcal{D}}}$,
-we call this a \emph{copresheaf} (covariant presheaf) on $\mathcal{D}$.
-Therefore, a copresheaf on $\SCtx$ is a functor $\SCtx\ra\Sets$; colloquially,
-such a copresheaf can be thought of as a set that \emph{varies over} symbol
-contexts.
-
-This leads to a very intuitive setting in which to consider sets
-that vary over collections of $\Sorts$-sorted symbols. In particular, for any
-morphism
-%
-\begin{tikzcd}[cramped]
-  \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
-\end{tikzcd}
-%
-in $\SCtx$, any copresheaf $\Of{X}{\Sets^\SCtx}$ has a symbol-renaming action
-$\Of{X(\varrho)}{X(\Upsilon)\ra X(\Upsilon')}$. When the copresheaf $X$ is
-clear from context, we will write $m\cdot\varrho$ for $X(\varrho)(m)$.
-
-Now, when $\Member{m}{X(\Upsilon)}$ for a copresheaf $\Of{X}{\Sets^\SCtx}$, we
-can define the notion that $m$ depends on \emph{at most} a subcontext
-$\Upsilon'$ of $\Upsilon$; this is called \emph{support}.
-
-\begin{definition}[Support]
-  \label{def:support}
-  When
-  \begin{tikzcd}[cramped]
-    \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
-  \end{tikzcd},
-  we say that $\Upsilon$ supports $\Member{m}{\Upsilon'}$ (written
-  $\Supports{\Upsilon}{\varrho}{m}$) when, for all
-  \begin{tikzcd}[cramped]
-    \IMode{\Upsilon'}\arrow[r, hook, "\IMode{\varrho_1}{,} \IMode{\varrho_2}"] & \IMode{\Upsilon''}
-  \end{tikzcd},
-  if $\IsEq{\varrho_1\circ\varrho}{\varrho_2\circ\varrho}$ then $\IsEq{m\cdot\varrho_1}{m\cdot\varrho_2}$.
-\end{definition}
-
-We can also always calculate the \emph{least support} of a term
-$\Member{m}{X(\Upsilon')}$, written $\Supp{m}$:
-\[
-  \Define{\Supp{m}}{
-    \bigcap_{
-      \begin{tikzcd}[cramped, ampersand replacement = \&]
-        \Upsilon\arrow[r, hook, "\varrho"] \& \Upsilon'
-      \end{tikzcd}
-    }
-    \MkSet{
-      \Upsilon
-      \mid
-      \SupportsUnmoded{\Upsilon}{\varrho}{m}
-    }
-  }
-\]
-Intuitively, $\Supp{m}$ is the exact symbol context that $m$ depends on.
-
-\subsection{Sheaves on the atomic site}
-
-When
-  \begin{tikzcd}[cramped]
-    \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
-  \end{tikzcd}
-such that $\Supports{\Upsilon}{\varrho}{m}$ for some $\Member{m}{X(\Upsilon')}$,
-we would expect that we can work backward to a unique $\Of{m'}{\Upsilon}$
-such that $\IsEq{m'\cdot\varrho}{m}$; whilst it is not in general the case for
-presheaves $X$, this is precisely the sheaf condition for copresheaves on
-$\SCtx$ under the atomic topology \cite[p.~126]{mac-lane-moerdijk:1992}, which
-we will define in Definition~$\ref{def:atomic-topology}$.
-
-The sheaf condition ensures that the notion of support is well-behaved in the
-following sense: if some term lies in two different fibers, then it may be
-\emph{strengthened} uniquely into the fiber of its support.
-
-\begin{definition}[Sieves and matching families]
-  A \emph{sieve} $S$ on an object $C$ is a subfunctor of the Yoneda embedding
-  $\Define{\Yoneda{C}}{\Hom{\mathcal{C}}{-}{C}}$, i.e.\
-  $\IMode{S}\sqsubseteq\IMode{\Yoneda{C}}$. For a presheaf $X$ on the category
-  $\mathcal{C}$, a \emph{matching family} for a sieve $S$ is a natural
-  transformation $\Of{\phi}{S\ra X}$. An \emph{amalgamation} for $\phi$ is an
-  object $\Member{m}{X(C)}$ such that $\IsEq{m\cdot f}{\phi_D(f)}$ for each
-  $\Member{f}{S(D)}$, for all objects $D$.
-\end{definition}
-
-\begin{definition}[Sites and coverages]
-  A \emph{site} is a pair $\Pair{\mathcal{C}}{J}$, with $\mathcal{C}$ a category
-  of ``worlds'' and $J$ a Grothendieck topology (or ``coverage''). A Grothendieck
-  topology $J$ (or \emph{coverage}) is a family of sets of sieves, indexed by
-  $\mathcal{C}$-objects; that is, for each object $\Of{C}{\mathcal{C}}$, $J(C)$
-  is the set of $C$-sieves which will be said to $\emph{cover}$ $C$. A topology
-  $J$ must satisfy several axioms (see \cite{mac-lane-moerdijk:1992} for
-  details).
-\end{definition}
-
-\begin{definition}[Separation and sheafhood]
-  A presheaf on a site $\Pair{\mathcal{C}}{J}$ is a presheaf on $\mathcal{C}$; we
-  say that such a presheaf $X$ is \emph{separated} in case for every sieve
-  $\Member{S}{J(C)}$ and matching family $\Of{\phi}{S\ra X}$, there is at most
-  one amalgamation $\Member{m}{X(C)}$ for $\phi$. Additionally, $X$ is a
-  \emph{sheaf} when there is \emph{exactly one} such amalgamation.
-\end{definition}
-
-\begin{definition}[The atomic topology]
-  \label{def:atomic-topology}
-  The atomic coverage $\AtomicCoverage$ is the one in which all non-empty sieves
-  cover; in other words:
-  \[
-    \Define{
-      \AtomicCoverage(C)
-    }{
-      \MkSet{
-        S\sqsubseteq\Yoneda{C}
-        \mid
-        \textstyle
-        \bigcup_D S(D)\ \textit{non-empty}
-      }
-    }
-  \]
-\end{definition}
-
-In the future, we will write $\SCtxSite$ for the Grothendieck site
-$\Pair{\OpCat{\left(\SCtx\right)}}{\AtomicCoverage}$.
-Sheafhood on the atomic site $\SCtxSite$ amounts to the
-following condition, using Definition~\ref{def:support}:
-\begin{quote}
-  A copresheaf $\Of{X}{\Sets^\SCtx}$ is a sheaf on $\SCtxSite$ just when, for all $\Member{n}{X(\Upsilon')}$, if we have
-  \begin{tikzcd}[cramped]
-    \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
-  \end{tikzcd}
-  and $\Supports{\Upsilon}{\varrho}{n}$, then there is a unique
-  $\Member{m}{X(\Upsilon)}$ such that $\IsEq{m\cdot\varrho}{n}$.
-\end{quote}
-
-Therefore, a sheaf on $\SCtxSite$ is a set that varies over $\Sorts$-sorted
-symbol contexts, which has a well-behaved notion of support. Another way to put
-it is, if $\Member{m}{X(\Upsilon)}$ only ``uses'' some of the symbols available
-at world $\Upsilon$, there's a canonical way to restrict $m$ to a term in the
-world which contains \emph{only} the symbols that it uses.
-
-\paragraph{Constructions on sheaves}
-
-%% For any Grothendieck site $\Match{\mathfrak{S}}{\Pair{\mathcal{C}}{J}}$, the
-%% inclusion $\Of{\iota}{\Sheaves{\mathfrak{S}}\ra\Presheaves{\mathcal{C}}}$ of
-%% sheaves into presheaves has a left adjoint
-%% $\IMode{\Sheafify{(-)}}\dashv\IMode{\iota}$, which takes any presheaf
-%% $\Of{X}{\Presheaves{\mathcal{C}}}$ to a unique sheaf
-%% $\Of{\Sheafify{X}}{\Sheaves{\mathfrak{S}}}$, called the \emph{sheafification}
-%% of $X$. Sheafification is the twice-iterated application of the
-%% \emph{+-construction} $X^+$, which turns a presheaf into a separated presheaf,
-%% and a separated presheaf into a sheaf \cite{mac-lane-moerdijk:1992}:
-%% \begin{align*}
-%%   \ADefine{
-%%     X^+(C)
-%%   }{
-%%     \int^{S\in J(C)}
-%%     \Hom{\Presheaves{\mathcal{C}}}{S}{X}
-%%   }\\
-%%   \ADefine{
-%%     \Sheafify{X}
-%%   }{
-%%     X^{++}
-%%   }
-%% \end{align*}
-%%
-%% We also have a canonical natural transformation $\Of{\eta}{X\ra X^+}$, defined
-%% using the maximal sieve $\Define{t_C}{\Yoneda{C}}$:
-%% \[
-%%   \HypJ{
-%%     \Define{\eta_C(m)}{
-%%       \Pair{t_C}{
-%%         X(-)(m)
-%%       }
-%%     }
-%%   }{
-%%     \Member{m}{X(C)}
-%%   }
-%% \]
-
-Every category of sheaves on a site gives rise to a topos, which equips us with
-a number of standard constructions, including (among other things) the disjoint
-union of sheaves $X\oplus Y$, the product of sheaves $X\otimes Y$, and the terminal sheaf $\mathbb{1}$.
-
-\section{Operators and signatures}
-
-A valence $\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}$ specifies an expression
-of sort $\tau$ which binds symbols in $\vec{\sigma}$ and variables in
-$\vec{\tau}$.
-
-\[
-  \infer{
-    \IsValence{
-      \MkValence
-        {\sigma_0,\dots,\sigma_m}
-        {\tau_0,\dots,\tau_n}
-        {\tau}
-    }
-  }{
-    \IsSort{\tau}
-&
-    \HypJ{\IsSort{\sigma_i}}{\Leq{i}{m}}
-&
-    \HypJ{\IsSort{\tau_i}}{\Leq{i}{n}}
-  }
-\]
-
-An arity $\MkArity{\vec{v}}{\tau}$ specifies an operator of sort $\tau$ with
-arguments of valences $\vec{v}$. We will call the set of valences $\Valences$,
-and the set of arities $\Arities$.
-
-\[
-  \infer{
-    \IsArity{
-      \MkArity
-        {v_0,\ldots,v_n}
-        {\tau}
-    }
-  }{
-    \IsSort{\tau}
-&
-    \HypJ
-      {\IsValence{v_i}}
-      {\Leq{i}{n}}
-  }
-\]
-
-Fix a family of sheaves of operators
-$\Member{\Operators_a}{\Sheaves{\SCtxSite}}$, indexed by arities
-$\Member{a}{\Arities}$; for each $\Operators_a$, arrows in $\SCtx$ will lift to
-renamings in operators' symbolic parameters. Together, $\Sorts$ and $\Operators$ are said to form a \emph{signature}
-$\Define{\Sigma}{\Pair{\Sorts}{\Operators}}$.
-
-We will write $\IsOperator{\Upsilon}{\vartheta}{a}$ in case
-$\Member{\vartheta}{\Operators_a(\Upsilon)}$; note that this judgment enjoys
-the structural properties of weakening and exchange via functoriality of
-$\Operators$, and strengthening via the sheaf condition. An operator signature
-is defined by specifying, for each arity $a$, the sheaf $\Operators_a$, whose
-``elements'' are the operators of arity $a$.
-
-\paragraph{Examples}
-
-For instance, consider a $\lambda$-calculus with a single sort, $\SortExp$; we
-give its signature $\Sigma_{\lambda}$ by asserting the following about its
-operators:
-\begin{gather*}
-  \IsOperator{\Upsilon}{\mathtt{lam}}{
-    \MkArity{
-      \MkValence{}{\SortExp}{\SortExp}
-    }{\SortExp}
-  }\\
-  \IsOperator{\Upsilon}{\mathtt{fix}}{
-    \MkArity{
-      \MkValence{}{\SortExp}{\SortExp}
-    }{\SortExp}
-  }\\
-  \IsOperator{\Upsilon}{\mathtt{ap}}{
-    \MkArity{
-      \MkValence{}{}{\SortExp},
-      \MkValence{}{}{\SortExp}
-    }{\SortExp}
-  }
-\end{gather*}
-
-These rules correspond to the following definition of $\Operators$:
-\begin{align*}
-  \ADefine{
-    \Operators_{
-       \MkArity{
-        \MkValence{}{\SortExp}{\SortExp}
-      }{\SortExp}
-    }
-  }{
-    \mathbb{1}\oplus\mathbb{1}
-  }
-  \\
-  %
-  \ADefine{
-    \Operators_{
-      \MkArity{
-        \MkValence{}{}{\SortExp},
-        \MkValence{}{}{\SortExp}
-      }{\SortExp}
-    }
-  }{
-    \mathbb{1}
-  }
-\end{align*}
-
-So far, we have made no use of symbols and parameters; however, consider the
-extension of the calculus with assignables (references):
-%
-\begin{gather*}
-  \IsOperator{\Upsilon}{\mathtt{decl}}{
-    \MkArity{
-      \MkValence{}{}{\SortExp},
-      \MkValence{\SortExp}{}{\SortExp}
-    }{
-      \SortExp
-    }
-  }
-\\
-  \IsOperator{\Upsilon,u:\SortExp}{\mathtt{get}[u]}{
-    \MkArity{}{\SortExp}
-  }
-\\
-  \IsOperator{\Upsilon,u:\SortExp}{\mathtt{set}[u]}{
-    \MkArity{
-      \MkValence{}{}{\SortExp}
-    }{
-      \SortExp
-    }
-  }
-\end{gather*}
-
-Let $\mathbf{S}_\tau$ be the sheaf of symbols of sort $\tau$, a subobject of
-the Yoneda embedding of the empty symbol context, $\Yoneda{\cdot}$:
-\[
-  \Define{\mathbf{S}_\tau(\Upsilon)}{
-    \MkSet{
-      u\in\Dom{\Upsilon}
-      \mid
-      \Upsilon(u) \equiv \tau
-    }
-  }
-\]
-%
-Then the above rules correspond to the following family of sheaves:
-\begin{align*}
-  \ADefine{
-    \Operators_{
-      \MkArity{
-        \MkValence{}{}{\SortExp},
-        \MkValence{\SortExp}{}{\SortExp}
-      }{
-        \SortExp
-      }
-    }
-  }{
-    \mathbb{1}
-  }\\
-  \ADefine{
-    \Operators_{
-      \MkArity{}{\SortExp}
-    }
-  }{
-    \mathbf{S}_\SortExp
-  }\\
-  \ADefine{
-    \Operators_{
-      \MkArity{
-        \MkValence{}{}{\SortExp}
-      }{
-        \SortExp
-      }
-    }
-  }{
-    \mathbf{S}_\SortExp
-  }
-\end{align*}
-
-Declaring a new assignable consists in providing an initial value, and an
-expression binding a symbol (which shall represent the assignable in scope).
-Note that the functoriality of $\Operators$ guarantees for any renaming
-$\Of{\varrho}{\Upsilon\inj\Upsilon'}$, a family of lifted maps
-$\Of{\Operators_a(\varrho)}{\Operators_a(\Upsilon)\inj\Operators_a(\Upsilon')}$
-natural in $a$, such that
-$\IsOperator{\Upsilon'}{\Operators(\varrho)(\vartheta)}{a}$ when
-$\IsOperator{\Upsilon}{\vartheta}{a}$. In particular, the renaming
-$\Upsilon,u\mapsto\Upsilon,v$ shall take $\mathtt{get}[u]$ to
-$\mathtt{get}[v]$. In the future, we will write $\Rename{\varrho}{\vartheta}$
-for $\Operators_a(\varrho)(\vartheta)$.
-
-\section{Contexts}
-
-In general, we will have three kinds of context: symbolic (parameter) contexts,
-variable contexts, and metavariable contexts. The symbol contexts have already
-been defined via the comma construction $\SCtx$, but they also admit a
-syntactic characterization,
-\begin{gather*}
-  \infer{
-    \IsSymCtx{\cdot}
-  }{
-  }
-\qquad
-  \infer{
-    \IsSymCtx{\Upsilon,u:\tau}
-  }{
-    \IsSymCtx{\Upsilon}
-&
-    \IsSort{\tau}
-&
-    \NotIn{u}{\Dom\Upsilon}
-  }
-\end{gather*}
-%
-Because, modulo notation, we have $\Member{\Upsilon}{\SCtx}$ just when
-$\IsSymCtx{\Upsilon}$, we will use the syntactic view when it is convenient.
-
-Contexts of variables are similar to contexts of symbols, except that they
-admit \emph{any} renamings, not just the injective ones. As such, when
-$\FinSets$ is the category of finite cardinals and all functions between them, the
-comma construction $\Ctx$ is the category of variable contexts. As above, we
-can give them an equivalent syntactic treatment:
-\begin{gather*}
-  \infer{
-    \IsVarCtx{\cdot}
-  }{
-  }
-\qquad
-  \infer{
-    \IsVarCtx{\Gamma,x:\tau}
-  }{
-    \IsVarCtx{\Gamma}
-&
-    \IsSort{\tau}
-&
-    \NotIn{x}{\Dom\Gamma}
-  }
-\end{gather*}
-
-A metavariable context consists of bindings of $\Sorts$-\emph{valences} to
-metavariables; let $\Define{\Valences}{\MkSet{v\mid \IsValenceUnmoded{v}}}$ be the set
-of valences. Then, the category of metavariable contexts is the comma
-construction $\MCtx$, which likewise admits an equivalent inductive definition:
-\begin{gather*}
-  \infer{
-    \IsMetaCtx{\cdot}
-  }{
-  }
-\qquad
-  \infer{
-    \IsMetaCtx{\Theta,\MV{m}:v}
-  }{
-    \IsMetaCtx{\Theta}
-&
-    \IsValence{v}
-&
-    \NotIn{\MV{m}}{\Dom\Theta}
-  }
-\end{gather*}
-
-\section{Nominal Abstract Binding Trees}
-
-Let the judgment $\IsAbt{\Theta}{\Upsilon}{\Gamma}{M}{\tau}$ presuppose
-\footnote{
-  In the \emph{judgmental method}, as pioneered by Per Martin-L\"of, a form of
-  judgment is propounded first by specifying its range of significance (i.e.\
-  the circumstances under which it shall have a meaning), and then, a meaning
-  explanation (definition) for the judgment is given. The range of significance
-  of a judgment is called its \emph{presupposition}, and the meaning of the
-  judgment may proceed by induction on the evidence that the presupposition
-  obtains. See \cite{schroeder-heister:1987} for a more detailed explanation of
-  presuppositions and the judgmental method.
-}
-$\IsMetaCtx{\Theta}$, $\IsSymCtx\Upsilon$, $\IsVarCtx\Gamma$ and
-$\IsSort{\tau}$, meaning that $M$ is an abstract binding tree of sort $s$, with
-metavariables in $\Theta$, parameters in $\Upsilon$, and variables in $\Gamma$.
-Let the judgment $\IsAbs{\Theta}{\Upsilon}{\Gamma}{E}{v}$ presuppose
-$\IsValence{v}$. Then, the syntax of abstract binding trees is inductively
-defined in four rules:
-
-\newcommand\RuleIsAbtVar{\RuleNameVar{\vdash}{}}
-\newcommand\RuleIsAbtMVar{\RuleNameMVar{\vdash}{}}
-\newcommand\RuleIsAbtApp{\RuleNameApp{\vdash}}
-\newcommand\RuleIsAbtAbs{\RuleNameAbs{\vdash}{}}
-
-\[
-  \infer[\RuleIsAbtVar]{
-    \IsAbt{\Theta}{\Upsilon}{\Gamma}{x}{\tau}
-  }{
-    \Lookup{\Gamma}{x}{\tau}
-  }
-\]
-\[
-  \infer[\RuleIsAbtMVar]{
-    \IsAbt{\Theta}{\Upsilon}{\Gamma}{
-      \MApp{\MV{m}}{u_0,\dots,u_m}{M_0,\dots,M_n}
-    }{\tau}
-  }{
-    \begin{array}{l}
-      \Lookup{\Theta}{\MV{m}}{
-        \MkValence{
-          \sigma_0,\dots,\sigma_m
-        }{
-          \tau_0,\dots,\tau_n
-        }{\tau}
-      }
-\\
-      \HypJ{
-        \Lookup{\Upsilon}{u_i}{\sigma_i}
-      }{\Leq{i}{m}}
-\\
-      \HypJ{
-        \IsAbt{\Theta}{\Upsilon}{\Gamma}{M_i}{\tau_i}
-      }{\Leq{i}{n}}
-    \end{array}
-  }
-\]
-\[
-  \infer[\RuleIsAbtApp]{
-    \IsAbt{\Theta}{\Upsilon}{\Gamma}{
-      \App{\vartheta}{E_0,\dots,E_n}
-    }{\tau}
-  }{
-    \begin{array}{l}
-      \IsOperator{\Upsilon}{\vartheta}{
-        \MkArity{v_1,\dots,v_n}{\tau}
-      }
-\\
-      \HypJ{
-        \IsAbs{\Theta}{\Upsilon}{\Gamma}{E_i}{v_i}
-      }{\Leq{i}{n}}
-    \end{array}
-  }
-\]
-\[
-  \infer[\RuleIsAbtAbs]{
-    \IsAbs{\Theta}{\Upsilon}{\Gamma}{
-      \MkAbs{\vec{u}}{\vec{x}}{M}
-    }{
-      \MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}
-    }
-  }{
-    \IsAbt{
-      \Theta
-    }{
-      \Upsilon, \vec{u}:\vec{\sigma}
-    }{
-      \Gamma,\vec{x}:{\vec{\tau}}
-    }{
-      M
-    }{\tau}
-  }
-\]
-
-Abts are identified up to $\alpha$-equivalence.
-
-\subsection{Calculating free variables}
-
-\newcommand\RuleFVVar{\RuleNameVar{\mathbf{FV}}{}}
-\newcommand\RuleFVMVar{\RuleNameMVar{\mathbf{FV}}{}}
-\newcommand\RuleFVApp{\RuleNameApp{\mathbf{FV}}}
-\newcommand\RuleFVAbs{\RuleNameAbs{\mathbf{FV}}{}}
-
-We can easily calculate the variables free in an expression by recursion on its
-structure:
-
-\begin{gather*}
-  \infer[\RuleFVVar]{
-    \CalcFV{x}{\MkSet{x}}
-  }{
-  }\\[6pt]
-  %
-  \infer[\RuleFVMVar]{
-    \CalcFV{
-      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
-    }{
-      \bigcup_{i\leq n} \vec{x}_i
-    }
-  }{
-    \HypJ{
-      \CalcFV{M_i}{\vec{x}_i}
-    }{
-      \Leq{i}{n}
-    }
-  }\\[6pt]
-  %
-  \infer[\RuleFVApp]{
-    \CalcFV{
-      \App{\vartheta}{E_0,\dots,E_n}
-    }{
-      \bigcup_{i\leq n} \vec{x}_i
-    }
-  }{
-    \HypJ{
-      \CalcFV{E_i}{\vec{x}_i}
-    }{
-      \Leq{i}{n}
-    }
-  }\\[6pt]
-  %
-  \infer[\RuleFVAbs]{
-    \CalcFV{
-      \MkAbs{\vec{u}}{\vec{y}}{M}
-    }{
-      \vec{x}\setminus\vec{y}
-    }
-  }{
-    \CalcFV{M}{\vec{x}}
-  }
-\end{gather*}
-
-Because this is a total relation, henceforth we will write $\FV{M}$ for
-$\vec{x}$ when $\CalcFV{M}{\vec{x}}$.
-
-\subsection{Calculating free symbols}
-
-\newcommand\RuleFSVar{\RuleNameVar{\mathbf{FS}}{}}
-\newcommand\RuleFSMVar{\RuleNameMVar{\mathbf{FS}}{}}
-\newcommand\RuleFSApp{\RuleNameApp{\mathbf{FS}}}
-\newcommand\RuleFSAbs{\RuleNameAbs{\mathbf{FS}}{}}
-
-Whereas the calculation of free variables pivoted on the $\RuleFVVar$ rule, the
-calculation of free symbols will pivot on the $\RuleFSApp$ and $\RuleFSMVar$ rules,
-because the only way a symbol can be introduced is as a parameter to an
-operator or as a parameter to a metavariable.
-
-\begin{gather*}
-  \infer[\RuleFSVar]{
-    \CalcFS{x}{\MkSet{}}
-  }{
-  }\\[6pt]
-  %
-  \infer[\RuleFSMVar]{
-    \CalcFS{
-      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
-    }{
-      \vec{u} \cup \bigcup_{i\leq n} \vec{u}_i
-    }
-  }{
-    \HypJ{
-      \CalcFS{M_i}{\vec{u}_i}
-    }{
-      \Leq{i}{n}
-    }
-  }\\[6pt]
-  %
-  \infer[\RuleFSApp]{
-    \CalcFS{
-      \App{\vartheta}{E_0,\dots,E_n}
-    }{
-      \Dom{\Supp{\vartheta}} \cup \bigcup_{i\leq n} \vec{u}_i
-    }
-  }{
-    \HypJ{
-      \CalcFS{E_i}{\vec{u}_i}
-    }{
-      \Leq{i}{n}
-    }
-  }\\[6pt]
-  %
-  \infer[\RuleFSAbs]{
-    \CalcFS{
-      \MkAbs{\vec{v}}{\vec{x}}{M}
-    }{
-      \vec{u}\setminus\vec{v}
-    }
-  }{
-    \CalcFS{M}{\vec{u}}
-  }
-\end{gather*}
-
-Because this is a total relation, henceforth we will write $\FS{M}$ for
-$\vec{u}$ when $\CalcFS{M}{\vec{u}}$.
-
-\subsection{Renaming of symbols}
-
-\newcommand\RuleRenPrefix{\bullet}
-\newcommand\RuleRenVar{\RuleNameVar{\RuleRenPrefix}{}}
-\newcommand\RuleRenMVar{\RuleNameMVar{\RuleRenPrefix}{}}
-\newcommand\RuleRenApp{\RuleNameApp{\RuleRenPrefix}}
-\newcommand\RuleRenAbs{\RuleNameAbs{\RuleRenPrefix}{}}
-
-The only place that symbols appear in our calculus is as parameters to operators
-(unlike variables, symbols are not terms). Therefore, the functorial action of
-the operator sheaf can be lifted into terms by recursion on their
-structure, via a pair of judgments $\IsRenaming{\varrho}{M}{N}$ and
-$\IsRenaming{\varrho}{E}{F}$, presupposing $\Of{\varrho}{\Upsilon\inj\Upsilon'}$,
-and $\IsAbt{\Theta}{\Upsilon}{\Gamma}{M}{\tau}$ and
-$\IsAbt{\Theta}{\Upsilon}{\Gamma}{E}{v}$ respectively:
-
-\begin{gather*}
-  \infer[\RuleRenVar]{
-    \IsRenaming{\varrho}{x}{x}
-  }{
-  }
-\\[6pt]
-  \infer[\RuleRenMVar]{
-    \IsRenaming{\varrho}{
-      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
-    }{
-      \MApp{\MV{m}}{\vec{v}}{N_0,\dots,N_n}
-    }
-  }{
-    \Match{\varrho(\vec{u})}{\vec{v}}
-&
-    \HypJ{
-      \IsRenaming{\varrho}{M_i}{N_i}
-    }{
-      \Leq{i}{n}
-    }
-  }
-\\[6pt]
-  \infer[\RuleRenApp]{
-    \IsRenaming{\varrho}{
-      \App{\vartheta}{E_0,\dots,E_n}
-    }{
-      \App{\Rename{\varrho}{\vartheta}}{F_0,\dots,F_n}
-    }
-  }{
-    \HypJ{
-      \IsRenaming{\varrho}{E_i}{F_i}
-    }{
-      \Leq{i}{n}
-    }
-  }
-\\[6pt]
-  \infer[\RuleRenAbs]{
-    \IsRenaming{\varrho}{
-      \MkAbs{\vec{u}}{\vec{x}}{M}
-    }{
-      \MkAbs{\vec{u}}{\vec{x}}{N}
-    }
-  }{
-    \IsRenaming{\varrho\setminus\vec{u}}{M}{N}
-  }
-\end{gather*}
-
-Above, the notation $\varrho\setminus\vec{u}$ means the omission of the
-variables $\vec{u}$ from the renaming $\varrho$. Because terms are identified up
-to $\alpha$-equivalence, the renaming judgment is functional in its input, and
-so we are justified in writing $\Rename{\varrho}{M}$ for $N$ when
-$\IsRenaming{\varrho}{M}{N}$.
-
-\subsection{Substitution of variables}
-
-\newcommand\RuleSubstPrefix{/}
-\newcommand\RuleSubstVar[1]{\RuleNameVar{\RuleSubstPrefix}{#1}}
-\newcommand\RuleSubstMVar{\RuleNameMVar{\RuleSubstPrefix}{}}
-\newcommand\RuleSubstApp{\RuleNameApp{\RuleSubstPrefix}}
-\newcommand\RuleSubstAbs[1]{\RuleNameAbs{\RuleSubstPrefix}{#1}}
-
-Variable substitution in abts is defined inductively by a pair of judgments,
-$\IsSubst{N}{x}{M}{M'}$ and $\IsSubst{N}{x}{E}{F}$:
-
-\begin{gather*}
-  \infer[\RuleSubstVar{1}]{
-    \IsSubst{N}{x}{y}{N}
-  }{
-    \IsEq{x}{y}
-  }
-  \qquad
-  \infer[\RuleSubstVar{2}]{
-      \IsSubst{N}{x}{y}{y}
-    }{
-      \IsApart{x}{y}
-    }
-\\[6pt]
-  \infer[\RuleSubstMVar]{
-    \IsSubst{N}{x}{
-      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
-    }{
-      \MApp{\MV{m}}{\vec{u}}{M'_0,\dots,M'_n}
-    }
-  }{
-    \HypJ{
-      \IsSubst{N}{x}{M_i}{M'_i}
-    }{
-      \Leq{i}{n}
-    }
-  }
-\\[6pt]
-  \infer[\RuleSubstApp]{
-    \IsSubst{N}{x}{
-      \App{\vartheta}{E_0,\dots,E_n}
-    }{
-      \App{\vartheta}{F_0,\dots,F_n}
-    }
-  }{
-    \HypJ{
-      \IsSubst{N}{x}{E_i}{F_i}
-    }{
-      \Leq{i}{n}
-    }
-  }
-\end{gather*}
-\begin{gather*}
-  \infer[\RuleSubstAbs{1}]{
-    \IsSubst{N}{x}{
-      \MkAbs{\vec{u}}{\vec{y}}{M}
-    }{
-      \MkAbs{\vec{u}}{\vec{y}}{M'}
-    }
-  }{
-    \NotIn{x}{\vec{y}}
-&
-    \IsApart{\vec{u}}{\FS{N}}
-&
-    \IsApart{\vec{y}}{\FV{N}}
-&
-    \IsSubst{N}{x}{M}{M'}
-  }
-\\[6pt]
-  \infer[\RuleSubstAbs{2}]{
-    \IsSubst{N}{x}{
-      \MkAbs{\vec{u}}{\vec{y}}{M}
-    }{
-      \MkAbs{\vec{u}}{\vec{y}}{M}
-    }
-  }{
-    \Member{x}{\vec{y}}
-&
-    \IsApart{\vec{u}}{\FS{N}}
-&
-    \IsApart{\vec{y}}{\FV{N}}
-  }
-\end{gather*}
-
-In the remainder, we will write $\Subst{N}{x}{M}$ for $M'$ when
-$\IsSubst{N}{x}{M}{M'}$, and $\Subst{\vec{N}}{\vec{x}}{M}$ for the simultaneous
-substitution of $\vec{N}$ for $\vec{x}$ in $M$.
-
-\subsection{Substitution of metavariables}
-
-\newcommand\RuleMSubstPrefix{/}
-\newcommand\RuleMSubstSuffix{\texttt{meta}}
-\newcommand\RuleMSubstVar{\RuleNameVar{\RuleMSubstPrefix}{}^\RuleMSubstSuffix}
-\newcommand\RuleMSubstMVar[1]{\RuleNameMVar{\RuleMSubstPrefix}{#1}^\RuleMSubstSuffix}
-\newcommand\RuleMSubstApp{\RuleNameApp{\RuleMSubstPrefix}^\RuleMSubstSuffix}
-\newcommand\RuleMSubstAbs{\RuleNameAbs{\RuleMSubstPrefix}{}^\RuleMSubstSuffix}
-
-Metavariables are substituted by abstractions; since a metavariable may only
-appear in an application expression $\MApp{\MV{m}}{\cdots}{\cdots}$, we will
-instantiate the abstraction at the supplied parameters and arguments.
-Substitution for metavariables is defined inductively by the judgments
-$\IsSubst{E}{\MV{m}}{M}{N}$ and $\IsSubst{E}{\MV{m}}{F}{F'}$:
-
-\begin{gather*}
-  \infer[\RuleMSubstVar]{
-    \IsSubst{E}{\MV{m}}{x}{x}
-  }{
-  }
-\\[6pt]
-  \infer[\RuleMSubstMVar{1}]{
-    \IsSubst{E}{\MV{m}}{
-      \MApp{\MV{n}}{\vec{u}}{M_0,\dots,M_n}
-    }{
-      \MApp{\MV{n}}{\vec{u}}{N_0,\dots,N_n}
-    }
-  }{
-    \IsApart{\MV{m}}{\MV{n}}
-&
-    \HypJ{
-      \IsSubst{E}{\MV{n}}{M_i}{N_i}
-    }{
-      \Leq{i}{n}
-    }
-  }
-\\[6pt]
-  \infer[\RuleMSubstMVar{2}]{
-    \IsSubst{
-      \MkAbs{\vec{u}}{\vec{x}}{N}
-    }{\MV{m}}{
-      \MApp{\MV{n}}{\vec{v}}{\vec{M}}
-    }{
-      N''
-    }
-  }{
-    \IsEq{\MV{m}}{\MV{n}}
-&
-    \IsSubst{\vec{M}}{\vec{x}}{N}{N'}
-&
-    \IsRenaming{\vec{u}\mapsto\vec{v}}{N'}{N''}
-  }\\[6pt]
-  %
-  \infer[\RuleMSubstApp]{
-    \IsSubst{E}{\MV{m}}{
-      \App{\vartheta}{F_0,\dots,F_n}
-    }{
-      \App{\vartheta}{F'_0,\dots,F'_n}
-    }
-  }{
-    \HypJ{
-      \IsSubst{E}{\MV{m}}{F_i}{F'_i}
-    }{
-      \Leq{i}{n}
-    }
-  }
-\\[6pt]
-  \infer[\RuleMSubstAbs]{
-    \IsSubst{E}{\MV{m}}{
-      \MkAbs{\vec{u}}{\vec{x}}{M}
-    }{
-      \MkAbs{\vec{u}}{\vec{x}}{N}
-    }
-  }{
-    \IsApart{\vec{u}}{\FS{E}}
-&
-    \IsApart{\vec{x}}{\FV{E}}
-&
-    \IsSubst{E}{\MV{m}}{M}{N}
-  }
-\end{gather*}
-
-The rules for metavariable substitution are reminiscent of the ``hereditary
-substitutions'' first devised for the Concurrent Logical Framework by Watkins
-in \cite{watkins:2004}.
-%
-As usual, we will write $\Subst{E}{\MV{m}}{M}$ for $N$ when
-$\IsSubst{E}{\MV{m}}{M}{N}$.
-
-\section{Model Theory}
-
-\newcommand\SymPsh{\mathbf{S}}
-\newcommand\VarPsh{\mathbf{V}}
-
-Let $\Define{\HypCat}{\OpCat{(\SCtx\times\Ctx)}}$; then we fix the functor
-category $\widehat{\HypCat}^\Sorts$ as our semantic universe, using the
-notation of the French school,
-$\Define{\widehat{\mathcal{C}}}{\Sets^{\OpCat{\mathcal{C}}}}$.
-%
-Let $\Define{\VarPsh_\tau(\Upsilon\parallel\Gamma)}{\MkSet{x\in\Dom\Gamma\mid
-    \Gamma(x)\equiv\tau}}$ be called the presheaf of variables; additionally, we have
-a presheaf of symbols
-$\Define{\SymPsh_\tau(\Upsilon\parallel\Gamma)}{\MkSet{u\in\Dom\Upsilon\mid
-    \Upsilon(u)\equiv\tau}}$. Lastly, we have the presheaf of operators with arity
-$a$,
-$\Define{\Operators_a(\Upsilon\parallel\Gamma)}{\Operators_a(\Upsilon)}$
-
-\subsection{Substitution monoidal structures}
-
-For an object $\Of{P}{\widehat{\HypCat}^\Sorts}$, we will use the notation
-$P^{[\Gamma]}$ to mean $\prod_{x\in\Dom\Gamma}P_{\Gamma(x)}$; likewise,
-$\SymPsh^{\{\Upsilon\}}$ shall mean $\prod_{u\in\Dom\Upsilon}S_{\Upsilon(u)}$. For a
-presheaf $\Of{A}{\widehat{\HypCat}}$ and a sort-indexed family of presheaves
-$\Of{P}{\widehat{\HypCat}^\Sorts}$, we have an operation $A\bullet P$, defined
-as a coend in the following way:
-\[
-  \Define{(A\bullet P)(\Upsilon\parallel\Gamma)}{
-    \int^{(\Upsilon'\parallel\Delta)\in\HypCat}
-      A(\Upsilon'\parallel\Delta)
-      \times \SymPsh^{\{\Upsilon'\}}(\Upsilon\parallel\Gamma)
-      \times P^{[\Delta]}(\Upsilon\parallel\Gamma)
-  }
-\]
-
-Intuitively, the coend construction $A\bullet P$ provides the data of a
-suspended substitution of an $A$-term's variables by $P$-terms.
-%
-Using this, we can define a tensor $P\odot Q$ for
-$\Of{P,Q}{\widehat{\HypCat}^\Sorts}$ as follows:
-\[
-  \HypJ{
-    \Define{{(P\odot Q)}_{\tau}}{
-      P_\tau\bullet Q
-    }
-\quad
-  }{
-    \Member{\tau}{\Sorts}
-  }
-\]
-
-Then, $V$ is the unit to this tensor. We will say that an object
-$\Of{P}{\widehat{\HypCat}^\Sorts}$ is a $\Sigma$-monoid in case it is equipped
-with the following natural transformations where $\nu$ embeds variables into $P$
-and $\varsigma$ equips $P$ with an operation for simultaneous substitutions of
-variables. Furthermore, $\nu$ and $\varsigma$ induce maps ${\nu}_{\Gamma}$ and
-${\varsigma}^{\tau}_{\Upsilon\parallel\Gamma}$:
-\begin{gather*}
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{V}} \arrow[r, "\IMode\nu"]
-    \&
-    {\IMode{P}}
-    \&
-    {\IMode{P\odot P}}
-      \arrow[l, swap, "\IMode\varsigma"]
-  \end{tikzcd}\\[6pt]
-  %
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{{V}^{[\Gamma]}}}
-      \arrow[r, "\IMode{{\nu}_{\Gamma}}"]
-    \&
-    {\IMode{{P}^{[\Gamma]}}}
-  \end{tikzcd}
-  \qquad
-  \begin{tikzcd}[cramped, sep = large, ampersand replacement = \&]
-    {\IMode{{P_\tau^{\Yoneda{\Upsilon\parallel\Gamma}}\times \SymPsh^{\{\Upsilon\}}\times P^{[\Gamma]}}}}
-      \arrow[r, "\IMode{\varsigma^\tau_{\Upsilon\parallel\Gamma}}"]
-    \&
-    {\IMode{P_\tau}}
-  \end{tikzcd}
-\end{gather*}
-
-\subsection{The signature endofunctor and its algebras}
-
-For each signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we have an
-endofunctor
-$\Of{\SigFun{\Sigma}}{\widehat{\HypCat}^\Sorts\ra\widehat{\HypCat}^\Sorts}$,
-which is defined as follows:
-\begin{align*}
-  \IMode{{\SigFun{\Sigma}(X)}_{\tau}}
-&
-    \triangleq
-    \OMode{
-      \coprod_{
-        \vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}
-      }
-      \prod_{
-        \MkValence
-          {\vec{\sigma}}
-          {\vec{\tau}}
-          {\tau_i}
-        \in\vec{v}
-      }
-      X_{\tau_i}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}
-    }
-\end{align*}
-
-Then, a $\Sigma$-model is a $\Sigma$-monoid $P$ which is equipped with an
-algebra $\Of{\alpha}{\SigFun{\Sigma}(P)\ra P}$, which shall interpret
-applications of each operator.
-
-\subsection{Interpretation of terms}
-
-The metavariable, symbol and variable contexts are interpreted for a model $P$ as
-an environment presheaf in the following way:
-\[
-  \Define{
-    \SemBrackets{
-      \Theta\triangleright\Upsilon\parallel\Gamma
-    }_P
-  }{
-    \left(
-      \prod_{
-        (m:%
-        \MkValence
-          {\vec{\sigma}}
-          {\vec{\tau}}
-          {\tau})
-        \in\Theta
-      }{{P_{\tau}}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}}
-    \right)
-    \times
-    \SymPsh^{\{\Upsilon\}}
-    \times
-    \VarPsh^{[\Gamma]}
-  }
-\]
-
-Then, the interpretation of a term in a model $P$ is a map from its environment
-to $P$:
-\[
-  \Of{
-    \SemBrackets{
-      \Theta\triangleright\Upsilon\parallel\Gamma\vdash M:\tau
-    }_{P}
-  }{
-    \SemBrackets{
-      \Theta\triangleright\Upsilon\parallel\Gamma
-    }_{P}\ra {P}_{\tau}
-  }
-\]
-
-Variables are interpreted by the map
-${\SemBrackets{
-  \Theta\triangleright\Upsilon\parallel\Gamma\vdash x:\tau
-}}_{P}$ which
-projects them from the environment and embeds them into the model. Metavariables
-are resolved by the map
-${\SemBrackets{
-  \Theta\triangleright\Upsilon\parallel\Gamma\vdash\MApp{\MV{m}}{\vec{u}}{\vec{M}}
-  : \tau
-}}_{P}$ (where
-$\Lookup{\Theta}{\MV{m}}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$) which
-projects their interpretation from the environment and instantiates it via
-substitution:
-\[
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{P_\tau}}
-\&[8.0em]
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-    }}
-      \arrow[
-        , rr
-        , dashed
-        , "
-          \OMode{
-            {\SemBrackets{
-              \Theta\triangleright\Upsilon\parallel\Gamma
-              \vdash
-              x:\tau
-            }}_{P}
-          }
-          "
-      ]
-      \ar{drr}[
-        , rotate = -16.5
-        , xshift = +1.5em
-        , yshift = -0.2em
-        , swap
-      ]{
-        {\nu}_{\Gamma} {\pi}_{3}
-      }
-      \ar{dl}[
-        , rotate = +11.5
-        , xshift = -2.75em
-      ]{
-        \IMode{
-          \Tuple{
-            \pi_\MV{m}\,
-            \pi_1,
-            \phi\,
-            \pi_2,
-            \psi
-          }
-        }
-      }
-      \arrow[
-        , l
-        , dashed
-        , swap
-        , "
-          \OMode{
-            {\SemBrackets{
-              \Theta\triangleright\Upsilon\parallel\Gamma
-              \vdash
-              \MApp{\MV{m}}{\vec{u}}{\vec{M}}: \tau
-            }}_{P}
-          }"
-      ]
-\&[-2.0em]
-    \null
-\&[+3.0em]
-    {\IMode{P_\tau}}
-\\
-    {\IMode{
-      {P}_{\tau}^{
-        \Yoneda{\vec{\sigma}\parallel\vec{\tau}}
-      }
-      \times \SymPsh^{\{\vec{\sigma}\}}
-      \times P^{[\vec{\tau}]}
-    }}
-      \arrow[
-        , u
-        , "
-          \IMode{
-            \varsigma^\tau_{\vec{\sigma}\parallel\vec{\tau}}
-          }
-          "
-        ]
-\&
-    \null
-\&
-    \null
-\&
-    {\IMode{P^{[\Gamma]}}}
-      \arrow[
-        , u
-        , swap
-        , "{\pi}_{x}"
-      ]
-\\
-    {\IMode{P^{[\vec{\tau}]}}}
-\&
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-    }}
-      \arrow[
-        , dashed
-        , l
-        , swap
-        , "
-          \Define{
-            \psi\,
-          }{
-            \,\Tuple{
-              {\SemBrackets{
-                \Theta\triangleright\Upsilon\parallel\Gamma\vdash M:\tau
-              }}_{P}
-            }_{(M,\tau)\in (\vec{M},\vec{\tau})}
-          }
-          "]
-\&
-    {\IMode{\SymPsh^{\{\Upsilon\}}}}
-      \arrow[
-        , dashed
-        , r
-        , "\Define{\phi\,}{\,\Tuple{\pi_u}_{u\in\vec{u}}}"
-      ]
-\&
-    {\IMode{\SymPsh^{\{\vec{\sigma}\}}}}
-  \end{tikzcd}
-\]
-
-Interpretation of operator applications is the most complicated. Recall that,
-unlike in standard treatments of universal algebra, our operators are indexed by
-symbol collections; therefore, operators must pass through suitable renamings in
-order to be used in the interpretation. Let us begin by constructing for each
-operator $\IsOperator{\Upsilon}{\vartheta}{a}$ the morphism
-$\SemBrackets{\vartheta}_P$ which shall rename the parameters of the operator
-using the environment:
-\[
-  \begin{tikzcd}[cramped, ampersand replacement = \&, sep = huge]
-    |[alias = Src]|
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-    }}
-      \arrow[
-        , r
-        , swap
-        , "\IMode{\pi_2}"
-      ]
-\&
-    {\IMode{\SymPsh^{\{\Upsilon\}}}}
-      \arrow[
-        , r
-        , swap
-        , "\IMode{\Operators_a(\pi_{(-)})(\vartheta)}"
-      ]
-\&[1.5em]
-    |[alias = Tgt]|
-    {\IMode{\Operators_a}}
-      \arrow[
-        , from = Src
-        , to = Tgt
-        , bend left
-        , dashed
-        , "\OMode{\SemBrackets{\vartheta}_{P}}"
-      ]
-  \end{tikzcd}
-\]
-
-We will proceed using the $\SigFun{\Sigma}$-algebra $\alpha$, as
-follows, by composing it with a morphism $\beta$ from the environment into
-the signature endofunctor, which interprets the syntax of operator applications:
-\[
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{
-      {\SemBrackets{\Theta\triangleright\Upsilon\parallel\Gamma}}_{P}
-    }}
-      \arrow[r, "\IMode{\beta}"]
-\&
-    {\IMode{
-      {\SigFun{\Sigma}(P)}_{\tau}
-    }}
-      \arrow[r, "\IMode{\alpha_\tau}"]
-\&
-    {\IMode{P_\tau}}
-  \end{tikzcd}
-\]
-
-The construction of $\beta$ proceeds by renaming the parameters of the operator
-$\vartheta$ and constructing the (bound) exponentiated arguments $\gamma$ of the
-operator.
-\[
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-    }}
-      \arrow[
-        , r
-        , dashed
-        , "
-          \Define{
-            \beta\,
-          }{
-            \,\Tuple{
-              \SemBrackets{\vartheta}_P, \lambda\gamma
-            }
-          }
-          "
-      ]
-\&[4.0em]
-    {\IMode{
-      \coprod_{
-        \vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}
-      }
-      \prod_{
-        \MkValence{\vec{\sigma}_i}{\vec{\tau}_i}{\tau_i}\in\vec{v}
-      }
-      P_{\tau_i}^{\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}}
-    }}
-  \end{tikzcd}
-\]
-Arguments $\lambda\IMode{{\gamma}_{i}}$ are the exponential transposes (curried
-form) of the composites $\IMode{{\gamma}_{i}}$:
-\[
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-    }}
-      \arrow[
-        , dashed
-        , d
-        , swap
-        , "\OMode{\lambda{\gamma}_{i}}"
-      ]
-\&
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-      \times
-      \Yoneda{
-        \vec{\sigma}_{i}\parallel\vec{\tau}_{i}}
-      }
-    }
-      \arrow[
-        , r
-        , "
-          \IMode{
-            \Tuple{
-              {\pi}_{[1,1]},
-              {\phi}_{i},
-              {\psi}_{i}
-            }
-          }
-          "
-        ]
-      \arrow[
-        , dashed
-        , d
-        , swap
-        , "\OMode{{\gamma}_{i}}"
-      ]
-\&[3.5em]
-    {\IMode{
-      \SemBrackets{
-        \Theta\triangleright\Upsilon,
-        \vec{\sigma}_i\parallel\Gamma,
-        \vec{\tau}_{i}
-      }_{P}
-    }}
-      \ar{dl}[
-        , rotate = +19.25
-        , xshift = -4.5em
-        , yshift = -0.2em
-      ]{\IMode{\SemBrackets{\Theta\triangleright\Upsilon,\vec{\sigma}_{i}\parallel\Gamma,\vec{\tau}_{i}\vdash M_i:\tau_i}_P}}
-\\[2.0em]
-    {\IMode{{P}_{{\tau}_{i}}^{\Yoneda{\vec{\sigma}_{i}\parallel\vec{\tau}_{i}}}}}
-\&
-    {\IMode{P_{{\tau}_{i}}}}
-\&
-  \end{tikzcd}
-\]
-where $\phi_i,\psi_i$ are defined as follows:
-\begin{gather*}
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{
-      {\SemBrackets{\Theta\triangleright\Upsilon\parallel\Gamma}}_{P}
-      \times
-      \Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}
-    }}
-      \arrow[
-        , r
-        , dashed
-        , "
-          \Define
-            {
-              \phi_i\,
-            }
-            {
-              \,
-                \Tuple{\pi_u\circ\pi_1}_{u\in\Dom{\Upsilon}}
-              ,
-                \,\Tuple{\pi_u\circ\pi_{[2,1]}}_{u\in\Dom{\vec{\sigma}_i}}
-            }
-          "
-      ]
-\&[12.0em]
-    {\IMode{\SymPsh^{\{\Upsilon,\vec{\sigma}_i\}}}}
-\\[-1.0em]
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-      \times
-      \Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}
-    }}
-      \arrow[
-        , r
-        , dashed
-        , "
-          \Define{
-            \psi_i\,
-          }{
-            \,
-              \Tuple{\pi_x\circ\pi_1}_{x\in\Dom{\Gamma}}
-            ,
-              \,\Tuple{\pi_x\circ\pi_{[2,1]}}_{x\in\Dom{\vec{\tau}_i}}
-          }
-          "
-      ]
-\&[12.0em]
-    {\IMode{\VarPsh^{[\Gamma,\vec{\tau}_i]}}}
-  \end{tikzcd}
-\end{gather*}
-
-This concludes the interpretation of well-sorted terms into any $\Sigma$-model.
-
-\section{Case Study: Wellformed Sequents}
-
-The representation of telescopes and sequents in a logical framework is
-notoriously difficult; whilst it is possible to use higher-order abstract
-syntax or abts to encode the binding-structure of telescopes and sequents, the
-encoding is sufficiently laborious and obscure that it is not used in practice.
-Crary has demonstrated a first-order encoding of contexts in the Edinburgh
-Logical Framework (ELF) in bijection with actual LF-contexts~\cite{crary:2009},
-which has been successfully used in large-scale mechanization efforts,
-including that of Standard ML~\cite{lee-crary-harper:2007} and the ELF
-itself~\cite{martens-crary:2012}.
-
-We will approach the problem of encoding telescopes and sequents from the
-\emph{refinements} perspective, where a conservative approximation of the
-grammar is first given using the abt logical framework, and then the
-correctness of a code is expressed separately in a judgment that refines the
-existing specification.  Because we have not committed to using the built-in
-binding machinery to express the well-scopedness of telescopes and sequents, we
-are free to use \emph{symbols} in order to model the variables in the context.
-This is in fact quite sensible if we are actually trying to faithfully
-represent the syntax of telescopes and sequents, rather than replace them with
-their counterparts on the meta-level.
-
-The approach outlined above is actually quite similar to the standard practice
-in the Abella proof assistant \cite{gacek:2008}; in Abella, in addition to the
-canonical forms, each type is inhabited by an infinite collection of nominal
-constants or symbols, which can be used in the syntax of contexts. In order to
-specify where such symbols are allowed to occur, a wellformedness predicate (a
-``scheme'') is separately defined. Our setting is similar in that we may also
-use symbols to encode the syntax of contexts, except that here, we have better
-control over their proliferation because, contrary to the state of
-affairs in Abella, symbols are \emph{not} terms.
-
-There are two reasons that one might want to model sequents in a (meta)-logical
-framework. Perhaps the most obvious reason is to prove things about a logic
-with sequent judgments; this is what motivated Crary's encoding of contexts in
-LF, for instance. However, another common use-case is the practical development
-of a proof assistant in which users can define new sequent rules; such
-user-defined rules may only be compiled into well-behaved tactics in the proof
-system in case they obey certain wellformedness properties including (among
-other things) correct scoping of names. MetaPRL is a salient example of such a
-proof assistant, whose implementation might have benefited from the more
-principled treatment of symbols and metavariables which we present here.
-
-This insight leads the way to a simple abt signature for the theory of
-telescopes and sequents.
-%
-\newcommand\SortTele{\mathtt{tele}}
-\newcommand\SortJdg{\mathtt{jdg}}
-\newcommand\SortProp{\mathtt{prop}}
-\newcommand\SortExpr{\mathtt{exp}}
-\newcommand\OpNil{\mathtt{nil}}
-\newcommand\OpSnoc[1]{\mathtt{snoc}[#1]}
-\newcommand\OpHyp[1]{\mathtt{hyp}[#1]}
-\newcommand\OpSequent{\mathtt{sequent}}
-% %
-\[
-  \begin{array}{lr}
-    \IsSort{\SortTele}&\textit{telescopes}\\
-    \IsSort{\SortExpr}&\textit{expressions}\\
-    \IsSort{\SortProp}&\textit{propositions}\\
-    \IsSort{\SortJdg}&\textit{judgments}\\\\
-    \IsOperator{\Upsilon, u:\SortExpr}{\OpHyp{u}}{
-      \MkArity{}{\SortExpr}
-    }\\\\
-    \IsOperator{\Upsilon}{\OpNil}{
-      \MkArity{}{\SortTele}
-    }\\
-    \IsOperator{\Upsilon, u:\SortExpr}{\OpSnoc{u}}{
-      \MkArity{
-        \MkValence{}{}{\SortTele},
-        \MkValence{}{}{\SortProp}
-      }{\SortTele}
-    }\\\\
-    \IsOperator{\Upsilon}{\OpSequent}{
-      \MkArity{
-        \MkValence{}{}{\SortTele},
-        \MkValence{}{}{\SortProp}
-      }{\SortJdg}
-    }
-  \end{array}
-\]
-
-Suppose we have encoded a fragment of type theory as well:
-\[
-  \begin{array}{l}
-    \IsOperator{\Upsilon}{\mathtt{P}}{
-      \MkArity{}{\SortProp}
-    }
-\\
-    \IsOperator{\Upsilon}{\mathtt{pred}}{
-      \MkArity{
-        \MkValence{}{}{\SortExp}
-      }{\SortProp}
-    }
-\\
-  \end{array}
-\]
-
-Terms written using the abstract syntax will be difficult to read, so let us
-define some notation:
-\begin{align*}
-  \IMode{\diamond}
-&
-  \triangleq \OMode{\OpNil}
-\\
-  \IMode{H, u:P}
-&
-  \triangleq \OMode{
-    \App{\OpSnoc{u}}{H, P}
-  }
-\\
-  \IMode{H\gg A}
-&
-  \triangleq \OMode{
-    \App{\OpSequent}{
-      H, A
-    }
-  }
-\\
-  \IMode{`u}
-&
-  \triangleq \OMode{\OpHyp{u}}
-\end{align*}
-
-Now, we can write the following sequent:
-\[
-  \IsAbt{\cdot}{
-    u:\SortExp,
-    v:\SortExp
-  }{\cdot}{
-    \diamond,u:\mathtt{P},v:\App{\mathtt{pred}}{`u}
-    \gg
-    \App{\mathtt{pred}}{`u}
-  }{\SortJdg}
-\]
-
-\newcommand\OpNabla[1]{\nabla[#1]}
-
-The above sequent has free symbols, but we can close over them by adding a form
-of parametric higher-order judgment to our object language, indexed by a
-collection of sorts $\vec{\sigma}$:
-\[
-  \IsOperator{\Upsilon}{\OpNabla{\vec{\sigma}}}{
-    \MkArity{
-      \MkValence{\vec{\sigma}}{}{\SortJdg}
-    }{\SortJdg}
-  }
-\]
-
-Then, we may write a closed sequent judgment as follows:
-\[
-  \IsAbt{\cdot}{\cdot}{\cdot}{
-    \App{\OpNabla{\SortExp,\SortExp}}{
-      \MkAbs{u,v}{}{
-        \diamond,u:\mathtt{P},v:\App{\mathtt{pred}}{`u}
-        \gg
-        \App{\mathtt{pred}}{`u}
-      }
-    }
-  }{\SortJdg}
-\]
-
-\subsection{Refinements for wellformedness}
-
-Having specified an approximation of the grammar of telescopes and sequents in
-the abt logical framework, we can proceed to define proper wellformedness via
-\emph{inductive refinement}~\cite{harper:2016}. The basic idea is to introduce a
-new form of (meta)-judgment $\IsWf{\Upsilon}{\Gamma}{M}{\tau}$ which
-expresses the extrinsic wellformedness properties we wish to verify,
-presupposing $\IsAbt{\cdot}{\Upsilon}{\Gamma}{M}{\tau}$. Additionally, we
-introduce an analogous judgment on abstractions,
-$\IsWf{\Upsilon}{\Gamma}{E}{v}$ presupposing
-$\IsAbt{\cdot}{\Upsilon}{\Gamma}{E}{v}$, defined uniformly as follows:
-\[
-  \infer{
-    \IsWf{\Upsilon}{\Gamma}{
-      \MkAbs{\vec{u}}{\vec{x}}{M}
-    }{
-      \MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}
-    }
-  }{
-    \IsWf{\Upsilon,\vec{u}:\vec{\sigma}}{\Gamma,\vec{x}:\vec{\tau}}{M}{\tau}
-  }
-\]
-
-Likewise, wellformedness for variables is defined uniformly:
-\[
-  \infer{
-    \IsWf{\Upsilon}{\Gamma}{x}{\tau}
-  }{
-  }
-\]
-Note that the refinement for variables $x$ is not trivial, since it is only
-defined in case the presupposition $\IsAbt{\cdot}{\Upsilon}{\Gamma}{x}{\tau}$
-is satisfied.
-
-The remainder of the definition of refinement proceeds by induction on sorts and
-operators. For the sake of this example, we will just stipulate that anything of
-sort $\SortExp$ or $\SortProp$ is grammatical if its subterms are grammatical:
-
-\[
-  \infer[
-    \text{for $\Member{\tau}{\MkSet{\SortExp,\SortProp}}$}
-  ]{
-    \IsWf{\Upsilon}{\Gamma}{
-      \App{\vartheta}{E_0,\dots,E_n}
-    }{\tau}
-  }{
-    \begin{array}{l}
-      \IsOperator{\Upsilon}{\vartheta}{
-        \MkArity{
-          v_0,\dots,v_n
-        }{\tau}
-      }
-\\
-      \HypJ{
-        \IsWf{\Upsilon}{\Gamma}{E_i}{v_i}
-      }{
-        \Leq{i}{n}
-      }
-    \end{array}
-  }
-\]
-
-The refinements for parametric judgment and sequents simply delegate to their
-subterms as well:
-\begin{gather*}
-  \infer{
-    \IsWf{\Upsilon}{\Gamma}{
-      \App{\OpNabla{\vec{\sigma}}}{
-        \MkAbs{\vec{u}}{}{J}
-      }
-    }{\SortJdg}
-  }{
-    \IsWf{\Upsilon,\vec{u}:\vec{\sigma}}{\Gamma}{J}{\SortJdg}
-  }
-\\[6pt]
-  \infer{
-    \IsWf{\Upsilon}{\Gamma}{
-      H\gg A
-    }{\SortJdg}
-  }{
-    \IsWf{\Upsilon}{\Gamma}{H}{\SortTele}
-&
-    \IsWf{\Upsilon}{\Gamma}{A}{\SortProp}
-  }
-\end{gather*}
-
-The refinement for telescopes proceeds by induction:
-\begin{gather*}
-  \infer{
-    \IsWf{\Upsilon}{\Gamma}{\diamond}{\SortTele}
-  }{
-  }
-\qquad
-  \infer{
-    \IsWf{\Upsilon}{\Gamma}{H,u:A}{\SortTele}
-  }{
-    \begin{array}{l}
-      \IsWf{\Upsilon\setminus\MkSet{u}}{\Gamma}{H}{\SortTele}
-\\
-      \IsWf{\Upsilon\setminus\MkSet{u}}{\Gamma}{A}{\SortProp}
-    \end{array}
-  }
-\end{gather*}
-
-Based on these rules, it is easy to show that the example sequent from the
-previous section is wellformed:
-\[
-  \IsWf{\cdot}{\cdot}{
-    \App{\OpNabla{\SortExp,\SortExp}}{
-      \MkAbs{u,v}{}{
-        \diamond,u:\mathtt{P},v:\App{\mathtt{pred}}{`u}
-        \gg
-        \App{\mathtt{pred}}{`u}
-      }
-    }
-  }{\SortJdg}
-\]
-
-\begin{remark}
-  In the full development of a theory of behavioral refinements, the ad-hoc
-  judgment $\IsWf{\Upsilon}{\Gamma}{M}{\tau}$ would be replaced with a general
-  notion of refinement $\IMode{\phi}\sqsubset\tau$ ($\phi$ refines sort
-  $\tau$), and a new judgment $\IMode{\Upsilon}\parallel\IMode{\Phi}\vDash
-  \IMode{M} \in \IMode{\phi}$, such that $\phi$ refines $\tau$ and the
-  refinement context $\Phi$ refines some sort context $\Gamma$. Furthermore, in
-  addition to restricting membership, refinements might also be designed to
-  coarsen equivalence.
-\end{remark}
-
-\section*{Acknowledgements}
-
-The first author wishes to thank Robert Harper for numerous conversations about
-abstract binding trees and symbolic parameters, and Andy Pitts for his help in
-understanding multi-sorted nominal sets and atomic sheaves.
+\input{body}
 
@@
 
-\section{$\SCtxSite$, Nominal Sets, and the Schanuel Topos}
-
-Pitts defines a category $\mathbf{Nom}$ of nominal sets in \cite{pitts:2013}
-which function equivalently to the sheaves that we considered above, in the
-case of $\IMode{\Sorts}\cong\IMode{\mathbb{1}}$. Fix a countably infinite set
-of atoms $\mathbb{A}$; then let $\Perm\mathbb{A}$ be the group of permutations
-(i.e.\ autoequivalences) on $\mathbb{A}$.  Considered as a category,
-$\Perm\mathbb{A}$ has a single object $\bullet$ with morphisms
-$\Of{\pi}{\bullet\ra\bullet}$ for each $\Member{\pi}{\Dom{\Perm\mathbb{A}}}$.
-Then, the category of nominal sets $\mathbf{Nom}$ is the subcategory of
-$\Sets^{\Perm\mathbb{A}}$ containing just the presheaves which satisfy a
-\emph{finite support} condition.
-
-The category $\mathbf{Nom}$ is equivalent to the category of covariant sheaves
-on $\SymSets$ under the atomic coverage, called the Schanuel Topos
-$\Define{\mathbb{S}}{\Sheaves{\OpCat{\SymSets}}}$ \cite{pitts:2013,
-fiore-staton:2006}; equivalently, the Schanuel topos is the subcategory of
-$\Sets^\SymSets$ containing only pullbacks-preserving functors.
-
-In addition to the unisorted nominal sets framework, Pitts also briefly
-discusses a multi-sorted version of the apparatus in which the countably
-infinite set of atoms $\mathbb{A}$ is equipped with a sort assignment
-$\Of{\mathfrak{s}}{\mathbb{A}\to\Sorts}$ such that all the fibers of
-$\mathfrak{s}$ are countably infinite; then, he defines for any such assignment
-the category $\mathbf{Nom}_\mathfrak{s}$ of $\mathfrak{s}$-sorted nominal sets
-as a subcategory of the category of presheaves on the group of
-$\mathfrak{s}$-respecting permutations on $\mathbb{A}$,
-$\Perm_\mathfrak{s}\mathbb{A}$.
-
-In the same way as the Schanuel topos is equivalent to the unisorted nominal
-sets, we expect to find an equivalence between $\Sheaves{\SCtxSite}$ and the
-limit $\int_{\mathfrak{s}}\mathbf{Nom}_\mathfrak{s}$. At the very least, the
-sheaf topos $\Sheaves{\SCtxSite}$ may serve as a multi-sorted generalization of
-the Schanuel topos, for which we may carry over certain useful results,
-including its characterization as the subcategory of $\Sets^\SCtx$ which
-contains just the pullbacks-preserving functors.
-
-\begin{thm}
-  \label{thm:pullback-preservation}
-  A presheaf $\Of{X}{\Sets^{\SCtx}}$ is a sheaf on $\SCtxSite$ just when it
-  preserves pullbacks.
-\end{thm}
-
-\begin{proof}
-
-  The proof is essentially the same as that of the analogous lemma for the
-  Schanuel topos as presented in \cite[A.2.1.11.h]{johnstone:2002}, but we will
-  give a slightly more detailed version here. The presheaf $X$ is a sheaf on the
-  atomic site $\SCtxSite$ just when, for any renaming
-  $\Of{\varrho}{C\inj D}$ and any $\Member{n}{X(D)}$ if
-  $\IsEq{n\cdot \varrho_0}{n\cdot \varrho_1}$ for all diagrams
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{C} \arrow[r, hook, "\IMode{\varrho}"] &
-      \IMode{D}
-        \arrow[r, hook, shift left, "\IMode{\varrho_0}"]
-        \arrow[r, hook, swap, shift right, "\IMode{\varrho_1}"]
-      &
-      \IMode{E}
-    \end{tikzcd}
-  \]
-  such that $\IsEq{\varrho_0\circ\varrho}{\varrho_1\circ\varrho}$, then there
-  exists a unique $\Member{m}{X(C)}$ such that $\IsEq{m\cdot\varrho}{n}$. In other
-  words, the arrow $X(\varrho)$ is an equalizer, as in the following:
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{X(C)} \arrow[r, "\IMode{X(\varrho)}"] &
-      \IMode{X(D)}
-        \arrow[r, shift left, "\IMode{X(\varrho_0)}"]
-        \arrow[r, swap, shift right, "\IMode{X(\varrho_1)}"]&
-      \IMode{X(E)}
-      \\
-      \IMode{\mathbb{1}}
-        \arrow[u, densely dotted, "\OMode{m}"]
-        \arrow[ur, "\IMode{n}"]
-    \end{tikzcd}
-  \]
-
-  Now, because $\Sets$ is a regular category, $X(\varrho)$ is an equalizer
-  precisely when it is a monomorphism. Therefore, if we have assumed that $X$
-  preserves pullbacks, in order to demonstrate that $X$ is a sheaf, it suffices
-  to show that $X(\varrho)$ is monic. If $X$ preserves pullbacks, then it also
-  preserves monomorphisms, because in any category with pullbacks, $\Of{f}{A\ra
-  B}$ is monic just when the square
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{A}
-        \arrow[r, "\IMode{\ArrId{A}}"]
-        \arrow[d, "\IMode{\ArrId{A}}"] &
-      \IMode{A} \arrow[d, "\IMode{f}"]\\
-      \IMode{A} \arrow[r, "\IMode{f}"] &
-      \IMode{B}
-    \end{tikzcd}
-  \]
-  is a pullback \cite[p.\ 16]{mac-lane-moerdijk:1992}. Because $X$ preserves
-  monomorphisms, and all arrows in $\SCtx$ are monic, then $X(\varrho)$ is
-  monic. Hence, $X$ is a sheaf.
-
-  Next, we must show that if $X$ is a sheaf, then it preserves pullbacks; we
-  will loosely track the proof of Proposition~2.6.15 in \cite{biering:2004}.
-  Fix the following diagram such that it is a pullback square in $\SCtx$:
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{C}
-        \arrow[r, "\IMode{f_j}"]
-        \arrow[d, "\IMode{f_i}"] &
-      \IMode{C_j} \arrow[d, "\IMode{s_j}"]\\
-      \IMode{C_i} \arrow[r, "\IMode{s_i}"] &
-      \IMode{D}
-    \end{tikzcd}
-  \]
-
-  Then it suffices to show that the following diagram is also pullback square
-  in $\Sets$:
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{X(C)}
-        \arrow[r, "\IMode{X(f_j)}"]
-        \arrow[d, "\IMode{X(f_i)}"] &
-      \IMode{X(C_j)} \arrow[d, "\IMode{X(s_j)}"]\\
-      \IMode{X(C_i)} \arrow[r, "\IMode{X(s_i)}"] &
-      \IMode{X(D)}
-    \end{tikzcd}
-  \]
-
-  We have to show that $X(C)$ is, up to equivalence, the pullback of $X(s_i)$
-  along $X(s_j)$ in $\Sets$, namely:
-  \[
-    \IMode{X(C_i)\times_{X(D)}X(C_j)}
-      \cong
-    \OMode{
-      \MkSet{
-        (m_i, m_j)\in X(C_i)\times X(C_j)
-          \mid
-        m_i \cdot s_i = m_j \cdot s_j
-      }
-    }
-  \]
-  The first direction of the equivalence is trivial: for each
-  $\Member{m}{X(C)}$, we can easily exhibit $\Member{(m\cdot f_i,m\cdot
-  f_j)}{X(C_i)\times_{X(D)}X(C_j)}$.
-
-  For the other direction, fix $\Member{(m_i,
-  m_j)}{X(C_i)\times_{X(D)}X(C_j)}$.  Consider the singleton cover on $C$ that
-  contains $\Define{h}{s_i\circ f_i}=\OMode{s_j\circ f_j}$; a matching family
-  $\phi$ for the cover $\MkSet{h}$ is a map that takes $h$ to an object in
-  $X(D)$, and an amalgamation for $\phi$ consists in an object
-  $\Member{m}{X(C)}$ such that $\IsEq{\phi(h)}{m\cdot h}$. Now recall that $X$
-  is a sheaf on $\SCtxSite$ if and only if every matching family has a unique
-  amalgamation; for the matching family $\Define{\phi(h)}{m_i\cdot
-  s_i}=\OMode{m_j\cdot s_j}$, then, we must have a unique amalgamation
-  $\Member{m}{X(C)}$ since $X$ is a sheaf. Therefore, $X$ preserves pullbacks.
-
-\end{proof}
+\input{appendices}
 

### 2016-01-23 ijcar-abts.tex
[ijcar] lost double struck lambda, so remove
The double struck lambda is fine, but I can't figure out how to get it
to show up here; no lambda's less confusing than a plain one.
[ijcar] update some notation
IJCAR version: import body & appendices

@@
 \usepackage[h]{esvect}
-
-\definecolor{dark-gray}{gray}{0.20}
+\usepackage{ssabt}
 
@@
 \numberwithin{equation}{section}
+
 \renewcommand\vec[1]{\protect\vv{#1}}
-
-\newcommand\HypCat{\mathbf{H}}
-\newcommand\Supp[1]{\mathbf{supp}(#1)}
-\newcommand\SupportsUnmoded[3]{#1 \blacktriangleright_{#2} #3}
-\newcommand\Supports[3]{\SupportsUnmoded{\IMode{#1}}{\IMode{#2}}{\IMode{#3}}}
-
-\newcommand\CalcFV[2]{
-  \FV{\IMode{#1}}\leadsto \OMode{#2}
-}
-
-\newcommand\CalcFS[2]{
-  \FS{\IMode{#1}}\leadsto \OMode{#2}
-}
-
-\newcommand\SortExp{\mathtt{exp}}
-
-\newcommand\IsWf[4]{%
-  \IMode{#1}\parallel%
-  \IMode{#2}\vdash%
-  \IMode{#3}\in_{\mathbf{wf}}\IMode{#4}%
-}
-
-\newcommand\Sheaves[1]{\mathbf{Sh}\left(#1\right)}
-\newcommand\Presheaves[1]{\mathbf{Psh}\left(#1\right)}
-\newcommand\Sheafify[1]{#1^{\#}}
-\newcommand\ConstPsh[1]{\Delta[#1]}
-\newcommand\ConstSh[1]{\Delta^\#[#1]}
-\newcommand\SCtxSite{\SymSets[\Sorts]}
-\newcommand\AtomicCoverage{J_{\mathtt{atm}}}
-
-\DeclareMathOperator{\Perm}{\mathbf{Perm}}
+\renewcommand\MkAbs[3]{\{#1\}[#2].\, #3}
 
@@
 
-Abstract binding trees (abts) are a generalization of abstract syntax trees
-where operators may express variable binding structure as part of their arities.
-Originally formulated by Peter Aczel~\cite{aczel:1978}, unisorted abts have been
-deployed successfully as the uniform syntactic framework for several
-implementations of Constructive Type Theory, including
-Nuprl~\cite{constable:1986}, MetaPRL~\cite{hickey:2003} and
-JonPRL~\cite{jonprl:2015}.
-
-In \emph{Practical Foundations for Programming Languages}~\cite{harper:2016},
-Robert Harper develops the multi-sorted version of abstract binding trees and
-proposes an extension to include families of operators indexed by
-\emph{symbols}, which are, unlike variables, subject to only
-distinctness-preserving renaming and not substitution; furthermore, symbols do
-not appear in the syntax of abts, and are only introduced as parameters to
-operators. This extension to support symbols is essential for a correct
-treatment of programming languages with open sums (e.g. ML's \texttt{exn} type),
-as well as assignable references.
-
-In parallel, M.\ Fiore and his collaborators have developed the categorical
-semantics for several variations of second-order algebraic
-theories~\cite{fiore:1999,fiore:2005,fiore-hur:2010,fiore-mamoud:2010}; of
-these, the simply sorted variants are equivalent to Harper's abstract binding
-trees augmented with a notion of second-order variable (metavariable).
-
-The contribution of this paper is the development of the syntax and semantics
-of multi-sorted nominal abts, an extension of second order universal algebra to
-support symbol-indexed families of operators. Additionally, we have developed
-the categorical semantics for abts formally in Constructive Type Theory using
-the Agda proof assistant~\cite{norell:2007}; we have also developed an
-implementation for abstract binding trees in Standard ML, which we intend to
-integrate into the JonPRL proof assistant \cite{jonprl:2015}.
-
-\section{Categorical Preliminaries}
-
-Fix a set $\Sorts$ of \emph{sorts}; we will say $\IsSort{\tau}$ when
-$\Member{\tau}{\Sorts}$. In this section, we will develop a theory of sets
-varying over collections of $\Sorts$-sorted symbols.
-
-Let $\SymSets$ be the category of finite cardinals and their injective maps;
-then the comma construction $\SCtx$, is the category of contexts of symbols
-whose objects are finite sets of symbols $U$ and sort-assignments
-\begin{tikzcd}[cramped]
-  \IMode{U}\arrow[r, "\IMode{\mathfrak{s}}"] &\IMode{\Sorts}
-\end{tikzcd}, and whose morphisms are
-sort-preserving renamings; we will write $\Upsilon$ for a symbol context
-$(U,\mathfrak{s})$.
-
-\begin{remark}
-
-  To be precise, $\SCtx$ is an abuse of notation for the comma category
-  $\Dom{-}_\SymSets\downarrow \Delta[\Sorts]$, with $\Dom{-}_\SymSets$ the
-  forgetful functor from $\SymSets$ to $\Sets$, and
-  $\Define{\Delta[\Sorts](*)}{\Sorts}$ a constant functor from $\mathbb{1}$ to $\Sets$.
-  Formally, an object of $\SCtx$ is, then, a triple $\Tuple{U, *,
-  \mathfrak{s}}$ with $U$ an object in $\SymSets$, $*$ the unique object in
-  $\mathbb{1}$, and $\mathfrak{s}$ a function $\Dom{U}_\SymSets\ra
-  \Sorts$; an arrow $\Of{\varrho}{\Tuple{U,*,\mathfrak{s}}\ra
-  \Tuple{V,*,\mathfrak{t}}}$ is a commuting triangle like the following:
-  \[
-    \begin{tikzcd}[cramped]
-      \IMode{\Dom{U}_\SymSets}
-        \arrow[rr, "\IMode{\Dom{\rho}_\SymSets}"]
-        \arrow[dr, swap, "\IMode{\mathfrak{s}}"] &&
-      \IMode{\Dom{V}_\SymSets}
-      \arrow[dl, "\IMode{\mathfrak{t}}"] \\
-      &\IMode{\Sorts}
-    \end{tikzcd}
-  \]
-
-  Because the object $\Of{*}{\mathbb{1}}$ is unique, we will always omit it
-  from the data of an object in $\SCtx$.
-\end{remark}
-
-\paragraph{Copresheaves on $\SCtx$}
-
-A presheaf on a category $\mathcal{C}$ is a contravariant functor
-$\OpCat{\mathcal{C}}\ra\Sets$; when $\Match{\mathcal{C}}{\OpCat{\mathcal{D}}}$,
-we call this a \emph{copresheaf} (covariant presheaf) on $\mathcal{D}$.
-Therefore, a copresheaf on $\SCtx$ is a functor $\SCtx\ra\Sets$; colloquially,
-such a copresheaf can be thought of as a set that \emph{varies over} symbol
-contexts.
-
-This leads to a very intuitive setting in which to consider sets
-that vary over collections of $\Sorts$-sorted symbols. In particular, for any
-morphism
-%
-\begin{tikzcd}[cramped]
-  \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
-\end{tikzcd}
-%
-in $\SCtx$, any copresheaf $\Of{X}{\Sets^\SCtx}$ has a symbol-renaming action
-$\Of{X(\varrho)}{X(\Upsilon)\ra X(\Upsilon')}$. When the copresheaf $X$ is
-clear from context, we will write $m\cdot\varrho$ for $X(\varrho)(m)$.
-
-Now, when $\Member{m}{X(\Upsilon)}$ for a copresheaf $\Of{X}{\Sets^\SCtx}$, we
-can define the notion that $m$ depends on \emph{at most} a subcontext
-$\Upsilon'$ of $\Upsilon$; this is called \emph{support}.
-
-\begin{definition}[Support]
-  \label{def:support}
-  When
-  \begin{tikzcd}[cramped]
-    \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
-  \end{tikzcd},
-  we say that $\Upsilon$ supports $\Member{m}{\Upsilon'}$ (written
-  $\Supports{\Upsilon}{\varrho}{m}$) when, for all
-  \begin{tikzcd}[cramped]
-    \IMode{\Upsilon'}\arrow[r, hook, "\IMode{\varrho_1}{,} \IMode{\varrho_2}"] & \IMode{\Upsilon''}
-  \end{tikzcd},
-  if $\IsEq{\varrho_1\circ\varrho}{\varrho_2\circ\varrho}$ then $\IsEq{m\cdot\varrho_1}{m\cdot\varrho_2}$.
-\end{definition}
-
-We can also always calculate the \emph{least support} of a term
-$\Member{m}{X(\Upsilon')}$, written $\Supp{m}$:
-\[
-  \Define{\Supp{m}}{
-    \bigcap_{
-      \begin{tikzcd}[cramped, ampersand replacement = \&]
-        \Upsilon\arrow[r, hook, "\varrho"] \& \Upsilon'
-      \end{tikzcd}
-    }
-    \MkSet{
-      \Upsilon
-      \mid
-      \SupportsUnmoded{\Upsilon}{\varrho}{m}
-    }
-  }
-\]
-Intuitively, $\Supp{m}$ is the exact symbol context that $m$ depends on.
-
-\subsection{Sheaves on the atomic site}
-
-When
-  \begin{tikzcd}[cramped]
-    \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
-  \end{tikzcd}
-such that $\Supports{\Upsilon}{\varrho}{m}$ for some $\Member{m}{X(\Upsilon')}$,
-we would expect that we can work backward to a unique $\Of{m'}{\Upsilon}$
-such that $\IsEq{m'\cdot\varrho}{m}$; whilst it is not in general the case for
-presheaves $X$, this is precisely the sheaf condition for copresheaves on
-$\SCtx$ under the atomic topology \cite[p.~126]{mac-lane-moerdijk:1992}, which
-we will define in Definition~$\ref{def:atomic-topology}$.
-
-The sheaf condition ensures that the notion of support is well-behaved in the
-following sense: if some term lies in two different fibers, then it may be
-\emph{strengthened} uniquely into the fiber of its support.
-
-\begin{definition}[Sieves and matching families]
-  A \emph{sieve} $S$ on an object $C$ is a subfunctor of the Yoneda embedding
-  $\Define{\Yoneda{C}}{\Hom{\mathcal{C}}{-}{C}}$, i.e.\
-  $\IMode{S}\sqsubseteq\IMode{\Yoneda{C}}$. For a presheaf $X$ on the category
-  $\mathcal{C}$, a \emph{matching family} for a sieve $S$ is a natural
-  transformation $\Of{\phi}{S\ra X}$. An \emph{amalgamation} for $\phi$ is an
-  object $\Member{m}{X(C)}$ such that $\IsEq{m\cdot f}{\phi_D(f)}$ for each
-  $\Member{f}{S(D)}$, for all objects $D$.
-\end{definition}
-
-\begin{definition}[Sites and coverages]
-  A \emph{site} is a pair $\Pair{\mathcal{C}}{J}$, with $\mathcal{C}$ a category
-  of ``worlds'' and $J$ a Grothendieck topology (or ``coverage''). A Grothendieck
-  topology $J$ (or \emph{coverage}) is a family of sets of sieves, indexed by
-  $\mathcal{C}$-objects; that is, for each object $\Of{C}{\mathcal{C}}$, $J(C)$
-  is the set of $C$-sieves which will be said to $\emph{cover}$ $C$. A topology
-  $J$ must satisfy several axioms (see \cite{mac-lane-moerdijk:1992} for
-  details).
-\end{definition}
-
-\begin{definition}[Separation and sheafhood]
-  A presheaf on a site $\Pair{\mathcal{C}}{J}$ is a presheaf on $\mathcal{C}$; we
-  say that such a presheaf $X$ is \emph{separated} in case for every sieve
-  $\Member{S}{J(C)}$ and matching family $\Of{\phi}{S\ra X}$, there is at most
-  one amalgamation $\Member{m}{X(C)}$ for $\phi$. Additionally, $X$ is a
-  \emph{sheaf} when there is \emph{exactly one} such amalgamation.
-\end{definition}
-
-\begin{definition}[The atomic topology]
-  \label{def:atomic-topology}
-  The atomic coverage $\AtomicCoverage$ is the one in which all non-empty sieves
-  cover; in other words:
-  \[
-    \Define{
-      \AtomicCoverage(C)
-    }{
-      \MkSet{
-        S\sqsubseteq\Yoneda{C}
-        \mid
-        \textstyle
-        \bigcup_D S(D)\ \textit{non-empty}
-      }
-    }
-  \]
-\end{definition}
-
-In the future, we will write $\SCtxSite$ for the Grothendieck site
-$\Pair{\OpCat{\left(\SCtx\right)}}{\AtomicCoverage}$.
-Sheafhood on the atomic site $\SCtxSite$ amounts to the
-following condition, using Definition~\ref{def:support}:
-\begin{quote}
-  A copresheaf $\Of{X}{\Sets^\SCtx}$ is a sheaf on $\SCtxSite$ just when, for all $\Member{n}{X(\Upsilon')}$, if we have
-  \begin{tikzcd}[cramped]
-    \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
-  \end{tikzcd}
-  and $\Supports{\Upsilon}{\varrho}{n}$, then there is a unique
-  $\Member{m}{X(\Upsilon)}$ such that $\IsEq{m\cdot\varrho}{n}$.
-\end{quote}
-
-Therefore, a sheaf on $\SCtxSite$ is a set that varies over $\Sorts$-sorted
-symbol contexts, which has a well-behaved notion of support. Another way to put
-it is, if $\Member{m}{X(\Upsilon)}$ only ``uses'' some of the symbols available
-at world $\Upsilon$, there's a canonical way to restrict $m$ to a term in the
-world which contains \emph{only} the symbols that it uses.
-
-\paragraph{Constructions on sheaves}
-
-%% For any Grothendieck site $\Match{\mathfrak{S}}{\Pair{\mathcal{C}}{J}}$, the
-%% inclusion $\Of{\iota}{\Sheaves{\mathfrak{S}}\ra\Presheaves{\mathcal{C}}}$ of
-%% sheaves into presheaves has a left adjoint
-%% $\IMode{\Sheafify{(-)}}\dashv\IMode{\iota}$, which takes any presheaf
-%% $\Of{X}{\Presheaves{\mathcal{C}}}$ to a unique sheaf
-%% $\Of{\Sheafify{X}}{\Sheaves{\mathfrak{S}}}$, called the \emph{sheafification}
-%% of $X$. Sheafification is the twice-iterated application of the
-%% \emph{+-construction} $X^+$, which turns a presheaf into a separated presheaf,
-%% and a separated presheaf into a sheaf \cite{mac-lane-moerdijk:1992}:
-%% \begin{align*}
-%%   \ADefine{
-%%     X^+(C)
-%%   }{
-%%     \int^{S\in J(C)}
-%%     \Hom{\Presheaves{\mathcal{C}}}{S}{X}
-%%   }\\
-%%   \ADefine{
-%%     \Sheafify{X}
-%%   }{
-%%     X^{++}
-%%   }
-%% \end{align*}
-%%
-%% We also have a canonical natural transformation $\Of{\eta}{X\ra X^+}$, defined
-%% using the maximal sieve $\Define{t_C}{\Yoneda{C}}$:
-%% \[
-%%   \HypJ{
-%%     \Define{\eta_C(m)}{
-%%       \Pair{t_C}{
-%%         X(-)(m)
-%%       }
-%%     }
-%%   }{
-%%     \Member{m}{X(C)}
-%%   }
-%% \]
-
-Every category of sheaves on a site gives rise to a topos, which equips us with
-a number of standard constructions, including (among other things) the disjoint
-union of sheaves $X\oplus Y$, the product of sheaves $X\otimes Y$, and the terminal sheaf $\mathbb{1}$.
-
-\section{Operators and signatures}
-
-A valence $\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}$ specifies an expression
-of sort $\tau$ which binds symbols in $\vec{\sigma}$ and variables in
-$\vec{\tau}$.
-
-\[
-  \infer{
-    \IsValence{
-      \MkValence
-        {\sigma_0,\dots,\sigma_m}
-        {\tau_0,\dots,\tau_n}
-        {\tau}
-    }
-  }{
-    \IsSort{\tau}
-&
-    \HypJ{\IsSort{\sigma_i}}{\Leq{i}{m}}
-&
-    \HypJ{\IsSort{\tau_i}}{\Leq{i}{n}}
-  }
-\]
-
-An arity $\MkArity{\vec{v}}{\tau}$ specifies an operator of sort $\tau$ with
-arguments of valences $\vec{v}$. We will call the set of valences $\Valences$,
-and the set of arities $\Arities$.
-
-\[
-  \infer{
-    \IsArity{
-      \MkArity
-        {v_0,\ldots,v_n}
-        {\tau}
-    }
-  }{
-    \IsSort{\tau}
-&
-    \HypJ
-      {\IsValence{v_i}}
-      {\Leq{i}{n}}
-  }
-\]
-
-Fix a family of sheaves of operators
-$\Member{\Operators_a}{\Sheaves{\SCtxSite}}$, indexed by arities
-$\Member{a}{\Arities}$; for each $\Operators_a$, arrows in $\SCtx$ will lift to
-renamings in operators' symbolic parameters. Together, $\Sorts$ and $\Operators$ are said to form a \emph{signature}
-$\Define{\Sigma}{\Pair{\Sorts}{\Operators}}$.
-
-We will write $\IsOperator{\Upsilon}{\vartheta}{a}$ in case
-$\Member{\vartheta}{\Operators_a(\Upsilon)}$; note that this judgment enjoys
-the structural properties of weakening and exchange via functoriality of
-$\Operators$, and strengthening via the sheaf condition. An operator signature
-is defined by specifying, for each arity $a$, the sheaf $\Operators_a$, whose
-``elements'' are the operators of arity $a$.
-
-\paragraph{Examples}
-
-For instance, consider a $\lambda$-calculus with a single sort, $\SortExp$; we
-give its signature $\Sigma_{\lambda}$ by asserting the following about its
-operators:
-\begin{gather*}
-  \IsOperator{\Upsilon}{\mathtt{lam}}{
-    \MkArity{
-      \MkValence{}{\SortExp}{\SortExp}
-    }{\SortExp}
-  }\\
-  \IsOperator{\Upsilon}{\mathtt{fix}}{
-    \MkArity{
-      \MkValence{}{\SortExp}{\SortExp}
-    }{\SortExp}
-  }\\
-  \IsOperator{\Upsilon}{\mathtt{ap}}{
-    \MkArity{
-      \MkValence{}{}{\SortExp},
-      \MkValence{}{}{\SortExp}
-    }{\SortExp}
-  }
-\end{gather*}
-
-These rules correspond to the following definition of $\Operators$:
-\begin{align*}
-  \ADefine{
-    \Operators_{
-       \MkArity{
-        \MkValence{}{\SortExp}{\SortExp}
-      }{\SortExp}
-    }
-  }{
-    \mathbb{1}\oplus\mathbb{1}
-  }
-  \\
-  %
-  \ADefine{
-    \Operators_{
-      \MkArity{
-        \MkValence{}{}{\SortExp},
-        \MkValence{}{}{\SortExp}
-      }{\SortExp}
-    }
-  }{
-    \mathbb{1}
-  }
-\end{align*}
-
-So far, we have made no use of symbols and parameters; however, consider the
-extension of the calculus with assignables (references):
-%
-\begin{gather*}
-  \IsOperator{\Upsilon}{\mathtt{decl}}{
-    \MkArity{
-      \MkValence{}{}{\SortExp},
-      \MkValence{\SortExp}{}{\SortExp}
-    }{
-      \SortExp
-    }
-  }
-\\
-  \IsOperator{\Upsilon,u:\SortExp}{\mathtt{get}[u]}{
-    \MkArity{}{\SortExp}
-  }
-\\
-  \IsOperator{\Upsilon,u:\SortExp}{\mathtt{set}[u]}{
-    \MkArity{
-      \MkValence{}{}{\SortExp}
-    }{
-      \SortExp
-    }
-  }
-\end{gather*}
-
-Let $\mathbf{S}_\tau$ be the sheaf of symbols of sort $\tau$, a subobject of
-the Yoneda embedding of the empty symbol context, $\Yoneda{\cdot}$:
-\[
-  \Define{\mathbf{S}_\tau(\Upsilon)}{
-    \MkSet{
-      u\in\Dom{\Upsilon}
-      \mid
-      \Upsilon(u) \equiv \tau
-    }
-  }
-\]
-%
-Then the above rules correspond to the following family of sheaves:
-\begin{align*}
-  \ADefine{
-    \Operators_{
-      \MkArity{
-        \MkValence{}{}{\SortExp},
-        \MkValence{\SortExp}{}{\SortExp}
-      }{
-        \SortExp
-      }
-    }
-  }{
-    \mathbb{1}
-  }\\
-  \ADefine{
-    \Operators_{
-      \MkArity{}{\SortExp}
-    }
-  }{
-    \mathbf{S}_\SortExp
-  }\\
-  \ADefine{
-    \Operators_{
-      \MkArity{
-        \MkValence{}{}{\SortExp}
-      }{
-        \SortExp
-      }
-    }
-  }{
-    \mathbf{S}_\SortExp
-  }
-\end{align*}
-
-Declaring a new assignable consists in providing an initial value, and an
-expression binding a symbol (which shall represent the assignable in scope).
-Note that the functoriality of $\Operators$ guarantees for any renaming
-$\Of{\varrho}{\Upsilon\inj\Upsilon'}$, a family of lifted maps
-$\Of{\Operators_a(\varrho)}{\Operators_a(\Upsilon)\inj\Operators_a(\Upsilon')}$
-natural in $a$, such that
-$\IsOperator{\Upsilon'}{\Operators(\varrho)(\vartheta)}{a}$ when
-$\IsOperator{\Upsilon}{\vartheta}{a}$. In particular, the renaming
-$\Upsilon,u\mapsto\Upsilon,v$ shall take $\mathtt{get}[u]$ to
-$\mathtt{get}[v]$. In the future, we will write $\Rename{\varrho}{\vartheta}$
-for $\Operators_a(\varrho)(\vartheta)$.
-
-\section{Contexts}
-
-In general, we will have three kinds of context: symbolic (parameter) contexts,
-variable contexts, and metavariable contexts. The symbol contexts have already
-been defined via the comma construction $\SCtx$, but they also admit a
-syntactic characterization,
-\begin{gather*}
-  \infer{
-    \IsSymCtx{\cdot}
-  }{
-  }
-\qquad
-  \infer{
-    \IsSymCtx{\Upsilon,u:\tau}
-  }{
-    \IsSymCtx{\Upsilon}
-&
-    \IsSort{\tau}
-&
-    \NotIn{u}{\Dom\Upsilon}
-  }
-\end{gather*}
-%
-Because, modulo notation, we have $\Member{\Upsilon}{\SCtx}$ just when
-$\IsSymCtx{\Upsilon}$, we will use the syntactic view when it is convenient.
-
-Contexts of variables are similar to contexts of symbols, except that they
-admit \emph{any} renamings, not just the injective ones. As such, when
-$\FinSets$ is the category of finite cardinals and all functions between them, the
-comma construction $\Ctx$ is the category of variable contexts. As above, we
-can give them an equivalent syntactic treatment:
-\begin{gather*}
-  \infer{
-    \IsVarCtx{\cdot}
-  }{
-  }
-\qquad
-  \infer{
-    \IsVarCtx{\Gamma,x:\tau}
-  }{
-    \IsVarCtx{\Gamma}
-&
-    \IsSort{\tau}
-&
-    \NotIn{x}{\Dom\Gamma}
-  }
-\end{gather*}
-
-A metavariable context consists of bindings of $\Sorts$-\emph{valences} to
-metavariables; let $\Define{\Valences}{\MkSet{v\mid \IsValenceUnmoded{v}}}$ be the set
-of valences. Then, the category of metavariable contexts is the comma
-construction $\MCtx$, which likewise admits an equivalent inductive definition:
-\begin{gather*}
-  \infer{
-    \IsMetaCtx{\cdot}
-  }{
-  }
-\qquad
-  \infer{
-    \IsMetaCtx{\Theta,\MV{m}:v}
-  }{
-    \IsMetaCtx{\Theta}
-&
-    \IsValence{v}
-&
-    \NotIn{\MV{m}}{\Dom\Theta}
-  }
-\end{gather*}
-
-\section{Nominal Abstract Binding Trees}
-
-Let the judgment $\IsAbt{\Theta}{\Upsilon}{\Gamma}{M}{\tau}$ presuppose
-\footnote{
-  In the \emph{judgmental method}, as pioneered by Per Martin-L\"of, a form of
-  judgment is propounded first by specifying its range of significance (i.e.\
-  the circumstances under which it shall have a meaning), and then, a meaning
-  explanation (definition) for the judgment is given. The range of significance
-  of a judgment is called its \emph{presupposition}, and the meaning of the
-  judgment may proceed by induction on the evidence that the presupposition
-  obtains. See \cite{schroeder-heister:1987} for a more detailed explanation of
-  presuppositions and the judgmental method.
-}
-$\IsMetaCtx{\Theta}$, $\IsSymCtx\Upsilon$, $\IsVarCtx\Gamma$ and
-$\IsSort{\tau}$, meaning that $M$ is an abstract binding tree of sort $s$, with
-metavariables in $\Theta$, parameters in $\Upsilon$, and variables in $\Gamma$.
-Let the judgment $\IsAbs{\Theta}{\Upsilon}{\Gamma}{E}{v}$ presuppose
-$\IsValence{v}$. Then, the syntax of abstract binding trees is inductively
-defined in four rules:
-
-\[
-  \infer[\RuleIsAbtVar]{
-    \IsAbt{\Theta}{\Upsilon}{\Gamma}{x}{\tau}
-  }{
-    \Lookup{\Gamma}{x}{\tau}
-  }
-\]
-\[
-  \infer[\RuleIsAbtMVar]{
-    \IsAbt{\Theta}{\Upsilon}{\Gamma}{
-      \MApp{\MV{m}}{u_0,\dots,u_m}{M_0,\dots,M_n}
-    }{\tau}
-  }{
-    \begin{array}{l}
-      \Lookup{\Theta}{\MV{m}}{
-        \MkValence{
-          \sigma_0,\dots,\sigma_m
-        }{
-          \tau_0,\dots,\tau_n
-        }{\tau}
-      }
-\\
-      \HypJ{
-        \Lookup{\Upsilon}{u_i}{\sigma_i}
-      }{\Leq{i}{m}}
-\\
-      \HypJ{
-        \IsAbt{\Theta}{\Upsilon}{\Gamma}{M_i}{\tau_i}
-      }{\Leq{i}{n}}
-    \end{array}
-  }
-\]
-\[
-  \infer[\RuleIsAbtApp]{
-    \IsAbt{\Theta}{\Upsilon}{\Gamma}{
-      \App{\vartheta}{E_0,\dots,E_n}
-    }{\tau}
-  }{
-    \begin{array}{l}
-      \IsOperator{\Upsilon}{\vartheta}{
-        \MkArity{v_1,\dots,v_n}{\tau}
-      }
-\\
-      \HypJ{
-        \IsAbs{\Theta}{\Upsilon}{\Gamma}{E_i}{v_i}
-      }{\Leq{i}{n}}
-    \end{array}
-  }
-\]
-\[
-  \infer[\RuleIsAbtAbs]{
-    \IsAbs{\Theta}{\Upsilon}{\Gamma}{
-      \MkAbs{\vec{u}}{\vec{x}}{M}
-    }{
-      \MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}
-    }
-  }{
-    \IsAbt{
-      \Theta
-    }{
-      \Upsilon, \vec{u}:\vec{\sigma}
-    }{
-      \Gamma,\vec{x}:{\vec{\tau}}
-    }{
-      M
-    }{\tau}
-  }
-\]
-
-Abts are identified up to $\alpha$-equivalence.
-
-\subsection{Calculating free variables}
-
-\newcommand\RuleFVVar{\RuleNameVar{\mathbf{FV}}{}}
-\newcommand\RuleFVMVar{\RuleNameMVar{\mathbf{FV}}{}}
-\newcommand\RuleFVApp{\RuleNameApp{\mathbf{FV}}}
-\newcommand\RuleFVAbs{\RuleNameAbs{\mathbf{FV}}{}}
-
-We can easily calculate the variables free in an expression by recursion on its
-structure:
-
-\begin{gather*}
-  \infer[\RuleFVVar]{
-    \CalcFV{x}{\MkSet{x}}
-  }{
-  }\\[6pt]
-  %
-  \infer[\RuleFVMVar]{
-    \CalcFV{
-      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
-    }{
-      \bigcup_{i\leq n} \vec{x}_i
-    }
-  }{
-    \HypJ{
-      \CalcFV{M_i}{\vec{x}_i}
-    }{
-      \Leq{i}{n}
-    }
-  }\\[6pt]
-  %
-  \infer[\RuleFVApp]{
-    \CalcFV{
-      \App{\vartheta}{E_0,\dots,E_n}
-    }{
-      \bigcup_{i\leq n} \vec{x}_i
-    }
-  }{
-    \HypJ{
-      \CalcFV{E_i}{\vec{x}_i}
-    }{
-      \Leq{i}{n}
-    }
-  }\\[6pt]
-  %
-  \infer[\RuleFVAbs]{
-    \CalcFV{
-      \MkAbs{\vec{u}}{\vec{y}}{M}
-    }{
-      \vec{x}\setminus\vec{y}
-    }
-  }{
-    \CalcFV{M}{\vec{x}}
-  }
-\end{gather*}
-
-Because this is a total relation, henceforth we will write $\FV{M}$ for
-$\vec{x}$ when $\CalcFV{M}{\vec{x}}$.
-
-\subsection{Calculating free symbols}
-
-\newcommand\RuleFSVar{\RuleNameVar{\mathbf{FS}}{}}
-\newcommand\RuleFSMVar{\RuleNameMVar{\mathbf{FS}}{}}
-\newcommand\RuleFSApp{\RuleNameApp{\mathbf{FS}}}
-\newcommand\RuleFSAbs{\RuleNameAbs{\mathbf{FS}}{}}
-
-Whereas the calculation of free variables pivoted on the $\RuleFVVar$ rule, the
-calculation of free symbols will pivot on the $\RuleFSApp$ and $\RuleFSMVar$ rules,
-because the only way a symbol can be introduced is as a parameter to an
-operator or as a parameter to a metavariable.
-
-\begin{gather*}
-  \infer[\RuleFSVar]{
-    \CalcFS{x}{\MkSet{}}
-  }{
-  }\\[6pt]
-  %
-  \infer[\RuleFSMVar]{
-    \CalcFS{
-      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
-    }{
-      \vec{u} \cup \bigcup_{i\leq n} \vec{u}_i
-    }
-  }{
-    \HypJ{
-      \CalcFS{M_i}{\vec{u}_i}
-    }{
-      \Leq{i}{n}
-    }
-  }\\[6pt]
-  %
-  \infer[\RuleFSApp]{
-    \CalcFS{
-      \App{\vartheta}{E_0,\dots,E_n}
-    }{
-      \Dom{\Supp{\vartheta}} \cup \bigcup_{i\leq n} \vec{u}_i
-    }
-  }{
-    \HypJ{
-      \CalcFS{E_i}{\vec{u}_i}
-    }{
-      \Leq{i}{n}
-    }
-  }\\[6pt]
-  %
-  \infer[\RuleFSAbs]{
-    \CalcFS{
-      \MkAbs{\vec{v}}{\vec{x}}{M}
-    }{
-      \vec{u}\setminus\vec{v}
-    }
-  }{
-    \CalcFS{M}{\vec{u}}
-  }
-\end{gather*}
-
-Because this is a total relation, henceforth we will write $\FS{M}$ for
-$\vec{u}$ when $\CalcFS{M}{\vec{u}}$.
-
-\subsection{Renaming of symbols}
-
-\newcommand\RuleRenPrefix{\bullet}
-\newcommand\RuleRenVar{\RuleNameVar{\RuleRenPrefix}{}}
-\newcommand\RuleRenMVar{\RuleNameMVar{\RuleRenPrefix}{}}
-\newcommand\RuleRenApp{\RuleNameApp{\RuleRenPrefix}}
-\newcommand\RuleRenAbs{\RuleNameAbs{\RuleRenPrefix}{}}
-
-The only place that symbols appear in our calculus is as parameters to operators
-(unlike variables, symbols are not terms). Therefore, the functorial action of
-the operator sheaf can be lifted into terms by recursion on their
-structure, via a pair of judgments $\IsRenaming{\varrho}{M}{N}$ and
-$\IsRenaming{\varrho}{E}{F}$, presupposing $\Of{\varrho}{\Upsilon\inj\Upsilon'}$,
-and $\IsAbt{\Theta}{\Upsilon}{\Gamma}{M}{\tau}$ and
-$\IsAbt{\Theta}{\Upsilon}{\Gamma}{E}{v}$ respectively:
-
-\begin{gather*}
-  \infer[\RuleRenVar]{
-    \IsRenaming{\varrho}{x}{x}
-  }{
-  }
-\\[6pt]
-  \infer[\RuleRenMVar]{
-    \IsRenaming{\varrho}{
-      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
-    }{
-      \MApp{\MV{m}}{\vec{v}}{N_0,\dots,N_n}
-    }
-  }{
-    \Match{\varrho(\vec{u})}{\vec{v}}
-&
-    \HypJ{
-      \IsRenaming{\varrho}{M_i}{N_i}
-    }{
-      \Leq{i}{n}
-    }
-  }
-\\[6pt]
-  \infer[\RuleRenApp]{
-    \IsRenaming{\varrho}{
-      \App{\vartheta}{E_0,\dots,E_n}
-    }{
-      \App{\Rename{\varrho}{\vartheta}}{F_0,\dots,F_n}
-    }
-  }{
-    \HypJ{
-      \IsRenaming{\varrho}{E_i}{F_i}
-    }{
-      \Leq{i}{n}
-    }
-  }
-\\[6pt]
-  \infer[\RuleRenAbs]{
-    \IsRenaming{\varrho}{
-      \MkAbs{\vec{u}}{\vec{x}}{M}
-    }{
-      \MkAbs{\vec{u}}{\vec{x}}{N}
-    }
-  }{
-    \IsRenaming{\varrho\setminus\vec{u}}{M}{N}
-  }
-\end{gather*}
-
-Above, the notation $\varrho\setminus\vec{u}$ means the omission of the
-variables $\vec{u}$ from the renaming $\varrho$. Because terms are identified up
-to $\alpha$-equivalence, the renaming judgment is functional in its input, and
-so we are justified in writing $\Rename{\varrho}{M}$ for $N$ when
-$\IsRenaming{\varrho}{M}{N}$.
-
-\subsection{Substitution of variables}
-
-\newcommand\RuleSubstPrefix{/}
-\newcommand\RuleSubstVar[1]{\RuleNameVar{\RuleSubstPrefix}{#1}}
-\newcommand\RuleSubstMVar{\RuleNameMVar{\RuleSubstPrefix}{}}
-\newcommand\RuleSubstApp{\RuleNameApp{\RuleSubstPrefix}}
-\newcommand\RuleSubstAbs[1]{\RuleNameAbs{\RuleSubstPrefix}{#1}}
-
-Variable substitution in abts is defined inductively by a pair of judgments,
-$\IsSubst{N}{x}{M}{M'}$ and $\IsSubst{N}{x}{E}{F}$:
-
-\begin{gather*}
-  \infer[\RuleSubstVar{1}]{
-    \IsSubst{N}{x}{y}{N}
-  }{
-    \IsEq{x}{y}
-  }
-  \qquad
-  \infer[\RuleSubstVar{2}]{
-      \IsSubst{N}{x}{y}{y}
-    }{
-      \IsApart{x}{y}
-    }
-\\[6pt]
-  \infer[\RuleSubstMVar]{
-    \IsSubst{N}{x}{
-      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
-    }{
-      \MApp{\MV{m}}{\vec{u}}{M'_0,\dots,M'_n}
-    }
-  }{
-    \HypJ{
-      \IsSubst{N}{x}{M_i}{M'_i}
-    }{
-      \Leq{i}{n}
-    }
-  }
-\\[6pt]
-  \infer[\RuleSubstApp]{
-    \IsSubst{N}{x}{
-      \App{\vartheta}{E_0,\dots,E_n}
-    }{
-      \App{\vartheta}{F_0,\dots,F_n}
-    }
-  }{
-    \HypJ{
-      \IsSubst{N}{x}{E_i}{F_i}
-    }{
-      \Leq{i}{n}
-    }
-  }
-\end{gather*}
-\begin{gather*}
-  \infer[\RuleSubstAbs{1}]{
-    \IsSubst{N}{x}{
-      \MkAbs{\vec{u}}{\vec{y}}{M}
-    }{
-      \MkAbs{\vec{u}}{\vec{y}}{M'}
-    }
-  }{
-    \NotIn{x}{\vec{y}}
-&
-    \IsApart{\vec{u}}{\FS{N}}
-&
-    \IsApart{\vec{y}}{\FV{N}}
-&
-    \IsSubst{N}{x}{M}{M'}
-  }
-\\[6pt]
-  \infer[\RuleSubstAbs{2}]{
-    \IsSubst{N}{x}{
-      \MkAbs{\vec{u}}{\vec{y}}{M}
-    }{
-      \MkAbs{\vec{u}}{\vec{y}}{M}
-    }
-  }{
-    \Member{x}{\vec{y}}
-&
-    \IsApart{\vec{u}}{\FS{N}}
-&
-    \IsApart{\vec{y}}{\FV{N}}
-  }
-\end{gather*}
-
-In the remainder, we will write $\Subst{N}{x}{M}$ for $M'$ when
-$\IsSubst{N}{x}{M}{M'}$, and $\Subst{\vec{N}}{\vec{x}}{M}$ for the simultaneous
-substitution of $\vec{N}$ for $\vec{x}$ in $M$.
-
-\subsection{Substitution of metavariables}
-
-\newcommand\RuleMSubstPrefix{/}
-\newcommand\RuleMSubstSuffix{\texttt{meta}}
-\newcommand\RuleMSubstVar{\RuleNameVar{\RuleMSubstPrefix}{}^\RuleMSubstSuffix}
-\newcommand\RuleMSubstMVar[1]{\RuleNameMVar{\RuleMSubstPrefix}{#1}^\RuleMSubstSuffix}
-\newcommand\RuleMSubstApp{\RuleNameApp{\RuleMSubstPrefix}^\RuleMSubstSuffix}
-\newcommand\RuleMSubstAbs{\RuleNameAbs{\RuleMSubstPrefix}{}^\RuleMSubstSuffix}
-
-Metavariables are substituted by abstractions; since a metavariable may only
-appear in an application expression $\MApp{\MV{m}}{\cdots}{\cdots}$, we will
-instantiate the abstraction at the supplied parameters and arguments.
-Substitution for metavariables is defined inductively by the judgments
-$\IsSubst{E}{\MV{m}}{M}{N}$ and $\IsSubst{E}{\MV{m}}{F}{F'}$:
-
-\begin{gather*}
-  \infer[\RuleMSubstVar]{
-    \IsSubst{E}{\MV{m}}{x}{x}
-  }{
-  }
-\\[6pt]
-  \infer[\RuleMSubstMVar{1}]{
-    \IsSubst{E}{\MV{m}}{
-      \MApp{\MV{n}}{\vec{u}}{M_0,\dots,M_n}
-    }{
-      \MApp{\MV{n}}{\vec{u}}{N_0,\dots,N_n}
-    }
-  }{
-    \IsApart{\MV{m}}{\MV{n}}
-&
-    \HypJ{
-      \IsSubst{E}{\MV{n}}{M_i}{N_i}
-    }{
-      \Leq{i}{n}
-    }
-  }
-\\[6pt]
-  \infer[\RuleMSubstMVar{2}]{
-    \IsSubst{
-      \MkAbs{\vec{u}}{\vec{x}}{N}
-    }{\MV{m}}{
-      \MApp{\MV{n}}{\vec{v}}{\vec{M}}
-    }{
-      N''
-    }
-  }{
-    \IsEq{\MV{m}}{\MV{n}}
-&
-    \IsSubst{\vec{M}}{\vec{x}}{N}{N'}
-&
-    \IsRenaming{\vec{u}\mapsto\vec{v}}{N'}{N''}
-  }\\[6pt]
-  %
-  \infer[\RuleMSubstApp]{
-    \IsSubst{E}{\MV{m}}{
-      \App{\vartheta}{F_0,\dots,F_n}
-    }{
-      \App{\vartheta}{F'_0,\dots,F'_n}
-    }
-  }{
-    \HypJ{
-      \IsSubst{E}{\MV{m}}{F_i}{F'_i}
-    }{
-      \Leq{i}{n}
-    }
-  }
-\\[6pt]
-  \infer[\RuleMSubstAbs]{
-    \IsSubst{E}{\MV{m}}{
-      \MkAbs{\vec{u}}{\vec{x}}{M}
-    }{
-      \MkAbs{\vec{u}}{\vec{x}}{N}
-    }
-  }{
-    \IsApart{\vec{u}}{\FS{E}}
-&
-    \IsApart{\vec{x}}{\FV{E}}
-&
-    \IsSubst{E}{\MV{m}}{M}{N}
-  }
-\end{gather*}
-
-The rules for metavariable substitution are reminiscent of the ``hereditary
-substitutions'' first devised for the Concurrent Logical Framework by Watkins
-in \cite{watkins:2004}.
-%
-As usual, we will write $\Subst{E}{\MV{m}}{M}$ for $N$ when
-$\IsSubst{E}{\MV{m}}{M}{N}$.
-
-\section{Model Theory}
-
-\newcommand\SymPsh{\mathbf{S}}
-\newcommand\VarPsh{\mathbf{V}}
-
-Let $\Define{\HypCat}{\OpCat{(\SCtx\times\Ctx)}}$; then we fix the functor
-category $\widehat{\HypCat}^\Sorts$ as our semantic universe, using the
-notation of the French school,
-$\Define{\widehat{\mathcal{C}}}{\Sets^{\OpCat{\mathcal{C}}}}$.
-%
-Let $\Define{\VarPsh_\tau(\Upsilon\parallel\Gamma)}{\MkSet{x\in\Dom\Gamma\mid
-    \Gamma(x)\equiv\tau}}$ be called the presheaf of variables; additionally, we have
-a presheaf of symbols
-$\Define{\SymPsh_\tau(\Upsilon\parallel\Gamma)}{\MkSet{u\in\Dom\Upsilon\mid
-    \Upsilon(u)\equiv\tau}}$. Lastly, we have the presheaf of operators with arity
-$a$,
-$\Define{\Operators_a(\Upsilon\parallel\Gamma)}{\Operators_a(\Upsilon)}$
-
-\subsection{Substitution monoidal structures}
-
-For an object $\Of{P}{\widehat{\HypCat}^\Sorts}$, we will use the notation
-$P^{[\Gamma]}$ to mean $\prod_{x\in\Dom\Gamma}P_{\Gamma(x)}$; likewise,
-$\SymPsh^{\{\Upsilon\}}$ shall mean $\prod_{u\in\Dom\Upsilon}S_{\Upsilon(u)}$. For a
-presheaf $\Of{A}{\widehat{\HypCat}}$ and a sort-indexed family of presheaves
-$\Of{P}{\widehat{\HypCat}^\Sorts}$, we have an operation $A\bullet P$, defined
-as a coend in the following way:
-\[
-  \Define{(A\bullet P)(\Upsilon\parallel\Gamma)}{
-    \int^{(\Upsilon'\parallel\Delta)\in\HypCat}
-      A(\Upsilon'\parallel\Delta)
-      \times \SymPsh^{\{\Upsilon'\}}(\Upsilon\parallel\Gamma)
-      \times P^{[\Delta]}(\Upsilon\parallel\Gamma)
-  }
-\]
-
-Intuitively, the coend construction $A\bullet P$ provides the data of a
-suspended substitution of an $A$-term's variables by $P$-terms.
-%
-Using this, we can define a tensor $P\odot Q$ for
-$\Of{P,Q}{\widehat{\HypCat}^\Sorts}$ as follows:
-\[
-  \HypJ{
-    \Define{{(P\odot Q)}_{\tau}}{
-      P_\tau\bullet Q
-    }
-\quad
-  }{
-    \Member{\tau}{\Sorts}
-  }
-\]
-
-Then, $V$ is the unit to this tensor. We will say that an object
-$\Of{P}{\widehat{\HypCat}^\Sorts}$ is a $\Sigma$-monoid in case it is equipped
-with the following natural transformations where $\nu$ embeds variables into $P$
-and $\varsigma$ equips $P$ with an operation for simultaneous substitutions of
-variables. Furthermore, $\nu$ and $\varsigma$ induce maps ${\nu}_{\Gamma}$ and
-${\varsigma}^{\tau}_{\Upsilon\parallel\Gamma}$:
-\begin{gather*}
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{V}} \arrow[r, "\IMode\nu"]
-    \&
-    {\IMode{P}}
-    \&
-    {\IMode{P\odot P}}
-      \arrow[l, swap, "\IMode\varsigma"]
-  \end{tikzcd}\\[6pt]
-  %
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{{V}^{[\Gamma]}}}
-      \arrow[r, "\IMode{{\nu}_{\Gamma}}"]
-    \&
-    {\IMode{{P}^{[\Gamma]}}}
-  \end{tikzcd}
-  \qquad
-  \begin{tikzcd}[cramped, sep = large, ampersand replacement = \&]
-    {\IMode{{P_\tau^{\Yoneda{\Upsilon\parallel\Gamma}}\times \SymPsh^{\{\Upsilon\}}\times P^{[\Gamma]}}}}
-      \arrow[r, "\IMode{\varsigma^\tau_{\Upsilon\parallel\Gamma}}"]
-    \&
-    {\IMode{P_\tau}}
-  \end{tikzcd}
-\end{gather*}
-
-\subsection{The signature endofunctor and its algebras}
-
-For each signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we have an
-endofunctor
-$\Of{\SigFun{\Sigma}}{\widehat{\HypCat}^\Sorts\ra\widehat{\HypCat}^\Sorts}$,
-which is defined as follows:
-\begin{align*}
-  \IMode{{\SigFun{\Sigma}(X)}_{\tau}}
-&
-    \triangleq
-    \OMode{
-      \coprod_{
-        \vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}
-      }
-      \prod_{
-        \MkValence
-          {\vec{\sigma}}
-          {\vec{\tau}}
-          {\tau_i}
-        \in\vec{v}
-      }
-      X_{\tau_i}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}
-    }
-\end{align*}
-
-Then, a $\Sigma$-model is a $\Sigma$-monoid $P$ which is equipped with an
-algebra $\Of{\alpha}{\SigFun{\Sigma}(P)\ra P}$, which shall interpret
-applications of each operator.
-
-\subsection{Interpretation of terms}
-
-The metavariable, symbol and variable contexts are interpreted for a model $P$ as
-an environment presheaf in the following way:
-\[
-  \Define{
-    \SemBrackets{
-      \Theta\triangleright\Upsilon\parallel\Gamma
-    }_P
-  }{
-    \left(
-      \prod_{
-        (m:%
-        \MkValence
-          {\vec{\sigma}}
-          {\vec{\tau}}
-          {\tau})
-        \in\Theta
-      }{{P_{\tau}}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}}
-    \right)
-    \times
-    \SymPsh^{\{\Upsilon\}}
-    \times
-    \VarPsh^{[\Gamma]}
-  }
-\]
-
-Then, the interpretation of a term in a model $P$ is a map from its environment
-to $P$:
-\[
-  \Of{
-    \SemBrackets{
-      \Theta\triangleright\Upsilon\parallel\Gamma\vdash M:\tau
-    }_{P}
-  }{
-    \SemBrackets{
-      \Theta\triangleright\Upsilon\parallel\Gamma
-    }_{P}\ra {P}_{\tau}
-  }
-\]
-
-Variables are interpreted by the map
-${\SemBrackets{
-  \Theta\triangleright\Upsilon\parallel\Gamma\vdash x:\tau
-}}_{P}$ which
-projects them from the environment and embeds them into the model. Metavariables
-are resolved by the map
-${\SemBrackets{
-  \Theta\triangleright\Upsilon\parallel\Gamma\vdash\MApp{\MV{m}}{\vec{u}}{\vec{M}}
-  : \tau
-}}_{P}$ (where
-$\Lookup{\Theta}{\MV{m}}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$) which
-projects their interpretation from the environment and instantiates it via
-substitution:
-\[
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{P_\tau}}
-\&[8.0em]
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-    }}
-      \arrow[
-        , rr
-        , dashed
-        , "
-          \OMode{
-            {\SemBrackets{
-              \Theta\triangleright\Upsilon\parallel\Gamma
-              \vdash
-              x:\tau
-            }}_{P}
-          }
-          "
-      ]
-      \ar{drr}[
-        , rotate = -16.5
-        , xshift = +1.5em
-        , yshift = -0.2em
-        , swap
-      ]{
-        {\nu}_{\Gamma} {\pi}_{3}
-      }
-      \ar{dl}[
-        , rotate = +11.5
-        , xshift = -2.75em
-      ]{
-        \IMode{
-          \Tuple{
-            \pi_\MV{m}\,
-            \pi_1,
-            \phi\,
-            \pi_2,
-            \psi
-          }
-        }
-      }
-      \arrow[
-        , l
-        , dashed
-        , swap
-        , "
-          \OMode{
-            {\SemBrackets{
-              \Theta\triangleright\Upsilon\parallel\Gamma
-              \vdash
-              \MApp{\MV{m}}{\vec{u}}{\vec{M}}: \tau
-            }}_{P}
-          }"
-      ]
-\&[-2.0em]
-    \null
-\&[+3.0em]
-    {\IMode{P_\tau}}
-\\
-    {\IMode{
-      {P}_{\tau}^{
-        \Yoneda{\vec{\sigma}\parallel\vec{\tau}}
-      }
-      \times \SymPsh^{\{\vec{\sigma}\}}
-      \times P^{[\vec{\tau}]}
-    }}
-      \arrow[
-        , u
-        , "
-          \IMode{
-            \varsigma^\tau_{\vec{\sigma}\parallel\vec{\tau}}
-          }
-          "
-        ]
-\&
-    \null
-\&
-    \null
-\&
-    {\IMode{P^{[\Gamma]}}}
-      \arrow[
-        , u
-        , swap
-        , "{\pi}_{x}"
-      ]
-\\
-    {\IMode{P^{[\vec{\tau}]}}}
-\&
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-    }}
-      \arrow[
-        , dashed
-        , l
-        , swap
-        , "
-          \Define{
-            \psi\,
-          }{
-            \,\Tuple{
-              {\SemBrackets{
-                \Theta\triangleright\Upsilon\parallel\Gamma\vdash M:\tau
-              }}_{P}
-            }_{(M,\tau)\in (\vec{M},\vec{\tau})}
-          }
-          "]
-\&
-    {\IMode{\SymPsh^{\{\Upsilon\}}}}
-      \arrow[
-        , dashed
-        , r
-        , "\Define{\phi\,}{\,\Tuple{\pi_u}_{u\in\vec{u}}}"
-      ]
-\&
-    {\IMode{\SymPsh^{\{\vec{\sigma}\}}}}
-  \end{tikzcd}
-\]
-
-Interpretation of operator applications is the most complicated. Recall that,
-unlike in standard treatments of universal algebra, our operators are indexed by
-symbol collections; therefore, operators must pass through suitable renamings in
-order to be used in the interpretation. Let us begin by constructing for each
-operator $\IsOperator{\Upsilon}{\vartheta}{a}$ the morphism
-$\SemBrackets{\vartheta}_P$ which shall rename the parameters of the operator
-using the environment:
-\[
-  \begin{tikzcd}[cramped, ampersand replacement = \&, sep = huge]
-    |[alias = Src]|
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-    }}
-      \arrow[
-        , r
-        , swap
-        , "\IMode{\pi_2}"
-      ]
-\&
-    {\IMode{\SymPsh^{\{\Upsilon\}}}}
-      \arrow[
-        , r
-        , swap
-        , "\IMode{\Operators_a(\pi_{(-)})(\vartheta)}"
-      ]
-\&[1.5em]
-    |[alias = Tgt]|
-    {\IMode{\Operators_a}}
-      \arrow[
-        , from = Src
-        , to = Tgt
-        , bend left
-        , dashed
-        , "\OMode{\SemBrackets{\vartheta}_{P}}"
-      ]
-  \end{tikzcd}
-\]
-
-We will proceed using the $\SigFun{\Sigma}$-algebra $\alpha$, as
-follows, by composing it with a morphism $\beta$ from the environment into
-the signature endofunctor, which interprets the syntax of operator applications:
-\[
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{
-      {\SemBrackets{\Theta\triangleright\Upsilon\parallel\Gamma}}_{P}
-    }}
-      \arrow[r, "\IMode{\beta}"]
-\&
-    {\IMode{
-      {\SigFun{\Sigma}(P)}_{\tau}
-    }}
-      \arrow[r, "\IMode{\alpha_\tau}"]
-\&
-    {\IMode{P_\tau}}
-  \end{tikzcd}
-\]
-
-The construction of $\beta$ proceeds by renaming the parameters of the operator
-$\vartheta$ and constructing the (bound) exponentiated arguments $\gamma$ of the
-operator.
-\[
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-    }}
-      \arrow[
-        , r
-        , dashed
-        , "
-          \Define{
-            \beta\,
-          }{
-            \,\Tuple{
-              \SemBrackets{\vartheta}_P, \lambda\gamma
-            }
-          }
-          "
-      ]
-\&[4.0em]
-    {\IMode{
-      \coprod_{
-        \vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}
-      }
-      \prod_{
-        \MkValence{\vec{\sigma}_i}{\vec{\tau}_i}{\tau_i}\in\vec{v}
-      }
-      P_{\tau_i}^{\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}}
-    }}
-  \end{tikzcd}
-\]
-Arguments $\lambda\IMode{{\gamma}_{i}}$ are the exponential transposes (curried
-form) of the composites $\IMode{{\gamma}_{i}}$:
-\[
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-    }}
-      \arrow[
-        , dashed
-        , d
-        , swap
-        , "\OMode{\lambda{\gamma}_{i}}"
-      ]
-\&
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-      \times
-      \Yoneda{
-        \vec{\sigma}_{i}\parallel\vec{\tau}_{i}}
-      }
-    }
-      \arrow[
-        , r
-        , "
-          \IMode{
-            \Tuple{
-              {\pi}_{[1,1]},
-              {\phi}_{i},
-              {\psi}_{i}
-            }
-          }
-          "
-        ]
-      \arrow[
-        , dashed
-        , d
-        , swap
-        , "\OMode{{\gamma}_{i}}"
-      ]
-\&[3.5em]
-    {\IMode{
-      \SemBrackets{
-        \Theta\triangleright\Upsilon,
-        \vec{\sigma}_i\parallel\Gamma,
-        \vec{\tau}_{i}
-      }_{P}
-    }}
-      \ar{dl}[
-        , rotate = +19.25
-        , xshift = -4.5em
-        , yshift = -0.2em
-      ]{\IMode{\SemBrackets{\Theta\triangleright\Upsilon,\vec{\sigma}_{i}\parallel\Gamma,\vec{\tau}_{i}\vdash M_i:\tau_i}_P}}
-\\[2.0em]
-    {\IMode{{P}_{{\tau}_{i}}^{\Yoneda{\vec{\sigma}_{i}\parallel\vec{\tau}_{i}}}}}
-\&
-    {\IMode{P_{{\tau}_{i}}}}
-\&
-  \end{tikzcd}
-\]
-where $\phi_i,\psi_i$ are defined as follows:
-\begin{gather*}
-  \begin{tikzcd}[cramped, ampersand replacement = \&]
-    {\IMode{
-      {\SemBrackets{\Theta\triangleright\Upsilon\parallel\Gamma}}_{P}
-      \times
-      \Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}
-    }}
-      \arrow[
-        , r
-        , dashed
-        , "
-          \Define
-            {
-              \phi_i\,
-            }
-            {
-              \,
-                \Tuple{\pi_u\circ\pi_1}_{u\in\Dom{\Upsilon}}
-              ,
-                \,\Tuple{\pi_u\circ\pi_{[2,1]}}_{u\in\Dom{\vec{\sigma}_i}}
-            }
-          "
-      ]
-\&[12.0em]
-    {\IMode{\SymPsh^{\{\Upsilon,\vec{\sigma}_i\}}}}
-\\[-1.0em]
-    {\IMode{
-      {\SemBrackets{
-        \Theta\triangleright\Upsilon\parallel\Gamma
-      }}_{P}
-      \times
-      \Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}
-    }}
-      \arrow[
-        , r
-        , dashed
-        , "
-          \Define{
-            \psi_i\,
-          }{
-            \,
-              \Tuple{\pi_x\circ\pi_1}_{x\in\Dom{\Gamma}}
-            ,
-              \,\Tuple{\pi_x\circ\pi_{[2,1]}}_{x\in\Dom{\vec{\tau}_i}}
-          }
-          "
-      ]
-\&[12.0em]
-    {\IMode{\VarPsh^{[\Gamma,\vec{\tau}_i]}}}
-  \end{tikzcd}
-\end{gather*}
-
-This concludes the interpretation of well-sorted terms into any $\Sigma$-model.
-
-\section{Case Study: Wellformed Sequents}
-
-The representation of telescopes and sequents in a logical framework is
-notoriously difficult; whilst it is possible to use higher-order abstract
-syntax or abts to encode the binding-structure of telescopes and sequents, the
-encoding is sufficiently laborious and obscure that it is not used in practice.
-Crary has demonstrated a first-order encoding of contexts in the Edinburgh
-Logical Framework (ELF) in bijection with actual LF-contexts~\cite{crary:2009},
-which has been successfully used in large-scale mechanization efforts,
-including that of Standard ML~\cite{lee-crary-harper:2007} and the ELF
-itself~\cite{martens-crary:2012}.
-
-We will approach the problem of encoding telescopes and sequents from the
-\emph{refinements} perspective, where a conservative approximation of the
-grammar is first given using the abt logical framework, and then the
-correctness of a code is expressed separately in a judgment that refines the
-existing specification.  Because we have not committed to using the built-in
-binding machinery to express the well-scopedness of telescopes and sequents, we
-are free to use \emph{symbols} in order to model the variables in the context.
-This is in fact quite sensible if we are actually trying to faithfully
-represent the syntax of telescopes and sequents, rather than replace them with
-their counterparts on the meta-level.
-
-The approach outlined above is actually quite similar to the standard practice
-in the Abella proof assistant \cite{gacek:2008}; in Abella, in addition to the
-canonical forms, each type is inhabited by an infinite collection of nominal
-constants or symbols, which can be used in the syntax of contexts. In order to
-specify where such symbols are allowed to occur, a wellformedness predicate (a
-``scheme'') is separately defined. Our setting is similar in that we may also
-use symbols to encode the syntax of contexts, except that here, we have better
-control over their proliferation because, contrary to the state of affairs in
-Abella, symbols are \emph{not} expressions; following
-Harper~\cite{harper:2016}, symbols are meant only to serve as an open-ended
-indexing scheme for operators, not as actual expressions.
-
-There are two reasons that one might want to model sequents in a (meta)-logical
-framework. Perhaps the most obvious reason is to prove things about a logic
-with sequent judgments; this is what motivated Crary's encoding of contexts in
-LF, for instance. However, another common use-case is the practical development
-of a proof assistant in which users can define new sequent rules; such
-user-defined rules may only be compiled into well-behaved tactics in the proof
-system in case they obey certain wellformedness properties including (among
-other things) correct scoping of names. MetaPRL is a salient example of such a
-proof assistant, whose implementation might have benefited from the more
-principled treatment of symbols and metavariables which we present here.
-
-This insight leads the way to a simple abt signature for the theory of
-telescopes and sequents.
-%
-\newcommand\SortTele{\mathtt{tele}}
-\newcommand\SortJdg{\mathtt{jdg}}
-\newcommand\SortProp{\mathtt{prop}}
-\newcommand\SortExpr{\mathtt{exp}}
-\newcommand\OpNil{\mathtt{nil}}
-\newcommand\OpSnoc[1]{\mathtt{snoc}[#1]}
-\newcommand\OpHyp[1]{\mathtt{hyp}[#1]}
-\newcommand\OpSequent{\mathtt{sequent}}
-% %
-\[
-  \begin{array}{lr}
-    \IsSort{\SortTele}&\textit{telescopes}\\
-    \IsSort{\SortExpr}&\textit{expressions}\\
-    \IsSort{\SortProp}&\textit{propositions}\\
-    \IsSort{\SortJdg}&\textit{judgments}\\\\
-    \IsOperator{\Upsilon, u:\SortExpr}{\OpHyp{u}}{
-      \MkArity{}{\SortExpr}
-    }\\\\
-    \IsOperator{\Upsilon}{\OpNil}{
-      \MkArity{}{\SortTele}
-    }\\
-    \IsOperator{\Upsilon, u:\SortExpr}{\OpSnoc{u}}{
-      \MkArity{
-        \MkValence{}{}{\SortTele},
-        \MkValence{}{}{\SortProp}
-      }{\SortTele}
-    }\\\\
-    \IsOperator{\Upsilon}{\OpSequent}{
-      \MkArity{
-        \MkValence{}{}{\SortTele},
-        \MkValence{}{}{\SortProp}
-      }{\SortJdg}
-    }
-  \end{array}
-\]
-
-Suppose we have encoded a fragment of type theory as well:
-\[
-  \begin{array}{l}
-    \IsOperator{\Upsilon}{\mathtt{P}}{
-      \MkArity{}{\SortProp}
-    }
-\\
-    \IsOperator{\Upsilon}{\mathtt{pred}}{
-      \MkArity{
-        \MkValence{}{}{\SortExp}
-      }{\SortProp}
-    }
-\\
-  \end{array}
-\]
-
-Terms written using the abstract syntax will be difficult to read, so let us
-define some notation:
-\begin{align*}
-  \IMode{\diamond}
-&
-  \triangleq \OMode{\OpNil}
-\\
-  \IMode{H, u:P}
-&
-  \triangleq \OMode{
-    \App{\OpSnoc{u}}{H, P}
-  }
-\\
-  \IMode{H\gg A}
-&
-  \triangleq \OMode{
-    \App{\OpSequent}{
-      H, A
-    }
-  }
-\\
-  \IMode{`u}
-&
-  \triangleq \OMode{\OpHyp{u}}
-\end{align*}
-
-Now, we can write the following sequent:
-\[
-  \IsAbt{\cdot}{
-    u:\SortExp,
-    v:\SortExp
-  }{\cdot}{
-    \diamond,u:\mathtt{P},v:\App{\mathtt{pred}}{`u}
-    \gg
-    \App{\mathtt{pred}}{`u}
-  }{\SortJdg}
-\]
-
-\newcommand\OpNabla[1]{\nabla[#1]}
-
-The above sequent has free symbols, but we can close over them by adding a form
-of parametric higher-order judgment (as in \cite{harper:2016}) to our object
-language, indexed by a collection of sorts $\vec{\sigma}$:
-\[
-  \IsOperator{\Upsilon}{\OpNabla{\vec{\sigma}}}{
-    \MkArity{
-      \MkValence{\vec{\sigma}}{}{\SortJdg}
-    }{\SortJdg}
-  }
-\]
-
-Then, we may write a closed sequent judgment as follows:
-\[
-  \IsAbt{\cdot}{\cdot}{\cdot}{
-    \App{\OpNabla{\SortExp,\SortExp}}{
-      \MkAbs{u,v}{}{
-        \diamond,u:\mathtt{P},v:\App{\mathtt{pred}}{`u}
-        \gg
-        \App{\mathtt{pred}}{`u}
-      }
-    }
-  }{\SortJdg}
-\]
-
-\subsection{Refinements for wellformedness}
-
-Having specified an approximation of the grammar of telescopes and sequents in
-the abt logical framework, we can proceed to define proper wellformedness via
-\emph{inductive refinement}~\cite{harper:2016}. The basic idea is to introduce a
-new form of (meta)-judgment $\IsWf{\Upsilon}{\Gamma}{M}{\tau}$ which
-expresses the extrinsic wellformedness properties we wish to verify,
-presupposing $\IsAbt{\cdot}{\Upsilon}{\Gamma}{M}{\tau}$. Additionally, we
-introduce an analogous judgment on abstractions,
-$\IsWf{\Upsilon}{\Gamma}{E}{v}$ presupposing
-$\IsAbt{\cdot}{\Upsilon}{\Gamma}{E}{v}$, defined uniformly as follows:
-\[
-  \infer{
-    \IsWf{\Upsilon}{\Gamma}{
-      \MkAbs{\vec{u}}{\vec{x}}{M}
-    }{
-      \MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}
-    }
-  }{
-    \IsWf{\Upsilon,\vec{u}:\vec{\sigma}}{\Gamma,\vec{x}:\vec{\tau}}{M}{\tau}
-  }
-\]
-
-Likewise, wellformedness for variables is defined uniformly:
-\[
-  \infer{
-    \IsWf{\Upsilon}{\Gamma}{x}{\tau}
-  }{
-  }
-\]
-Note that the refinement for variables $x$ is not trivial, since it is only
-defined in case the presupposition $\IsAbt{\cdot}{\Upsilon}{\Gamma}{x}{\tau}$
-is satisfied.
-
-The remainder of the definition of refinement proceeds by induction on sorts and
-operators. For the sake of this example, we will just stipulate that anything of
-sort $\SortExp$ or $\SortProp$ is grammatical if its subterms are grammatical:
-
-\[
-  \infer[
-    \text{for $\Member{\tau}{\MkSet{\SortExp,\SortProp}}$}
-  ]{
-    \IsWf{\Upsilon}{\Gamma}{
-      \App{\vartheta}{E_0,\dots,E_n}
-    }{\tau}
-  }{
-    \begin{array}{l}
-      \IsOperator{\Upsilon}{\vartheta}{
-        \MkArity{
-          v_0,\dots,v_n
-        }{\tau}
-      }
-\\
-      \HypJ{
-        \IsWf{\Upsilon}{\Gamma}{E_i}{v_i}
-      }{
-        \Leq{i}{n}
-      }
-    \end{array}
-  }
-\]
-
-The refinements for parametric judgment and sequents simply delegate to their
-subterms as well:
-\begin{gather*}
-  \infer{
-    \IsWf{\Upsilon}{\Gamma}{
-      \App{\OpNabla{\vec{\sigma}}}{
-        \MkAbs{\vec{u}}{}{J}
-      }
-    }{\SortJdg}
-  }{
-    \IsWf{\Upsilon,\vec{u}:\vec{\sigma}}{\Gamma}{J}{\SortJdg}
-  }
-\\[6pt]
-  \infer{
-    \IsWf{\Upsilon}{\Gamma}{
-      H\gg A
-    }{\SortJdg}
-  }{
-    \IsWf{\Upsilon}{\Gamma}{H}{\SortTele}
-&
-    \IsWf{\Upsilon}{\Gamma}{A}{\SortProp}
-  }
-\end{gather*}
-
-The refinement for telescopes proceeds by induction:
-\begin{gather*}
-  \infer{
-    \IsWf{\Upsilon}{\Gamma}{\diamond}{\SortTele}
-  }{
-  }
-\qquad
-  \infer{
-    \IsWf{\Upsilon}{\Gamma}{H,u:A}{\SortTele}
-  }{
-    \begin{array}{l}
-      \IsWf{\Upsilon\setminus\MkSet{u}}{\Gamma}{H}{\SortTele}
-\\
-      \IsWf{\Upsilon\setminus\MkSet{u}}{\Gamma}{A}{\SortProp}
-    \end{array}
-  }
-\end{gather*}
-
-Based on these rules, it is easy to show that the example sequent from the
-previous section is wellformed:
-\[
-  \IsWf{\cdot}{\cdot}{
-    \App{\OpNabla{\SortExp,\SortExp}}{
-      \MkAbs{u,v}{}{
-        \diamond,u:\mathtt{P},v:\App{\mathtt{pred}}{`u}
-        \gg
-        \App{\mathtt{pred}}{`u}
-      }
-    }
-  }{\SortJdg}
-\]
-
-\begin{remark}
-  In the full development of a theory of behavioral refinements, the ad-hoc
-  judgment $\IsWf{\Upsilon}{\Gamma}{M}{\tau}$ would be replaced with a general
-  notion of refinement $\IMode{\phi}\sqsubset\tau$ ($\phi$ refines sort
-  $\tau$), and a new judgment $\IMode{\Upsilon}\parallel\IMode{\Phi}\vDash
-  \IMode{M} \in \IMode{\phi}$, such that $\phi$ refines $\tau$ and the
-  refinement context $\Phi$ refines some sort context $\Gamma$. Furthermore, in
-  addition to restricting membership, refinements might also be designed to
-  coarsen equivalence as in \cite{denney:1998}.
-\end{remark}
-
-\section*{Acknowledgements}
-
-The first author wishes to thank Robert Harper for numerous conversations about
-abstract binding trees and symbolic parameters, and for his feedback on a draft
-of this paper; and Andy Pitts for his help in understanding multi-sorted
-nominal sets and atomic sheaves.
+\input{body}
 
@@
 \appendixpage
-
-\section{$\SCtxSite$, Nominal Sets, and the Schanuel Topos}
-
-Pitts defines a category $\mathbf{Nom}$ of nominal sets in \cite{pitts:2013}
-which function equivalently to the sheaves that we considered above, in the
-case of $\IMode{\Sorts}\cong\IMode{\mathbb{1}}$. Fix a countably infinite set
-of atoms $\mathbb{A}$; then let $\Perm\mathbb{A}$ be the group of permutations
-(i.e.\ autoequivalences) on $\mathbb{A}$.  Considered as a category,
-$\Perm\mathbb{A}$ has a single object $\bullet$ with morphisms
-$\Of{\pi}{\bullet\ra\bullet}$ for each $\Member{\pi}{\Dom{\Perm\mathbb{A}}}$.
-Then, the category of nominal sets $\mathbf{Nom}$ is the subcategory of
-$\Sets^{\Perm\mathbb{A}}$ containing just the presheaves which satisfy a
-\emph{finite support} condition.
-
-The category $\mathbf{Nom}$ is equivalent to the category of covariant sheaves
-on $\SymSets$ under the atomic coverage, called the Schanuel Topos
-$\Define{\mathbb{S}}{\Sheaves{\OpCat{\SymSets}}}$
-\cite{pitts:2013,fiore-staton:2006}; equivalently, the Schanuel topos is the
-subcategory of $\Sets^\SymSets$ containing only pullbacks-preserving functors.
-
-In addition to the unisorted nominal sets framework, Pitts also briefly
-discusses a multi-sorted version of the apparatus in which the countably
-infinite set of atoms $\mathbb{A}$ is equipped with a sort assignment
-$\Of{\mathfrak{s}}{\mathbb{A}\to\Sorts}$ such that all the fibers of
-$\mathfrak{s}$ are countably infinite; then, he defines for any such assignment
-the category $\mathbf{Nom}_\mathfrak{s}$ of $\mathfrak{s}$-sorted nominal sets
-as a subcategory of the category of presheaves on the group of
-$\mathfrak{s}$-respecting permutations on $\mathbb{A}$,
-$\Perm_\mathfrak{s}\mathbb{A}$.
-
-In the same way as the Schanuel topos is equivalent to the unisorted nominal
-sets, we expect to find an equivalence between $\Sheaves{\SCtxSite}$ and the
-limit $\int_{\mathfrak{s}}\mathbf{Nom}_\mathfrak{s}$. At the very least, the
-sheaf topos $\Sheaves{\SCtxSite}$ may serve as a multi-sorted generalization of
-the Schanuel topos, for which we may carry over certain useful results,
-including its characterization as the subcategory of $\Sets^\SCtx$ which
-contains just the pullbacks-preserving functors.
-
-\begin{thm}
-  \label{thm:pullback-preservation}
-  A presheaf $\Of{X}{\Sets^{\SCtx}}$ is a sheaf on $\SCtxSite$ just when it
-  preserves pullbacks.
-\end{thm}
-
-\begin{proof}
-
-  The proof is essentially the same as that of the analogous lemma for the
-  Schanuel topos as presented in \cite[A.2.1.11.h]{johnstone:2002}, but we will
-  give a slightly more detailed version here. The presheaf $X$ is a sheaf on the
-  atomic site $\SCtxSite$ just when, for any renaming
-  $\Of{\varrho}{C\inj D}$ and any $\Member{n}{X(D)}$ if
-  $\IsEq{n\cdot \varrho_0}{n\cdot \varrho_1}$ for all diagrams
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{C} \arrow[r, hook, "\IMode{\varrho}"] &
-      \IMode{D}
-        \arrow[r, hook, shift left, "\IMode{\varrho_0}"]
-        \arrow[r, hook, swap, shift right, "\IMode{\varrho_1}"]
-      &
-      \IMode{E}
-    \end{tikzcd}
-  \]
-  such that $\IsEq{\varrho_0\circ\varrho}{\varrho_1\circ\varrho}$, then there
-  exists a unique $\Member{m}{X(C)}$ such that $\IsEq{m\cdot\varrho}{n}$. In other
-  words, the arrow $X(\varrho)$ is an equalizer, as in the following:
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{X(C)} \arrow[r, "\IMode{X(\varrho)}"] &
-      \IMode{X(D)}
-        \arrow[r, shift left, "\IMode{X(\varrho_0)}"]
-        \arrow[r, swap, shift right, "\IMode{X(\varrho_1)}"]&
-      \IMode{X(E)}
-      \\
-      \IMode{\mathbb{1}}
-        \arrow[u, densely dotted, "\OMode{m}"]
-        \arrow[ur, "\IMode{n}"]
-    \end{tikzcd}
-  \]
-
-  Now, because $\Sets$ is a regular category, $X(\varrho)$ is an equalizer
-  precisely when it is a monomorphism. Therefore, if we have assumed that $X$
-  preserves pullbacks, in order to demonstrate that $X$ is a sheaf, it suffices
-  to show that $X(\varrho)$ is monic. If $X$ preserves pullbacks, then it also
-  preserves monomorphisms, because in any category with pullbacks, $\Of{f}{A\ra
-  B}$ is monic just when the square
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{A}
-        \arrow[r, "\IMode{\ArrId{A}}"]
-        \arrow[d, "\IMode{\ArrId{A}}"] &
-      \IMode{A} \arrow[d, "\IMode{f}"]\\
-      \IMode{A} \arrow[r, "\IMode{f}"] &
-      \IMode{B}
-    \end{tikzcd}
-  \]
-  is a pullback \cite[p.\ 16]{mac-lane-moerdijk:1992}. Because $X$ preserves
-  monomorphisms, and all arrows in $\SCtx$ are monic, then $X(\varrho)$ is
-  monic. Hence, $X$ is a sheaf.
-
-  Next, we must show that if $X$ is a sheaf, then it preserves pullbacks; we
-  will loosely track the proof of Proposition~2.6.15 in \cite{biering:2004}.
-  Fix the following diagram such that it is a pullback square in $\SCtx$:
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{C}
-        \arrow[r, "\IMode{f_j}"]
-        \arrow[d, "\IMode{f_i}"] &
-      \IMode{C_j} \arrow[d, "\IMode{s_j}"]\\
-      \IMode{C_i} \arrow[r, "\IMode{s_i}"] &
-      \IMode{D}
-    \end{tikzcd}
-  \]
-
-  Then it suffices to show that the following diagram is also pullback square
-  in $\Sets$:
-  \[
-    \begin{tikzcd}[sep=large]
-      \IMode{X(C)}
-        \arrow[r, "\IMode{X(f_j)}"]
-        \arrow[d, "\IMode{X(f_i)}"] &
-      \IMode{X(C_j)} \arrow[d, "\IMode{X(s_j)}"]\\
-      \IMode{X(C_i)} \arrow[r, "\IMode{X(s_i)}"] &
-      \IMode{X(D)}
-    \end{tikzcd}
-  \]
-
-  We have to show that $X(C)$ is, up to equivalence, the pullback of $X(s_i)$
-  along $X(s_j)$ in $\Sets$, namely:
-  \[
-    \IMode{X(C_i)\times_{X(D)}X(C_j)}
-      \cong
-    \OMode{
-      \MkSet{
-        (m_i, m_j)\in X(C_i)\times X(C_j)
-          \mid
-        m_i \cdot s_i = m_j \cdot s_j
-      }
-    }
-  \]
-  The first direction of the equivalence is trivial: for each
-  $\Member{m}{X(C)}$, we can easily exhibit $\Member{(m\cdot f_i,m\cdot
-  f_j)}{X(C_i)\times_{X(D)}X(C_j)}$.
-
-  For the other direction, fix $\Member{(m_i,
-  m_j)}{X(C_i)\times_{X(D)}X(C_j)}$.  Consider the singleton cover on $C$ that
-  contains $\Define{h}{s_i\circ f_i}=\OMode{s_j\circ f_j}$; a matching family
-  $\phi$ for the cover $\MkSet{h}$ is a map that takes $h$ to an object in
-  $X(D)$, and an amalgamation for $\phi$ consists in an object
-  $\Member{m}{X(C)}$ such that $\IsEq{\phi(h)}{m\cdot h}$. Now recall that $X$
-  is a sheaf on $\SCtxSite$ if and only if every matching family has a unique
-  amalgamation; for the matching family $\Define{\phi(h)}{m_i\cdot
-  s_i}=\OMode{m_j\cdot s_j}$, then, we must have a unique amalgamation
-  $\Member{m}{X(C)}$ since $X$ is a sheaf. Therefore, $X$ preserves pullbacks.
-
-\end{proof}
+\input{appendices}
 
